### 2012-04-26 Paper/paper.tex
initial commit with example to start latex paper

% vim:set textwidth=100:
% vim:set fo+=t:

\documentclass[10pt]{article}

\usepackage{amsmath}
\usepackage[nofiglist,notablist]{endfloat}
\usepackage[usenames,dvipsnames]{color}
\usepackage{color}
\usepackage{authblk}
\usepackage{graphicx}
\usepackage{palatino}
\usepackage[activate={true,nocompatibility},final]{microtype}
\usepackage[super,sort&compress]{natbib}
\pagenumbering{arabic}
\parskip = 0.08in \parindent = 0.0in

% Custom macros for author comments
\newcommand{\Alberto}[1]{\color{ForestGreen}Alberto: #1 \normalcolor }
\newcommand{\Justin}[1]{\color{blue}Justin: #1 \normalcolor}
\newcommand{\Arijit}[1]{\color{yellow}Arijit: #1 \normalcolor}
\newcommand{\Ken}[1]{\color{red}Ken: #1 \normalcolor}

\author{Alberto~Perez$^2$}
\author{Ivet~Bahar$^1$}
\author{Ken~A.~Dill$^2$}
\author{Justin~L.~MacCallum$^2$}
\affil{$^1$Department of Computational and Systems Biology, and Clinical \& Translational Science
    Institute, School of Medicine, University of Pittsburgh, 3064 BST3, 3501 Fifth Ave, Pittsburgh,
    PA 15213}
\affil{$^2$Laufer Center for Physical and Quantitative Biology\\
    Stony Brook University\\
    Stony Brook, NY 11794-5252}

\title{Title}

\begin{document}

\maketitle
\begin{abstract}
%How to cite something in paper.bib
    Abstract \citep{CASP9}

\end{abstract}

\section{Introduction}

%this is how you put a figure in document
\begin{figure} \centering \includegraphics[width=3in]{rmsd_energy_schematic2.pdf}

    \caption{Schematic}

    \label{fig:conf_schematic} 
\end{figure}

\section{Methods}

\subsection{ENM}

%this is how you put a table
\begin{table} 
    \centering
    \caption{Structures used for large motions}
    \label{tbl:targets}
    \begin{tabular}{l l l}
    \hline
    Target                           & PDB1 & PDB2 \\
    \hline
	Scallop myosin                   & 1kk7 & 1kk8 \\
	Acyl Carrier Protein             & 1acp & 2fae \\
	Aspartate Aminotransferase       & 1ama & 8aat \\
	Bcl-xl                           & 1bxl & 1ysn \\
	Calcium Sensor                   & 1k9k & 1k9p \\
	Calmodulin                       & 1cll & 1ctr \\
	Cyclin Dependent Kinase Inhibitor & 1dc2 & 2a5e \\
	Cystatin                         & 1a67 & 1cew \\
	%CytCOxidase                     & 1oco & 2occ \\
	Hydrolase                        & 1qz3 & 1u4n \\
	LacRepressor                     & 1lcc & 1lqc \\
	LambdaCro                        & 5cro & 6cro \\
	Lupin Hydrolase                  & 1f3y & 1jkn \\
	Maltose Binding Protein          & 1omp & 3mbp \\
	Pin1                             & 1f8a & 1pin \\
	%Thrombin                        & 1avg & 1ett \\
        \hline
    \end{tabular}
\end{table}

%This tells latex to reference the figure or table
See figure \ref{fig:conf_schematic}.

See table \ref{tbl:targets}.

\bibliographystyle{achemso} 
\bibliography{paper}

\end{document}

### 2012-09-05 Paper/paper.tex
first draft

@@
 
-\documentclass[10pt]{article}
+%\documentclass[preprint,preprintnumbers,amsmath,amssymb]{revtex4}
+\documentclass[12pt]{article}
 
+%\usepackage{fontspec} 
+%\setmainfont{TeX Gyre Pagella}
 \usepackage{amsmath}
@@
 
-\author{Alberto~Perez$^2$}
-\author{Ivet~Bahar$^1$}
-\author{Ken~A.~Dill$^2$}
-\author{Justin~L.~MacCallum$^2$}
-\affil{$^1$Department of Computational and Systems Biology, and Clinical \& Translational Science
-    Institute, School of Medicine, University of Pittsburgh, 3064 BST3, 3501 Fifth Ave, Pittsburgh,
-    PA 15213}
-\affil{$^2$Laufer Center for Physical and Quantitative Biology\\
+\author{Arijit~Roy$^1$}
+\author{Justin~L.~MacCallum$^1$}
+\author{Alberto~Perez$^1$}
+\author{Ken~A.~Dill$^1$}
+\affil{$^1$Laufer Center for Physical and Quantitative Biology\\
     Stony Brook University\\
-    Stony Brook, NY 11794-5252}
+    Stony Brook, NY 11794-5252.}
 
-\title{Title}
+\title{Ranking protein structures using confinement free energy method}
 
@@
 \begin{abstract}
-%How to cite something in paper.bib
-    Abstract \citep{CASP9}
+%How to cite something in paper.bib Abstract \citep{CASP9}
+
+The calculation of free energy differences is of central importance in the simulation of biochemical systems.  The computation of the 
+free energy difference between pairs of macromolecule with large conformational change is particularly a difficult task and computationally 
+expensive with existing methods. In this work, an improved version of the confinement approach is used to calculate absolute free energies 
+of bio molecular systems. The method does not require a reaction coordinate or 
+transition path. It is fast to compute. We show that the method correctly picks out the state of lowest free energy of a pair of 
+structures having similar sequence but different fold. We also show using models from CASP9 that the method picks out native-like 
+structures from misfolded or decoys, provided at least one good structure is in the input set.
 
@@
 
-%this is how you put a figure in document
-\begin{figure} \centering \includegraphics[width=3in]{rmsd_energy_schematic2.pdf}
+In computational structural biology there are numerous cases where free energy between well defined states are necessary. Examples
+ranges from small to large conformational change of protein due to ligand binding, change of pH etc. Free energy also play an important 
+role in case of protein folding. The ground breaking
+work of Christian B. Anfinsen and coworkers showed that the native structures of small globular proteins have a unique, thermodynamically
+stable native structure with their conformation at the global free energy minimum. Regardless of the starting point or unfolding it by
+changing different condition most proteins will finally assume the same structure. It is very important for proteins to achieve 
+their native conformation since defects in protein folding may be the molecular cause of a range of human genetic disorder. For 
+example a misfolded protein known as prion when enters any healthy organism it converts properly folded protein into misfolded one. 
+Thus, Free energy can also act as a scale to identify between the misfolded and the native state of the protein. These 
+observations further give free energy a special importance in structural biology.
 
-    \caption{Schematic}
+However, the theoretical calculation of conformational free energy change become difficult if the states of interest are very 
+different. In such scenario timescale involved in such tranformation may be beyond timescale of
+direct classical molecular dynamics simulation.  Generally, in order to calculate free energy a pathway or reaction coordinate is created with
+the intermediate structures. Method like umbrella sampling, thermodynamic integration along with classical MD can suffer from overlap problem
+and require large computational effort. The idea of reaction coordinate became more complicated in case of protein folding. 
+As free energy landscape of the protein folding is rough, it can be trapped in a energy well during its visit in the conformational 
+space. Even with the advanced method like replica exchange molecular dynamics the three dimensional structure may be trapped in a 
+local minima for a considerable time. Due to various reasons it can also be misfolded and trapped
+in a local energy minima. And one can mistakenly identify the misfolded state as the native state (give the example of human pin1ww domain
+from Benout Roux and Klaus Schultan).
 
-    \label{fig:conf_schematic} 
+Calculation of free energy has been successfully attempted by a number of groups (give examples). In recent years methods like 
+Refrence system method, Decativated Morphing, Orthogonal Space Random Work, Confinement Method have came out for calculation of 
+free energy method. Some of the groups  relies on one of the 
+great strength of free energy calculation, that it is a state function and thus it is independent of path.
+
+In this work we have applied the confinement method which was originally devoled by Tyka et. al. and Cecchini et. al.
+This method relies on a thermodynamic cycle where first, non-harmonic degrees of freedom are removed by applying a
+series of restraint to the system. Then the free energy of the remaining harmonic system is obtained using a normal mode 
+calculation and combined with the non-harmonic part to give the total absolute free energy. 
+Previously, this method was applied to calculate the conformational free energy of a 16 amino acid residue peptide, known as BHP. 
+We have first reproduced that known result. Then we have applied this method to larger proteins for the first time. The free energy 
+of conformation of pair of proteins with similar sequence but completely different fold was calculated. Encouraged by the result, 
+we applied the method of different targets of CASP9 competition.
+
+%of protein can be obtained to atomic resolution by X-ray crystallography or NMR. Unfortunately, the sheer amount
+%of time and economic investment for structure determination cannot keep up with the rate at which
+%proteins of unknown structure are sequenced. Therefore, computational models can be a good way to bridge the
+%increasing gap between sequence and structure. In this respect CASP (Critical Assessment of Structure Prediction) has played a 
+%crucial role which is a community wide experiments for the protein structure prediction.
+
+\section{Method}
+
+The confinement method has been described in details in ref. by Tyka et.al. and Cecchini et. al. Here we briefly describe the methodology.
+We compute free energy $\Delta G_{AB}$ between A and B conformation of the protein. For this purpose we use a thermodynamic cycle.
+\begin{enumerate}
+
+\item  We minimized both the conformation. These minimized conformation(A* and B*) are the restrained conformation of that state
+
+\item Backbone dihedral angle of the whole protein is restrained by a harmonic potential to define the configurational state.
+
+\item  In order to calculate the free energy $\Delta G_{A,A*}$ and  $\Delta G_{B,B*}$ A and B are gradually transformed in A* and B* by applying 
+a harmonic restraint potential to all atom. For this purpose around 21 simulation were run each of 20 ns with increasing harmonic 
+restraint potential from 0.00005 to 81.92 untill the free and the restrained state overlap well. The final restrained state was chosen so high so that
+the rotational contributional of the protein frozen out and only remaining contribution remain that of vibrational free enrgy. The fluctuation from 
+the reference structure was recorded and the free energy is calculated using a numerical approach developed by Tyka et. al.
+
+\item  Finally the thermodynamic cycle is closed by calculating the free energy between the final restrained state using normal mode analysis.
+
+\end{enumerate}
+
+In all the calculation amber 10 program is used with ff99SB with GB/SA implicit solvent.
+
+\section{Result}
+
+As a proof of priciple we first reproduced the examples from previous work where this method has been applied to a 
+16 amino acid residue peptide, $\beta$ hairpin from protein G popularly known as BHP to calculate the free energy difference 
+between two of its conformation.
+
+\subsection{BHP}
+
+First we calculated the free energy difference between two conformation of the polypeptide $\beta$ hairpin from protein G.  
+This peptide contain 16 residue. One of the conformation is the 
+native structure, has a two stranded $\beta$ sheet known as bhp1 and the other one is known as bhp3 which has a three stranded $\beta$ sheet.
+The free energy difference between them was calculated using 50 independent MD simulation each of which has a lengh of 4 $\mu$ second totaling
+200 $\mu$ second. The free energy calculated was 1.8 Kcal/mol at 360 K. On the other hand using confinement method, we found that the 
+free energy difference between bhp2 and bhp1 is 1.67 $\pm$  .
+
+\subsection{Protein with similar sequence but different fold}
+
+In general, the protein structure with similar kind of sequences tends to have similar 3D structure and are usually assumed to be 
+homologous. This idea is the basis of comparative modeling and fold recognition in protein structure prediction.
+Although there are examples in literature that protein can undergo changes in their fold due to cleavage of a peptide bond 
+or due to few changes in amino acid sequence but a more dramatic event happen for the case of two small engineered protein 
+consisting 56 amino acid residues from domain of
+protein G, known as protein GA and GB. The protein GA is serum binding domain with $3 \alpha$ fold and the protein GB is an 
+IgG-binding protein with $4 \beta + \alpha$ fold. Two pairs of such protein are engineered by step by step mutation until they 
+change fold. One pair that have 88 \% similar sequence are known as GA88 and GB88 and another pair with 95 \% similar sequence 
+are known as GA95 and GB95. There are differences in 7 and 3 positions with 88 \% and 95 \% similar sequence respectively and also can 
+be visible from Figure \ref{fig:G88} and Figure \ref{fig:G95}. The difference between GA95 and GB95 is only at residue positions 20, 
+30 and 45. While GA95 have amino acid residues Leu , Ile and Leu at positions 20, 30 and 45, in case of GB95 the amino acid residues in
+these positions are Ala, Phe and Tyr. On the other hand there are amino acid residues Gly-24, Ile-25, Ile-30, Ile-33,Leu-45,Ile-49 and 
+Leu-50 in GA88 and in GB88 in similar positions there are residues Ala-24, Thr-25, Phe-30, Tyr-33, Tyr-45, Thr-49 and Lys-50. 
+We have calculated the free energy difference between GA and GB with 
+both 88 \% and 95 \% sequence similarity. In order to do such calculation, we have first taken the sequence of GA95 and generated the both 
+GA ($3 \alpha$) fold and GB $4 \beta + \alpha$ fold with the same sequence using a specialized restraint based replica exchange 
+molecular dynamic simulation. The similar procedure was followed to generate the structure with the sequence of GB95, GA88 and GB88.
+% write why we used this %
+In the process we generated total of four pairs of structure and we also checked that the structures were withing 2 \AA rmsd of the
+NMR generated structure in all the cases. We expect the free energy should be lowest in case of sequence with the native 
+conformation. The result is presented in Table~\protect\ref{tab:GA_GB}.  
+In all cases the sequence with the native structure has the lowest free energy value. Also, the low free energy difference support
+the hypothesis that due to such low free energy difference these proteins can switch their fold.
+
+\subsection{Application on different targets of CASP9 experiments}
+
+Next we have applied the confinement method to different targets of CASP(Critical Assessment of Structure Prediction) 9 experiments. 
+CASP is a world wide competition aiming at finding the state of the art methodologies in protein folding. Groups, both from knowledge 
+based methods and physics based methods participate in this blind competition and for each sequence target they can submit 
+five models. The predictors were also asked to calculate their best model among the set of five they submitted. However, in many cases
+predictors were unable to predict/rank their models. Using the confinement method, we aim to rank protein structures using free 
+energy as a scale. In recent CASP experiments Global distance test scose (GDT\_TS) method is used for final analysis of different
+submitted structures with the NMR/crystallographically solved protein native structure. GDT\_TS represent the average 
+percentage of residues that are in close proximity in two structures optimally superimposed
+using four different distance cutoff. Although GDT\_TS alone is not a reliable measure for the difficult modeling cases, it has 
+been well accepted method in recent years among the community. We have also compared our free energy result with the 
+GDT\_TS score for different cases we studied.
+
+%The GDTTS value of a model is determined as follows. A large sample of possible structure superpositions of the model on the corresponding
+%experimental structure is generated by superposing all sets of three, five, and seven consecutive C$\alpha$ atoms along the backbone (each
+%peptide segment provides one superposition). Each of these initial superpositions is iteratively extended, including all residue pairs under
+%a specified threshold in the next iteration, and continuing until there is no change in included residues. The procedure is carried out using
+%thresholds of 1, 2, 4, and 8 $\AA$, and the final superposition that includes the maximum number of residues is selected for each threshold.
+
+\subsubsection{T0559}
+
+The native structure is a  67 residue NMR solved structure. We first reviewed all the five models submitted by the group 
+"BAKER-ROSETTASERVER", which was the best predictor group for this particular target. 
+Among them we choose only the three structures, model 1, 3 and 5 to calculate the free energy and subsequently identify the most stable structure 
+among them. The rest of the two structures are 
+close to either of the selected structure and thus they are ommited from our calculation. According to the GDT\_TS score, the model 1 was 
+picked up correctly from the set of structure generated . However, the order of model 3 and 5 was not 
+correct as can be seen from the Table 1. It will be interesting to know whether confinement method can order the strcutures in terms of 
+free energy scale. The result is presented in Table~\protect\ref{tab:T0559}.
+
+\subsubsection{T0560}
+
+Target T0560 is a 74 amino acid residue target from CASP9. However residues 3-66 were kept for final result.
+For this target we have compared the free energy of the model 2 and model 4 submitted by the group "Splicer" along with the crystal structure. 
+The main difference between the model 2 and 4 is the orientation of the ....
+The final result is presented in Table~\protect\ref{tab:T0560}.
+
+\subsubsection{T0538}
+
+Target T0538 was a 54 residue target. The native structure was generated using NMR. The best model that is closest to the native structure was
+generated by the group PconsR. However, their model 3 was better than their model 1. This structure was quite close to the NMR structure and 
+had GDT\_TS value of 96.23. We have calculated the free energy between this structure and the best structure submitted by the group "Shell",
+"FOLDIT" and "BAKER-ROSETTASERVER". The calculated free energy value is presented in Table~\protect\ref{tab:T0538}.
+
+\subsubsection{T0540}
+
+Target T0540 was a 90 residue target. We have taken the best models from the group "LTB" which was the best performing group for this target and
+also from the group Mufold and calculated the free energy difference between these models with that of the native crystal structure.
+As expected, the crystal structure has the lowest free energy value, followed by Model 1 with GDT\_TS score of $69.72$ and model 2
+with GDT\_TS value of $53.789$. The final result is presented in Table~\protect\ref{tab:T0540}.
+
+\subsubsection{T0531}
+
+This is a 65 amino acid residue target with residues 6-63 were kept for final analysis. We have calculated the free energy
+difference between all the five models submitted by the group "MUFOLD-MD", which was one of the best performed group for this target.
+The result is presented in Table~\protect\ref{tab:T0531}
+
+\subsubsection{T0605}
+This is a 72 amino acid residue protein but for the final calculation only the residues 18-66 amino acids were kept for the final analysis.
+The best performing group was "Baker" for this target. We have carried out our calculation with the model first, second and fifth for the
+final calculation. The rest two model had closer GDT\_TS value compared to the previously mentioned models and thus they were omitted from 
+furthe calculation. The result is presented in Table~\protect\ref{tab:T0605}. (This is a case where we can demand the confinement 
+method is better than GDT\_TS).
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=14cm,height=14cm]{method.pdf}
+\end{center}
+\caption{Graphical representation of the thermodynamic cycle involving confinement method.}
+\label{fig:method}
 \end{figure}
 
-\section{Methods}
+\begin{figure} 
+\begin{center} 
+\includegraphics[width=12cm,height=10cm]{bhp.pdf}
+\end{center}
+\caption{Two conformations from $\beta$ hairpin from protein G, bhp1 and bhp2. The two stranded $\beta$ sheet, bhp1 is the native structure and the three stranded $\beta$ sheet is known as bhp3.}
+\label{fig:bhp_conf}
+\end{figure}
 
-\subsection{ENM}
+\begin{figure}
+\begin{center}
+\includegraphics[width=12cm,height=10cm]{G95.pdf}
+\end{center}
+\caption{Two protein with 95 \% similar sequence but different folds}
+\label{fig:G95}
+\end{figure}
 
-%this is how you put a table
+\begin{figure}
+\begin{center}
+\includegraphics[width=12cm,height=10cm]{G88.pdf}
+\end{center}
+\caption{Two protein with 88 \% similar sequence but different folds}
+\label{fig:G88}
+\end{figure}
+
 \begin{table}
-    \centering
-    \caption{Structures used for large motions}
-    \label{tbl:targets}
-    \begin{tabular}{l l l}
-    \hline
-    Target                           & PDB1 & PDB2 \\
-    \hline
-	Scallop myosin                   & 1kk7 & 1kk8 \\
-	Acyl Carrier Protein             & 1acp & 2fae \\
-	Aspartate Aminotransferase       & 1ama & 8aat \\
-	Bcl-xl                           & 1bxl & 1ysn \\
-	Calcium Sensor                   & 1k9k & 1k9p \\
-	Calmodulin                       & 1cll & 1ctr \\
-	Cyclin Dependent Kinase Inhibitor & 1dc2 & 2a5e \\
-	Cystatin                         & 1a67 & 1cew \\
-	%CytCOxidase                     & 1oco & 2occ \\
-	Hydrolase                        & 1qz3 & 1u4n \\
-	LacRepressor                     & 1lcc & 1lqc \\
-	LambdaCro                        & 5cro & 6cro \\
-	Lupin Hydrolase                  & 1f3y & 1jkn \\
-	Maltose Binding Protein          & 1omp & 3mbp \\
-	Pin1                             & 1f8a & 1pin \\
-	%Thrombin                        & 1avg & 1ett \\
-        \hline
+\caption{The comparison of free energy values of proteins with 95 \% and 88 \% similar sequences in $3 \alpha$ and $4 \beta + \alpha$ fold.}
+\label{tab:GA_GB}
+\begin{center}
+\begin{tabular}{l l l}\hline
+Changes in   &     $\Delta G$ (Kcal/Mol) &  $\Delta G$ (Kcal/Mol) \\
+ Sequences   &     ($3 \alpha$ fold)       & ($4 \beta + \alpha$ fold) \\ \hline
+L20I30L45    &     $0.000$            &  $2.901 \pm 0.475$   \\
+A20F30Y45    &     $4.365 \pm 0.460$ &  $0.000$    \\
+G24I25I30I33 &     $0.000$            &  $ 3.941 \pm 0.517$ \\
+L45I49L50    &                        &              \\
+A24T25F30Y33 &     $5.576 \pm 0.464$  &  $0.000$     \\
+Y45T49K50    &                        &             \\ \hline
 \end{tabular}
+\end{center}
 \end{table}
 
-%This tells latex to reference the figure or table
-See figure \ref{fig:conf_schematic}.
+\begin{figure}
+\begin{center}
+\includegraphics[width=12cm,height=10cm]{T0559.pdf}
+\end{center}
+\caption{Target T0559 from CASP9}
+\label{fig:T0559}
+\end{figure}
 
-See table \ref{tbl:targets}.
+\begin{table}
+\caption{The value of free energy and GDT\_TS for the CASP9 target T0559}
+\label{tab:T0559}
+\begin{center}
+\begin{tabular}{l l l}\hline
+Model   &     $\Delta G$ (Kcal/Mol) &  GDT\_TS \\ \hline
+Model 1 &     0.000            &  94.03    \\
+Model 3 &     $14.484 \pm 0.633$ &  48.13    \\         
+Model 5 &     $4.526  \pm 0.384$ &  82.46    \\ \hline
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=12cm,height=10cm]{T0538.pdf}
+\end{center}
+\caption{Target T0538}
+\label{fig:T0538}
+\end{figure}
+
+\begin{table}
+\caption{The Free energy and GDT\_TS values of the CASP9 Target T0538}
+\label{tab:T0538}
+\begin{center}
+\begin{tabular}{l l l l}\hline
+Model   &     $\Delta G$ (Kcal/Mol) &  GDT\_TS & Group Submitted \\ \hline
+Model 1 &     0.000              &  96.23    & PconsR  \\ 
+Model 2 &     $15.891 \pm 0.475$ &  90.09  & Shell  \\ 
+Model 3 &     $23.243 \pm 0.393$ &  86.32  & FOLDIT  \\ 
+Model 4 &     $26.002 \pm 0.756$ &  80.66  & BAKER-ROSETTASERVER \\ \hline
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=12cm,height=10cm]{T0560.pdf}
+\end{center}
+\caption{Target T0560}
+\label{fig:T0560}
+\end{figure}
+
+\begin{table}
+\caption{The free energy and the GDT\_TS value of the CASP9 target T0560.}
+\label{tab:T0560}
+\begin{center}
+\begin{tabular}{l l l}\hline
+Model   &     $\Delta G$ (Kcal/Mol) &  GDT\_TS \\ \hline
+Crystal &     0.000              & 100.00    \\
+Model 2 &     $22.153 \pm 0.491$ &  94.14    \\
+Model 4 &     $27.428  \pm 0.477$ &  76.56    \\ \hline
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=12cm,height=10cm]{T0540.pdf}
+\end{center}
+\caption{T0540}
+\label{fig:}
+\end{figure}
+
+\begin{table}
+\caption{The free energy and the GDT\_TS value of the CASP9 target T0540.}
+\label{tab:T0540}
+\begin{center}
+\begin{tabular}{l l l l}\hline
+Model   &     $\Delta G$ (Kcal/Mol) &  GDT\_TS & Group Submitted \\ \hline
+Crystal &     0.000              &  100.00   & -      \\ 
+Model 1 &     $22.003 \pm 0.575$ &  69.72    & LTB    \\
+Model 2 &     $53.786 \pm 0.553$ &  63.89    & Mufold \\ \hline
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=12cm,height=10cm]{T0531.pdf}
+\end{center}
+\caption{T0531}
+\label{fig:T0531}
+\end{figure}
+
+\begin{table}
+\caption{The free energy and the GDT\_TS value of the CASP9 target T0531.}
+\label{tab:T0531}
+\begin{center}
+\begin{tabular}{l l l}\hline
+Model   &     $\Delta G$ (Kcal/Mol) &  GDT\_TS \\ \hline
+Crystal &     0.000              & 100.00    \\
+Model 1 &     $12.496 \pm 0.704$ &  43.53    \\
+Model 2 &     $27.374 \pm 0.691$ &  34.48    \\
+Model 3 &     $13.914 \pm 0.673$ &  34.48    \\
+Model 4 &     $53.055 \pm 0.684$ &  28.02    \\
+Model 5 &     $29.508 \pm 0.698$ &  25.86   \\ \hline
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=12cm,height=10cm]{T0605.pdf}
+\end{center}
+\caption{T0605}
+\label{fig:T0605}
+\end{figure}
+
+\begin{table}
+\caption{The free energy and the GDT\_TS value of the CASP9 target T0605.}
+\label{tab:T0605}
+\begin{center}
+\begin{tabular}{l l l}\hline
+Model   &     $\Delta G$ (Kcal/Mol) &  GDT\_TS \\ \hline
+Model 1 &     $0.000$ &  $97.96$    \\
+Model 2 &     $1.689 \pm ?$ &  $89.29$    \\
+Model 5 &     $16.064 \pm ?$ & $93.88$   \\ \hline
+\end{tabular}
+\end{center}
+\end{table}
+
+\section{Conclusion}
 
@@
 
+Ytreberg, F.; Zuckerman, D. Simple estimation of absolute free energies for biomolecules. J. Chem. Phys. 2006, 124, 104105.
+
+Park, S.; Lau, A.; Roux, B. Computing conformational free energy by deactivated morphing. J. Chem. Phys. 2008, 129, 134102
+
+Zheng, L.; Chen, M.; Yang, W. Random walk in orthogonal space to achieve efficient free-energy simulation of complex systems, Proc. Natl. Acad. Sci. 2008, 105 (51), 20227.
+
+Tyka, M.; Clarke, A.; Sessions, R. An Efficient, Path-Independent Method for Free-Energy Calculations. J.Phys.Chem. B 2006, 110, 17212–17220.
+
 \end{document}

### 2012-09-09 Paper/paper.tex
Reworked Orban section

@@
 
-%\documentclass[preprint,preprintnumbers,amsmath,amssymb]{revtex4}
 \documentclass[12pt]{article}
 
-%\usepackage{fontspec} 
-%\setmainfont{TeX Gyre Pagella}
 \usepackage{amsmath}
@@
 % Custom macros for author comments
-\newcommand{\Alberto}[1]{\color{ForestGreen}Alberto: #1 \normalcolor }
-\newcommand{\Justin}[1]{\color{blue}Justin: #1 \normalcolor}
-\newcommand{\Arijit}[1]{\color{yellow}Arijit: #1 \normalcolor}
-\newcommand{\Ken}[1]{\color{red}Ken: #1 \normalcolor}
+\newcommand{\Alberto}[1]{\color{ForestGreen}#1\normalcolor }
+\newcommand{\Justin}[1]{\color{blue}#1\normalcolor}
+\newcommand{\Arijit}[1]{\color{yellow}#1\normalcolor}
+\newcommand{\Ken}[1]{\color{red}#1\normalcolor}
 
-\author{Arijit~Roy$^1$}
-\author{Justin~L.~MacCallum$^1$}
-\author{Alberto~Perez$^1$}
-\author{Ken~A.~Dill$^1$}
-\affil{$^1$Laufer Center for Physical and Quantitative Biology\\
+\author{Arijit~Roy}
+\author{Justin~L.~MacCallum}
+\author{Alberto~Perez}
+\author{Ken~A.~Dill}
+\affil{Laufer Center for Physical and Quantitative Biology\\
     Stony Brook University\\
@@
 
-\title{Ranking protein structures using confinement free energy method}
+\title{Physics-based protein structure prediction and design using the confinement method}
 
@@
 \maketitle
+
 \begin{abstract}
-%How to cite something in paper.bib Abstract \citep{CASP9}
 
-The calculation of free energy differences is of central importance in the simulation of biochemical systems.  The computation of the 
-free energy difference between pairs of macromolecule with large conformational change is particularly a difficult task and computationally 
-expensive with existing methods. In this work, an improved version of the confinement approach is used to calculate absolute free energies 
-of bio molecular systems. The method does not require a reaction coordinate or 
-transition path. It is fast to compute. We show that the method correctly picks out the state of lowest free energy of a pair of 
-structures having similar sequence but different fold. We also show using models from CASP9 that the method picks out native-like 
-structures from misfolded or decoys, provided at least one good structure is in the input set.
+The calculation of free energy differences is of central importance in the simulation of biochemical
+systems.  The computation of the free energy difference between pairs of macromolecule with large
+conformational change is particularly a difficult task and computationally expensive with existing
+methods. In this work, an improved version of the confinement approach is used to calculate absolute
+free energies of bio molecular systems. The method does not require a reaction coordinate or
+transition path. It is fast to compute. We show that the method correctly picks out the state of
+lowest free energy of a pair of structures having similar sequence but different fold. We also show
+using models from CASP9 that the method picks out native-like structures from misfolded or decoys,
+provided at least one good structure is in the input set.
 
@@
 
-In computational structural biology there are numerous cases where free energy between well defined states are necessary. Examples
-ranges from small to large conformational change of protein due to ligand binding, change of pH etc. Free energy also play an important 
-role in case of protein folding. The ground breaking
-work of Christian B. Anfinsen and coworkers showed that the native structures of small globular proteins have a unique, thermodynamically
-stable native structure with their conformation at the global free energy minimum. Regardless of the starting point or unfolding it by
-changing different condition most proteins will finally assume the same structure. It is very important for proteins to achieve 
-their native conformation since defects in protein folding may be the molecular cause of a range of human genetic disorder. For 
-example a misfolded protein known as prion when enters any healthy organism it converts properly folded protein into misfolded one. 
-Thus, Free energy can also act as a scale to identify between the misfolded and the native state of the protein. These 
-observations further give free energy a special importance in structural biology.
+In computational structural biology there are numerous cases where free energy between well defined
+states are necessary. Examples ranges from small to large conformational change of protein due to
+ligand binding, change of pH etc. Free energy also play an important role in case of protein
+folding. The ground breaking work of Christian B. Anfinsen and coworkers showed that the native
+structures of small globular proteins have a unique, thermodynamically stable native structure with
+their conformation at the global free energy minimum. Regardless of the starting point or unfolding
+it by changing different condition most proteins will finally assume the same structure. It is very
+important for proteins to achieve their native conformation since defects in protein folding may be
+the molecular cause of a range of human genetic disorder. For example a misfolded protein known as
+prion when enters any healthy organism it converts properly folded protein into misfolded one.
+Thus, Free energy can also act as a scale to identify between the misfolded and the native state of
+the protein. These observations further give free energy a special importance in structural biology.
 
-However, the theoretical calculation of conformational free energy change become difficult if the states of interest are very 
-different. In such scenario timescale involved in such tranformation may be beyond timescale of
-direct classical molecular dynamics simulation.  Generally, in order to calculate free energy a pathway or reaction coordinate is created with
-the intermediate structures. Method like umbrella sampling, thermodynamic integration along with classical MD can suffer from overlap problem
-and require large computational effort. The idea of reaction coordinate became more complicated in case of protein folding. 
-As free energy landscape of the protein folding is rough, it can be trapped in a energy well during its visit in the conformational 
-space. Even with the advanced method like replica exchange molecular dynamics the three dimensional structure may be trapped in a 
-local minima for a considerable time. Due to various reasons it can also be misfolded and trapped
-in a local energy minima. And one can mistakenly identify the misfolded state as the native state (give the example of human pin1ww domain
-from Benout Roux and Klaus Schultan).
+However, the theoretical calculation of conformational free energy change become difficult if the
+states of interest are very different. In such scenario timescale involved in such tranformation may
+be beyond timescale of direct classical molecular dynamics simulation.  Generally, in order to
+calculate free energy a pathway or reaction coordinate is created with the intermediate structures.
+Method like umbrella sampling, thermodynamic integration along with classical MD can suffer from
+overlap problem and require large computational effort. The idea of reaction coordinate became more
+complicated in case of protein folding.  As free energy landscape of the protein folding is rough,
+it can be trapped in a energy well during its visit in the conformational space. Even with the
+advanced method like replica exchange molecular dynamics the three dimensional structure may be
+trapped in a local minima for a considerable time. Due to various reasons it can also be misfolded
+and trapped in a local energy minima. And one can mistakenly identify the misfolded state as the
+native state (give the example of human pin1ww domain from Benout Roux and Klaus Schultan).
 
-Calculation of free energy has been successfully attempted by a number of groups (give examples). In recent years methods like 
-Refrence system method, Decativated Morphing, Orthogonal Space Random Work, Confinement Method have came out for calculation of 
-free energy method. Some of the groups  relies on one of the 
-great strength of free energy calculation, that it is a state function and thus it is independent of path.
+Calculation of free energy has been successfully attempted by a number of groups (give examples). In
+recent years methods like Refrence system method, Decativated Morphing, Orthogonal Space Random
+Work, Confinement Method have came out for calculation of free energy method. Some of the groups
+relies on one of the great strength of free energy calculation, that it is a state function and thus
+it is independent of path.
 
-In this work we have applied the confinement method which was originally devoled by Tyka et. al. and Cecchini et. al.
-This method relies on a thermodynamic cycle where first, non-harmonic degrees of freedom are removed by applying a
-series of restraint to the system. Then the free energy of the remaining harmonic system is obtained using a normal mode 
-calculation and combined with the non-harmonic part to give the total absolute free energy. 
-Previously, this method was applied to calculate the conformational free energy of a 16 amino acid residue peptide, known as BHP. 
-We have first reproduced that known result. Then we have applied this method to larger proteins for the first time. The free energy 
-of conformation of pair of proteins with similar sequence but completely different fold was calculated. Encouraged by the result, 
-we applied the method of different targets of CASP9 competition.
+In this work we have applied the confinement method which was originally devoled by Tyka et. al. and
+Cecchini et. al.  This method relies on a thermodynamic cycle where first, non-harmonic degrees of
+freedom are removed by applying a series of restraint to the system. Then the free energy of the
+remaining harmonic system is obtained using a normal mode calculation and combined with the
+non-harmonic part to give the total absolute free energy.  Previously, this method was applied to
+calculate the conformational free energy of a 16 amino acid residue peptide, known as BHP.  We have
+first reproduced that known result. Then we have applied this method to larger proteins for the
+first time. The free energy of conformation of pair of proteins with similar sequence but completely
+different fold was calculated. Encouraged by the result, we applied the method of different targets
+of CASP9 competition.
 
-%of protein can be obtained to atomic resolution by X-ray crystallography or NMR. Unfortunately, the sheer amount
-%of time and economic investment for structure determination cannot keep up with the rate at which
-%proteins of unknown structure are sequenced. Therefore, computational models can be a good way to bridge the
-%increasing gap between sequence and structure. In this respect CASP (Critical Assessment of Structure Prediction) has played a 
-%crucial role which is a community wide experiments for the protein structure prediction.
+%of protein can be obtained to atomic resolution by X-ray crystallography or NMR. Unfortunately, the
+%sheer amount of time and economic investment for structure determination cannot keep up with the
+%rate at which proteins of unknown structure are sequenced. Therefore, computational models can be a
+%good way to bridge the increasing gap between sequence and structure. In this respect CASP
+%(Critical Assessment of Structure Prediction) has played a crucial role which is a community wide
+%experiments for the protein structure prediction.
 
-\section{Method}
+\section{Results and Discussion}
 
-The confinement method has been described in details in ref. by Tyka et.al. and Cecchini et. al. Here we briefly describe the methodology.
-We compute free energy $\Delta G_{AB}$ between A and B conformation of the protein. For this purpose we use a thermodynamic cycle.
-\begin{enumerate}
+\subsection{The confinement method produces correct results in control experiments}
 
-\item  We minimized both the conformation. These minimized conformation(A* and B*) are the restrained conformation of that state
+As a first step, we performed several control experiments to verify that our implementation of the
+confinement method produces results compatible with previous calculations reported in the
+literature.
 
-\item Backbone dihedral angle of the whole protein is restrained by a harmonic potential to define the configurational state.
+The method has previously been applied to a 16 amino acid residue $\beta$-hairpin from
+protein G, known as BHP[ref]. We calculated the free energy difference between two different
+conformations of the peptide: (1) the native conformation, called bhp1, with a two-stranded
+$\beta$-sheet; and (2) a conformation, called bhp3, which has a three-stranded $\beta$-sheet.
+Analysis of long (4 $\mu$s) equilibrium simulations[ref] shows that bhp1 is the more favorable
+configuration by 1.8 kcal/mol. Using the confinement method, we obtain a value 1.7 kcal/mol, which
+is in good agreement with the equilibrium simulations and with previous calculations using the
+confinement method[ref] (see Supporting Information for further details).
 
-\item  In order to calculate the free energy $\Delta G_{A,A*}$ and  $\Delta G_{B,B*}$ A and B are gradually transformed in A* and B* by applying 
-a harmonic restraint potential to all atom. For this purpose around 21 simulation were run each of 20 ns with increasing harmonic 
-restraint potential from 0.00005 to 81.92 untill the free and the restrained state overlap well. The final restrained state was chosen so high so that
-the rotational contributional of the protein frozen out and only remaining contribution remain that of vibrational free enrgy. The fluctuation from 
-the reference structure was recorded and the free energy is calculated using a numerical approach developed by Tyka et. al.
+\Justin{Arijit, Al: did we do any other control experiments on small peptides that we should report
+here?}
 
-\item  Finally the thermodynamic cycle is closed by calculating the free energy between the final restrained state using normal mode analysis.
+Previous applications of the confiment method have focused on relatively short peptides, up to X
+residues in length \Justin{Arijit: fill in X}. In this work we apply the confinement method to
+larger proteins. One of the  applications we explore in the present work is the re-scoring, or
+metaprediction, of structures submitted during the Critical Assessment of Structure Prediction
+(CASP) experiment (described in detail later). We computed the relative free energies of predictions
+submitted during CASP9 and 10, with the expectation that the most native-like prediction will have
+the lowest free energy. As a control, for several targets we also calculated the relative free
+energy of the experimental structure (once it was available from the PDB). As
+Table~\ref{table:casp_control} shows, in X of Y cases, the confinement method correctly assigns a
+lower free energy to the experimentally determined structure than to any of the decoys. Although
+differentiating between experimentally determined structures and computer generated predictions is
+not a stringent test of a scoring method---for example, examining the residue-residue packing can
+easily distinguish between the two [ref Rosetta Holes]---the results none the less serve as a useful
+"sanity check".
 
-\end{enumerate}
+\begin{table}
+\label{table:casp_control}
 
-In all the calculation amber 10 program is used with ff99SB with GB/SA implicit solvent.
+\begin{center}
+\begin{tabular}{l l l}\hline
+    CASP Target  & PDB Identifier & $\Delta\Delta G_{native \to best decoy}$ (kcal/mol) \\ \hline
+\end{tabular}
+\end{center}
 
-\section{Result}
+\caption{The confinement method assigns a more favorable free energy to the experimentally
+determined structure than to computer-generated predictions. For each target, we examined as many as
+five predictions submitted by CASP participants. Positive $\Delta\Delta G$ values indicate that the
+experimental structure is predicted to be more favorable than any of the decoys.}
 
-As a proof of priciple we first reproduced the examples from previous work where this method has been applied to a 
-16 amino acid residue peptide, $\beta$ hairpin from protein G popularly known as BHP to calculate the free energy difference 
-between two of its conformation.
+\end{table}
 
-\subsection{BHP}
+\subsection{The confinement method correctly predicts the structural preferences of chameleon
+sequences}
 
-First we calculated the free energy difference between two conformation of the polypeptide $\beta$ hairpin from protein G.  
-This peptide contain 16 residue. One of the conformation is the 
-native structure, has a two stranded $\beta$ sheet known as bhp1 and the other one is known as bhp3 which has a three stranded $\beta$ sheet.
-The free energy difference between them was calculated using 50 independent MD simulation each of which has a lengh of 4 $\mu$ second totaling
-200 $\mu$ second. The free energy calculated was 1.8 Kcal/mol at 360 K. On the other hand using confinement method, we found that the 
-free energy difference between bhp2 and bhp1 is 1.67 $\pm$  .
+In general, proteins with similar sequences tend to have similar structures. This idea is the basis
+of comparative modeling and fold recognition in protein structure prediction. There are, however,
+examples---often referred to as chameleon sequences---of proteins with similar sequences that have
+remarkably different structures. Orban and co-workers have designed a series of 56-residue proteins
+(based on Protein G) that adopt one of two different folds depending on small changes in sequence
+(see Figure~\ref{fig:orban}). Sequences that adopt a mixed alpha/beta structure similar to Protein G
+are denoted as ``GB'', while sequences that form a three-helix bundle are denoted as ``GA''. One
+pair of sequences (GA88/GB88) are 88 percent identical and differ in seven positions. Another pair
+(GA95/GA95) are 95 percent identical and differ in three positions. The last pair (GA98/GB98) differ
+only in a single tyrosine to alanine mutation.
 
-\subsection{Protein with similar sequence but different fold}
+The fact such small changes in sequence can lead to such dramatic changes in structure is rather
+remarkable. Accurately predicting the structural preferences of these structures presents a serious
+challenge for computational methods [ref].
 
-In general, the protein structure with similar kind of sequences tends to have similar 3D structure and are usually assumed to be 
-homologous. This idea is the basis of comparative modeling and fold recognition in protein structure prediction.
-Although there are examples in literature that protein can undergo changes in their fold due to cleavage of a peptide bond 
-or due to few changes in amino acid sequence but a more dramatic event happen for the case of two small engineered protein 
-consisting 56 amino acid residues from domain of
-protein G, known as protein GA and GB. The protein GA is serum binding domain with $3 \alpha$ fold and the protein GB is an 
-IgG-binding protein with $4 \beta + \alpha$ fold. Two pairs of such protein are engineered by step by step mutation until they 
-change fold. One pair that have 88 \% similar sequence are known as GA88 and GB88 and another pair with 95 \% similar sequence 
-are known as GA95 and GB95. There are differences in 7 and 3 positions with 88 \% and 95 \% similar sequence respectively and also can 
-be visible from Figure \ref{fig:G88} and Figure \ref{fig:G95}. The difference between GA95 and GB95 is only at residue positions 20, 
-30 and 45. While GA95 have amino acid residues Leu , Ile and Leu at positions 20, 30 and 45, in case of GB95 the amino acid residues in
-these positions are Ala, Phe and Tyr. On the other hand there are amino acid residues Gly-24, Ile-25, Ile-30, Ile-33,Leu-45,Ile-49 and 
-Leu-50 in GA88 and in GB88 in similar positions there are residues Ala-24, Thr-25, Phe-30, Tyr-33, Tyr-45, Thr-49 and Lys-50. 
-We have calculated the free energy difference between GA and GB with 
-both 88 \% and 95 \% sequence similarity. In order to do such calculation, we have first taken the sequence of GA95 and generated the both 
-GA ($3 \alpha$) fold and GB $4 \beta + \alpha$ fold with the same sequence using a specialized restraint based replica exchange 
-molecular dynamic simulation. The similar procedure was followed to generate the structure with the sequence of GB95, GA88 and GB88.
-% write why we used this %
-In the process we generated total of four pairs of structure and we also checked that the structures were withing 2 \AA rmsd of the
-NMR generated structure in all the cases. We expect the free energy should be lowest in case of sequence with the native 
-conformation. The result is presented in Table~\protect\ref{tab:GA_GB}.  
-In all cases the sequence with the native structure has the lowest free energy value. Also, the low free energy difference support
-the hypothesis that due to such low free energy difference these proteins can switch their fold.
+We initially approached this problem by making a model of each sequence with the same backbone
+structure as its partner chameleon sequence. For example, we took the sequence of GA88 and built a
+model with the same overall structure as GB88. We then used the confinement method to assess the
+free energy difference between the experimentally determined structure of GA88 and the model (with
+the GA88 sequence and the GB88 structure). The confinement method was able to predict the
+conformational preferences correctly for all six sequences (data not shown). There is, however, a
+serious problem with this analysis: we are comparing an experimentally determined structure with a
+computer generated model. It might be possible that we were able to make correct predictions simply
+because artefacts of our modeling procedure always lead to the model having a higher free energy
+than the experimental structure.
 
-\subsection{Application on different targets of CASP9 experiments}
+To avoid this potential problem, we instead computed the relative free energy two different
+structural models for each sequence. One model is based on the GA structure and the other on the GB
+structure (see Supporting Information for details on the modeling procedure). This is a much more
+realistic test of the confinement method's ability to accurately calculate relative free energies.
 
-Next we have applied the confinement method to different targets of CASP(Critical Assessment of Structure Prediction) 9 experiments. 
-CASP is a world wide competition aiming at finding the state of the art methodologies in protein folding. Groups, both from knowledge 
-based methods and physics based methods participate in this blind competition and for each sequence target they can submit 
-five models. The predictors were also asked to calculate their best model among the set of five they submitted. However, in many cases
-predictors were unable to predict/rank their models. Using the confinement method, we aim to rank protein structures using free 
-energy as a scale. In recent CASP experiments Global distance test scose (GDT\_TS) method is used for final analysis of different
-submitted structures with the NMR/crystallographically solved protein native structure. GDT\_TS represent the average 
-percentage of residues that are in close proximity in two structures optimally superimposed
-using four different distance cutoff. Although GDT\_TS alone is not a reliable measure for the difficult modeling cases, it has 
-been well accepted method in recent years among the community. We have also compared our free energy result with the 
-GDT\_TS score for different cases we studied.
+The results of these calculations are presented in Figure~\ref{fig:orban}. The confinement method
+identifies the correct structure for all six sequences. The free energy differences range from
+around 2.9 to 5.7 kcal/mol, which is consistent with small changes in sequences and with estimates
+made by Orban and co-workers [ref].
 
-%The GDTTS value of a model is determined as follows. A large sample of possible structure superpositions of the model on the corresponding
-%experimental structure is generated by superposing all sets of three, five, and seven consecutive C$\alpha$ atoms along the backbone (each
-%peptide segment provides one superposition). Each of these initial superpositions is iteratively extended, including all residue pairs under
-%a specified threshold in the next iteration, and continuing until there is no change in included residues. The procedure is carried out using
-%thresholds of 1, 2, 4, and 8 $\AA$, and the final superposition that includes the maximum number of residues is selected for each threshold.
+\begin{figure}
+\label{fig:orban}
+
+\caption{The confinement method correctly predicts the structural preferences of six chameleon
+sequences. (A) The six sequences used in this study. (B) Each sequence adopts either a Protein
+G-like fold (denoted GB) or a three-helix bundle fold (denoted GA). The relative free energies of
+the two folds are reported for each sequence.}
+
+\Justin{Arijit: we need to make this figure, which we've discussed before and move the data from the
+following table into the figure.}
+\end{figure}
+
+%\begin{table}
+%\caption{The confinement method correctly reproduces the structural preferences of a series of
+%designed peptides. Each sequence adopts either a Protein G-like fold ($4 \beta + \alpha$) or forms a
+%helix bundle ($3 \alpha$).}
+%\label{table:orban}
+%\begin{center}
+%\begin{tabular}{l l l}\hline
+%Changes in & $\Delta G$ (kcal/mol)    & $\Delta G$ (kcal/mol)     \\
+ %Sequences & ($3 \alpha$ fold)        & ($4 \beta + \alpha$ fold) \\ \hline
+%GA95       & $0.000$                  & $2.901 \pm 0.475$         \\
+%GB95       & $4.365 \pm 0.460$        & $0.000$                   \\
+%GA88       & $0.000$                  & $ 3.941 \pm 0.517$        \\
+%GB88       & $5.576 \pm 0.464$        & $0.000$                   \\
+%GA98       & \Justin{Arijit: fill in} &                           \\ \hline
+%GB98       &                          &                           \\ \hline
+%\end{tabular}
+%\end{center}
+%\end{table}
+
+\subsection{The confinement method correctly identifies the most native-like predictions from a
+subset of CASP predictions}
+
+Next, we have applied the confinement method to several targets from the CASP9 and CASP10
+experiments. CASP is a world-wide competition aiming at finding the state of the art methodologies
+in protein folding. Groups, both from knowledge based methods and physics based methods participate
+in this blind competition and for each sequence target they can submit five models. The predictors
+were also asked to calculate their best model among the set of five they submitted. However, in many
+cases predictors were unable to predict/rank their models. Using the confinement method, we aim to
+rank protein structures using free energy as a scale. In recent CASP experiments Global distance
+test scose (GDT\_TS) method is used for final analysis of different submitted structures with the
+NMR/crystallographically solved protein native structure. GDT\_TS represent the average percentage
+of residues that are in close proximity in two structures optimally superimposed using four
+different distance cutoff. Although GDT\_TS alone is not a reliable measure for the difficult
+modeling cases, it has been well accepted method in recent years among the community. We have also
+compared our free energy result with the GDT\_TS score for different cases we studied.
+
+%The GDTTS value of a model is determined as follows. A large sample of possible structure
+%superpositions of the model on the corresponding experimental structure is generated by superposing
+%all sets of three, five, and seven consecutive C$\alpha$ atoms along the backbone (each peptide
+%segment provides one superposition). Each of these initial superpositions is iteratively extended,
+%including all residue pairs under a specified threshold in the next iteration, and continuing until
+%there is no change in included residues. The procedure is carried out using thresholds of 1, 2, 4,
+%and 8 $\AA$, and the final superposition that includes the maximum number of residues is selected
+%for each threshold.
 
@@
 
-The native structure is a  67 residue NMR solved structure. We first reviewed all the five models submitted by the group 
-"BAKER-ROSETTASERVER", which was the best predictor group for this particular target. 
-Among them we choose only the three structures, model 1, 3 and 5 to calculate the free energy and subsequently identify the most stable structure 
-among them. The rest of the two structures are 
-close to either of the selected structure and thus they are ommited from our calculation. According to the GDT\_TS score, the model 1 was 
-picked up correctly from the set of structure generated . However, the order of model 3 and 5 was not 
-correct as can be seen from the Table 1. It will be interesting to know whether confinement method can order the strcutures in terms of 
-free energy scale. The result is presented in Table~\protect\ref{tab:T0559}.
+The native structure is a  67 residue NMR solved structure. We first reviewed all the five models
+submitted by the group "BAKER-ROSETTASERVER", which was the best predictor group for this particular
+target.  Among them we choose only the three structures, model 1, 3 and 5 to calculate the free
+energy and subsequently identify the most stable structure among them. The rest of the two
+structures are close to either of the selected structure and thus they are ommited from our
+calculation. According to the GDT\_TS score, the model 1 was picked up correctly from the set of
+structure generated . However, the order of model 3 and 5 was not correct as can be seen from the
+Table 1. It will be interesting to know whether confinement method can order the strcutures in terms
+of free energy scale. The result is presented in Table~\protect\ref{tab:T0559}.
 
@@
 
-Target T0560 is a 74 amino acid residue target from CASP9. However residues 3-66 were kept for final result.
-For this target we have compared the free energy of the model 2 and model 4 submitted by the group "Splicer" along with the crystal structure. 
-The main difference between the model 2 and 4 is the orientation of the ....
-The final result is presented in Table~\protect\ref{tab:T0560}.
+Target T0560 is a 74 amino acid residue target from CASP9. However residues 3-66 were kept for final
+result.  For this target we have compared the free energy of the model 2 and model 4 submitted by
+the group "Splicer" along with the crystal structure.  The main difference between the model 2 and 4
+is the orientation of the ....  The final result is presented in Table~\protect\ref{tab:T0560}.
 
@@
 
-Target T0538 was a 54 residue target. The native structure was generated using NMR. The best model that is closest to the native structure was
-generated by the group PconsR. However, their model 3 was better than their model 1. This structure was quite close to the NMR structure and 
-had GDT\_TS value of 96.23. We have calculated the free energy between this structure and the best structure submitted by the group "Shell",
-"FOLDIT" and "BAKER-ROSETTASERVER". The calculated free energy value is presented in Table~\protect\ref{tab:T0538}.
+Target T0538 was a 54 residue target. The native structure was generated using NMR. The best model
+that is closest to the native structure was generated by the group PconsR. However, their model 3
+was better than their model 1. This structure was quite close to the NMR structure and had GDT\_TS
+value of 96.23. We have calculated the free energy between this structure and the best structure
+submitted by the group "Shell", "FOLDIT" and "BAKER-ROSETTASERVER". The calculated free energy value
+is presented in Table~\protect\ref{tab:T0538}.
 
@@
 
-Target T0540 was a 90 residue target. We have taken the best models from the group "LTB" which was the best performing group for this target and
-also from the group Mufold and calculated the free energy difference between these models with that of the native crystal structure.
-As expected, the crystal structure has the lowest free energy value, followed by Model 1 with GDT\_TS score of $69.72$ and model 2
-with GDT\_TS value of $53.789$. The final result is presented in Table~\protect\ref{tab:T0540}.
+Target T0540 was a 90 residue target. We have taken the best models from the group "LTB" which was
+the best performing group for this target and also from the group Mufold and calculated the free
+energy difference between these models with that of the native crystal structure.  As expected, the
+crystal structure has the lowest free energy value, followed by Model 1 with GDT\_TS score of
+$69.72$ and model 2 with GDT\_TS value of $53.789$. The final result is presented in
+Table~\protect\ref{tab:T0540}.
 
@@
 
-This is a 65 amino acid residue target with residues 6-63 were kept for final analysis. We have calculated the free energy
-difference between all the five models submitted by the group "MUFOLD-MD", which was one of the best performed group for this target.
-The result is presented in Table~\protect\ref{tab:T0531}
+This is a 65 amino acid residue target with residues 6-63 were kept for final analysis. We have
+calculated the free energy difference between all the five models submitted by the group
+"MUFOLD-MD", which was one of the best performed group for this target.  The result is presented in
+Table~\protect\ref{tab:T0531}
 
 \subsubsection{T0605}
+
 This is a 72 amino acid residue protein but for the final calculation only the residues 18-66 amino acids were kept for the final analysis.
+
 The best performing group was "Baker" for this target. We have carried out our calculation with the model first, second and fifth for the
@@
 \end{figure}
-
-\begin{table}
-\caption{The comparison of free energy values of proteins with 95 \% and 88 \% similar sequences in $3 \alpha$ and $4 \beta + \alpha$ fold.}
-\label{tab:GA_GB}
-\begin{center}
-\begin{tabular}{l l l}\hline
-Changes in   &     $\Delta G$ (Kcal/Mol) &  $\Delta G$ (Kcal/Mol) \\
- Sequences   &     ($3 \alpha$ fold)       & ($4 \beta + \alpha$ fold) \\ \hline
-L20I30L45    &     $0.000$            &  $2.901 \pm 0.475$   \\
-A20F30Y45    &     $4.365 \pm 0.460$ &  $0.000$    \\
-G24I25I30I33 &     $0.000$            &  $ 3.941 \pm 0.517$ \\
-L45I49L50    &                        &              \\
-A24T25F30Y33 &     $5.576 \pm 0.464$  &  $0.000$     \\
-Y45T49K50    &                        &             \\ \hline
-\end{tabular}
-\end{center}
-\end{table}
 
@@
 
+\section{Method}
+
+The confinement method has been described in details in ref. by Tyka et.al. and Cecchini et. al.
+Here we briefly describe the methodology.  We compute free energy $\Delta G_{AB}$ between A and B
+conformation of the protein. For this purpose we use a thermodynamic cycle.
+
+\begin{enumerate}
+
+\item  We minimized both the conformation. These minimized conformation(A* and B*) are the
+    restrained conformation of that state
+
+\item Backbone dihedral angle of the whole protein is restrained by a harmonic potential to define
+    the configurational state.
+
+\item  In order to calculate the free energy $\Delta G_{A,A*}$ and  $\Delta G_{B,B*}$ A and B are
+    gradually transformed in A* and B* by applying a harmonic restraint potential to all atom. For
+    this purpose around 21 simulation were run each of 20 ns with increasing harmonic restraint
+    potential from 0.00005 to 81.92 untill the free and the restrained state overlap well. The final
+    restrained state was chosen so high so that the rotational contributional of the protein frozen
+    out and only remaining contribution remain that of vibrational free enrgy. The fluctuation from
+    the reference structure was recorded and the free energy is calculated using a numerical
+    approach developed by Tyka et. al.
+
+\item  Finally the thermodynamic cycle is closed by calculating the free energy between the final
+    restrained state using normal mode analysis.
+
+\end{enumerate}
+
+In all the calculation amber 10 program is used with ff99SB with GB/SA implicit solvent.
+
 \bibliographystyle{achemso}

### 2012-10-01 Paper/paper.tex
second draft arijit

@@
 overlap problem and require large computational effort. The idea of reaction coordinate became more
-complicated in case of protein folding.  As free energy landscape of the protein folding is rough,
+complicated in case of protein folding.  As free energy landscape of the protein folding is rough \cite{Dill1997},
 it can be trapped in a energy well during its visit in the conformational space. Even with the
@@
 
-Calculation of free energy has been successfully attempted by a number of groups (give examples). In
-recent years methods like Refrence system method, Decativated Morphing, Orthogonal Space Random
+Calculation of free energy has been successfully attempted by a number of groups \cite{Ytreberg2006} - \cite{Zheng2008}. In
+recent years methods like Refrence system method, Decativated Morphing \cite{Park2008}, Orthogonal Space Random
 Work, Confinement Method have came out for calculation of free energy method. Some of the groups
@@
 
-In this work we have applied the confinement method which was originally devoled by Tyka et. al. and
-Cecchini et. al.  This method relies on a thermodynamic cycle where first, non-harmonic degrees of
+In this work we have applied the confinement method which was originally devoled by Tyka et. al. \cite{Tyka2006} and
+Cecchini et. al. \cite{Cecchini2009}. This method relies on a thermodynamic cycle where first, non-harmonic degrees of
 freedom are removed by applying a series of restraint to the system. Then the free energy of the
@@
 The method has previously been applied to a 16 amino acid residue $\beta$-hairpin from
-protein G, known as BHP[ref]. We calculated the free energy difference between two different
+protein G, known as BHP \cite{Cecchini2009}. We calculated the free energy difference between two different
 conformations of the peptide: (1) the native conformation, called bhp1, with a two-stranded
 $\beta$-sheet; and (2) a conformation, called bhp3, which has a three-stranded $\beta$-sheet.
-Analysis of long (4 $\mu$s) equilibrium simulations[ref] shows that bhp1 is the more favorable
+Analysis of long (4 $\mu$s) equilibrium simulations \cite{Cecchini2009},\cite{Krivov2004}  shows that bhp1 is the more favorable
 configuration by 1.8 kcal/mol. Using the confinement method, we obtain a value 1.7 kcal/mol, which
 is in good agreement with the equilibrium simulations and with previous calculations using the
-confinement method[ref] (see Supporting Information for further details).
+confinement method \cite{Cecchini2009} (see Supporting Information for further details).
 
@@
 
-Previous applications of the confiment method have focused on relatively short peptides, up to X
-residues in length \Justin{Arijit: fill in X}. In this work we apply the confinement method to
+Previous applications of the confiment method have focused on relatively short peptides, up to $17$
+residues in length. In this work we apply the confinement method to
 larger proteins. One of the  applications we explore in the present work is the re-scoring, or
@@
 (CASP) experiment (described in detail later). We computed the relative free energies of predictions
-submitted during CASP9 and 10, with the expectation that the most native-like prediction will have
+submitted during CASP9 10, with the expectation that the most native-like prediction will have
 the lowest free energy. As a control, for several targets we also calculated the relative free
@@
     CASP Target  & PDB Identifier & $\Delta\Delta G_{native \to best decoy}$ (kcal/mol) \\ \hline
+     T0531       &    2KJX        &          $12.49 \pm 0.70$ \\ \hline
+     T0538       &    2L09        &          $7.70 \pm 0.47$ \\ \hline
+     T0540       &    3MX7        &          $22.00 \pm 0.49$ \\ \hline
+     T0559       &    2L01        &          $2.24 \pm 0.24$ \\ \hline
+     T0560       &    2L02        &          $22.15 \pm 0.49$ \\ \hline 
 \end{tabular}
@@
 remarkable. Accurately predicting the structural preferences of these structures presents a serious
-challenge for computational methods [ref].
+challenge for computational methods \cite{Alexander2007}$-$\cite{Shortle20009}.
 
@@
 around 2.9 to 5.7 kcal/mol, which is consistent with small changes in sequences and with estimates
-made by Orban and co-workers [ref].
+made by Orban and co-workers \cite{He2008}, \cite{{Alexander2009}}.
 
 \begin{figure}
+\includegraphics[width=14cm,height=14cm]{orban.pdf}
 \label{fig:orban}
-
 \caption{The confinement method correctly predicts the structural preferences of six chameleon
@@
 the two folds are reported for each sequence.}
-
-\Justin{Arijit: we need to make this figure, which we've discussed before and move the data from the
-following table into the figure.}
 \end{figure}
@@
 calculation. According to the GDT\_TS score, the model 1 was picked up correctly from the set of
-structure generated . However, the order of model 3 and 5 was not correct as can be seen from the
-Table 1. It will be interesting to know whether confinement method can order the strcutures in terms
-of free energy scale. The result is presented in Table~\protect\ref{tab:T0559}.
+structure generated . However, the order of model 3 and 5 was not correct.
+It will be interesting to know whether confinement method can order the strcutures in terms
+of free energy scale. The result is presented in Figure~\ref{fig:T0559}.
 
@@
 the group "Splicer" along with the crystal structure.  The main difference between the model 2 and 4
-is the orientation of the ....  The final result is presented in Table~\protect\ref{tab:T0560}.
+is the orientation of the ....  The final result is presented in Figure~\ref{fig:T0560}.
 
@@
 submitted by the group "Shell", "FOLDIT" and "BAKER-ROSETTASERVER". The calculated free energy value
-is presented in Table~\protect\ref{tab:T0538}.
+is presented in Figure~\ref{fig:T0538}.
 
@@
 $69.72$ and model 2 with GDT\_TS value of $53.789$. The final result is presented in
-Table~\protect\ref{tab:T0540}.
+Figure~\ref{fig:T0540}.
 
@@
 "MUFOLD-MD", which was one of the best performed group for this target.  The result is presented in
-Table~\protect\ref{tab:T0531}
+Figure~\ref{fig:T0531}.
 
@@
 final calculation. The rest two model had closer GDT\_TS value compared to the previously mentioned models and thus they were omitted from
-furthe calculation. The result is presented in Table~\protect\ref{tab:T0605}. (This is a case where we can demand the confinement
+furthe calculation. The result is presented in Figure~\ref{fig:T0605}. (This is a case where we can demand the confinement
 method is better than GDT\_TS).
@@
 \begin{center}
-\includegraphics[width=12cm,height=10cm]{G88.pdf}
-\end{center}
-\caption{Two protein with 88 \% similar sequence but different folds}
-\label{fig:G88}
-\end{figure}
-
-\begin{figure}
-\begin{center}
 \includegraphics[width=12cm,height=10cm]{T0559.pdf}
@@
 \end{figure}
-
-\begin{table}
-\caption{The value of free energy and GDT\_TS for the CASP9 target T0559}
-\label{tab:T0559}
-\begin{center}
-\begin{tabular}{l l l}\hline
-Model   &     $\Delta G$ (Kcal/Mol) &  GDT\_TS \\ \hline
-Model 1 &     0.000            &  94.03    \\
-Model 3 &     $14.484 \pm 0.633$ &  48.13    \\
-Model 5 &     $4.526  \pm 0.384$ &  82.46    \\ \hline
-\end{tabular}
-\end{center}
-\end{table}
 
@@
 
-\begin{table}
-\caption{The Free energy and GDT\_TS values of the CASP9 Target T0538}
-\label{tab:T0538}
-\begin{center}
-\begin{tabular}{l l l l}\hline
-Model   &     $\Delta G$ (Kcal/Mol) &  GDT\_TS & Group Submitted \\ \hline
-Model 1 &     0.000              &  96.23    & PconsR  \\
-Model 2 &     $15.891 \pm 0.475$ &  90.09  & Shell  \\
-Model 3 &     $23.243 \pm 0.393$ &  86.32  & FOLDIT  \\
-Model 4 &     $26.002 \pm 0.756$ &  80.66  & BAKER-ROSETTASERVER \\ \hline
-\end{tabular}
-\end{center}
-\end{table}
-
 \begin{figure}
@@
 
-\begin{table}
-\caption{The free energy and the GDT\_TS value of the CASP9 target T0560.}
-\label{tab:T0560}
-\begin{center}
-\begin{tabular}{l l l}\hline
-Model   &     $\Delta G$ (Kcal/Mol) &  GDT\_TS \\ \hline
-Crystal &     0.000              & 100.00    \\
-Model 2 &     $22.153 \pm 0.491$ &  94.14    \\
-Model 4 &     $27.428  \pm 0.477$ &  76.56    \\ \hline
-\end{tabular}
-\end{center}
-\end{table}
-
 \begin{figure}
@@
 \caption{T0540}
-\label{fig:}
+\label{fig:T0540}
 \end{figure}
-
-\begin{table}
-\caption{The free energy and the GDT\_TS value of the CASP9 target T0540.}
-\label{tab:T0540}
-\begin{center}
-\begin{tabular}{l l l l}\hline
-Model   &     $\Delta G$ (Kcal/Mol) &  GDT\_TS & Group Submitted \\ \hline
-Crystal &     0.000              &  100.00   & -      \\
-Model 1 &     $22.003 \pm 0.575$ &  69.72    & LTB    \\
-Model 2 &     $53.786 \pm 0.553$ &  63.89    & Mufold \\ \hline
-\end{tabular}
-\end{center}
-\end{table}
 
@@
 
-\begin{table}
-\caption{The free energy and the GDT\_TS value of the CASP9 target T0531.}
-\label{tab:T0531}
-\begin{center}
-\begin{tabular}{l l l}\hline
-Model   &     $\Delta G$ (Kcal/Mol) &  GDT\_TS \\ \hline
-Crystal &     0.000              & 100.00    \\
-Model 1 &     $12.496 \pm 0.704$ &  43.53    \\
-Model 2 &     $27.374 \pm 0.691$ &  34.48    \\
-Model 3 &     $13.914 \pm 0.673$ &  34.48    \\
-Model 4 &     $53.055 \pm 0.684$ &  28.02    \\
-Model 5 &     $29.508 \pm 0.698$ &  25.86   \\ \hline
-\end{tabular}
-\end{center}
-\end{table}
-
 \begin{figure}
@@
 \end{figure}
-
-\begin{table}
-\caption{The free energy and the GDT\_TS value of the CASP9 target T0605.}
-\label{tab:T0605}
-\begin{center}
-\begin{tabular}{l l l}\hline
-Model   &     $\Delta G$ (Kcal/Mol) &  GDT\_TS \\ \hline
-Model 1 &     $0.000$ &  $97.96$    \\
-Model 2 &     $1.689 \pm ?$ &  $89.29$    \\
-Model 5 &     $16.064 \pm ?$ & $93.88$   \\ \hline
-\end{tabular}
-\end{center}
-\end{table}
 
@@
 
-\bibliographystyle{achemso}
-\bibliography{paper}
+\begin{thebibliography}{99}
 
+\bibitem{Dill1997}
+Dill, K.A., and H.S. Chan.  From Levinthal to Pathways to Funnels:  The "New View" of Protein Folding Kinetics.  Nature Structural Biology 4: 10-19 (1997)
+
+\bibitem{Ytreberg2006}
 Ytreberg, F.; Zuckerman, D. Simple estimation of absolute free energies for biomolecules. J. Chem. Phys. 2006, 124, 104105.
 
+\bibitem{Park2008}
 Park, S.; Lau, A.; Roux, B. Computing conformational free energy by deactivated morphing. J. Chem. Phys. 2008, 129, 134102
 
+\bibitem{Zheng2008}
 Zheng, L.; Chen, M.; Yang, W. Random walk in orthogonal space to achieve efficient free-energy simulation of complex systems, Proc. Natl. Acad. Sci. 2008, 105 (51), 20227.
 
-Tyka, M.; Clarke, A.; Sessions, R. An Efficient, Path-Independent Method for Free-Energy Calculations. J.Phys.Chem. B 2006, 110, 17212–17220.
+\bibitem{Tyka2006}
+Tyka, M.; Clarke, A.; Sessions, R. An Efficient, Path-Independent Method for Free-Energy Calculations. J.Phys.Chem. B 2006, 110, 17212-17220.
+
+\bibitem{Cecchini2009}
+Cecchini, M., Krivov, S.V., Spichty, M., Karplus, M. Calculation of free-energy differences by confinement simulations. Application to peptide conformers. 
+J. Phys. Chem. B 113, p. 9728-9740 (2009).
+
+\bibitem{Krivov2004} 
+Krivov, S.; Karplus, M. Hidden complexity of free energy surfaces for peptide (protein) folding Proc. Natl. Acad. Sci. U.S.A. 2004, 101, (41), 14766.
+
+\bibitem{Alexander2007}
+Alexander, P.A.; He, Y.; Chen, Y.; Orban, J. Bryan, P. The design and characterization of two proteins with $88 \%$ sequence identity but different 
+structure and function. Proc. Natl. Acad. Sci. 2007, 104 (29), 11963-11968.
+
+\bibitem{He2008}
+He, Y.; Chen, Y.; Alexander, P.A.; Orban, J. NMR structures of two designed proteins with high sequence identity but different fold and function. Proc. Natl. Acad. Sci. 2008, 105 (38), 14412-14417.
+
+\bibitem{Alexander2009}
+Alexander, P.A.; He, Y.; Chen, Y.; Orban, J. Bryan, P. A minimal sequence code for switching protein structure and function. Proc. Natl. Acad. Sci. 2009, 106(50), 21149-21154.
+
+\bibitem{Shortle20009}
+Shortle, D. One sequence plus one mutation equals two folds. Proc. Natl. Acad. Sci. 2009, 106(50), 21011-21012.
+
+\bibitem{Sheffler2009}
+Sheffler, W.; Baker, D. RosettaHoles: Rapid assessment of protein core packing for structure prediction, refinement, design, and validation. Protein Science. 2009, 18(1), 229-239.
+
+\end{thebibliography}
 

### 2012-10-04 Paper/paper.tex
Rewrite of CASP section

--- 2012-09-09 Paper/paper.tex
@@
 subset of CASP predictions}
+\Alberto{Arijit: we need a global table of success, rather than individual tables. The simplest form
+    would have 3 columns: name of the PDB, number of structures tested, percent of success of
+    picking 1st model}
+\Alberto{Arijit: CASP notation should not be used (T????) instead, pdb code should be used}
 
-Next, we have applied the confinement method to several targets from the CASP9 and CASP10
-experiments. CASP is a world-wide competition aiming at finding the state of the art methodologies
-in protein folding. Groups, both from knowledge based methods and physics based methods participate
-in this blind competition and for each sequence target they can submit five models. The predictors
-were also asked to calculate their best model among the set of five they submitted. However, in many
-cases predictors were unable to predict/rank their models. Using the confinement method, we aim to
-rank protein structures using free energy as a scale. In recent CASP experiments Global distance
-test scose (GDT\_TS) method is used for final analysis of different submitted structures with the
-NMR/crystallographically solved protein native structure. GDT\_TS represent the average percentage
-of residues that are in close proximity in two structures optimally superimposed using four
-different distance cutoff. Although GDT\_TS alone is not a reliable measure for the difficult
-modeling cases, it has been well accepted method in recent years among the community. We have also
-compared our free energy result with the GDT\_TS score for different cases we studied.
+This method is particularly well suited for events such as CASP (critical assessment of structure
+prediction). CASP is a blind test world wide competition in which different groups apply their
+methods to predict the 3D structure of proteins given their sequence. This is done with a strict 3
+week limit on each target (3 day for servers) and each group is allowed to submit 5 possible
+structures (ranked from better to worst). In our role as assessors during the CASP8 and CASP9 [CITE]
+competition we observed no correlation at all between the ordering of structures submited by the
+groups and the real ranking compared to an experimental model. The consequences of this go beyond
+those five structures; the deeper meaning is that groups producing ensembles of structures are
+generating structures that are better than the ones they submit, but they do not know about it. In
+fact, when querrying different groups after the results were known, most groups agree that this is
+the case. Beryond the CASP problem, this reflects on the modelers ability to correctly rank order
+models in many different environments, from structures for drug design leads to designing more
+stable proteins or peptide mimetics such as peptoids. One of the main culprits of this lack of
+accuracy is the fact that ranking is often done via a potential energy function, which in many cases
+lacks an entropy component. Other initiatives including knowledge based potentials do use some sort
+of free energy to rank order, but its accuracy is not enough. Our method provides a physics based
+solution to this problem. We have tried two experiments centered on CASP. First we tried to rank
+order some structures from the previous CASP9 experiment. Our second test was to participate as a
+metapredictor group in CASP10. In both methods, our ranking is determined as a free energy measure,
+and it is compared with the ranking given by a geometrical comparison between native and submitted
+models called GDT\_TS (Global distance test score [CITE, look in CASP9 refinement paper for cite
+21]). GDT\_TS represent the average percentage of residues that are in close proximity in two
+structures optimally superimposed using four different distance cutoffs (1, 2, 4 and 8 Å).
+
+We have chosen the inital models on which to test the methods based on server groups that have
+traditionally done good in past CASP events. There is a high correlation between the structures that
+are chosen for this method and its predictive accuracy. In particular, when GDT\_TS scores are below
+50, it is difficult for us to say anything about the models. Our approach is simple, with every
+subset of structures we choose, we use the confinement method in a pairwise procedure, and then rank
+the structures. Note that we do not need to make comparisons to native in order to rank order.
+
+\Subsubsection{CASP9}
+Can we distinguish between native and models? Identifying the native state from a set of models is a very easy test, that most metapredictors
+already handle correctly, but it is also a first proof of principle. We exemplify this with
+T0660\Alberto{Arijit:PDB code please, here can you complete the text in subsection T0560 or added it here
+    directly?} by selecting the native state and comparing it to two models from the "Splicer"
+server. Here, the main difference between the two models is the orientation of the ... As expected,
+the native state is identified correctly, and the two other models is correctly identified. Our next test was to see if we could rank order within a given server's predictions. We chose
+T0559 \Alberto{Arijit: change to PDB code, say CASP code T0559} and the BAKER-ROSETTASERVER. We
+initlally checked similarity between the models and discarded two of them from the analysis on the
+basis of being too similar to some of the other models. For the remaining structures, we correctly
+ranked order the results as seen from Table~\protect\ref{tab:T0559}. This was a relatively easy test
+as the differences in GDT\_TS score are fairly large. Therefore, our next test was between models
+from the best submissions of different groups (range of GDT\_TS
+from 96.23 to 80.66). Again, our method was able to correctly order the results
+(Table~\protect\ref{tab:T0538}). \Alberto{I don't like so many small tables. Have to find out a way
+    to combine into something bigger, so can check this whole paragraph in one table or figure.}
+There are however some instances where the method can fail, specially when close GDT scores are
+given. This is the case in T605, Table~\protect\ref{tab:T0605}, where the first model from the BAKER
+group is correctly identified, put the 2nd and 3rd are given in inverse order. \Alberto{Airjit:
+    Reasons for this???}
+
+What can we say about low resolution models?
+We mentioned earlier that our predictive abilities are greatly decreased with low model qualities.
+We performed such a test with T0531, in which we compared the crystal structure to five models from
+the MUFOLD server. Looking at table ~\protect\ref{tab:T0531} shows: 1.) native is correctly
+identified as expected and 2.) Surprisingly there is a high level of correlation between the GDT
+ranking and the free energy ranking, only two structures whith GDT\_TS < 30Å are ordered
+incorrectly. It is worth to note that models 2 and 3 have the same GDT and very different free
+energies, meaning that the actual ordering could change a lot [Cite FlexE].
+
+\Subsubsection{CASP10}
+We participated in CASP10 as a metapredictor group. Given the tigh schedule of the CASP competition,
+we only attempted proteins under 100 residues in length. Our main problem during this part of the
+competition was that proteins were trimmed from the actual calculation.
+\Alberto{Justin,Arijit: Probably should not comment on metapredicting since we were not able to do
+    good due to the trimming factor. Instead, we should move forward with Justin's idea of using
+    structures that other meta predictors did badly at. What stage is that in?}
 
@@
 
+\Alberto{This subsections have to go. I'm momentarily leaving them on so you can complete the text
+    with number of residues and maybe experimental origin. I would put the number of residues in the
+    tables}
 \subsubsection{T0559}
@@
 \subsubsection{T0540}
+\Alberto{Arijit: between T0560 and T0540, which one is better? I would keep only one of them, and my
+    preference is T0560, we are showing the same test}
 

### 2012-10-19 Paper/bibl.tex
third draft arijit

--- 2012-10-01 Paper/paper.tex
@@
-% vim:set textwidth=100:
-% vim:set fo+=t:
-
-\documentclass[12pt]{article}
-
-\usepackage{amsmath}
-\usepackage[nofiglist,notablist]{endfloat}
-\usepackage[usenames,dvipsnames]{color}
-\usepackage{color}
-\usepackage{authblk}
-\usepackage{graphicx}
-\usepackage{palatino}
-\usepackage[activate={true,nocompatibility},final]{microtype}
-\usepackage[super,sort&compress]{natbib}
-\pagenumbering{arabic}
-\parskip = 0.08in \parindent = 0.0in
-
-% Custom macros for author comments
-\newcommand{\Alberto}[1]{\color{ForestGreen}#1\normalcolor }
-\newcommand{\Justin}[1]{\color{blue}#1\normalcolor}
-\newcommand{\Arijit}[1]{\color{yellow}#1\normalcolor}
-\newcommand{\Ken}[1]{\color{red}#1\normalcolor}
-
-\author{Arijit~Roy}
-\author{Justin~L.~MacCallum}
-\author{Alberto~Perez}
-\author{Ken~A.~Dill}
-\affil{Laufer Center for Physical and Quantitative Biology\\
-    Stony Brook University\\
-    Stony Brook, NY 11794-5252.}
-
-\title{Physics-based protein structure prediction and design using the confinement method}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-
-The calculation of free energy differences is of central importance in the simulation of biochemical
-systems.  The computation of the free energy difference between pairs of macromolecule with large
-conformational change is particularly a difficult task and computationally expensive with existing
-methods. In this work, an improved version of the confinement approach is used to calculate absolute
-free energies of bio molecular systems. The method does not require a reaction coordinate or
-transition path. It is fast to compute. We show that the method correctly picks out the state of
-lowest free energy of a pair of structures having similar sequence but different fold. We also show
-using models from CASP9 that the method picks out native-like structures from misfolded or decoys,
-provided at least one good structure is in the input set.
-
-\end{abstract}
-
-\section{Introduction}
-
-In computational structural biology there are numerous cases where free energy between well defined
-states are necessary. Examples ranges from small to large conformational change of protein due to
-ligand binding, change of pH etc. Free energy also play an important role in case of protein
-folding. The ground breaking work of Christian B. Anfinsen and coworkers showed that the native
-structures of small globular proteins have a unique, thermodynamically stable native structure with
-their conformation at the global free energy minimum. Regardless of the starting point or unfolding
-it by changing different condition most proteins will finally assume the same structure. It is very
-important for proteins to achieve their native conformation since defects in protein folding may be
-the molecular cause of a range of human genetic disorder. For example a misfolded protein known as
-prion when enters any healthy organism it converts properly folded protein into misfolded one.
-Thus, Free energy can also act as a scale to identify between the misfolded and the native state of
-the protein. These observations further give free energy a special importance in structural biology.
-
-However, the theoretical calculation of conformational free energy change become difficult if the
-states of interest are very different. In such scenario timescale involved in such tranformation may
-be beyond timescale of direct classical molecular dynamics simulation.  Generally, in order to
-calculate free energy a pathway or reaction coordinate is created with the intermediate structures.
-Method like umbrella sampling, thermodynamic integration along with classical MD can suffer from
-overlap problem and require large computational effort. The idea of reaction coordinate became more
-complicated in case of protein folding.  As free energy landscape of the protein folding is rough \cite{Dill1997},
-it can be trapped in a energy well during its visit in the conformational space. Even with the
-advanced method like replica exchange molecular dynamics the three dimensional structure may be
-trapped in a local minima for a considerable time. Due to various reasons it can also be misfolded
-and trapped in a local energy minima. And one can mistakenly identify the misfolded state as the
-native state (give the example of human pin1ww domain from Benout Roux and Klaus Schultan).
-
-Calculation of free energy has been successfully attempted by a number of groups \cite{Ytreberg2006} - \cite{Zheng2008}. In
-recent years methods like Refrence system method, Decativated Morphing \cite{Park2008}, Orthogonal Space Random
-Work, Confinement Method have came out for calculation of free energy method. Some of the groups
-relies on one of the great strength of free energy calculation, that it is a state function and thus
-it is independent of path.
-
-In this work we have applied the confinement method which was originally devoled by Tyka et. al. \cite{Tyka2006} and
-Cecchini et. al. \cite{Cecchini2009}. This method relies on a thermodynamic cycle where first, non-harmonic degrees of
-freedom are removed by applying a series of restraint to the system. Then the free energy of the
-remaining harmonic system is obtained using a normal mode calculation and combined with the
-non-harmonic part to give the total absolute free energy.  Previously, this method was applied to
-calculate the conformational free energy of a 16 amino acid residue peptide, known as BHP.  We have
-first reproduced that known result. Then we have applied this method to larger proteins for the
-first time. The free energy of conformation of pair of proteins with similar sequence but completely
-different fold was calculated. Encouraged by the result, we applied the method of different targets
-of CASP9 competition.
-
-%of protein can be obtained to atomic resolution by X-ray crystallography or NMR. Unfortunately, the
-%sheer amount of time and economic investment for structure determination cannot keep up with the
-%rate at which proteins of unknown structure are sequenced. Therefore, computational models can be a
-%good way to bridge the increasing gap between sequence and structure. In this respect CASP
-%(Critical Assessment of Structure Prediction) has played a crucial role which is a community wide
-%experiments for the protein structure prediction.
-
-\section{Results and Discussion}
-
-\subsection{The confinement method produces correct results in control experiments}
-
-As a first step, we performed several control experiments to verify that our implementation of the
-confinement method produces results compatible with previous calculations reported in the
-literature.
-
-The method has previously been applied to a 16 amino acid residue $\beta$-hairpin from
-protein G, known as BHP \cite{Cecchini2009}. We calculated the free energy difference between two different
-conformations of the peptide: (1) the native conformation, called bhp1, with a two-stranded
-$\beta$-sheet; and (2) a conformation, called bhp3, which has a three-stranded $\beta$-sheet.
-Analysis of long (4 $\mu$s) equilibrium simulations \cite{Cecchini2009},\cite{Krivov2004}  shows that bhp1 is the more favorable
-configuration by 1.8 kcal/mol. Using the confinement method, we obtain a value 1.7 kcal/mol, which
-is in good agreement with the equilibrium simulations and with previous calculations using the
-confinement method \cite{Cecchini2009} (see Supporting Information for further details).
-
-\Justin{Arijit, Al: did we do any other control experiments on small peptides that we should report
-here?}
-
-Previous applications of the confiment method have focused on relatively short peptides, up to $17$
-residues in length. In this work we apply the confinement method to
-larger proteins. One of the  applications we explore in the present work is the re-scoring, or
-metaprediction, of structures submitted during the Critical Assessment of Structure Prediction
-(CASP) experiment (described in detail later). We computed the relative free energies of predictions
-submitted during CASP9 10, with the expectation that the most native-like prediction will have
-the lowest free energy. As a control, for several targets we also calculated the relative free
-energy of the experimental structure (once it was available from the PDB). As
-Table~\ref{table:casp_control} shows, in X of Y cases, the confinement method correctly assigns a
-lower free energy to the experimentally determined structure than to any of the decoys. Although
-differentiating between experimentally determined structures and computer generated predictions is
-not a stringent test of a scoring method---for example, examining the residue-residue packing can
-easily distinguish between the two [ref Rosetta Holes]---the results none the less serve as a useful
-"sanity check".
-
-\begin{table}
-\label{table:casp_control}
-
-\begin{center}
-\begin{tabular}{l l l}\hline
-    CASP Target  & PDB Identifier & $\Delta\Delta G_{native \to best decoy}$ (kcal/mol) \\ \hline
-     T0531       &    2KJX        &          $12.49 \pm 0.70$ \\ \hline
-     T0538       &    2L09        &          $7.70 \pm 0.47$ \\ \hline
-     T0540       &    3MX7        &          $22.00 \pm 0.49$ \\ \hline
-     T0559       &    2L01        &          $2.24 \pm 0.24$ \\ \hline
-     T0560       &    2L02        &          $22.15 \pm 0.49$ \\ \hline 
-\end{tabular}
-\end{center}
-
-\caption{The confinement method assigns a more favorable free energy to the experimentally
-determined structure than to computer-generated predictions. For each target, we examined as many as
-five predictions submitted by CASP participants. Positive $\Delta\Delta G$ values indicate that the
-experimental structure is predicted to be more favorable than any of the decoys.}
-
-\end{table}
-
-\subsection{The confinement method correctly predicts the structural preferences of chameleon
-sequences}
-
-In general, proteins with similar sequences tend to have similar structures. This idea is the basis
-of comparative modeling and fold recognition in protein structure prediction. There are, however,
-examples---often referred to as chameleon sequences---of proteins with similar sequences that have
-remarkably different structures. Orban and co-workers have designed a series of 56-residue proteins
-(based on Protein G) that adopt one of two different folds depending on small changes in sequence
-(see Figure~\ref{fig:orban}). Sequences that adopt a mixed alpha/beta structure similar to Protein G
-are denoted as ``GB'', while sequences that form a three-helix bundle are denoted as ``GA''. One
-pair of sequences (GA88/GB88) are 88 percent identical and differ in seven positions. Another pair
-(GA95/GA95) are 95 percent identical and differ in three positions. The last pair (GA98/GB98) differ
-only in a single tyrosine to alanine mutation.
-
-The fact such small changes in sequence can lead to such dramatic changes in structure is rather
-remarkable. Accurately predicting the structural preferences of these structures presents a serious
-challenge for computational methods \cite{Alexander2007}$-$\cite{Shortle20009}.
-
-We initially approached this problem by making a model of each sequence with the same backbone
-structure as its partner chameleon sequence. For example, we took the sequence of GA88 and built a
-model with the same overall structure as GB88. We then used the confinement method to assess the
-free energy difference between the experimentally determined structure of GA88 and the model (with
-the GA88 sequence and the GB88 structure). The confinement method was able to predict the
-conformational preferences correctly for all six sequences (data not shown). There is, however, a
-serious problem with this analysis: we are comparing an experimentally determined structure with a
-computer generated model. It might be possible that we were able to make correct predictions simply
-because artefacts of our modeling procedure always lead to the model having a higher free energy
-than the experimental structure.
-
-To avoid this potential problem, we instead computed the relative free energy two different
-structural models for each sequence. One model is based on the GA structure and the other on the GB
-structure (see Supporting Information for details on the modeling procedure). This is a much more
-realistic test of the confinement method's ability to accurately calculate relative free energies.
-
-The results of these calculations are presented in Figure~\ref{fig:orban}. The confinement method
-identifies the correct structure for all six sequences. The free energy differences range from
-around 2.9 to 5.7 kcal/mol, which is consistent with small changes in sequences and with estimates
-made by Orban and co-workers \cite{He2008}, \cite{{Alexander2009}}.
-
-\begin{figure}
-\includegraphics[width=14cm,height=14cm]{orban.pdf}
-\label{fig:orban}
-\caption{The confinement method correctly predicts the structural preferences of six chameleon
-sequences. (A) The six sequences used in this study. (B) Each sequence adopts either a Protein
-G-like fold (denoted GB) or a three-helix bundle fold (denoted GA). The relative free energies of
-the two folds are reported for each sequence.}
-\end{figure}
-
-%\begin{table}
-%\caption{The confinement method correctly reproduces the structural preferences of a series of
-%designed peptides. Each sequence adopts either a Protein G-like fold ($4 \beta + \alpha$) or forms a
-%helix bundle ($3 \alpha$).}
-%\label{table:orban}
-%\begin{center}
-%\begin{tabular}{l l l}\hline
-%Changes in & $\Delta G$ (kcal/mol)    & $\Delta G$ (kcal/mol)     \\
- %Sequences & ($3 \alpha$ fold)        & ($4 \beta + \alpha$ fold) \\ \hline
-%GA95       & $0.000$                  & $2.901 \pm 0.475$         \\
-%GB95       & $4.365 \pm 0.460$        & $0.000$                   \\
-%GA88       & $0.000$                  & $ 3.941 \pm 0.517$        \\
-%GB88       & $5.576 \pm 0.464$        & $0.000$                   \\
-%GA98       & \Justin{Arijit: fill in} &                           \\ \hline
-%GB98       &                          &                           \\ \hline
-%\end{tabular}
-%\end{center}
-%\end{table}
-
-\subsection{The confinement method correctly identifies the most native-like predictions from a
-subset of CASP predictions}
-
-Next, we have applied the confinement method to several targets from the CASP9 and CASP10
-experiments. CASP is a world-wide competition aiming at finding the state of the art methodologies
-in protein folding. Groups, both from knowledge based methods and physics based methods participate
-in this blind competition and for each sequence target they can submit five models. The predictors
-were also asked to calculate their best model among the set of five they submitted. However, in many
-cases predictors were unable to predict/rank their models. Using the confinement method, we aim to
-rank protein structures using free energy as a scale. In recent CASP experiments Global distance
-test scose (GDT\_TS) method is used for final analysis of different submitted structures with the
-NMR/crystallographically solved protein native structure. GDT\_TS represent the average percentage
-of residues that are in close proximity in two structures optimally superimposed using four
-different distance cutoff. Although GDT\_TS alone is not a reliable measure for the difficult
-modeling cases, it has been well accepted method in recent years among the community. We have also
-compared our free energy result with the GDT\_TS score for different cases we studied.
-
-%The GDTTS value of a model is determined as follows. A large sample of possible structure
-%superpositions of the model on the corresponding experimental structure is generated by superposing
-%all sets of three, five, and seven consecutive C$\alpha$ atoms along the backbone (each peptide
-%segment provides one superposition). Each of these initial superpositions is iteratively extended,
-%including all residue pairs under a specified threshold in the next iteration, and continuing until
-%there is no change in included residues. The procedure is carried out using thresholds of 1, 2, 4,
-%and 8 $\AA$, and the final superposition that includes the maximum number of residues is selected
-%for each threshold.
-
-\subsubsection{T0559}
-
-The native structure is a  67 residue NMR solved structure. We first reviewed all the five models
-submitted by the group "BAKER-ROSETTASERVER", which was the best predictor group for this particular
-target.  Among them we choose only the three structures, model 1, 3 and 5 to calculate the free
-energy and subsequently identify the most stable structure among them. The rest of the two
-structures are close to either of the selected structure and thus they are ommited from our
-calculation. According to the GDT\_TS score, the model 1 was picked up correctly from the set of
-structure generated . However, the order of model 3 and 5 was not correct.
-It will be interesting to know whether confinement method can order the strcutures in terms
-of free energy scale. The result is presented in Figure~\ref{fig:T0559}.
-
-\subsubsection{T0560}
-
-Target T0560 is a 74 amino acid residue target from CASP9. However residues 3-66 were kept for final
-result.  For this target we have compared the free energy of the model 2 and model 4 submitted by
-the group "Splicer" along with the crystal structure.  The main difference between the model 2 and 4
-is the orientation of the ....  The final result is presented in Figure~\ref{fig:T0560}.
-
-\subsubsection{T0538}
-
-Target T0538 was a 54 residue target. The native structure was generated using NMR. The best model
-that is closest to the native structure was generated by the group PconsR. However, their model 3
-was better than their model 1. This structure was quite close to the NMR structure and had GDT\_TS
-value of 96.23. We have calculated the free energy between this structure and the best structure
-submitted by the group "Shell", "FOLDIT" and "BAKER-ROSETTASERVER". The calculated free energy value
-is presented in Figure~\ref{fig:T0538}.
-
-\subsubsection{T0540}
-
-Target T0540 was a 90 residue target. We have taken the best models from the group "LTB" which was
-the best performing group for this target and also from the group Mufold and calculated the free
-energy difference between these models with that of the native crystal structure.  As expected, the
-crystal structure has the lowest free energy value, followed by Model 1 with GDT\_TS score of
-$69.72$ and model 2 with GDT\_TS value of $53.789$. The final result is presented in
-Figure~\ref{fig:T0540}.
-
-\subsubsection{T0531}
-
-This is a 65 amino acid residue target with residues 6-63 were kept for final analysis. We have
-calculated the free energy difference between all the five models submitted by the group
-"MUFOLD-MD", which was one of the best performed group for this target.  The result is presented in
-Figure~\ref{fig:T0531}.
-
-\subsubsection{T0605}
-
-This is a 72 amino acid residue protein but for the final calculation only the residues 18-66 amino acids were kept for the final analysis.
-
-The best performing group was "Baker" for this target. We have carried out our calculation with the model first, second and fifth for the
-final calculation. The rest two model had closer GDT\_TS value compared to the previously mentioned models and thus they were omitted from
-furthe calculation. The result is presented in Figure~\ref{fig:T0605}. (This is a case where we can demand the confinement
-method is better than GDT\_TS).
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=14cm,height=14cm]{method.pdf}
-\end{center}
-\caption{Graphical representation of the thermodynamic cycle involving confinement method.}
-\label{fig:method}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=12cm,height=10cm]{bhp.pdf}
-\end{center}
-\caption{Two conformations from $\beta$ hairpin from protein G, bhp1 and bhp2. The two stranded $\beta$ sheet, bhp1 is the native structure and the three stranded $\beta$ sheet is known as bhp3.}
-\label{fig:bhp_conf}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=12cm,height=10cm]{G95.pdf}
-\end{center}
-\caption{Two protein with 95 \% similar sequence but different folds}
-\label{fig:G95}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=12cm,height=10cm]{T0559.pdf}
-\end{center}
-\caption{Target T0559 from CASP9}
-\label{fig:T0559}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=12cm,height=10cm]{T0538.pdf}
-\end{center}
-\caption{Target T0538}
-\label{fig:T0538}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=12cm,height=10cm]{T0560.pdf}
-\end{center}
-\caption{Target T0560}
-\label{fig:T0560}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=12cm,height=10cm]{T0540.pdf}
-\end{center}
-\caption{T0540}
-\label{fig:T0540}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=12cm,height=10cm]{T0531.pdf}
-\end{center}
-\caption{T0531}
-\label{fig:T0531}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=12cm,height=10cm]{T0605.pdf}
-\end{center}
-\caption{T0605}
-\label{fig:T0605}
-\end{figure}
-
-\section{Conclusion}
-
-\section{Method}
-
-The confinement method has been described in details in ref. by Tyka et.al. and Cecchini et. al.
-Here we briefly describe the methodology.  We compute free energy $\Delta G_{AB}$ between A and B
-conformation of the protein. For this purpose we use a thermodynamic cycle.
-
-\begin{enumerate}
-
-\item  We minimized both the conformation. These minimized conformation(A* and B*) are the
-    restrained conformation of that state
-
-\item Backbone dihedral angle of the whole protein is restrained by a harmonic potential to define
-    the configurational state.
-
-\item  In order to calculate the free energy $\Delta G_{A,A*}$ and  $\Delta G_{B,B*}$ A and B are
-    gradually transformed in A* and B* by applying a harmonic restraint potential to all atom. For
-    this purpose around 21 simulation were run each of 20 ns with increasing harmonic restraint
-    potential from 0.00005 to 81.92 untill the free and the restrained state overlap well. The final
-    restrained state was chosen so high so that the rotational contributional of the protein frozen
-    out and only remaining contribution remain that of vibrational free enrgy. The fluctuation from
-    the reference structure was recorded and the free energy is calculated using a numerical
-    approach developed by Tyka et. al.
-
-\item  Finally the thermodynamic cycle is closed by calculating the free energy between the final
-    restrained state using normal mode analysis.
-
-\end{enumerate}
-
-In all the calculation amber 10 program is used with ff99SB with GB/SA implicit solvent.
-
 \begin{thebibliography}{99}
 
+\bibitem{Meirovitch2007}
+Meirovitch, H. Recent developments in methodologies for calculating the entropy and free energy of biological systems by computer simulation.
+Current Opinion in Structural Biology, 2007, 17, 181-186.
+
+\bibitem{Chipot2007}
+Chipot, C.; Shell, M.S.; Pohorille, A. Introduction, in Chipot, C., Pohorille, A., editors. Free Energy
+Calculations: Theory and Applications in Chemistry and Biology. Springer Series in Chemical
+Physics, vol. 86. Berlin and Heidelberg: Springer; 2007, p. 1–32.
+
+\bibitem{Jorgensen2004}
+Jorgensen, W.L. The many roles of computation in drug discovery, Science 2004, 303, 1813–8.
+
+\bibitem{Gilson2007}
+Gilson, M.K.; Zhou, H.X. Calculation of protein-ligand binding affinities. Annu Rev Biophys Biomol Struct. (2007) 36, 21-42.
+
+\bibitem{Torrie1977}
+Torrie, G. M.; Valleau, J. P. iNonphysical sampling distributions in Monte Carlo free-energy estimation: Umbrella sampling 
+(1977) J. Comput. Phys. 23, 187
+
+\bibitem{Tironi1994}
+Tironi, I.G.; van Gunsteren, W.F. A molecular-dynamics simulation study of chloroform. Mol. Phys. (1994) 83, 381-403.
+
 \bibitem{Dill1997}
-Dill, K.A., and H.S. Chan.  From Levinthal to Pathways to Funnels:  The "New View" of Protein Folding Kinetics.  Nature Structural Biology 4: 10-19 (1997)
+Dill, K.A.; H.S. Chan.  From Levinthal to Pathways to Funnels:  The "New View" of Protein Folding Kinetics.  Nature Structural Biology 4, 10-19 (1997)
+
+\bibitem{Dill2008}
+Dill, K.A.; Ozkan, S.B.; Shell, M.S.; Weikl, T.R. The protein folding problem. Annual Review of Biophysics (2008), 37, 289-316.
+
+\bibitem{Anfinsen1973}
+Anfinsen. C.B. Principles that Govern the Folding of Protein Chains. Science (1973) 181, 223-230.
+
+\bibitem{Christ2007}
+Christ, C.D.; van Gunsteren, W.F. Enveloping distribution sampling: A method to calculate free energy differences from a single simulation,
+J. Chem. Phys. (2007), 126, 184110.
 
@@
 J. Phys. Chem. B 113, p. 9728-9740 (2009).
+
+\bibitem{Strajbl2000}
+Strajbl, M.; Sham, Y.Y.; Villà, J.; Chu, Z.-T.; Warshel, A. Calculations of Activation Entropies of Chemical Reactions 
+in Solution. (2000) 104, 4578-4584.
 
@@
 
+\bibitem{MacCallum2011}
+MacCallum, J.; Pérez, A.; Schnieders, MJ.; Hua, L.; Jacobson, M.P.; Dill, K.A. Assessment of protein structure refinement 
+in CASP9. Proteins, 2011, 79, 74-90.
+
+\bibitem{Kryshtafovych2011}
+Kryshtafovych, A.; Fidelis, K; and Tramontano, A. Evaluation of model quality predictions in CASP9. Proteins, 2011, 79, 91–106
+
+\bibitem{Zemla2003}
+Zemla, A. LGA: a method for finding 3D similarities in protein structures. Nucleic Acids Res 2003, 31, 3370–3374.
+
+\bibitem{Perez2012}
+Perez, A.; Yang, Z.; Bahar, I.; Dill, K.A.; MacCallum, J.L.; FlexE: Using Elastic Network Models to Compare Models of Protein Structure. J. Chem. Theory Comput., 2012, 8, 3985-3991.
+
+\bibitem{Wang2011}
+Wang, Q.; Vantasin, K.; Xu, D.; Shang, Y. MUFOLD-WQA: A new selective consensus method for quality assessment in protein structure prediction. 
+Proteins, 2011, 79: 185–195.
+
+\bibitem{Case2005}
+Case, D.A.; Cheatham, III, T.E.; Darden, T.; Gohlke, Luo, H.R.; Merz, Jr., K.M.;  Onufriev, A; Simmerling, C.; 
+Wang, B.; R. Woods, R. The Amber biomolecular simulation programs. J. Computat. Chem. (20005) 26, 1668-1688.
+
 \end{thebibliography}

### 2012-10-19 Paper/paper.tex
third draft arijit

--- 2012-10-01 Paper/paper.tex
@@
 states are necessary. Examples ranges from small to large conformational change of protein due to
-ligand binding, change of pH etc. Free energy also play an important role in case of protein
+ligand binding, change of pH etc \cite{Meirovitch2007}, \cite{Chipot2007}, \cite{Jorgensen2004}. 
+Free energy also play an important role in case of protein
 folding. The ground breaking work of Christian B. Anfinsen and coworkers showed that the native
 structures of small globular proteins have a unique, thermodynamically stable native structure with
-their conformation at the global free energy minimum. Regardless of the starting point or unfolding
+their conformation at the global free energy minimum \cite{Anfinsen1973}. Regardless of the starting point or unfolding
 it by changing different condition most proteins will finally assume the same structure. It is very
@@
 However, the theoretical calculation of conformational free energy change become difficult if the
-states of interest are very different. In such scenario timescale involved in such tranformation may
+states of interest are very different \cite{Meirovitch2007}. In such scenario timescale involved in such tranformation may
 be beyond timescale of direct classical molecular dynamics simulation.  Generally, in order to
 calculate free energy a pathway or reaction coordinate is created with the intermediate structures.
-Method like umbrella sampling, thermodynamic integration along with classical MD can suffer from
+Method like umbrella sampling \cite{Torrie1977}, thermodynamic integration \cite{Tironi1994} along with classical MD can suffer from
 overlap problem and require large computational effort. The idea of reaction coordinate became more
-complicated in case of protein folding.  As free energy landscape of the protein folding is rough \cite{Dill1997},
+complicated in case of protein folding.  As free energy landscape of the protein folding 
+is rough \cite{Dill1997}, \cite{Dill2008}
 it can be trapped in a energy well during its visit in the conformational space. Even with the
@@
 
-Calculation of free energy has been successfully attempted by a number of groups \cite{Ytreberg2006} - \cite{Zheng2008}. In
-recent years methods like Refrence system method, Decativated Morphing \cite{Park2008}, Orthogonal Space Random
-Work, Confinement Method have came out for calculation of free energy method. Some of the groups
+Calculation of free energy has been successfully attempted by a number of groups \cite{Meirovitch2007}, 
+\cite{Ytreberg2006} - \cite{Zheng2008}, . In recent years methods like Refrence system method \cite{Ytreberg2006}, 
+Decativated Morphing \cite{Park2008}, Orthogonal Space Random 
+Work, Confinement Method have came out for calculation of free energy method.
+
+Some of the groups
 relies on one of the great strength of free energy calculation, that it is a state function and thus
@@
 In this work we have applied the confinement method which was originally devoled by Tyka et. al. \cite{Tyka2006} and
-Cecchini et. al. \cite{Cecchini2009}. This method relies on a thermodynamic cycle where first, non-harmonic degrees of
+Cecchini et. al. \cite{Cecchini2009}. The free energy is a state function and thus it can be independent of path. 
+Confinement method relies on this fact and uses a thermodynamic cycle where first, non-harmonic degrees of
 freedom are removed by applying a series of restraint to the system. Then the free energy of the
@@
 
-\Justin{Arijit, Al: did we do any other control experiments on small peptides that we should report
-here?}
-
 Previous applications of the confiment method have focused on relatively short peptides, up to $17$
 residues in length. In this work we apply the confinement method to
-larger proteins. One of the  applications we explore in the present work is the re-scoring, or
+larger proteins. On that direction, We first test the confinement method for a chamelon sequence and 
+found that it can correctly pick out the prefential structure of the chamelon sequence. Another of the 
+applications we explore in the present work is the re-scoring, or
 metaprediction, of structures submitted during the Critical Assessment of Structure Prediction
@@
 energy of the experimental structure (once it was available from the PDB). As
-Table~\ref{table:casp_control} shows, in X of Y cases, the confinement method correctly assigns a
-lower free energy to the experimentally determined structure than to any of the decoys. Although
-differentiating between experimentally determined structures and computer generated predictions is
-not a stringent test of a scoring method---for example, examining the residue-residue packing can
-easily distinguish between the two [ref Rosetta Holes]---the results none the less serve as a useful
-"sanity check".
+Table~\ref{table:casp_control} shows, in most cases, the confinement method correctly assigns a
+lower free energy to the experimentally determined structure than to any of the decoys. 
+%Although differentiating between experimentally determined structures and computer generated predictions is
+%not a stringent test of a scoring method---for example, examining the residue-residue packing can
+%easily distinguish between the two [ref Rosetta Holes]---the results none the less serve as a useful
+%"sanity check".
 
@@
      T0531       &    2KJX        &          $12.49 \pm 0.70$ \\ \hline
-     T0538       &    2L09        &          $7.70 \pm 0.47$ \\ \hline
+     T0538       &    2L09        &          $-2.48 \pm 0.47$ \\ \hline
      T0540       &    3MX7        &          $22.00 \pm 0.49$ \\ \hline
@@
 The results of these calculations are presented in Figure~\ref{fig:orban}. The confinement method
-identifies the correct structure for all six sequences. The free energy differences range from
-around 2.9 to 5.7 kcal/mol, which is consistent with small changes in sequences and with estimates
-made by Orban and co-workers \cite{He2008}, \cite{{Alexander2009}}.
+identifies the correct structure for all five sequences. One of the hypothesis for fold switching
+of chamelon sequence is that the structural transitions require states with diminished stability.
+It is widely believed that if the free energy of the native state and the alternative state is within
+ a range of $-5 Kcal/Mol$ then  it can quickly change fold when the stability of the native state 
+decreases. The stability of the native state can decrease for a number reason ranging from chemical 
+modification, breaking of disulphide bond or mutation as in this case. The free energy differences, that
+came out from our calculation ranges from
+around 2.9 to 5.7 kcal/mol, which is consistent with the above hypothesis \cite{He2008}, \cite{{Alexander2009}}.
 
 \begin{figure}
-\includegraphics[width=14cm,height=14cm]{orban.pdf}
+\includegraphics[width=3.5 in,height=4 in]{orban.pdf}
 \label{fig:orban}
@@
 %GB88       & $5.576 \pm 0.464$        & $0.000$                   \\
-%GA98       & \Justin{Arijit: fill in} &                           \\ \hline
+%GA98       &  &                           \\ \hline
 %GB98       &                          &                           \\ \hline
@@
 subset of CASP predictions}
+\Alberto{Arijit: we need a global table of success, rather than individual tables. The simplest form
+    would have 3 columns: name of the PDB, number of structures tested, percent of success of
+    picking 1st model}
 
-Next, we have applied the confinement method to several targets from the CASP9 and CASP10
-experiments. CASP is a world-wide competition aiming at finding the state of the art methodologies
-in protein folding. Groups, both from knowledge based methods and physics based methods participate
-in this blind competition and for each sequence target they can submit five models. The predictors
-were also asked to calculate their best model among the set of five they submitted. However, in many
-cases predictors were unable to predict/rank their models. Using the confinement method, we aim to
-rank protein structures using free energy as a scale. In recent CASP experiments Global distance
-test scose (GDT\_TS) method is used for final analysis of different submitted structures with the
-NMR/crystallographically solved protein native structure. GDT\_TS represent the average percentage
-of residues that are in close proximity in two structures optimally superimposed using four
-different distance cutoff. Although GDT\_TS alone is not a reliable measure for the difficult
-modeling cases, it has been well accepted method in recent years among the community. We have also
-compared our free energy result with the GDT\_TS score for different cases we studied.
+This method is particularly well suited to pick out the native/native like structure from misfolded or decoys. We have
+tested this using different models from CASP (critical assessment of structure
+prediction). CASP is a blind test world wide competition in which different groups apply their
+methods to predict the 3D structure of proteins given their sequence. This is done with a strict 3
+week limit on each target (3 day for servers) and each group is allowed to submit 5 possible
+structures (ranked from better to worst). In our role as assessors during the CASP8 and CASP9 \cite{MacCallum2011}
+competition we observed no correlation at all between the ordering of structures submited by the
+groups and the real ranking compared to an experimental model \cite{Kryshtafovych2011}}. The consequences of this go beyond
+those five structures; the deeper meaning is that groups producing ensembles of structures are
+generating structures that are better than the ones they submit, but they do not know about it. In
+fact, when querrying different groups after the results were known, most groups agree that this is
+the case. Beyond the CASP problem, this reflects on the modelers ability to correctly rank order
+models in many different environments, from structures for drug design leads to designing more
+stable proteins or peptide mimetics such as peptoids. One of the main culprits of this lack of
+accuracy is the fact that ranking is often done via a potential energy function, which in many cases
+lacks an entropy component. Other initiatives including knowledge based potentials do use some sort
+of free energy to rank order, but its accuracy is not enough. Our method provides a physics based
+solution to this problem. We have tried two experiments centered on CASP. First we tried to rank
+order some structures from the previous CASP9 experiment. Our second test was to participate as a
+metapredictor group in CASP10. In both methods, our ranking is determined as a free energy measure,
+and it is compared with the ranking given by a geometrical comparison between native and submitted
+models called GDT\_TS (Global distance test score \cite{Zemla2003}. 
+GDT\_TS represent the average percentage of residues that are in close proximity in two
+structures optimally superimposed using four different distance cutoffs (1, 2, 4 and 8 Å).
 
-%The GDTTS value of a model is determined as follows. A large sample of possible structure
-%superpositions of the model on the corresponding experimental structure is generated by superposing
-%all sets of three, five, and seven consecutive C$\alpha$ atoms along the backbone (each peptide
-%segment provides one superposition). Each of these initial superpositions is iteratively extended,
-%including all residue pairs under a specified threshold in the next iteration, and continuing until
-%there is no change in included residues. The procedure is carried out using thresholds of 1, 2, 4,
-%and 8 $\AA$, and the final superposition that includes the maximum number of residues is selected
-%for each threshold.
+We have chosen the inital models on which to test the methods based on server groups that have
+traditionally done good in past CASP events. There is a high correlation between the structures that
+are chosen for this method and its predictive accuracy. In particular, when GDT\_TS scores are below
+50, it is difficult for us to say anything about the models. Our approach is simple, with every
+subset of structures we choose, we use the confinement method in a pairwise procedure, and then rank
+the structures. In all cases, we have calculated the free energy with respect to the native structure either obtained from NMR or 
+x-ray crystallography but we do not need to make comparisons to native in order to do rank ordering.
 
-\subsubsection{T0559}
+\subsubsection{CASP9}
 
-The native structure is a  67 residue NMR solved structure. We first reviewed all the five models
-submitted by the group "BAKER-ROSETTASERVER", which was the best predictor group for this particular
-target.  Among them we choose only the three structures, model 1, 3 and 5 to calculate the free
-energy and subsequently identify the most stable structure among them. The rest of the two
-structures are close to either of the selected structure and thus they are ommited from our
-calculation. According to the GDT\_TS score, the model 1 was picked up correctly from the set of
-structure generated . However, the order of model 3 and 5 was not correct.
-It will be interesting to know whether confinement method can order the strcutures in terms
-of free energy scale. The result is presented in Figure~\ref{fig:T0559}.
+The question that we want to answer here is whether we can distinguish between native structure and the submitted models? 
+%Identifying the native state from a set of models is a very easy test, that most metapredictors
+%already handle correctly, but it is also a first proof of principle. 
+First we choose a protein BVU3908 from Bacteroides vulgatus whose PDB id and CASP target code are
+2L01 and T0559 respectively. The native structure of this 69 residue protein was solved using NMR.
+The best predictor group for this target was "BAKER-ROSETTASERVER".
+We initlally checked similarity between the models submitted by "BAKER-ROSETTASERVER" and discarded two of them from the 
+analysis on the basis of being too similar to some of the other models. The GDT\_TS score and rmsd
+values as shown in Figure~\ref{fig:T0559} indicate that the model 1 was predicted correctly, 
+whereas the order of model 3 and 5 was wrong. The main difference
+between the model 3 and the rest of the model is that the orientation of first alpha helix of model 3 is opposite.
+On the other hand, as shown in Figure~\ref{fig:T0559} confine and release method not only can differentiate the native 
+structure from the submitted models , it can also correctly rank all the models.
 
-\subsubsection{T0560}
+To further test the method we have taken the example of protein BT2368 from Bacteroides thetaiotaomicron. The PDB id of this 
+74 amino acid residue protein is 2L02 and CASP target code is T0560. We have compared the free energy difference between 
+the native structure and the two of the five submitted models from the group "Splicer". The remaining three models are discarded as 
+again they are similar to the rest of the models. The comparison of GDT\_TS score as shown in 
+Figure~\ref{fig:T0560} and published in final CASP result is only for models with residues 3 to 66. In order to keep
+consistency, we have also do our analysis with models and crystal structure consisting residues 3 to 66. As expected,
+the native state is identified correctly, and the two other models is ranked in the correct order.
 
-Target T0560 is a 74 amino acid residue target from CASP9. However residues 3-66 were kept for final
-result.  For this target we have compared the free energy of the model 2 and model 4 submitted by
-the group "Splicer" along with the crystal structure.  The main difference between the model 2 and 4
-is the orientation of the ....  The final result is presented in Figure~\ref{fig:T0560}.
+Our next test was between models from the best submissions of different groups. 
+For this purpose we have chosen the x-ray crystallographic structure of 
+fas apoptosis inhibitory protein molecule whose pdb id and CASP target codes are 3MX7 and T0540 respectively. 
+This protein contain 90 amino acid residues with 8 beta strands. We have chosen best models from groups 
+"LTB" and "Mufold" for our analysis, which are labelled as Model 1 and Model 2 in Figure~\ref{fig:T0540}. 
+As presented in  Figure~\ref{fig:T0540} we found that this method is able to correctly order the models which matches 
+with the GDT\_TS score.
 
-\subsubsection{T0538}
+There are however some instances where the method can fail, specially when GDT scores are
+very close. We have found this for an engineered protein from Asr4154 protein (PDB ID: 2L09 and CASP Target T0538).
+This protein contains 54 amino acid residues with three alpha helixes and two very short beta strand. 
+The model that is closest to the native structure was generated by the group PconsR with a GDT\_TS value of 96.23.
+We have also considered one model from group "Shell" (GDT\_TS = 90.09) and "FOLDIT" (GDT\_TS = 86.32) for our analysis along with
+the native crystal struture. The models from PconsR, Shell and FOLDIT are labelled as model 1, model 2 and model 3 
+respectively in Figure~\ref{fig:T0538}. As presented from Figure~\ref{fig:T0538}, one can found that the 
+model from PconsR has lowest free energy value, making it a more stable structure. Although ranking of the other models
+remain consistent, it may happen that the theoretical model produced by the group "PconsR" much more stable than the 
+crystal structure. This is particularly difficult target for rank ordering as all the models those are considered for this 
+calculation are within RMSD values of 2.6 \AA.
 
-Target T0538 was a 54 residue target. The native structure was generated using NMR. The best model
-that is closest to the native structure was generated by the group PconsR. However, their model 3
-was better than their model 1. This structure was quite close to the NMR structure and had GDT\_TS
-value of 96.23. We have calculated the free energy between this structure and the best structure
-submitted by the group "Shell", "FOLDIT" and "BAKER-ROSETTASERVER". The calculated free energy value
-is presented in Figure~\ref{fig:T0538}.
+\subsubsection{What can we say about low resolution models?}
 
-\subsubsection{T0540}
+We mentioned earlier that our predictive abilities are greatly decreased with low model qualities. It is
+better to have atleast one good model to pick out the native/native like structures from the rest of the decoy. 
+In oorder to test this hypothesis we performed another test with target T0531 of CASP9, in which we compared the crystal 
+structure to five models from the MUFOLD server. The pdb id of this extracellular domain of the jumping translocation 
+breakpoint protein is 2KJX and the CASP target code is T0531.
+Looking at Figure~\ref{fig:T0531} shows: 1.) native is correctly
+identified as expected and 2.) Surprisingly there is a high level of correlation between the GDT
+ranking and the free energy ranking, only two structures with GDT\_TS score less than  35  are ordered
+incorrectly. It is worth to note that models 2 and 3 have the same GDT and very different free
+energies, meaning that the actual ordering could change a lot \cite{Perez2012}.
 
-Target T0540 was a 90 residue target. We have taken the best models from the group "LTB" which was
-the best performing group for this target and also from the group Mufold and calculated the free
-energy difference between these models with that of the native crystal structure.  As expected, the
-crystal structure has the lowest free energy value, followed by Model 1 with GDT\_TS score of
-$69.72$ and model 2 with GDT\_TS value of $53.789$. The final result is presented in
-Figure~\ref{fig:T0540}.
+\subsubsection{Difficulty in CASP Experiments}
 
-\subsubsection{T0531}
+It is often difficult to model CASP targets using confinement and release method. There are many cases,
+where the sequence of a particular target are trimmed during actual result presentation. This is illustrated
+with the crystal structure of the leucine zipper domain of cGMP dependent protein kinase I (pdb id: 3NMD
+and CASP9 code: T0605) as shown in Figure~\ref{fig:T0605}. The actual CASP 9 target consisted of 72 amino acid residues, but
+as only 18-66 amino acid residues could be detected in actual crystal structure of this protein, the final result was trimmed 
+with respect to that crystal structure with 49 amino acid residues which consist of a single alpha helix. The amino acid residues 18 and 66 are shown in
+three models submitted by the group "Baker". It is important to note that blind prediction of model ranking on the basis of free energy can be hugely 
+affected in such scenario as the only major difference between those models are the trimmed parts, specially the region consisting residues 1-18.
 
-This is a 65 amino acid residue target with residues 6-63 were kept for final analysis. We have
-calculated the free energy difference between all the five models submitted by the group
-"MUFOLD-MD", which was one of the best performed group for this target.  The result is presented in
-Figure~\ref{fig:T0531}.
+\subsubsection{Can confinement method perform better quality assessment in protein structure prediction ?}
+A part of CASP experiment is dedicated to the assessment of the quality of predicted model \cite{Kryshtafovych2011}. 
+Most of the top performing group in this catagory use consensus approaches for quality assessment \cite{Wang2011}.
+A review of CASP9 quality assessment also pointed out that the method which are based on clustering technique perform
+much better compare to those methos which are based on analysis of individual model \cite{Kryshtafovych2011}. 
+In this background we try to answer, whether confinement method can do better quality assessment compare to the best performing
+group (MUFOLD-WQA) in this catagory.
+We have again picked up the CASP taget T0538 to answer this question. We have chosen the Model 3 submitted by PconsR (GDT\_TS = 96) 
+and model 5 from the MULTICOM-NOVEL ( GDT\_TS = 83). Both are server predicted model and they were further used for quality assessment 
+by the group MUFOLD-WQA \cite{Wang2011}. It is important to note that the model from PconsR was most accurate predicted model for this 
+target and the model from MULTICOM-NOVEL was predicted as best by the quality assessment experiments of MUFOLD-WQA. 
+Our free energy analysis we have shown that model from Pconsr is better than the model from 
+MULTICOM-NOVEL $(\Delta \Delta G (Model\_Pconsr - Model\_MULTICOM-NOVEL = -8.2 Kcal/Mol)$, which   
+matches well with the GDT\_TS score. On the other hand the MUFOLD-WQA predicted the ranking in the reverse order. 
+There is no doubt that the consensus approaches can predict the model quality in a faster manner. But we expect that confinement
+approach can predict the model quality in a relatively expensive but much accurate way.
 
-\subsubsection{T0605}
+Further Analysis is going on Target 560.
 
-This is a 72 amino acid residue protein but for the final calculation only the residues 18-66 amino acids were kept for the final analysis.
+\subsubsection{CASP10}
 
-The best performing group was "Baker" for this target. We have carried out our calculation with the model first, second and fifth for the
-final calculation. The rest two model had closer GDT\_TS value compared to the previously mentioned models and thus they were omitted from
-furthe calculation. The result is presented in Figure~\ref{fig:T0605}. (This is a case where we can demand the confinement
-method is better than GDT\_TS).
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=14cm,height=14cm]{method.pdf}
-\end{center}
-\caption{Graphical representation of the thermodynamic cycle involving confinement method.}
-\label{fig:method}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=12cm,height=10cm]{bhp.pdf}
-\end{center}
-\caption{Two conformations from $\beta$ hairpin from protein G, bhp1 and bhp2. The two stranded $\beta$ sheet, bhp1 is the native structure and the three stranded $\beta$ sheet is known as bhp3.}
-\label{fig:bhp_conf}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=12cm,height=10cm]{G95.pdf}
-\end{center}
-\caption{Two protein with 95 \% similar sequence but different folds}
-\label{fig:G95}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=12cm,height=10cm]{T0559.pdf}
-\end{center}
-\caption{Target T0559 from CASP9}
-\label{fig:T0559}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=12cm,height=10cm]{T0538.pdf}
-\end{center}
-\caption{Target T0538}
-\label{fig:T0538}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=12cm,height=10cm]{T0560.pdf}
-\end{center}
-\caption{Target T0560}
-\label{fig:T0560}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=12cm,height=10cm]{T0540.pdf}
-\end{center}
-\caption{T0540}
-\label{fig:T0540}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=12cm,height=10cm]{T0531.pdf}
-\end{center}
-\caption{T0531}
-\label{fig:T0531}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=12cm,height=10cm]{T0605.pdf}
-\end{center}
-\caption{T0605}
-\label{fig:T0605}
-\end{figure}
+We participated in CASP10 as a metapredictor group. Given the tigh schedule of the CASP competition,
+we only attempted proteins under 100 residues in length. Our main problem during this part of the
+competition. was that proteins were trimmed from the actual calculation.
+%\Alberto{Justin,Arijit: Probably should not comment on metapredicting since we were not able to do
+%    good due to the trimming factor. Instead, we should move forward with Justin's idea of using
+%    structures that other meta predictors did badly at. What stage is that in?}
 
 \section{Conclusion}
+
+Proteins are essential parts of organisms and participate in virtually every process within living cells. 
+In order to understand different functions of proteins, detect diseases caused by them and design more efficient drugs 
+it is often useful to have the three dimensional structure of the protein. Experimental studies using NMR and X-ray crystallography is 
+not enough to detect large number of structures. In recent past there are number of studies directed towards understanding protein 
+structure using theoretical tools. Experimentally, it is relatively easy to identify the native structure from the data obtained 
+from x-ray crystallography or NMR. On the other hand using theoretical tools it is often difficult to identify the native structure 
+using both physics based methods and knowledge based method. In order to rank structure from the available data, some groups use 
+clustering algorithms and the structure of the
+most populated cluster is picked out as the most probable structure. However, it may happen that the structure may be trapped and 
+spend most of its time in a kinetic trap, which may give a wrong interpretation of the clustering result. On the other hand, some groups
+also depend on potential energy function to pick out their best structure. In this paper, we have used a free energy method to identify
+the native/native like structure. First, we have tested this method for a chameleon sequence, which has a $95$ percent sequence identity
+but one of them has a $3 \alpha$ structure and another one has $4 \beta + \alpha$ structure. Using the current method, we can found that 
+the native structure always has most stable structure. We have also tested this method for different targets of CASP experiments. We found
+that in most cases, it can pick out the native/native like structure in terms of lowest free enrgy value.
+
+However, there may be some limitation of this method in terms of size of the protein molecule. The normal mode analysis may not be
+accurate for proteins larger than 100 amino acid residues. This can be approximated by removing the hydrogen molecules from 
+the system. 
+The method can be also give wrong result if all the models in the decoy are very poor. It may also be difficult to rank structures if 
+all the structures are very very close to the native structure. But that does not matter at all as our main aim is to differentiate
+between the good and the bad structure. In this article we have extensively tested this method using examples from CASP experiment.
+This method can be computationally expensive in actual CASP experiment if one want rank thousands of submitted models in CASP 
+experiments. However, in real life experiment with a particular target this method can give accurate result. The important 
+fact about Confinement method is that all the part of the calculation can be done independently. We have mostly carried out 
+our calculation with Amber program \cite{Case2005} in GPU computer. With a avarage 56 residue we found that it take only 4 hour to complete 
+20 ns of the confinement step. So, with available computer power this method will be easy to use.
+
+This method can be further used for calculating free energy due to very large conformational change. Another application may be
+for the case of ligand binding. It can also be used to distinguish the most stable structure during protein design. Another, important
+application may be to study the protein folding kinetics, specially to study the phi value analysis.
 
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.5 in,height=3.5 in]{method.pdf}
+\end{center}
+\caption{Graphical representation of the thermodynamic cycle involving confinement method.}
+\label{fig:method}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.5 in,height=3.0 in]{bhp.pdf}
+\end{center}
+\caption{Two conformations from $\beta$ hairpin from protein G, bhp1 and bhp2. The two stranded $\beta$ sheet, bhp1 is the native structure and the three stranded $\beta$ sheet is known as bhp3.}
+\label{fig:bhp_conf}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.5 in,height=3.0 in]{G95.pdf}
+\end{center}
+\caption{Two protein with 95 \% similar sequence but different folds}
+\label{fig:G95}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.8 in,height=3.0 in]{T0559.pdf}
+\end{center}
+\caption{The native and three submitted model structure, along with their GDT\_TS, RMSD and relative Free energy values 
+of protein BVU3908 from Bacteroides vulgatus (PDB id: 2L01 and CASP code: T0559).}
+\label{fig:T0559}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.8 in,height=3.0 in]{T0538.pdf}
+\end{center}
+\caption{The native and three model structure of engineered protein from Asr4154 protein (PDB ID: 2L09 and CASP code:T0538). The model 1,2 and 3 are
+from the group PconsR, Shell and FOLDIT respectively.}
+\label{fig:T0538}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.5 in,height=3.0 in]{T0560.pdf}
+\end{center}
+\caption{Native and two model structure of protein BT2368 from Bacteroides thetaiotaomicron (pdb id: 2L02 and CASP code: T0560). The two models
+were from the group "Splicer".}
+\label{fig:T0560}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.5 in,height=3.0 in]{T0540.pdf}
+\end{center}
+\caption{X-ray crystallographic structure and two submitted models of fas apoptosis inhibitory protein (pdb id: 3MX7 and CASP code: T0540). 
+Model 1 and Model 2 in this analysis were submitted by the group LTB and MUFOLD respectively.}
+\label{fig:T0540}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.5 in,height=5.7 in]{T0531.pdf}
+\end{center}
+\caption{The native structure and 5 models of extracellular domain of the jumping translocation
+breakpoint protein (pdb id: 2KJX and the CASP code: T0531).}
+\label{fig:T0531}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=12cm,height=10cm]{T0605.pdf}
+\end{center}
+\caption{ Three models of the leucine zipper domain of cGMP dependent protein
+kinase I (original pdb id: 3NMD and CASP target: T0605). The original target sequence of this protein was 72 residues.
+However only residues 3 to 66 is only kept during result announcement as that part can be found from crystallographic analysis. The
+region 3 to 66 are almost identical in all the 3 models  }
+\label{fig:T0605}
+\end{figure}
+
 \begin{thebibliography}{99}
 
+\bibitem{Meirovitch2007}
+Meirovitch, H. Recent developments in methodologies for calculating the entropy and free energy of biological systems by computer simulation.
+Current Opinion in Structural Biology, 2007, 17, 181-186.
+
+\bibitem{Chipot2007}
+Chipot, C.; Shell, M.S.; Pohorille, A. Introduction, in Chipot, C., Pohorille, A., editors. Free Energy
+Calculations: Theory and Applications in Chemistry and Biology. Springer Series in Chemical
+Physics, vol. 86. Berlin and Heidelberg: Springer; 2007, p. 1–32.
+
+\bibitem{Jorgensen2004}
+Jorgensen, W.L. The many roles of computation in drug discovery, Science 2004, 303, 1813–8.
+
+\bibitem{Gilson2007}
+Gilson, M.K.; Zhou, H.X. Calculation of protein-ligand binding affinities. Annu Rev Biophys Biomol Struct. (2007) 36, 21-42.
+
+\bibitem{Torrie1977}
+Torrie, G. M.; Valleau, J. P. iNonphysical sampling distributions in Monte Carlo free-energy estimation: Umbrella sampling 
+(1977) J. Comput. Phys. 23, 187
+
+\bibitem{Tironi1994}
+Tironi, I.G.; van Gunsteren, W.F. A molecular-dynamics simulation study of chloroform. Mol. Phys. (1994) 83, 381-403.
+
 \bibitem{Dill1997}
-Dill, K.A., and H.S. Chan.  From Levinthal to Pathways to Funnels:  The "New View" of Protein Folding Kinetics.  Nature Structural Biology 4: 10-19 (1997)
+Dill, K.A.; H.S. Chan.  From Levinthal to Pathways to Funnels:  The "New View" of Protein Folding Kinetics.  Nature Structural Biology 4, 10-19 (1997)
+
+\bibitem{Dill2008}
+Dill, K.A.; Ozkan, S.B.; Shell, M.S.; Weikl, T.R. The protein folding problem. Annual Review of Biophysics (2008), 37, 289-316.
+
+\bibitem{Anfinsen1973}
+Anfinsen. C.B. Principles that Govern the Folding of Protein Chains. Science (1973) 181, 223-230.
+
+\bibitem{Christ2007}
+Christ, C.D.; van Gunsteren, W.F. Enveloping distribution sampling: A method to calculate free energy differences from a single simulation,
+J. Chem. Phys. (2007), 126, 184110.
 

### 2012-10-29 Paper/paper.tex
Some edits on text

@@
 The calculation of free energy differences is of central importance in the simulation of biochemical
-systems.  The computation of the free energy difference between pairs of macromolecule with large
-conformational change is particularly a difficult task and computationally expensive with existing
-methods. In this work, an improved version of the confinement approach is used to calculate absolute
-free energies of bio molecular systems. The method does not require a reaction coordinate or
-transition path. It is fast to compute. We show that the method correctly picks out the state of
-lowest free energy of a pair of structures having similar sequence but different fold. We also show
-using models from CASP9 that the method picks out native-like structures from misfolded or decoys,
-provided at least one good structure is in the input set.
+systems.  The computation of the free energy difference between pairs of macromolecular
+conformations is currently a particularly difficult and computationally expensive task with existing
+methods. In this work, an improved \Alberto{what is the improvement} version of the confinement approach is used to calculate absolute
+free energies of bio molecular systems. The method provides two main advantages: it does not require a reaction coordinate or
+transition path and it is fast to compute. Throught out the article we show its use in different
+challenging modeling problems. In particular, we show its use in predicting the conformational
+preference of chamaleon sequences (sequences with high sequence identity and different
+folds). This sequence dependent conformational preference set the stage for the use of the method in
+protein design.
 
@@
 are denoted as ``GB'', while sequences that form a three-helix bundle are denoted as ``GA''. One
-pair of sequences (GA88/GB88) are 88 percent identical and differ in seven positions. Another pair
+pair of sequences (GA88/GB88) are 88 percent identical in sequence and differ in seven positions. Another pair
 (GA95/GA95) are 95 percent identical and differ in three positions. The last pair (GA98/GB98) differ
@@
 the GA88 sequence and the GB88 structure). The confinement method was able to predict the
-conformational preferences correctly for all six sequences (data not shown). There is, however, a
-serious problem with this analysis: we are comparing an experimentally determined structure with a
-computer generated model. It might be possible that we were able to make correct predictions simply
+conformational preferences correctly for all six sequences (data not shown). This is however not
+surprising. It is well known [CITE] that it is easy to distinguish computational models from native
+structures. Despite the fact that this models are built by having a huge amount of structural
+information, it might be possible that we were able to make correct predictions simply
 because artefacts of our modeling procedure always lead to the model having a higher free energy
@@
 decreases. The stability of the native state can decrease for a number reason ranging from chemical 
-modification, breaking of disulphide bond or mutation as in this case. The free energy differences, that
+modification, breaking of disulphide bonds or mutations as in this case. The free energy differences, that
 came out from our calculation ranges from
@@
 and it is compared with the ranking given by a geometrical comparison between native and submitted
-models called GDT\_TS (Global distance test score \cite{Zemla2003}. 
-GDT\_TS represent the average percentage of residues that are in close proximity in two
+models called GDT\_TS (Global distance test score \cite{Zemla2003}). 
+GDT\_TS represents the average percentage of residues that are in close proximity in two
 structures optimally superimposed using four different distance cutoffs (1, 2, 4 and 8 Å).
@@
 Figure~\ref{fig:T0560} and published in final CASP result is only for models with residues 3 to 66. In order to keep
-consistency, we have also do our analysis with models and crystal structure consisting residues 3 to 66. As expected,
-the native state is identified correctly, and the two other models is ranked in the correct order.
+consistency, we have also done our analysis with models and crystal structure consisting residues 3 to 66. As expected,
+the native state is identified correctly, and the two other models are ranked in the correct order.
 
@@
 the native crystal struture. The models from PconsR, Shell and FOLDIT are labelled as model 1, model 2 and model 3 
-respectively in Figure~\ref{fig:T0538}. As presented from Figure~\ref{fig:T0538}, one can found that the 
+respectively in Figure~\ref{fig:T0538}. As seen in Figure~\ref{fig:T0538}, one can find that the 
 model from PconsR has lowest free energy value, making it a more stable structure. Although ranking of the other models
@@
 We mentioned earlier that our predictive abilities are greatly decreased with low model qualities. It is
-better to have atleast one good model to pick out the native/native like structures from the rest of the decoy. 
-In oorder to test this hypothesis we performed another test with target T0531 of CASP9, in which we compared the crystal 
+better to have at least one good model to pick out the native/native like structures from the rest of the decoy. 
+In order to test this hypothesis we performed another test with target T0531 of CASP9, in which we compared the crystal 
 structure to five models from the MUFOLD server. The pdb id of this extracellular domain of the jumping translocation 
@@
 
-It is often difficult to model CASP targets using confinement and release method. There are many cases,
-where the sequence of a particular target are trimmed during actual result presentation. This is illustrated
+It is often difficult to model CASP targets using the confinement and release method. There are many cases,
+where the sequence of a particular target is trimmed during actual result presentation. This is illustrated
 with the crystal structure of the leucine zipper domain of cGMP dependent protein kinase I (pdb id: 3NMD
@@
 Most of the top performing group in this catagory use consensus approaches for quality assessment \cite{Wang2011}.
-A review of CASP9 quality assessment also pointed out that the method which are based on clustering technique perform
-much better compare to those methos which are based on analysis of individual model \cite{Kryshtafovych2011}. 
-In this background we try to answer, whether confinement method can do better quality assessment compare to the best performing
+A review of CASP9 quality assessment also pointed out that the methods which are based on clustering
+techniques perform
+much better compared to those methos which are based on analysis of individual models \cite{Kryshtafovych2011}. 
+In this background we try to answer, whether the confinement method can do better quality assessment compare to the best performing
 group (MUFOLD-WQA) in this catagory.
 We have again picked up the CASP taget T0538 to answer this question. We have chosen the Model 3 submitted by PconsR (GDT\_TS = 96) 
-and model 5 from the MULTICOM-NOVEL ( GDT\_TS = 83). Both are server predicted model and they were further used for quality assessment 
-by the group MUFOLD-WQA \cite{Wang2011}. It is important to note that the model from PconsR was most accurate predicted model for this 
+and model 5 from the MULTICOM-NOVEL ( GDT\_TS = 83). Both are server predicted models and they were further used for quality assessment 
+by the group MUFOLD-WQA \cite{Wang2011}. It is important to note that the model from PconsR was the most accurate predicted model for this 
 target and the model from MULTICOM-NOVEL was predicted as best by the quality assessment experiments of MUFOLD-WQA. 
@@
 
-\subsubsection{CASP10}
-
-We participated in CASP10 as a metapredictor group. Given the tigh schedule of the CASP competition,
-we only attempted proteins under 100 residues in length. Our main problem during this part of the
-competition. was that proteins were trimmed from the actual calculation.
-%\Alberto{Justin,Arijit: Probably should not comment on metapredicting since we were not able to do
-%    good due to the trimming factor. Instead, we should move forward with Justin's idea of using
-%    structures that other meta predictors did badly at. What stage is that in?}
-
 \section{Conclusion}
@@
 In order to understand different functions of proteins, detect diseases caused by them and design more efficient drugs 
-it is often useful to have the three dimensional structure of the protein. Experimental studies using NMR and X-ray crystallography is 
-not enough to detect large number of structures. In recent past there are number of studies directed towards understanding protein 
+it is necessary to have the three dimensional structure of the protein. Experimental studies using NMR and X-ray crystallography is 
+not able to keep up with the sheer number of protein sequences available. In recent past there are number of studies directed towards understanding protein 
 structure using theoretical tools. Experimentally, it is relatively easy to identify the native structure from the data obtained 
@@
 spend most of its time in a kinetic trap, which may give a wrong interpretation of the clustering result. On the other hand, some groups
-also depend on potential energy function to pick out their best structure. In this paper, we have used a free energy method to identify
+also depend on potential energy functions to pick out their best structure, lacking an entropic
+component. In this paper, we have used a free energy method to identify
 the native/native like structure. First, we have tested this method for a chameleon sequence, which has a $95$ percent sequence identity
-but one of them has a $3 \alpha$ structure and another one has $4 \beta + \alpha$ structure. Using the current method, we can found that 
-the native structure always has most stable structure. We have also tested this method for different targets of CASP experiments. We found
+but one of them has a $3 \alpha$ structure and another one has $4 \beta + \alpha$ structure. Using the current method, we found that 
+the native structure is always the most stable \Alberto{This is not true!!}. We have also tested this method for different targets of CASP experiments. We found
 that in most cases, it can pick out the native/native like structure in terms of lowest free enrgy value.
 
-However, there may be some limitation of this method in terms of size of the protein molecule. The normal mode analysis may not be
+However, there may be some limitations to this method in terms of size of the protein molecule. The normal mode analysis may not be
 accurate for proteins larger than 100 amino acid residues. This can be approximated by removing the hydrogen molecules from 
@@
 In all the calculation amber 10 program is used with ff99SB with GB/SA implicit solvent.
+
+\Alberto{Somewhere in the introduction or methods}
+There are two big bottlenecks in this calculation. The first has to do with the amount of
+simulations that have to be performed to confine the ensemble of conformations close to a microstate
+into a single microstate, the second bottleneck has to do with the size of the system in the normal
+mode calculation. We have tackled the first issue by using GPU (Graphical Processing Unit) technology as opposed to the classical
+CPUs, which gives us a two order of magnitude boost in computational efficiency. Normal mode
+analysis calculations are Oh(N**3) and require large amounts of memory, effectively limiting the
+size of the systems under study. Where as this can be overcome by larger memory machines, these are
+often not available, and so we propose an alternative by using quasiharmonic analysis. This is still
+an Oh(N**3) problem, but we now the motions of atoms are already determined in trajectories, and so
+we can choose to do quasiharmonic analysis excluding hydrogens, which will reduce the number of
+atoms N to almost N/2. This means that we can use this method for systems twice as large.
 

### 2012-10-31 Paper/new.tex
fourth draft

We have also studied the solution NMR solved of a domain of adhesion exoprotein from Pediococcus pentosaceus. The
pdb id and CASP code of this 79 residue protein are 2KWY and T0569 respectively. We have compared the native structure with the
best model structure submitted for this target in CASP9. This model structute was submitted by the group " Mufold". ALl the
five model structures submitted by this group was found to be close. Thus we proceed by comparing only the native structure and 
the best predicted model of this target. Visualization upon superimposition of two structures reveals that 
the difference between the two structure is at the region consist of residues 48 to 65, where model1 is much more disordered compare to the 
native structure(see Figure~\ref{fig:fig:T0569}) . Other regions look similar for both the cases with of model1 is within 2.6 \AA 
rmsd value of the native structure. In this case we found that the confinement method predict that the predicted model is stable than the
native structure with a free energy difference of $9.4 \pm 0.7$ Kcal/Mol.

\begin{figure}
\begin{center}
\includegraphics[width=3.2 in,height=3.8 in]{T0569.pdf}
\end{center}
\caption{Native and best model structure of a domain of adhesion exoprotein from Pediococcus pentosaceus (pdb id: 2KWY and CASP code: T0569). 
The best model was from the group "Mufold".}
\label{fig:T0569}
\end{figure}

\begin{table}
\label{table:casp_control}
\begin{center}
\begin{tabular}{l l l}\hline
    CASP Target  & PDB Identifier & $\Delta\Delta G_{native \to best decoy}$ (kcal/mol) \\ \hline
     T0531       &    2KJX        &          $12.49 \pm 0.70$ \\ \hline
     T0538       &    2L09        &          $-2.48 \pm 0.47$ \\ \hline
     T0540       &    3MX7        &          $22.00 \pm 0.49$ \\ \hline
     T0559       &    2L01        &          $2.24 \pm 0.24$ \\ \hline
     T0560       &    2L02        &          $22.15 \pm 0.49$ \\ \hline
     T0569       &    2KWY        &          $-9.40 \pm 0.69$  \\ \hline 
\end{tabular}
\end{center}

### 2012-10-31 Paper/paper.tex
fourth draft

--- 2012-10-19 Paper/paper.tex
@@
 \label{table:casp_control}
-
 \begin{center}
@@
      T0560       &    2L02        &          $22.15 \pm 0.49$ \\ \hline
+     T0569       &    2KWY        &          $-9.40 \pm 0.69$  \\ \hline
 \end{tabular}
 \end{center}
-
 \caption{The confinement method assigns a more favorable free energy to the experimentally
@@
 
+There is another instance where this method fails. This is the case of solution NMR structure of a domain of adhesion exoprotein 
+from Pediococcus pentosaceus. The
+pdb id and CASP code of this 79 residue protein are 2KWY and T0569 respectively. We have compared the native structure with the
+best model structure submitted for this target in CASP9. This model structute was submitted by the group " Mufold". ALl the
+five model structures submitted by this group was found to be close. Thus we proceed by comparing only the native structure and
+the best predicted model of this target. Visualization upon superimposition of two structures reveals that
+the difference between the two structure is at the region consist of residues 48 to 65, where model1 is much more disordered compare to the
+native structure(see Figure~\ref{fig:T0569}) . Other regions look similar for both the cases with of model1 is within 2.6 \AA
+rmsd value of the native structure. In this case we found that the confinement method predict that the predicted model is stable than the
+native structure with a free energy difference of $9.4 \pm 0.7$ Kcal/Mol.
+
 \subsubsection{What can we say about low resolution models?}
@@
 
-Further Analysis is going on Target 560.
+\Arijit{Will upload the result from Target 560 once cluster is back.}
 
@@
 
-The confinement method has been described in details in ref. by Tyka et.al. and Cecchini et. al.
-Here we briefly describe the methodology.  We compute free energy $\Delta G_{AB}$ between A and B
-conformation of the protein. For this purpose we use a thermodynamic cycle.
+The confinement method has been described in details in ref. by Tyka et.al. \cite{Tyka2006} and 
+Cecchini et. al. \cite{Cecchini2009}. The basic approach of confinement approach is same in both
+these paper. However, there are some technical differences. Here we briefly describe what we follow
+to compute free energy $\Delta G_{AB}$ between A and B
+conformation of the protein.
 
@@
 
-\item  We minimized both the conformation. These minimized conformation(A* and B*) are the
-    restrained conformation of that state
+\item  We first minimized both the conformation, A and B. These minimized conformation(A* and B*) 
+       are the reference conformation of that state
 
-\item Backbone dihedral angle of the whole protein is restrained by a harmonic potential to define
-    the configurational state.
+\item Backbone dihedral angle of both the full protein is restrained by a harmonic potential. This is required to keep
+    the configurational state of the protein. This approach is particularly important, so that the any misfolded part 
+    of the protein does not chage its configuration.
 
-\item  In order to calculate the free energy $\Delta G_{A,A*}$ and  $\Delta G_{B,B*}$ A and B are
-    gradually transformed in A* and B* by applying a harmonic restraint potential to all atom. For
-    this purpose around 21 simulation were run each of 20 ns with increasing harmonic restraint
-    potential from 0.00005 to 81.92 untill the free and the restrained state overlap well. The final
-    restrained state was chosen so high so that the rotational contributional of the protein frozen
-    out and only remaining contribution remain that of vibrational free enrgy. The fluctuation from
+\item  The free energy is calculated using a thermodynamic cycle. The first part of this approach is the confinement, where
+       the state A and B are confined to the reference state A* and B*. This is done gradually by applying larger and larger
+       harmonic restraint on all the atoms of the biomolecule. This is done by running 21 molecular dynamics simulation 
+       each of 20 ns long, where the harmonic restraint constant was varied from 0.00005 to 81.92 untill the free and the 
+       restrained state overlap well. The final restrained state was chosen so high so that the rotational contributional 
+       of the protein frozen out and only remaining contribution remain that of vibrational free enrgy. The fluctuation from
        the reference structure was recorded and the free energy is calculated using a numerical
-    approach developed by Tyka et. al.
+       approach developed by Tyka et. al. \cite{Tyka2006}. The confinement free energy calculated in this way recorder as 
+       $\Delta G_{A,A*}$ and $\Delta G_{B,B*}$ as shown in Figure~\ref{fig:method}.
 
 \item  Finally the thermodynamic cycle is closed by calculating the free energy between the final
-    restrained state using normal mode analysis.
+       restrained state A* and B* using normal mode analysis or quasiharmonic analysis. The free energy calculated in 
+       this way is shown as $\Delta G_{A*,B*}$ in Figure~\ref{fig:method}.
+
+\item  The full free energy, $\Delta G_{A,B}$ between the two state A and B is calculated using the equation 
+       $\Delta G_{A,B}$ = $\Delta G_{A,A*}$ - $\Delta G_{B,B*}$ + $\Delta G_{A*,B*}$
 
@@
 \label{fig:T0531}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.2 in,height=3.8 in]{T0569.pdf}
+\end{center}
+\caption{Native and best model structure of a domain of adhesion exoprotein from Pediococcus pentosaceus (pdb id: 2KWY and CASP code: T0569).
+The best model was from the group "Mufold".}
+\label{fig:T0569}
 \end{figure}

### 2012-10-31 Paper/paper.tex
before merge

--- 2012-10-29 Paper/paper.tex
@@
 conformations is currently a particularly difficult and computationally expensive task with existing
-methods. In this work, an improved \Alberto{what is the improvement} version of the confinement approach is used to calculate absolute
+methods. In this work, an improved \Alberto{Arijit: what is the improvement??} version of the confinement approach is used to calculate absolute
 free energies of bio molecular systems. The method provides two main advantages: it does not require a reaction coordinate or
@@
 competition we observed no correlation at all between the ordering of structures submited by the
-groups and the real ranking compared to an experimental model \cite{Kryshtafovych2011}}. The consequences of this go beyond
+groups and the real ranking compared to an experimental model \cite{Kryshtafovych2011}. The consequences of this go beyond
 those five structures; the deeper meaning is that groups producing ensembles of structures are

### 2013-01-08 Paper/paper.tex
merged comments from Arijit

@@
 systems.  The computation of the free energy difference between pairs of macromolecular
-conformations is currently a particularly difficult and computationally expensive task with existing
-methods. In this work, an improved \Alberto{Arijit: what is the improvement??} version of the confinement approach is used to calculate absolute
+conformations is particularly difficult and computationally expensive task with existing
+methods. In this work, confinement approach is used to calculate absolute
 free energies of bio molecular systems. The method provides two main advantages: it does not require a reaction coordinate or
-transition path and it is fast to compute. Throught out the article we show its use in different
+transition path and it is fast to compute. The free energy calculated can be decomposed into per residue contribution in a 
+approximate way. Per residue free energy allow us
+to identify the reason behind conformational preference of the biomolecules. Through out the article we show its use in different
 challenging modeling problems. In particular, we show its use in predicting the conformational
 preference of chamaleon sequences (sequences with high sequence identity and different
-folds). This sequence dependent conformational preference set the stage for the use of the method in
-protein design.
+folds). This sequence dependent conformational preference and per residue free energy decomposition set the stage for the use of this 
+method in protein design.
 
@@
 Thus, Free energy can also act as a scale to identify between the misfolded and the native state of
-the protein. These observations further give free energy a special importance in structural biology.
+the protein. These facts further give free energy a special importance in structural biology.
 
@@
 advanced method like replica exchange molecular dynamics the three dimensional structure may be
-trapped in a local minima for a considerable time. Due to various reasons it can also be misfolded
-and trapped in a local energy minima. And one can mistakenly identify the misfolded state as the
-native state (give the example of human pin1ww domain from Benout Roux and Klaus Schultan).
+trapped in a local minima for a considerable time and trapped in a local energy minima
+and one can mistakenly identify the misfolded state as the native state.
+% (give the example of human pin1ww domain from Benout Roux and Klaus Schultan).
 
@@
 \cite{Ytreberg2006} - \cite{Zheng2008}, . In recent years methods like Refrence system method \cite{Ytreberg2006}, 
-Decativated Morphing \cite{Park2008}, Orthogonal Space Random 
-Work, Confinement Method have came out for calculation of free energy method.
-
-Some of the groups
-relies on one of the great strength of free energy calculation, that it is a state function and thus
-it is independent of path.
-
+Decativated Morphing \cite{Park2008}, Orthogonal Space Random Work \cite{Zheng2008}, Confinement 
+Method \cite{Tyka2006}, \cite{Cecchini2009} have came out for calculation of free energy method. 
 In this work we have applied the confinement method which was originally devoled by Tyka et. al. \cite{Tyka2006} and
-Cecchini et. al. \cite{Cecchini2009}. The free energy is a state function and thus it can be independent of path. 
-Confinement method relies on this fact and uses a thermodynamic cycle where first, non-harmonic degrees of
+Cecchini et. al. \cite{Cecchini2009}. This method relies on the fact that free energy is a state function and 
+thus independent of path. It uses a thermodynamic cycle where first, non-harmonic degrees of
 freedom are removed by applying a series of restraint to the system. Then the free energy of the
@@
 different fold was calculated. Encouraged by the result, we applied the method of different targets
-of CASP9 competition.
+of CASP9 competition and tested whether it can predict the most stable conformation of the protein based 
+on its free energy. The remarkable ability of confinement method to identify residues which 
+stabilizes/destabilizes a particular conformation from per residue free energy analysis set the stage
+for future application of this method in protein design.
 
@@
 larger proteins. On that direction, We first test the confinement method for a chamelon sequence and 
-found that it can correctly pick out the prefential structure of the chamelon sequence. Another of the 
+found that it can correctly pick out the prefential structure of the chamelon sequence. Other 
 applications we explore in the present work is the re-scoring, or
@@
 (CASP) experiment (described in detail later). We computed the relative free energies of predictions
-submitted during CASP9 10, with the expectation that the most native-like prediction will have
+submitted during CASP9, with the expectation that the most native-like prediction will have
 the lowest free energy. As a control, for several targets we also calculated the relative free
-energy of the experimental structure (once it was available from the PDB). As
+energy of the experimentally determined structure. As
 Table~\ref{table:casp_control} shows, in most cases, the confinement method correctly assigns a
 lower free energy to the experimentally determined structure than to any of the decoys. 
+Most interestingly, we decomposed the free energy into its per residue component, which help 
+us to identify the residues which stabilize or destabilize a particular conformation of the protein.
 %Although differentiating between experimentally determined structures and computer generated predictions is
@@
 \begin{table}
-\label{table:casp_control}
 \begin{center}
-\begin{tabular}{l l l}\hline
-    CASP Target  & PDB Identifier & $\Delta\Delta G_{native \to best decoy}$ (kcal/mol) \\ \hline
-     T0531       &    2KJX        &          $12.49 \pm 0.70$ \\ \hline
-     T0538       &    2L09        &          $-2.48 \pm 0.47$ \\ \hline
-     T0540       &    3MX7        &          $22.00 \pm 0.49$ \\ \hline
-     T0559       &    2L01        &          $2.24 \pm 0.24$ \\ \hline
-     T0560       &    2L02        &          $22.15 \pm 0.49$ \\ \hline
-     T0569       &    2KWY        &          $-9.40 \pm 0.69$  \\ \hline
-\end{tabular}
-\end{center}
 \caption{The confinement method assigns a more favorable free energy to the experimentally
@@
 experimental structure is predicted to be more favorable than any of the decoys.}
+\label{table:casp_control}
+\begin{tabular}{l l l}\hline
+    CASP Target  & PDB Identifier & $\Delta\Delta G_{native \to best decoy}$ (kcal/mol) \\ \hline
+     T0531       &    2KJX        &          $11.15 \pm 0.70$ \\ \hline
+     T0538       &    2L09        &          $-3.00 \pm 0.47$ \\ \hline
+     T0540       &    3MX7        &          $16.94 \pm 0.49$ \\ \hline
+     T0559       &    2L01        &          $2.10 \pm 0.24$ \\ \hline
+     T0560       &    2L02        &          $18.00 \pm 0.49$ \\ \hline
+     T0569       &    2KWY        &          $20.01 \pm 0.69$  \\ \hline
+\end{tabular}
+\end{center}
 
@@
 
+\subsection{Per residue free energy calculation can identify the mechanistic detail behind conformational pereference 
+of a particular residue}
+
+We decomposed the calculated free energy into per residue decomposition as described in method section. This help us for
+identification of important residues that stabilizes a particular conformation. 
+In this direction, We first calculate the per residue free energy for the GA95 and GB95 sequences. Similar to the total 
+free energy for each sequence, we calculate the free energy difference between $4 \beta + \alpha$ and $3 \alpha$ structure for each 
+residue. The free energy, $\Delta \Delta G (4 \beta + \alpha$ - $3 \alpha)$ is plotted in Figure~\ref{fig:orban_full}(F). 
+In this plot, the negative peak indicate the residue will stabilize the $4 \beta + \alpha$ conformation of a particular
+residue and the positive peak will stabilize the $3 \alpha$ structure. As, the differences of amino acid 
+residues are only at 3 position, there are some common features of residues in both GA95 and GB95 sequences in both 
+$3 \alpha$ and $4 \beta + \alpha$ conformation. They are reported in Figure~\ref{fig:orban_full} 
+and Table~\ref{table:orban_perresidue}. Before, we further go into that detail, it is important to note that 
+experimental observations classified the protein into two parts: Amino acids 9-51 are fully structured in both
+the fold, where as residues 1-8 and 52-56 are unstructured in $3 \alpha$, but form $\beta$ strand in $4 \beta + \alpha$ structure.
+Most of the amino acid residues in the region 1-9 favors $4 \beta + \alpha$ structure and can also seen from the plot 
+Figure~\ref{fig:orban_full}(F). Here, we try to explain some of the largest peak in that plot. The residues Leu-5 and 
+Leu-7 favors $4 \beta + \alpha$ structure as these hydrophobic residues are oriented towards the protein hydrophobic 
+core, whereas in $3 \alpha$ structure they are exposed to the solvent (Figure~\ref{fig:orban_full}(A)). One can argue that 
+hydrophilic residues in these positions can stabilize the $3 \alpha$ conformation.
+The next big peak is located at position 11. A close observation of the sidechain orientation of residue Gln at position 11 reveal that it
+is exposed to the protein surface in both the conformation, but in $3 \alpha$ structure it has an additional H-bond with Glu-15 which gives
+it more stability compare to the $4 \beta + \alpha$ structure (Figure~\ref{fig:orban_full}(B)). The peak 
+at position 14 is due to the hydrophilic Glu, which is completely
+exposed to the solvent in $4 \beta + \alpha$ conformation as shown in (Figure~\ref{fig:orban_full}(E)), but only partially exposed 
+to the solvent in $3 \alpha$ form. 
+This is the reason why Glu at position 14 stabilize $4 \beta + \alpha$ conformation. Another high peak is at position 26, which indicate
+that the residue in that position stabilizes $4 \beta + \alpha$ conformation. Again, the hydrophobic amino acid residue, Ala at position 26
+is oriented towards the hydrophobic core of the protein in $4 \beta + \alpha$ conformation , whereas it is exposed towards the protein surface 
+in $3 \alpha$ structure (Figure~\ref{fig:orban_full}(C)). Similar situations arise for residues at position 54 and 56. 
+These residues are in the region, where $4 \beta + \alpha$ form is fully structured and $3 \alpha$ form is unstructured. 
+As shown in Figure~\ref{fig:orban_full}(D), the sidechain of hydrophobic Val-54 is oriented towards the hydrophobic core in 
+$4 \beta + \alpha$ conformation and thus have hydrophobic interactions with other residues in this conformation.
+On the other hand, in $3 \alpha$ it is exposed to the solvent. The reverse happens for the residue at position 56.
+Here the hydrophilic residue Glu is partially or not at all exposed to the protein surface in $4 \beta + \alpha$ but fully exposed 
+to the solvent in $3 \alpha$ conformation (Figure~\ref{fig:orban_full}(G)). This is the reason, why one can see a positive peak for position 56. 
+Similarly, hydrophobic Ile-49 in $4 \beta + \alpha$ conformation is exposed to protein surface and solvent, whereas as shown in 
+Figure~\ref{fig:orban_full}(H)$3 \alpha$ conformation 
+it has hydrophobic interaction with other hydrophobic residues in protein core.   
+Now, let us discuss the differences of $3 \alpha$ and $4 \beta + \alpha$ fold due to specific difference of mutated amino acid 
+residues in GA and GB sequence. The differences are in positions 20, 30 and 45. While GA95 has Leu-20-Ile-30-Leu-45, GB95 sequence 
+has Ala-20-Phe-30-Tyr-45 at those position. One can find that the peak at position 49 stabilizes the GA sequence more than that of GB
+sequence. In $3 \alpha$ fold, Ile-49 has hydrophobic interactions with Leu-20 and Ile-30 with GA sequence, whereas with GB sequence it 
+has such interactions with smaller Ala-20 and Phe-30 residues (Figure~\ref{fig:orban_full}(I). In $3 \alpha$, with GB sequence, Phe-30 
+being a larger group can not be completely accomodated in the hydrophobic core(Figure~\ref{fig:orban_full}(J)) and therefore, 
+partially exposed to the protein surface. There is a very interesting difference for residue at position 45 as well. In GB95 sequence
+and $4 \beta + \alpha$ fold there is H-bonding between Tyr-45 and Asp-47 (Figure~\ref{fig:orban_full}(K)), which is absent in the
+$3 \alpha$ conformation. On the other hand with GA conformation, this H-bond is absent for both the conformation as there is a Leu at position
+45. Although, one can not find much difference between $4 \beta + \alpha$ and $3 \alpha$ conformation
+for position 20 for both GA95 and GB95 sequence. There are significant difference for the residues at position 30 and 45.  
+These residues have some long term effects as well. As already explained, residue Ile-49 has much more hydrophic interactions with residues
+at position 20 and 30 in GA sequence, which is less in GB sequence. There are difference for Asp-47 as well. With $3 \alpha$ conformation 
+and GA95 sequence, Asp-47 has H-bonding with Lys-50, which is absent in the other conformation.
+
+\begin{table}
+\begin{center}
+\caption{Analysis of per residue free energy decomposition of GA95 and GB95 sequence reveal that certain residues prefer
+either $3 \alpha$  or $4 \beta + \alpha$ structure. These observations and the reason behind such conformational preferences are listed in this
+table.}
+\label{table:orban_perresidue}
+\begin{tabular}{| l | l | l |}
+\hline
+    Observation                                & $3 \alpha$ structure           &    $4 \beta + \alpha$ structure                       \\ \hline
+    Leu-5 and Leu-7 favor                      & hydrophobic residues are       & oriented towards the                                  \\ 
+    $4 \beta + \alpha$                         & exposed to the solvent         & hydrophobic core                                      \\ \hline
+    Gln-11 favors $3 \alpha$                   & Gln-11 has H-bond with Glu-15  & no such H-bond                                        \\ \hline
+    Glu-14 favors $4 \beta + \alpha$           & hydrophilic residue partially  & completely exposed                                    \\       
+                                               & exposed to the solvent         & to the solvent                                        \\ \hline        
+    Ala-26 favors  $4 \beta + \alpha$          & hydrophobic residue oriented   & oriented towards the                                   \\ 
+                                               & towards surface                & hydrophobic core                                       \\ \hline
+    Ile-49 favors $3 \alpha$                   & hydrophobic interaction        & exposed to the                                          \\
+                                               & inside the protein             & solvent                                                \\ \hline 
+    Val-54 favors $4 \beta + \alpha$           & hydrophobic residue exposed    & hydrophobic interaction                                \\
+                                               & to the solvent                 & inside the protein                                      \\ \hline
+    Glu-56 favors $3 \alpha$                   & hydrophilic residue partially  & completely exposed                                    \\
+                                               & exposed to the solvent         & to the solvent                                        \\ \hline
+    Phe-30 in GB sequence                      & exposed to the solvent         & oriented towards the                                  \\
+    favors $4 \beta + \alpha$                  & in protein surface             & hydrophobic core                                      \\ \hline   
+    Tyr-45 in GB Sequence                      & No H-bonding                   & H-bond with Asp-47                                    \\  
+    favors $4 \beta + \alpha$                  &                                &                                                       \\ \hline
+    Asp-47 in GA sequence                      & H-bond with Lys-50             & No H-bond, No Tyr-45                                  \\  
+    favors $3 \alpha$                                 &                         & in GA sequence                                         \\ \hline
+\end{tabular}
+\end{center}
+\end{table}
+
 \begin{figure}
@@
 the two folds are reported for each sequence.}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=7 in,height=5.8 in]{orban_full.pdf}
+\label{fig:orban_full}
+\caption{Orientation of residues of different sidechains in $4 \beta + \alpha$ and $3 \alpha$ form
+A. Orientation of Hydrophobic residues Leu-5 and Leu-7.
+B. Gln-11 has H-bond with Glu-15 in $3 \alpha$, which is absent in $4 \beta + \alpha$
+C. Hydrophobic Ala-26
+D. Hydrophobic Val-54.
+E. Hydrophilic Glu-14.
+F. The plot of free energy difference for GA95 and GB95 sequence. Positive peak stabilize $3 \alpha$ form and negative peak 
+stabilize $4 \beta + \alpha$. 
+G. Hydrophilic Glu-56 .
+H. Orientation of Hydrophic Ile-49.
+I. With GA95 sequence, Ile-49 have hydrophic interaction with both Leu-20 and Ile-30.
+J. Larger Phe-30 in GB sequence, move towards the protein surface in $4 \beta + \alpha$.
+K. Tyr-45 form H-bond with Asp-47 in $4 \beta + \alpha$.}  
 \end{figure}
@@
 subset of CASP predictions}
-\Alberto{Arijit: we need a global table of success, rather than individual tables. The simplest form
-    would have 3 columns: name of the PDB, number of structures tested, percent of success of
-    picking 1st model}
+
+%\Alberto{Arijit: we need a global table of success, rather than individual tables. The simplest form
+%    would have 3 columns: name of the PDB, number of structures tested, percent of success of
+%    picking 1st model}
 
@@
 tested this using different models from CASP (critical assessment of structure
-prediction). CASP is a blind test world wide competition in which different groups apply their
+prediction). CASP is a blind test in which different groups world wide apply their
 methods to predict the 3D structure of proteins given their sequence. This is done with a strict 3
@@
 solution to this problem. We have tried two experiments centered on CASP. First we tried to rank
-order some structures from the previous CASP9 experiment. Our second test was to participate as a
-metapredictor group in CASP10. In both methods, our ranking is determined as a free energy measure,
+order some structures from the previous CASP9 experiment. We have tested different options for that purpose.
+First, we rank order submition from same group. Then we rank submition from different groups. Our second test was 
+to check whether this method can do better quality assessment/metaprediction than the best group in that catagory in CASP9. 
+In both methods, our ranking is determined as a free energy measure,
 and it is compared with the ranking given by a geometrical comparison between native and submitted
@@
 
-\subsubsection{CASP9}
+\subsubsection{confinement method can rank different models submitted by the same group and distinguish the native structure}
 
-The question that we want to answer here is whether we can distinguish between native structure and the submitted models? 
+The question that we want to answer here is whether we can rank models submitted by the same group.In all cases, we have calculated the free energy
+difference between the native structure and the submitted models as well. 
 %Identifying the native state from a set of models is a very easy test, that most metapredictors
@@
 On the other hand, as shown in Figure~\ref{fig:T0559} confine and release method not only can differentiate the native 
-structure from the submitted models , it can also correctly rank all the models.
+structure from the submitted models , it can also correctly rank all the models like their The GDT\_TS score.
 
@@
 consistency, we have also done our analysis with models and crystal structure consisting residues 3 to 66. As expected,
-the native state is identified correctly, and the two other models are ranked in the correct order.
+the native state is identified correctly, and the two other models are ranked in the correct order in accordance with the GDT\_TS score.
+
+\subsubsection{Confinement method can rank models submitted by different groups}
 
@@
 
-There are however some instances where the method can fail, specially when GDT scores are
+\subsubsection{Per residue free energy calculation identify the residues that are responsible for difference between two conformation}
+
+In order to understand the per residue contribution towards the free energy, we have first calculated the total free energy of NMR 
+structure of a domain of adhesion exoprotein from Pediococcus pentosaceus and its submitted CASP9 model. The 
+pdb id and CASP code of this 79 residue protein are 2KWY and T0569 respectively. We have compared the native structure with the
+best model structure submitted for this target in CASP9. This model structure was submitted by the group " Mufold". All the
+five model structures submitted by this group was found to be close, within a GDT\_TS score  of 72 to 78. Thus we proceed by 
+comparing only the native structure and
+the best predicted model of this target. Visualization upon superimposition of two structures reveals that
+the difference between the two structure is at the region consist of residues 48 to 65, where model1 is much more disordered compare to the
+native structure(see Figure~\ref{fig:T0569}). Other regions look similar for both the cases with of model1 is within 2.6 \AA backbone
+rmsd value of the native structure. We found that, confinement method can still differentiate between the native structure and 
+the generated model. We have also calculated the per residue free energy with a aim that this can identify the residue which are responsible
+for such free enerfy difference. The per residue free energy decomposition is shown in  Figure~\ref{fig:T0569_per_residue}(A). We indeed 
+found that two hydrophobic residues Val-59 and Ile-61 are oriented towards the solvent in the protein 
+surface in the generated model (Figure~\ref{fig:T0569_per_residue}(B). This the region, where there is a beta sheet in the original crystal structure, which is absent in the 
+model. There are also some region, which stabilizes the generated model compare to that of the crystal. A close inspection reveal that 
+the hydrophilic Glu in position 67 is completely exposed to the protein surface and solvent in the best model, whereas it is partially exposed in the 
+crystal. This is presented in Figure~\ref{fig:T0569_per_residue}(C). The peak at position 76 is due to the H-bond between Lys-76 and Asp-11, 
+which is absent in original crystalographic structure (Figure~\ref{fig:T0569_per_residue}(D)).
+
+\subsubsection{Confinement method fails ?}
+
+There are however some instances where the method can fail, specially when GDT scores and all the structures are
 very close. We have found this for an engineered protein from Asr4154 protein (PDB ID: 2L09 and CASP Target T0538).
@@
 crystal structure. This is particularly difficult target for rank ordering as all the models those are considered for this 
-calculation are within RMSD values of 2.6 \AA.
-
-There is another instance where this method fails. This is the case of solution NMR structure of a domain of adhesion exoprotein 
-from Pediococcus pentosaceus. The
-pdb id and CASP code of this 79 residue protein are 2KWY and T0569 respectively. We have compared the native structure with the
-best model structure submitted for this target in CASP9. This model structute was submitted by the group " Mufold". ALl the
-five model structures submitted by this group was found to be close. Thus we proceed by comparing only the native structure and
-the best predicted model of this target. Visualization upon superimposition of two structures reveals that
-the difference between the two structure is at the region consist of residues 48 to 65, where model1 is much more disordered compare to the
-native structure(see Figure~\ref{fig:T0569}) . Other regions look similar for both the cases with of model1 is within 2.6 \AA
-rmsd value of the native structure. In this case we found that the confinement method predict that the predicted model is stable than the
-native structure with a free energy difference of $9.4 \pm 0.7$ Kcal/Mol.
+calculation are within RMSD values of 1.6 - 2.0 \AA. In order to understand the reason behind the mismatch of GDT\_TS values
+of crystal and the model1, we have decomposed the total free energy into the per residue component. A close analysis of 
+the per residue free energy value, we found that although, crystal and different model have closed GDT\_TS and backbone RMSD 
+value, there are number of sidechains oriented in a very different way. As shown in Figure~\ref{fig:T0538compare}(A),
+there is a salt bridge between Arg-32 and Glu-35 present in the crystallographic structure, which is absent in the
+model 1. In model 1, Arg-32 has another salt bridge with Glu-28. Glu-28 in crystal has the salt bridge with Lys-24. In model 1, Lys-24 is
+exposed towards the protein surface. This is shown in Figure~\ref{fig:T0538compare}(A). There are more such interesting and may be crucial
+sidechain differences between the model and the crystallographic structure, which indicate the limitation of the GDT\_TS based 
+ranking of the protein structure. As, shown in Figure~\ref{fig:T0538compare}(B), Arg-26 and Glu-50 has a salt bridge in the crystallographic 
+structure. Also, Phe-51 is oriented towards the protein core in the crystallographic structure. In model 1, the salt bridge between Arg-26 and Glu-50
+is absent and the Phe-51 is oriented towards the protein surface.
 
@@
 
-\subsubsection{Difficulty in CASP Experiments}
+%\subsubsection{Difficulty in CASP Experiments}
 
-It is often difficult to model CASP targets using the confinement and release method. There are many cases,
-where the sequence of a particular target is trimmed during actual result presentation. This is illustrated
-with the crystal structure of the leucine zipper domain of cGMP dependent protein kinase I (pdb id: 3NMD
-and CASP9 code: T0605) as shown in Figure~\ref{fig:T0605}. The actual CASP 9 target consisted of 72 amino acid residues, but
-as only 18-66 amino acid residues could be detected in actual crystal structure of this protein, the final result was trimmed 
-with respect to that crystal structure with 49 amino acid residues which consist of a single alpha helix. The amino acid residues 18 and 66 are shown in
-three models submitted by the group "Baker". It is important to note that blind prediction of model ranking on the basis of free energy can be hugely 
-affected in such scenario as the only major difference between those models are the trimmed parts, specially the region consisting residues 1-18.
+%It is often difficult to model CASP targets using the confinement and release method. There are many cases,
+%where the sequence of a particular target is trimmed during actual result presentation. This is illustrated
+%with the crystal structure of the leucine zipper domain of cGMP dependent protein kinase I (pdb id: 3NMD
+%and CASP9 code: T0605) as shown in Figure~\ref{fig:T0605}. The actual CASP 9 target consisted of 72 amino acid residues, but
+%as only 18-66 amino acid residues could be detected in actual crystal structure of this protein, the final result was trimmed 
+%with respect to that crystal structure with 49 amino acid residues which consist of a single alpha helix. The amino acid residues 18 and 66 are shown in
+%three models submitted by the group "Baker". It is important to note that blind prediction of model ranking on the basis of free energy can be hugely 
+%affected in such scenario as the only major difference between those models are the trimmed parts, specially the region consisting residues 1-18.
 
@@
 In this background we try to answer, whether the confinement method can do better quality assessment compare to the best performing
-group (MUFOLD-WQA) in this catagory.
-We have again picked up the CASP taget T0538 to answer this question. We have chosen the Model 3 submitted by PconsR (GDT\_TS = 96) 
-and model 5 from the MULTICOM-NOVEL ( GDT\_TS = 83). Both are server predicted models and they were further used for quality assessment 
+group (MUFOLD-WQA) in this catagory. In CASP9, the overall reliability of models in quality assessment (QA) was accepted in a mode known as
+QMODE 1. Here, predictors were asked to score each model on a scale from 0 to 1, with higher values corresponding to better models \cite{Kryshtafovych2011}. 
+We have again picked up the CASP taget T0538 to answer this question. We have chosen the Model 3 submitted by PconsR (GDT\_TS = 96, QMODE 1 = 0.5434) 
+and model 5 from the MULTICOM-NOVEL ( GDT\_TS = 83, QMODE 1 = 0.5865). Both are server predicted models and they were further used for quality assessment 
 by the group MUFOLD-WQA \cite{Wang2011}. It is important to note that the model from PconsR was the most accurate predicted model for this 
-target and the model from MULTICOM-NOVEL was predicted as best by the quality assessment experiments of MUFOLD-WQA. 
-Our free energy analysis we have shown that model from Pconsr is better than the model from 
-MULTICOM-NOVEL $(\Delta \Delta G (Model\_Pconsr - Model\_MULTICOM-NOVEL = -8.2 Kcal/Mol)$, which   
+target and the model from MULTICOM-NOVEL was predicted as best by the quality assessment experiments of MUFOLD-WQA. Also, 
+the QMODE 1 value presented here is from the group MUFOLD-WQA. 
+Our free energy analysis predict that the model from Pconsr is indeed better than the model from 
+MULTICOM-NOVEL $(\Delta \Delta G (Model\_Pconsr - Model\_MULTICOM-NOVEL = -3.9 Kcal/Mol)$, which   
 matches well with the GDT\_TS score. On the other hand the MUFOLD-WQA predicted the ranking in the reverse order.
-There is no doubt that the consensus approaches can predict the model quality in a faster manner. But we expect that confinement
-approach can predict the model quality in a relatively expensive but much accurate way.
 
-\Arijit{Will upload the result from Target 560 once cluster is back.}
+We have also carried out the similar type of analysis for the case of Target 560 of CASP9 experiment. For that 
+we have chosen the model 1 from the group PconsR (GDT\_TS = 94, QMODE 1 = 0.5178), Bilab\-Enable (GDT\_TS = 87, QMODE 1 = 0.5344) 
+and Multicom\_Refine (GDT\_TS = 74, QMODE 1 = 0.4770). If, one want to rank order with respect to the GDT\_TS values, the model from 
+PconsR is the best predicted model, but according to the QA assessment by MUFOLD-WQA, the model from Bilab-Enable was found to be best predicted model. 
+All the QMODE 1 values are from MUFOLD\-WQA. In CASP 9, the sequence of 74 residues were given to predict the structure and for 
+quality assessment, whereas during GDT\_TS analysis only residues 3-66 were considered. When we considered the full 74 amino acid 
+residues, the free energy calculated were $(\Delta \Delta G (Model\_Pconsr - Model\_Bilab\-Enable = 3.3 Kcal/Mol)$ and 
+$(\Delta \Delta G (Model\_Pconsr - Model\_Multicom\_Refine = -30.0 Kcal/Mol)$ which 
+support the QA assessment by MUFOLD-WQA. On the other hand, if we calculate the same free energy with the trimmed version of the protein 
+(with residues 3-66), the values of the free energy difference become -1.7 and -31.1 Kcal/Mol respectively, which support the GDT\_TS score. 
+The summary of this whole observation is presented in Table~\ref{table:560QA}.
+There is no doubt that the consensus approach can predict the model quality in a faster manner. But we expect that confinement
+method can predict the model quality in a relatively expensive but much accurate way.
+
+\begin{table}
+\label{table:560QA}
+\begin{center}
+\begin{tabular}{| l | l | l | l | l |}
+\hline
+Model            &  $\Delta \Delta G$ (1-74)  & $\Delta \Delta G$ (3-66) &    GDT\_TS    & QMODE 1  \\ \hline
+PconsR           &         0.0                &      0.0                 &    94         &  0.5178   \\ \hline
+Bilab\-Enable     &        -3.3                &      1.7                 &    87         &  0.5344   \\ \hline
+Multicom\_Refine &        30.0                &     31.1                 &    74         &  0.4770   \\ \hline
+\end{tabular}
+\end{center}
+\caption{Comparison of quality assessment by the group MUFOLD\-WQA and the confinement method of the CASP9 target 560 (pdb id: 2L02). 
+The GDT\_TS score obtained from CASP9 website, was based on the trimmed model containing residue 3-66, whereas, the QMODE 1 values are from
+quality assessment by the group MUFOLD\-WQA and based on residues 1-74.}
+\end{table}
 
@@
 
-Proteins are essential parts of organisms and participate in virtually every process within living cells. 
-In order to understand different functions of proteins, detect diseases caused by them and design more efficient drugs 
-it is necessary to have the three dimensional structure of the protein. Experimental studies using NMR and X-ray crystallography is 
-not able to keep up with the sheer number of protein sequences available. In recent past there are number of studies directed towards understanding protein 
+In recent past there are number of studies directed towards understanding protein 
 structure using theoretical tools. Experimentally, it is relatively easy to identify the native structure from the data obtained 
@@
 also depend on potential energy functions to pick out their best structure, lacking an entropic
-component. In this paper, we have used a free energy method to identify
-the native/native like structure. First, we have tested this method for a chameleon sequence, which has a $95$ percent sequence identity
-but one of them has a $3 \alpha$ structure and another one has $4 \beta + \alpha$ structure. Using the current method, we found that 
-the native structure is always the most stable \Alberto{This is not true!!}. We have also tested this method for different targets of CASP experiments. We found
-that in most cases, it can pick out the native/native like structure in terms of lowest free enrgy value.
-
-However, there may be some limitations to this method in terms of size of the protein molecule. The normal mode analysis may not be
-accurate for proteins larger than 100 amino acid residues. This can be approximated by removing the hydrogen molecules from 
-the system. 
+component. In this work we attempted to identify the native/native like protein strcuture from the decoy using free energy as a scale.
+We have used previously developed confinement method to larger systems.
+We found that in majority of the case, the native structure is always the most stable and we can rank the protein structure upto 100 amino 
+acid residues long. This method works very well if atleast one of the strcuture in the decoy is good. The main advantage of this method is 
+that it does not require any reaction coordinate. An interesting application, that we have developed using this method is per residue
+free energy calculation. Although, we have calculated that in an approximate way, but it give important information regarding conformational 
+preference of a particular residue. In future, this may help for protein design. 
+First, we have tested this method for a series of chameleon sequences, which have $88$, $95$ and $98$ percent sequence identity
+but one of them has a $3 \alpha$ structure and another one has $4 \beta + \alpha$ structure. We found that the method can pick the 
+the conformational preferance of a particular sequence, which matches with the experimental results. Per residue free energy calculation 
+identify the residues that stabilize/destabilize a particular conformation and can help for better protein design. 
+Encouraged by that result we have tested this method for different targets of CASP experiments. We found 
+in most cases, it can pick out the native/native like structure in terms of lowest free enrgy value and rank the protein structure
+with terms of free energy. 
 The method can be also give wrong result if all the models in the decoy are very poor. It may also be difficult to rank structures if 
@@
 between the good and the bad structure. In this article we have extensively tested this method using examples from CASP experiment.
-This method can be computationally expensive in actual CASP experiment if one want rank thousands of submitted models in CASP 
-experiments. However, in real life experiment with a particular target this method can give accurate result. The important 
-fact about Confinement method is that all the part of the calculation can be done independently. We have mostly carried out 
-our calculation with Amber program \cite{Case2005} in GPU computer. With a avarage 56 residue we found that it take only 4 hour to complete 
-20 ns of the confinement step. So, with available computer power this method will be easy to use.
+This method can be computationally expensive in actual CASP experiment if one want rank hundreds of submitted models in CASP 
+experiments. However, in real life experiment with a particular target, we expect this method can give accurate result. The important 
+fact about confinement method is that all the part of the calculation can be done independently. 
+There is another big bottlenecks in this calculation. This has to do with the amount of
+simulations that have to be performed to confine the ensemble of conformations close to a microstate
+into a single microstate. We have tackled this issue by using GPU (Graphical Processing Unit) technology as opposed to the classical
+CPUs, which gives us a two order of magnitude boost in computational efficiency.
+For example, we have mostly carried out our calculation with Amber program \cite{Case2005} in GPU computer. With a avarage 56 residue 
+we found that it take only 4 hour to complete 20 ns of the confinement step. So, with available computer power this method will 
+be easy to use for real problems. 
+%There is another issue, which is normal mode
+%analysis calculations are Oh(N**3) and require large amounts of memory, effectively limiting the
+%size of the systems under study. In most cases quasiharmonic analysis.
 
@@
 
-In all the calculation amber 10 program is used with ff99SB with GB/SA implicit solvent.
+In all the calculation amber 11 program is used with ff99SB with GB/SA implicit solvent. Interestingly, we extend the
+method for calculation of per residue free energy in an approximate way. For this purpose, the confinement energy, 
+$\Delta G_{A,A*}$ and $\Delta G_{B,B*}$ of each residue is calculated in the usual numerical way as described above. The
+internal energy of each residue is calculated using the decomp module of amber from the final restrained trajectory. We 
+call this method approximate as we ignore the contribution from the normal mode or quasiharmonic analysis. 
+However, this contrbution to the total free energy is very less (less than $1.0 Kcal/Mol$), which allow us to 
+study the mechanistic details of conformational preference of each residue.
 
-\Alberto{Somewhere in the introduction or methods}
-There are two big bottlenecks in this calculation. The first has to do with the amount of
-simulations that have to be performed to confine the ensemble of conformations close to a microstate
-into a single microstate, the second bottleneck has to do with the size of the system in the normal
-mode calculation. We have tackled the first issue by using GPU (Graphical Processing Unit) technology as opposed to the classical
-CPUs, which gives us a two order of magnitude boost in computational efficiency. Normal mode
-analysis calculations are Oh(N**3) and require large amounts of memory, effectively limiting the
-size of the systems under study. Where as this can be overcome by larger memory machines, these are
-often not available, and so we propose an alternative by using quasiharmonic analysis. This is still
-an Oh(N**3) problem, but we now the motions of atoms are already determined in trajectories, and so
-we can choose to do quasiharmonic analysis excluding hydrogens, which will reduce the number of
-atoms N to almost N/2. This means that we can use this method for systems twice as large.
+%There are two big bottlenecks in this calculation. The first has to do with the amount of
+%simulations that have to be performed to confine the ensemble of conformations close to a microstate
+%into a single microstate, the second bottleneck has to do with the size of the system in the normal
+%mode calculation. We have tackled the first issue by using GPU (Graphical Processing Unit) technology as opposed to the classical
+%CPUs, which gives us a two order of magnitude boost in computational efficiency. Normal mode
+%analysis calculations are Oh(N**3) and require large amounts of memory, effectively limiting the
+%size of the systems under study. Where as this can be overcome by larger memory machines, these are
+%often not available, and so we propose an alternative by using quasiharmonic analysis. This is still
+%an Oh(N**3) problem, but we now the motions of atoms are already determined in trajectories, and so
+%we can choose to do quasiharmonic analysis excluding hydrogens, which will reduce the number of
+%atoms N to almost N/2. This means that we can use this method for systems twice as large.
 
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.5 in,height=3.0 in]{bhp.pdf}
-\end{center}
-\caption{Two conformations from $\beta$ hairpin from protein G, bhp1 and bhp2. The two stranded $\beta$ sheet, bhp1 is the native structure and the three stranded $\beta$ sheet is known as bhp3.}
-\label{fig:bhp_conf}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.5 in,height=3.0 in]{G95.pdf}
-\end{center}
-\caption{Two protein with 95 \% similar sequence but different folds}
-\label{fig:G95}
-\end{figure}
+%\begin{figure}
+%\begin{center}
+%\includegraphics[width=3.5 in,height=3.0 in]{G95.pdf}
+%\end{center}
+%\caption{Two protein with 95 \% similar sequence but different folds}
+%\label{fig:G95}
+%\end{figure}
 
@@
 \label{fig:T0538}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=4.9 in,height=4.0 in]{Target_538_compare.pdf}
+\end{center}
+\caption{Difference of the sidechain orientation between model 1 and crystallographic structure in Target-538.
+A. The salt bridge between Glu-35 and Arg-32 in crystal structure. In model 1, this is compensated by H-bonding between 
+Arg-32 and Glu-28. There is another salt bridge between Glu-28 and Lys-24.
+B. Salt bridge between Arg-26 and Glu-50 in crystal structure. Orientation of Phe-51 is different in crystal than in model1.}
+\label{fig:T0538compare}
 \end{figure}
@@
 \begin{center}
-\includegraphics[width=12cm,height=10cm]{T0605.pdf}
+\includegraphics[width=5 in,height=4 in]{Target_569_per_residue.pdf}
 \end{center}
-\caption{ Three models of the leucine zipper domain of cGMP dependent protein
-kinase I (original pdb id: 3NMD and CASP target: T0605). The original target sequence of this protein was 72 residues.
-However only residues 3 to 66 is only kept during result announcement as that part can be found from crystallographic analysis. The
-region 3 to 66 are almost identical in all the 3 models  }
-\label{fig:T0605}
+\caption{A. Plot of per residue free energy difference between the model and the crystal structure of Traget 569 (pdb id: 2KWY) of CASP9.
+B. In model 1, the beta sheet containing Val-59 and Ile-61 is disordered. These hydrophobic residues are exposed to the solvent.
+C. Glu-67 is partially exposed to the solvent in crystal and it is fully exposed in model 1.
+D. There is a H-bond between Lys-76 and Asp-11 in model 1, which is absent in the crystallaographic structure.}
+\label{fig:T0569_per_residue}
 \end{figure}
+
+%\begin{figure}
+%\begin{center}
+%\includegraphics[width=12cm,height=10cm]{T0605.pdf}
+%\end{center}
+%\caption{ Three models of the leucine zipper domain of cGMP dependent protein
+%kinase I (original pdb id: 3NMD and CASP target: T0605). The original target sequence of this protein was 72 residues.
+%However only residues 3 to 66 is only kept during result announcement as that part can be found from crystallographic analysis. The
+%region 3 to 66 are almost identical in all the 3 models  }
+%\label{fig:T0605}
+%\end{figure}
 

### 2013-01-09 Paper/paper.tex
Fleshed out Intro and abstract

@@
 The calculation of free energy differences is of central importance in the simulation of biochemical
-systems.  The computation of the free energy difference between pairs of macromolecular
-conformations is particularly difficult and computationally expensive task with existing
-methods. In this work, confinement approach is used to calculate absolute
-free energies of bio molecular systems. The method provides two main advantages: it does not require a reaction coordinate or
-transition path and it is fast to compute. The free energy calculated can be decomposed into per residue contribution in a 
-approximate way. Per residue free energy allow us
-to identify the reason behind conformational preference of the biomolecules. Through out the article we show its use in different
+systems. It is particularly difficult to calculate between pairs of macromolecular
+conformations as well as a computationally expensive task with existing
+methods. In this work, the confinement approach is used to calculate absolute
+free energies of biomolecular systems. The method provides two main advantages: it does not require a reaction coordinate or
+transition path and it is fast to compute. The free energy calculated can be decomposed into a per
+residue contribution in an 
+approximate way. Per residue free energy allows us
+to identify the reason behind conformational preferences in biomolecules. Through out the article we show its use in different
 challenging modeling problems. In particular, we show its use in predicting the conformational
 preference of chamaleon sequences (sequences with high sequence identity and different
-folds). This sequence dependent conformational preference and per residue free energy decomposition set the stage for the use of this 
+folds). This sequence dependent conformational preferences and per residue free energy decomposition set the stage for the use of this 
 method in protein design.
@@
 
-In computational structural biology there are numerous cases where free energy between well defined
-states are necessary. Examples ranges from small to large conformational change of protein due to
-ligand binding, change of pH etc \cite{Meirovitch2007}, \cite{Chipot2007}, \cite{Jorgensen2004}. 
-Free energy also play an important role in case of protein
+In computational structural biology there are numerous cases where the free energy difference
+between two well defined 
+states plays an important role. Examples range from small to large conformational changes of proteins due to
+ligand binding, change of pH or other conditions \cite{Meirovitch2007}, \cite{Chipot2007}, \cite{Jorgensen2004}. 
+Free energy also plays an important role in the case of protein
 folding. The ground breaking work of Christian B. Anfinsen and coworkers showed that the native
 structures of small globular proteins have a unique, thermodynamically stable native structure with
-their conformation at the global free energy minimum \cite{Anfinsen1973}. Regardless of the starting point or unfolding
-it by changing different condition most proteins will finally assume the same structure. It is very
-important for proteins to achieve their native conformation since defects in protein folding may be
-the molecular cause of a range of human genetic disorder. For example a misfolded protein known as
-prion when enters any healthy organism it converts properly folded protein into misfolded one.
-Thus, Free energy can also act as a scale to identify between the misfolded and the native state of
-the protein. These facts further give free energy a special importance in structural biology.
+their conformation at the global free energy minimum \cite{Anfinsen1973}. Regardless of the starting
+point or even after unfolding 
+it by changing different condition most proteins will go back to the native state after
+reinstituting native like conditions. Often, the changes in free energy between different
+conformations is small, but nonetheless crucial in favoring the native state
+ since defects in protein folding may be
+the molecular cause of a range of human genetic disorders. For example a misfolded protein known as
+a
+prion can misfold correctly folded proteins when entering a healthy organism.
+\Alberto{I'm not sure I understand this here} \Arijit{Thus, Free energy can also act as a scale to identify between the misfolded and the native state of
+the protein.} These facts further give free energy a special importance in structural biology.
 
-However, the theoretical calculation of conformational free energy change become difficult if the
-states of interest are very different \cite{Meirovitch2007}. In such scenario timescale involved in such tranformation may
-be beyond timescale of direct classical molecular dynamics simulation.  Generally, in order to
-calculate free energy a pathway or reaction coordinate is created with the intermediate structures.
-Method like umbrella sampling \cite{Torrie1977}, thermodynamic integration \cite{Tironi1994} along with classical MD can suffer from
-overlap problem and require large computational effort. The idea of reaction coordinate became more
-complicated in case of protein folding.  As free energy landscape of the protein folding 
-is rough \cite{Dill1997}, \cite{Dill2008}
-it can be trapped in a energy well during its visit in the conformational space. Even with the
-advanced method like replica exchange molecular dynamics the three dimensional structure may be
-trapped in a local minima for a considerable time and trapped in a local energy minima
-and one can mistakenly identify the misfolded state as the native state.
+However, the theoretical calculation of conformational free energy change becomes difficult if the
+states of interest are very different \cite{Meirovitch2007}. In such scenario timescales involved in
+such conformational transitions may
+be beyond the sampling ability of classical molecular dynamics simulation.  Generally, in order to
+calculate the free energy a pathway or reaction coordinate is needed with several intermediate structures.
+Methods like umbrella sampling \cite{Torrie1977}, thermodynamic integration \cite{Tironi1994} along with classical MD can suffer from
+overlap problem and require a large computational effort. The idea of a reaction coordinate became
+even more
+complicated in the case of protein folding.  As the free energy landscape for protein folding 
+is rough \cite{Dill1997}, \cite{Dill2008}, the molecule 
+can get trapped in an energy well during conformational sampling. Even with 
+advanced methods like replica exchange molecular dynamics, the three dimensional structure may be
+trapped in a local minima for a considerable time, giving rise to nonconverged distributions where
+misfolded states might be selected as native like
 % (give the example of human pin1ww domain from Benout Roux and Klaus Schultan).
 
-Calculation of free energy has been successfully attempted by a number of groups \cite{Meirovitch2007}, 
-\cite{Ytreberg2006} - \cite{Zheng2008}, . In recent years methods like Refrence system method \cite{Ytreberg2006}, 
-Decativated Morphing \cite{Park2008}, Orthogonal Space Random Work \cite{Zheng2008}, Confinement 
-Method \cite{Tyka2006}, \cite{Cecchini2009} have came out for calculation of free energy method. 
+The calculation of free energy has been successfully attempted by a number of groups \cite{Meirovitch2007}, 
+\cite{Ytreberg2006} - \cite{Zheng2008}. In recent years methods like the Reference system method \cite{Ytreberg2006}, 
+Decativated Morphing \cite{Park2008}, Orthogonal Space Random Work \cite{Zheng2008} or the Confinement 
+Method \cite{Tyka2006}, \cite{Cecchini2009} to name a few have been used to calculate free energy
+differences.
 In this work we have applied the confinement method which was originally devoled by Tyka et. al. \cite{Tyka2006} and
-Cecchini et. al. \cite{Cecchini2009}. This method relies on the fact that free energy is a state function and 
-thus independent of path. It uses a thermodynamic cycle where first, non-harmonic degrees of
-freedom are removed by applying a series of restraint to the system. Then the free energy of the
+Cecchini et. al. \cite{Cecchini2009}. This method relies on the fact that the free energy difference is a state function and 
+thus it is independent of the sampling path. This approach uses a thermodynamic cycle where first, non-harmonic degrees of
+freedom are removed by applying a series of restraints to the system. Then the free energy of the
 remaining harmonic system is obtained using a normal mode calculation and combined with the
@@
 first reproduced that known result. Then we have applied this method to larger proteins for the
-first time. The free energy of conformation of pair of proteins with similar sequence but completely
-different fold was calculated. Encouraged by the result, we applied the method of different targets
-of CASP9 competition and tested whether it can predict the most stable conformation of the protein based 
-on its free energy. The remarkable ability of confinement method to identify residues which 
-stabilizes/destabilizes a particular conformation from per residue free energy analysis set the stage
-for future application of this method in protein design.
+first time. The free energy difference of pairs of conformations in proteins with similar sequence but completely
+different fold was calculated. \Alberto{i made this part too long, have to summarize here and define
+better inside text?-->} Encouraged by the result, we applied the method to help tackle the
+problem of protein conformation ranking. We did this by selecting different targets from the
+Critical Assessment of Structure Prediction (CASP) event. In this event, different research groups
+try to find the 3D structure of a protein given a sequence in a blind way using their methodologies. Upon comparing
+with the real structure it is often seen that several groups are able to generate better structures
+than the ones they submit, but are not able to identify them. This is known as the ranking problem,
+and this method is able to correctly rank structures in the majority of cases we have tried.
+ Finally, we have been exploring the ability of working on a per residue basis to identify which
+ residues are the major players in the stabilization/destabilization of a pair of conformations. Our
+ initial experience makes us thinkg that this method might be helpful in identifying key residues
+ for protein design.
 

### 2013-01-09 Paper/paper.tex
Al. Done with Chamaleon sequences

@@
 
-As a first step, we performed several control experiments to verify that our implementation of the
+\Alberto{ This is benchmarking, good internal, but not relevant. Should go to SI,out, or summarized
+    in one sentence:  "After reproducing the results of Cecchini and coworkers we were interested in
+the ability to apply this method to larger systems"}
+
+\Alberto{As a first step, we performed several control experiments to verify that our implementation of the
 confinement method produces results compatible with previous calculations reported in the
@@
 is in good agreement with the equilibrium simulations and with previous calculations using the
-confinement method \cite{Cecchini2009} (see Supporting Information for further details).
+confinement method \cite{Cecchini2009} (see Supporting Information for further details).}
 
@@
 residues in length. In this work we apply the confinement method to
-larger proteins. On that direction, We first test the confinement method for a chamelon sequence and 
+larger proteins. On that direction, we first test the confinement method for a chamelon sequence and 
 found that it can correctly pick out the prefential structure of the chamelon sequence. Other 
@@
 the lowest free energy. As a control, for several targets we also calculated the relative free
-energy of the experimentally determined structure. As
+energy of the experimentally determined structure. \Alberto{should not start with results yet, first
+we are describing what we are going to show them} As
 Table~\ref{table:casp_control} shows, in most cases, the confinement method correctly assigns a
@@
 examples---often referred to as chameleon sequences---of proteins with similar sequences that have
-remarkably different structures. Orban and co-workers have designed a series of 56-residue proteins
-(based on Protein G) that adopt one of two different folds depending on small changes in sequence
-(see Figure~\ref{fig:orban}). Sequences that adopt a mixed alpha/beta structure similar to Protein G
+remarkably different structures. Orban and co-workers have designed a sequence of 56-residues that
+is marginally stable in one of two possible folds. By mutating key residues in this sequence they
+are able to stabilize one fold or the other (see Figure~\ref{fig:orban}). Sequences that adopt a mixed alpha/beta structure similar to Protein G
 are denoted as ``GB'', while sequences that form a three-helix bundle are denoted as ``GA''. One
 pair of sequences (GA88/GB88) are 88 percent identical in sequence and differ in seven positions. Another pair
-(GA95/GA95) are 95 percent identical and differ in three positions. The last pair (GA98/GB98) differ
+(GA95/GB95) are 95 percent identical and differ in three positions. The last pair (GA98/GB98) differ
 only in a single tyrosine to alanine mutation.
@@
 
-To avoid this potential problem, we instead computed the relative free energy two different
+To avoid this potential problem, we instead computed the relative free energy of two different
 structural models for each sequence. One model is based on the GA structure and the other on the GB
@@
 
-We decomposed the calculated free energy into per residue decomposition as described in method section. This help us for
-identification of important residues that stabilizes a particular conformation. 
-In this direction, We first calculate the per residue free energy for the GA95 and GB95 sequences. Similar to the total 
+We decomposed the calculated free energy into per residue decomposition as described in the methods
+section. This helps us to
+identify important residues that stabilize a particular conformation. 
+\Alberto{This is not Reader Friendly at all. we do not have to describe every little detail. I would
+    rather have a table where we have the pairing and the reason for favoring one structure or the
+    other. In the main text I would say that the main reason for preferences of one fold or another
+    has to do with the relative solvent exposure of the residue in one conformation or another, or
+    the possibility of stablishing hydrogen/salt bridges in one conformation. This would be my
+preference, ditching everything below until the next section. What do you think?}
+In this direction, we first calculate the per residue free energy for the GA95 and GB95 sequences. Similar to the total 
 free energy for each sequence, we calculate the free energy difference between $4 \beta + \alpha$ and $3 \alpha$ structure for each 
@@
 experimental observations classified the protein into two parts: Amino acids 9-51 are fully structured in both
-the fold, where as residues 1-8 and 52-56 are unstructured in $3 \alpha$, but form $\beta$ strand in $4 \beta + \alpha$ structure.
-Most of the amino acid residues in the region 1-9 favors $4 \beta + \alpha$ structure and can also seen from the plot 
+the folds, where as residues 1-8 and 52-56 are unstructured in $3 \alpha$, but form $\beta$ strand in $4 \beta + \alpha$ structure.
+Most of the amino acid residues in the region 1-9 favor $4 \beta + \alpha$ structure as can be seen from the plot 
 Figure~\ref{fig:orban_full}(F). Here, we try to explain some of the largest peak in that plot. The residues Leu-5 and 
-Leu-7 favors $4 \beta + \alpha$ structure as these hydrophobic residues are oriented towards the protein hydrophobic 
+Leu-7 favor $4 \beta + \alpha$ structure as these hydrophobic residues are oriented towards the protein hydrophobic 
 core, whereas in $3 \alpha$ structure they are exposed to the solvent (Figure~\ref{fig:orban_full}(A)). One can argue that 
@@
 is exposed to the protein surface in both the conformation, but in $3 \alpha$ structure it has an additional H-bond with Glu-15 which gives
-it more stability compare to the $4 \beta + \alpha$ structure (Figure~\ref{fig:orban_full}(B)). The peak 
+it more stability compared to the $4 \beta + \alpha$ structure (Figure~\ref{fig:orban_full}(B)). The peak 
 at position 14 is due to the hydrophilic Glu, which is completely
-exposed to the solvent in $4 \beta + \alpha$ conformation as shown in (Figure~\ref{fig:orban_full}(E)), but only partially exposed 
+exposed to the solvent in $4 \beta + \alpha$ conformation as shown in
+(Figure~\ref{fig:orban_full}(E)), but is only partially exposed 
 to the solvent in $3 \alpha$ form. 

### 2013-01-09 Paper/paper.tex
Casp

@@
 week limit on each target (3 day for servers) and each group is allowed to submit 5 possible
-structures (ranked from better to worst). In our role as assessors during the CASP8 and CASP9 \cite{MacCallum2011}
+structures (ranked from best to worst). In our role as assessors during the CASP8 and CASP9 \cite{MacCallum2011}
 competition we observed no correlation at all between the ordering of structures submited by the
-groups and the real ranking compared to an experimental model \cite{Kryshtafovych2011}. The consequences of this go beyond
+groups and the real ranking compared to an experimental model \cite{Kryshtafovych2011}. The
+consequences of this goes beyond
 those five structures; the deeper meaning is that groups producing ensembles of structures are
@@
 the case. Beyond the CASP problem, this reflects on the modelers ability to correctly rank order
-models in many different environments, from structures for drug design leads to designing more
+models in many different environments, such as structures for drug design leads or designing more
 stable proteins or peptide mimetics such as peptoids. One of the main culprits of this lack of
@@
 order some structures from the previous CASP9 experiment. We have tested different options for that purpose.
-First, we rank order submition from same group. Then we rank submition from different groups. Our second test was 
+First, we rank order submitions from same group and then from different groups. Our second test was 
 to check whether this method can do better quality assessment/metaprediction than the best group in that catagory in CASP9. 
@@
 are chosen for this method and its predictive accuracy. In particular, when GDT\_TS scores are below
-50, it is difficult for us to say anything about the models. Our approach is simple, with every
+50, it is difficult for us to say anything about the models. Our approach is simple, for every
 subset of structures we choose, we use the confinement method in a pairwise procedure, and then rank
-the structures. In all cases, we have calculated the free energy with respect to the native structure either obtained from NMR or 
-x-ray crystallography but we do not need to make comparisons to native in order to do rank ordering.
+the structures. In all cases, we have also calculated the free energy with respect to the native
+structure to check the stability of the models.
+However, this step is not needed if we just want to rank order.
 
-\subsubsection{confinement method can rank different models submitted by the same group and distinguish the native structure}
+\subsubsection{The confinement method can rank different models generated by the same methodology and distinguish the native structure}
 
-The question that we want to answer here is whether we can rank models submitted by the same group.In all cases, we have calculated the free energy
+The question that we want to answer here is whether we can rank models that have been generated with
+the same methodology. In all cases, we have calculated the free energy
 difference between the native structure and the submitted models as well. 
@@
 The best predictor group for this target was "BAKER-ROSETTASERVER".
-We initlally checked similarity between the models submitted by "BAKER-ROSETTASERVER" and discarded two of them from the 
+We initially checked similarity between the models submitted by "BAKER-ROSETTASERVER" and discarded two of them from the 
 analysis on the basis of being too similar to some of the other models. The GDT\_TS score and rmsd
-values as shown in Figure~\ref{fig:T0559} indicate that the model 1 was predicted correctly, 
+values as shown in Figure~\ref{fig:T0559} indicate that model 1 was predicted correctly, 
 whereas the order of model 3 and 5 was wrong. The main difference
-between the model 3 and the rest of the model is that the orientation of first alpha helix of model 3 is opposite.
-On the other hand, as shown in Figure~\ref{fig:T0559} confine and release method not only can differentiate the native 
-structure from the submitted models , it can also correctly rank all the models like their The GDT\_TS score.
+between model 3 and the rest of the models is that the orientation of the first alpha helix of model
+3 is \Alberto{opposite??}.
+On the other hand, as shown in Figure~\ref{fig:T0559} the confine and release method not only can differentiate the native 
+structure from the submitted models, the ranking also correlates perfectly with the GDT\_TS score.
 
@@
 again they are similar to the rest of the models. The comparison of GDT\_TS score as shown in 
-Figure~\ref{fig:T0560} and published in final CASP result is only for models with residues 3 to 66. In order to keep
+Figure~\ref{fig:T0560} and published in the final CASP result is only for models with residues 3 to 66. In order to keep
 consistency, we have also done our analysis with models and crystal structure consisting residues 3 to 66. As expected,
@@
 
-\subsubsection{Confinement method can rank models submitted by different groups}
+\subsubsection{The confinement method can rank models generated by different methods}
 
-Our next test was between models from the best submissions of different groups. 
+Our next test was between models that are produced with different methodologies. We do this by
+looking at some of the best CASP submissions from different research groups. 
 For this purpose we have chosen the x-ray crystallographic structure of 
@@
 
-\subsubsection{Per residue free energy calculation identify the residues that are responsible for difference between two conformation}
+\subsubsection{Per residue free energy calculation identifies the residues that are responsible for
+differences between two conformations}
 
-In order to understand the per residue contribution towards the free energy, we have first calculated the total free energy of NMR 
-structure of a domain of adhesion exoprotein from Pediococcus pentosaceus and its submitted CASP9 model. The 
-pdb id and CASP code of this 79 residue protein are 2KWY and T0569 respectively. We have compared the native structure with the
-best model structure submitted for this target in CASP9. This model structure was submitted by the group " Mufold". All the
-five model structures submitted by this group was found to be close, within a GDT\_TS score  of 72 to 78. Thus we proceed by 
-comparing only the native structure and
-the best predicted model of this target. Visualization upon superimposition of two structures reveals that
+We used one of the CASP targets (pdb id=2KWY, casp id=T0569) to study the per residue contribution
+between this NMR structure and the best CASP9 predicted model (from the 'Mufold' group, GDT\_TS=78).
+Visualization upon superimposition of two structures reveals that
 the difference between the two structure is at the region consist of residues 48 to 65, where model1 is much more disordered compare to the
 native structure(see Figure~\ref{fig:T0569}). Other regions look similar for both the cases with of model1 is within 2.6 \AA backbone
-rmsd value of the native structure. We found that, confinement method can still differentiate between the native structure and 
+rmsd value of the native structure. We found that, the confinement method can still differentiate between the native structure and 
 the generated model. We have also calculated the per residue free energy with a aim that this can identify the residue which are responsible
@@
 surface in the generated model (Figure~\ref{fig:T0569_per_residue}(B). This the region, where there is a beta sheet in the original crystal structure, which is absent in the 
-model. There are also some region, which stabilizes the generated model compare to that of the crystal. A close inspection reveal that 
+model. Interestinlgy, some regions are more stable in the generated model. A close inspection
+reveals that 
 the hydrophilic Glu in position 67 is completely exposed to the protein surface and solvent in the best model, whereas it is partially exposed in the 
-crystal. This is presented in Figure~\ref{fig:T0569_per_residue}(C). The peak at position 76 is due to the H-bond between Lys-76 and Asp-11, 
-which is absent in original crystalographic structure (Figure~\ref{fig:T0569_per_residue}(D)).
+crystal (Figure~\ref{fig:T0569_per_residue}(C)). The peak at position 76 is due to the H-bond between Lys-76 and Asp-11, 
+which is absent in the original crystalographic structure (Figure~\ref{fig:T0569_per_residue}(D)).
 

### 2013-01-09 Paper/paper.tex
Jan10 3 Draft

--- 2013-01-09 Paper/paper.tex
@@
 The calculation of free energy differences is of central importance in the simulation of biochemical
-systems. It is particularly difficult to calculate between pairs of macromolecular
-conformations as well as a computationally expensive task with existing
-methods. In this work, the confinement approach is used to calculate absolute
-free energies of biomolecular systems. The method provides two main advantages: it does not require a reaction coordinate or
-transition path and it is fast to compute. The free energy calculated can be decomposed into a per
-residue contribution in an 
-approximate way. Per residue free energy allows us
-to identify the reason behind conformational preferences in biomolecules. Through out the article we show its use in different
+systems.  The computation of the free energy difference between pairs of macromolecular
+conformations is particularly difficult and computationally expensive task with existing
+methods. In this work, confinement approach is used to calculate absolute
+free energies of bio molecular systems. The method provides two main advantages: it does not require a reaction coordinate or
+transition path and it is fast to compute. The free energy calculated can be decomposed into per residue contribution in a 
+approximate way. Per residue free energy allow us
+to identify the reason behind conformational preference of the biomolecules. Through out the article we show its use in different
 challenging modeling problems. In particular, we show its use in predicting the conformational
 preference of chamaleon sequences (sequences with high sequence identity and different
-folds). This sequence dependent conformational preferences and per residue free energy decomposition set the stage for the use of this 
+folds). This sequence dependent conformational preference and per residue free energy decomposition set the stage for the use of this 
 method in protein design.
@@
 
-In computational structural biology there are numerous cases where the free energy difference
-between two well defined 
-states plays an important role. Examples range from small to large conformational changes of proteins due to
-ligand binding, change of pH or other conditions \cite{Meirovitch2007}, \cite{Chipot2007}, \cite{Jorgensen2004}. 
-Free energy also plays an important role in the case of protein
+In computational structural biology there are numerous cases where free energy between well defined
+states are necessary. Examples ranges from small to large conformational change of protein due to
+ligand binding, change of pH etc \cite{Meirovitch2007}, \cite{Chipot2007}, \cite{Jorgensen2004}. 
+Free energy also play an important role in case of protein
 folding. The ground breaking work of Christian B. Anfinsen and coworkers showed that the native
 structures of small globular proteins have a unique, thermodynamically stable native structure with
-their conformation at the global free energy minimum \cite{Anfinsen1973}. Regardless of the starting
-point or even after unfolding 
-it by changing different condition most proteins will go back to the native state after
-reinstituting native like conditions. Often, the changes in free energy between different
-conformations is small, but nonetheless crucial in favoring the native state
- since defects in protein folding may be
-the molecular cause of a range of human genetic disorders. For example a misfolded protein known as
-a
-prion can misfold correctly folded proteins when entering a healthy organism.
-\Alberto{I'm not sure I understand this here} \Arijit{Thus, Free energy can also act as a scale to identify between the misfolded and the native state of
-the protein.} These facts further give free energy a special importance in structural biology.
+their conformation at the global free energy minimum \cite{Anfinsen1973}. Regardless of the starting point or unfolding
+it by changing different condition most proteins will finally assume the same structure. It is very
+important for proteins to achieve their native conformation since defects in protein folding may be
+the molecular cause of a range of human genetic disorder. For example a misfolded protein known as
+prion when enters any healthy organism it converts properly folded protein into misfolded one.
+Thus, Free energy can also act as a scale to identify between the misfolded and the native state of
+the protein. These facts further give free energy a special importance in structural biology.
 
-However, the theoretical calculation of conformational free energy change becomes difficult if the
-states of interest are very different \cite{Meirovitch2007}. In such scenario timescales involved in
-such conformational transitions may
-be beyond the sampling ability of classical molecular dynamics simulation.  Generally, in order to
-calculate the free energy a pathway or reaction coordinate is needed with several intermediate structures.
-Methods like umbrella sampling \cite{Torrie1977}, thermodynamic integration \cite{Tironi1994} along with classical MD can suffer from
-overlap problem and require a large computational effort. The idea of a reaction coordinate became
-even more
-complicated in the case of protein folding.  As the free energy landscape for protein folding 
-is rough \cite{Dill1997}, \cite{Dill2008}, the molecule 
-can get trapped in an energy well during conformational sampling. Even with 
-advanced methods like replica exchange molecular dynamics, the three dimensional structure may be
-trapped in a local minima for a considerable time, giving rise to nonconverged distributions where
-misfolded states might be selected as native like
+However, the theoretical calculation of conformational free energy change become difficult if the
+states of interest are very different \cite{Meirovitch2007}. In such scenario timescale involved in such tranformation may
+be beyond timescale of direct classical molecular dynamics simulation.  Generally, in order to
+calculate free energy a pathway or reaction coordinate is created with the intermediate structures.
+Method like umbrella sampling \cite{Torrie1977}, thermodynamic integration \cite{Tironi1994} along with classical MD can suffer from
+overlap problem and require large computational effort. The idea of reaction coordinate became more
+complicated in case of protein folding.  As free energy landscape of the protein folding 
+is rough \cite{Dill1997}, \cite{Dill2008}
+it can be trapped in a energy well during its visit in the conformational space. Even with the
+advanced method like replica exchange molecular dynamics the three dimensional structure may be
+trapped in a local minima for a considerable time and trapped in a local energy minima
+and one can mistakenly identify the misfolded state as the native state.
 % (give the example of human pin1ww domain from Benout Roux and Klaus Schultan).
 
-The calculation of free energy has been successfully attempted by a number of groups \cite{Meirovitch2007}, 
-\cite{Ytreberg2006} - \cite{Zheng2008}. In recent years methods like the Reference system method \cite{Ytreberg2006}, 
-Decativated Morphing \cite{Park2008}, Orthogonal Space Random Work \cite{Zheng2008} or the Confinement 
-Method \cite{Tyka2006}, \cite{Cecchini2009} to name a few have been used to calculate free energy
-differences.
+Calculation of free energy has been successfully attempted by a number of groups \cite{Meirovitch2007}, 
+\cite{Ytreberg2006} - \cite{Zheng2008}, . In recent years methods like Refrence system method \cite{Ytreberg2006}, 
+Decativated Morphing \cite{Park2008}, Orthogonal Space Random Work \cite{Zheng2008}, Confinement 
+Method \cite{Tyka2006}, \cite{Cecchini2009} have came out for calculation of free energy method. 
 In this work we have applied the confinement method which was originally devoled by Tyka et. al. \cite{Tyka2006} and
-Cecchini et. al. \cite{Cecchini2009}. This method relies on the fact that the free energy difference is a state function and 
-thus it is independent of the sampling path. This approach uses a thermodynamic cycle where first, non-harmonic degrees of
-freedom are removed by applying a series of restraints to the system. Then the free energy of the
+Cecchini et. al. \cite{Cecchini2009}. This method relies on the fact that free energy is a state function and 
+thus independent of path. It uses a thermodynamic cycle where first, non-harmonic degrees of
+freedom are removed by applying a series of restraint to the system. Then the free energy of the
 remaining harmonic system is obtained using a normal mode calculation and combined with the
@@
 first reproduced that known result. Then we have applied this method to larger proteins for the
-first time. The free energy difference of pairs of conformations in proteins with similar sequence but completely
-different fold was calculated. \Alberto{i made this part too long, have to summarize here and define
-better inside text?-->} Encouraged by the result, we applied the method to help tackle the
-problem of protein conformation ranking. We did this by selecting different targets from the
-Critical Assessment of Structure Prediction (CASP) event. In this event, different research groups
-try to find the 3D structure of a protein given a sequence in a blind way using their methodologies. Upon comparing
-with the real structure it is often seen that several groups are able to generate better structures
-than the ones they submit, but are not able to identify them. This is known as the ranking problem,
-and this method is able to correctly rank structures in the majority of cases we have tried.
- Finally, we have been exploring the ability of working on a per residue basis to identify which
- residues are the major players in the stabilization/destabilization of a pair of conformations. Our
- initial experience makes us thinkg that this method might be helpful in identifying key residues
- for protein design.
+first time. The free energy of conformation of pair of proteins with similar sequence but completely
+different fold was calculated. Encouraged by the result, we applied the method of different targets
+of CASP9 competition and tested whether it can predict the most stable conformation of the protein based 
+on its free energy. The remarkable ability of confinement method to identify residues which 
+stabilizes/destabilizes a particular conformation from per residue free energy analysis set the stage
+for future application of this method in protein design.
 
@@
 \begin{figure}
-\includegraphics[width=7 in,height=5.8 in]{orban_full.pdf}
+\includegraphics[width=6.5 in,height=5.4 in]{orban_full.pdf}
 \label{fig:orban_full}

### 2013-01-11 Paper/paper.tex
changed paper 11_1

--- 2013-01-09 Paper/paper.tex
@@
 of crystal and the model1, we have decomposed the total free energy into the per residue component. A close analysis of 
-the per residue free energy value, we found that although, crystal and different model have closed GDT\_TS and backbone RMSD 
-value, there are number of sidechains oriented in a very different way. As shown in Figure~\ref{fig:T0538compare}(A),
+the per residue free energy value i(data not shown) reveal that there are number of sidechains oriented in a very different way in crystal and different 
+model. This may account for differences in free energy values, inspite of close RMSD and GST\_TS values. As shown in Figure~\ref{fig:T0538compare}(A),
 there is a salt bridge between Arg-32 and Glu-35 present in the crystallographic structure, which is absent in the
@@
 \bibitem{MacCallum2011}
-MacCallum, J.; Pérez, A.; Schnieders, MJ.; Hua, L.; Jacobson, M.P.; Dill, K.A. Assessment of protein structure refinement 
+MacCallum, J.; Perez, A.; Schnieders, MJ.; Hua, L.; Jacobson, M.P.; Dill, K.A. Assessment of protein structure refinement 
 in CASP9. Proteins, 2011, 79, 74-90.

### 2013-01-11 Paper/paper.tex
changed paper 11_2

@@
 In order to test this hypothesis we performed another test with target T0531 of CASP9, in which we compared the crystal 
-structure to five models from the MUFOLD server. The pdb id of this extracellular domain of the jumping translocation 
-breakpoint protein is 2KJX and the CASP target code is T0531.
+structure to five models from the MUFOLD server. Pdb id of this extracellular domain of the jumping translocation 
+breakpoint protein is 2KJX and CASP target code is T0531.
 Looking at Figure~\ref{fig:T0531} shows: 1.) native is correctly
 identified as expected and 2.) Surprisingly there is a high level of correlation between the GDT
-ranking and the free energy ranking, only two structures with GDT\_TS score less than  35  are ordered
+ranking and the free energy ranking for model 1 and model 3, the rest three structures with GDT\_TS score less than  35  are ordered
 incorrectly. It is worth to note that models 2 and 3 have the same GDT and very different free
-energies, meaning that the actual ordering could change a lot \cite{Perez2012}.
+energies, meaning that the actual ordering could change a lot \cite{Perez2012}. It is encouraging that at least
+the method can pick out the best model with low GDT score of 44.
 

### 2013-01-11 Paper/paper.tex
changed paper 11_3

@@
 values as shown in Figure~\ref{fig:T0559} indicate that the model 1 was predicted correctly, 
-whereas the order of model 3 and 5 was wrong. The main difference
+whereas the order of model 3 and 5 was wrong. Main difference
 between the model 3 and the rest of the model is that the orientation of first alpha helix of model 3 is opposite.
@@
 We found that in majority of the case, the native structure is always the most stable and we can rank the protein structure upto 100 amino 
-acid residues long. This method works very well if atleast one of the strcuture in the decoy is good. The main advantage of this method is 
-that it does not require any reaction coordinate. An interesting application, that we have developed using this method is per residue
+acid residues long. This method works very well if atleast one of the strcuture in the decoy is good. Main advantage of this method is 
+that it does not require any reaction coordinate. Thus one can calculate the free energy difference between two very different conformations. 
+An interesting application, that we have developed using this method is per residue
 free energy calculation. Although, we have calculated that in an approximate way, but it give important information regarding conformational 
@@
 This method can be further used for calculating free energy due to very large conformational change. Another application may be
-for the case of ligand binding. It can also be used to distinguish the most stable structure during protein design. Another, important
-application may be to study the protein folding kinetics, specially to study the phi value analysis.
+for the case of ligand binding. It can also be used to distinguish the most stable structure during protein design. Decompition of 
+total free energy into the per residue component help identifying residues that give stability to a particular conformation. 
+Another, important application may be to study the protein folding kinetics, specially to study the phi value analysis.
 

### 2013-01-11 Paper/paper.tex
before merge

--- 2013-01-11 Paper/paper.tex
@@
 \begin{figure}
-\includegraphics[width=6.5 in,height=5.4 in]{orban_full.pdf}
+\includegraphics[width=7 in,height=5.8 in]{orban_full.pdf}
 \label{fig:orban_full}
@@
 surface in the generated model (Figure~\ref{fig:T0569_per_residue}(B). This the region, where there is a beta sheet in the original crystal structure, which is absent in the 
-model. There are also some region, which stabilizes the generated model compare to that of the crystal. A close inspection reveal that 
+model. Interestinlgy, some regions are more stable in the generated model. A close inspection
+reveals that 
 the hydrophilic Glu in position 67 is completely exposed to the protein surface and solvent in the best model, whereas it is partially exposed in the 
-crystal. This is presented in Figure~\ref{fig:T0569_per_residue}(C). The peak at position 76 is due to the H-bond between Lys-76 and Asp-11, 
-which is absent in original crystalographic structure (Figure~\ref{fig:T0569_per_residue}(D)).
+crystal (Figure~\ref{fig:T0569_per_residue}(C)). The peak at position 76 is due to the H-bond between Lys-76 and Asp-11, 
+which is absent in the original crystalographic structure (Figure~\ref{fig:T0569_per_residue}(D)).
 
-\subsubsection{Confinement method fails ?}
+\subsubsection{Failures in the confinement method}
 
-There are however some instances where the method can fail, specially when GDT scores and all the structures are
+Despite the great success in most of the studied systems, we found a few failures, specially when
+the GDT scores of the compared structures are
 very close. We have found this for an engineered protein from Asr4154 protein (PDB ID: 2L09 and CASP Target T0538).
-This protein contains 54 amino acid residues with three alpha helixes and two very short beta strand. 
+This protein contains 54 amino acid residues with three alpha helixes and two very short beta
+strands. 
 The model that is closest to the native structure was generated by the group PconsR with a GDT\_TS value of 96.23.
 We have also considered one model from group "Shell" (GDT\_TS = 90.09) and "FOLDIT" (GDT\_TS = 86.32) for our analysis along with
-the native crystal struture. The models from PconsR, Shell and FOLDIT are labelled as model 1, model 2 and model 3 
+the native crystal struture (RMSD values between 1.6 and 2.0\AA \Alberto{from native, right?all
+atom? CA?}). The models from PconsR, Shell and FOLDIT are labelled as model 1, model 2 and model 3 
 respectively in Figure~\ref{fig:T0538}. As seen in Figure~\ref{fig:T0538}, one can find that the 
 model from PconsR has lowest free energy value, making it a more stable structure. Although ranking of the other models
-remain consistent, it may happen that the theoretical model produced by the group "PconsR" much more stable than the 
-crystal structure. This is particularly difficult target for rank ordering as all the models those are considered for this 
-calculation are within RMSD values of 1.6 - 2.0 \AA. In order to understand the reason behind the mismatch of GDT\_TS values
-of crystal and the model1, we have decomposed the total free energy into the per residue component. A close analysis of 
-the per residue free energy value i(data not shown) reveal that there are number of sidechains oriented in a very different way in crystal and different 
-model. This may account for differences in free energy values, inspite of close RMSD and GST\_TS values. As shown in Figure~\ref{fig:T0538compare}(A),
-there is a salt bridge between Arg-32 and Glu-35 present in the crystallographic structure, which is absent in the
-model 1. In model 1, Arg-32 has another salt bridge with Glu-28. Glu-28 in crystal has the salt bridge with Lys-24. In model 1, Lys-24 is
-exposed towards the protein surface. This is shown in Figure~\ref{fig:T0538compare}(A). There are more such interesting and may be crucial
-sidechain differences between the model and the crystallographic structure, which indicate the limitation of the GDT\_TS based 
-ranking of the protein structure. As, shown in Figure~\ref{fig:T0538compare}(B), Arg-26 and Glu-50 has a salt bridge in the crystallographic 
-structure. Also, Phe-51 is oriented towards the protein core in the crystallographic structure. In model 1, the salt bridge between Arg-26 and Glu-50
-is absent and the Phe-51 is oriented towards the protein surface.
+remains consistent, model 1 is predicted to be more stable than the
+crystal structure. To rationalize this, we have decomposed the total free energy into the per
+residue components. The results show us that despite the small variation at the backbone level (as
+shown by high GDT\_TS scores and low RMSDs), the sidechains are oriented in very different ways,
+giving raise to large differences in the stabilization of certain residues.
+In particular, some of the differences arise from different salt bridge patterns (Arg-32 with Glu-35
+and Glu-28 with Lys-24
+in crystal versus Arg-32 with Glu-28 in model 1) and certain flexible polar residues exposed to the
+surface (Lys-24 in model 1, which has an entropic gain from not forming a salt bridge and is
+stabilized by interactions with the solvent). This is summarized in Figure~\ref{fig:T0538compare}(A) 
+Similar patterns arise in other places of the protein (see Figure ~\ref{fig:T0538compare}(B)), where
+the salt bridge between Arg-26 and Glu-50 is absent in model 1 and is further \Alberto{stabilized?} by Phe-51
+having a very different conformation that in the crystal structure.
+All in all, this unexpected result just shows us that this method is sensitive to local interactions
+such as those happening from side chain reorientation. At the same time, the results explained in
+past sections showed that for larger changes the accuracy of the method is very high.
 
@@
 
+So far we have seen that the method is good at predicting preferences when the structures are not
+very close to each other, how far from native can we go? In this section we want to explore what
+happens when we use the method to describe low quality models. We use one of the CASP targets (pdb
+id = 2KJX, CASP id = T0531) .
 We mentioned earlier that our predictive abilities are greatly decreased with low model qualities. It is
@@
 \bibitem{MacCallum2011}
-MacCallum, J.; Perez, A.; Schnieders, MJ.; Hua, L.; Jacobson, M.P.; Dill, K.A. Assessment of protein structure refinement 
+MacCallum, J.; Pérez, A.; Schnieders, MJ.; Hua, L.; Jacobson, M.P.; Dill, K.A. Assessment of protein structure refinement 
 in CASP9. Proteins, 2011, 79, 74-90.

### 2013-01-11 Paper/paper.tex
changed paper 11_4

--- 2013-01-11 Paper/paper.tex
@@
 
-The calculation of free energy differences is of central importance in the simulation of biochemical
+Calculation of free energy differences is of central importance in the simulation of biochemical
 systems. It is particularly difficult to calculate between pairs of macromolecular
 conformations as well as a computationally expensive task with existing
-methods. In this work, the confinement approach is used to calculate absolute
-free energies of biomolecular systems. The method provides two main advantages: it does not require a reaction coordinate or
-transition path and it is fast to compute. The free energy calculated can be decomposed into a per
+methods. In this work, confinement approach is used to calculate absolute
+free energies of biomolecular systems. This method provides two main advantages: it does not require a reaction coordinate or
+transition path and it is fast to compute. Free energy calculated can be decomposed into a per
 residue contribution in an 
@@
 Free energy also plays an important role in the case of protein
-folding. The ground breaking work of Christian B. Anfinsen and coworkers showed that the native
+folding. Ground breaking work of Christian B. Anfinsen and coworkers showed that the native
 structures of small globular proteins have a unique, thermodynamically stable native structure with
@@
 Methods like umbrella sampling \cite{Torrie1977}, thermodynamic integration \cite{Tironi1994} along with classical MD can suffer from
-overlap problem and require a large computational effort. The idea of a reaction coordinate became
+overlap problem and requires a larger computational effort. The idea of a reaction coordinate became
 even more
-complicated in the case of protein folding.  As the free energy landscape for protein folding 
+complicated in the case of protein folding.  As free energy landscape for protein folding 
 is rough \cite{Dill1997}, \cite{Dill2008}, the molecule 
@@
 
-The calculation of free energy has been successfully attempted by a number of groups \cite{Meirovitch2007}, 
+Calculation of free energy have been successfully attempted by a number of groups \cite{Meirovitch2007}, 
 \cite{Ytreberg2006} - \cite{Zheng2008}. In recent years methods like the Reference system method \cite{Ytreberg2006}, 
@@
 calculate the conformational free energy of a 16 amino acid residue peptide, known as BHP.  We have
-first reproduced that known result. Then we have applied this method to larger proteins for the
+first reproduced that known result. Next we have applied this method to larger proteins for the
 first time. The free energy difference of pairs of conformations in proteins with similar sequence but completely
@@
 
-\subsection{The confinement method produces correct results in control experiments}
+\subsection{Confinement method produces correct results in control experiments}
 
@@
 \begin{center}
-\caption{The confinement method assigns a more favorable free energy to the experimentally
+\caption{Confinement method assigns a more favorable free energy to the experimentally
 determined structure than to computer-generated predictions. For each target, we examined as many as
@@
 
-\subsection{The confinement method correctly predicts the structural preferences of chameleon
+\subsection{Confinement method correctly predicts the structural preferences of chameleon
 sequences}
@@
 
-The results of these calculations are presented in Figure~\ref{fig:orban}. The confinement method
+The results of these calculations are presented in Figure~\ref{fig:orban}. Confinement method
 identifies the correct structure for all five sequences. One of the hypothesis for fold switching
@@
 It is widely believed that if the free energy of the native state and the alternative state is within
- a range of $-5 Kcal/Mol$ then  it can quickly change fold when the stability of the native state 
+ a range of $5 Kcal/Mol$ then it can quickly change fold when the stability of the native state 
 decreases. The stability of the native state can decrease for a number reason ranging from chemical 
@@
 came out from our calculation ranges from
-around 2.9 to 5.7 kcal/mol, which is consistent with the above hypothesis \cite{He2008}, \cite{{Alexander2009}}.
+around 3.5 to 5.0 kcal/mol, which is consistent with the above hypothesis \cite{He2008}, \cite{{Alexander2009}}.
 
@@
 \hline
-    Observation                                & $3 \alpha$ structure           &    $4 \beta + \alpha$ structure                       \\ \hline
-    Leu-5 and Leu-7 favor                      & hydrophobic residues are       & oriented towards the                                  \\ 
-    $4 \beta + \alpha$                         & exposed to the solvent         & hydrophobic core                                      \\ \hline
-    Gln-11 favors $3 \alpha$                   & Gln-11 has H-bond with Glu-15  & no such H-bond                                        \\ \hline
+    Information from per                                 & What is happening in           & What is happening in                      \\ 
+    residue plot                                         & $3 \alpha$ conformation        & $4 \beta + \alpha$  conformation          \\ \hline
+    Leu-7 favors                                         & hydrophobic residue is         & oriented towards the                      \\ 
+    $4 \beta + \alpha$ (Figure~\ref{fig:orban_full}(A))  & exposed to the solvent         & hydrophobic core                          \\ \hline
+    Gln-11 stabilizes $3 \alpha$                   & Gln-11 has H-bond                         & no such H-bond                           \\ 
+    (Figure~\ref{fig:orban_full}(B))           &  with Glu-15                              &                                           \\ \hline 
     Glu-14 favors $4 \beta + \alpha$           & hydrophilic residue partially  & completely exposed                                    \\       
-                                               & exposed to the solvent         & to the solvent                                        \\ \hline        
-    Ala-26 favors  $4 \beta + \alpha$          & hydrophobic residue oriented   & oriented towards the                                   \\ 
-                                               & towards surface                & hydrophobic core                                       \\ \hline
-    Ile-49 favors $3 \alpha$                   & hydrophobic interaction        & exposed to the                                          \\
-                                               & inside the protein             & solvent                                                \\ \hline 
-    Val-54 favors $4 \beta + \alpha$           & hydrophobic residue exposed    & hydrophobic interaction                                \\
-                                               & to the solvent                 & inside the protein                                      \\ \hline
+    (Figure~\ref{fig:orban_full}(E))           & exposed to the solvent         & to the solvent                                        \\ \hline        
+    Ala-26 favors  $4 \beta + \alpha$          & hydrophobic residue            & oriented towards the                                   \\ 
+     (Figure~\ref{fig:orban_full}(C))          & oriented towards surface       & hydrophobic core                                       \\ \hline
+    Ile-49 favors $3 \alpha$                   & hydrophobic interaction        & sidechain exposed                                       \\
+    (Figure~\ref{fig:orban_full}(H) and (I))   & inside the protein             & to the solvent                                          \\ \hline 
+    Ile-49 favors GA more                      & more hydrophobic               &  unchanged                                              \\
+    than GB                                    & interactions with residues     &                                                         \\ 
+    (Figure~\ref{fig:orban_full}(H) and (I))  &  at 20, 30 in GA                &                                                         \\ \hline
+    Val-54 favors $4 \beta + \alpha$           & hydrophobic residue            & hydrophobic interaction                                \\
+    Figure~\ref{fig:orban_full}(D))             & exposed to the solvent        & inside the protein                                      \\ \hline
     Glu-56 favors $3 \alpha$                   & hydrophilic residue partially  & completely exposed                                    \\
-                                               & exposed to the solvent         & to the solvent                                        \\ \hline
+    (Figure~\ref{fig:orban_full}(G))            & exposed to the solvent         & to the solvent                                        \\ \hline
     Phe-30 in GB sequence                      & exposed to the solvent         & oriented towards the                                  \\
-    favors $4 \beta + \alpha$                  & in protein surface             & hydrophobic core                                      \\ \hline   
-    Tyr-45 in GB Sequence                      & No H-bonding                   & H-bond with Asp-47                                    \\  
-    favors $4 \beta + \alpha$                  &                                &                                                       \\ \hline
+    stabilizes $4 \beta + \alpha$                  & in protein surface             & hydrophobic core                                      \\ \hline   
+    Tyr-45 in GB favors                      & No H-bonding                   & H-bond with Asp-47                                    \\  
+    $4 \beta + \alpha$ (Figure~\ref{fig:orban_full}(K))&                 &                                                       \\ \hline
     Asp-47 in GA sequence                      & H-bond with Lys-50             & No H-bond, No Tyr-45                                  \\  
@@
 \label{fig:orban}
-\caption{The confinement method correctly predicts the structural preferences of six chameleon
+\caption{Confinement method correctly predicts the structural preferences of six chameleon
 sequences. (A) The six sequences used in this study. (B) Each sequence adopts either a Protein
@@
 \begin{figure}
-\includegraphics[width=6.5 in,height=5.4 in]{orban_full.pdf}
+\includegraphics[width=6.4 in,height=5.3 in]{orban_full.pdf}
 \label{fig:orban_full}
-\caption{Orientation of residues of different sidechains in $4 \beta + \alpha$ and $3 \alpha$ form
-A. Orientation of Hydrophobic residues Leu-5 and Leu-7.
+\caption{Sidechain orientation of some important residues which stabilize/destabilize either of the $4 \beta + \alpha$ or $3 \alpha$ conformation.
+The details are in Table~\ref{table:orban_perresidue} and the main text.  
+A. Leu-7.
 B. Gln-11 has H-bond with Glu-15 in $3 \alpha$, which is absent in $4 \beta + \alpha$
-C. Hydrophobic Ala-26
-D. Hydrophobic Val-54.
-E. Hydrophilic Glu-14.
-F. The plot of free energy difference for GA95 and GB95 sequence. Positive peak stabilize $3 \alpha$ form and negative peak 
+C. Ala-26
+D. Val-54.
+E. Glu-14.
+F. plot of free energy difference for GA95 and GB95 sequence. Positive peak stabilize $3 \alpha$ form and negative peak 
 stabilize $4 \beta + \alpha$. 
-G. Hydrophilic Glu-56 .
-H. Orientation of Hydrophic Ile-49.
-I. With GA95 sequence, Ile-49 have hydrophic interaction with both Leu-20 and Ile-30.
-J. Larger Phe-30 in GB sequence, move towards the protein surface in $4 \beta + \alpha$.
-K. Tyr-45 form H-bond with Asp-47 in $4 \beta + \alpha$.}  
+G. Glu-56 .
+H. sidechain orientation of Hydrophic Ile-49, along with Leu-20 and Ile-30 with GA95 sequence
+I. sidechain orientation of Hydrophic Ile-49, along with Ala-20 and Phe-30 with GB95 sequence
+J. Tyr-45 forms H-bond with Asp-47 with GB95 sequence and $4 \beta + \alpha$ conformation
+K. No H-bond possible between Leu-45 and Asp-47 with GB95 sequence and $4 \beta + \alpha$ conformation .}  
 \end{figure}
@@
 %\begin{table}
-%\caption{The confinement method correctly reproduces the structural preferences of a series of
+%\caption{Confinement method correctly reproduces the structural preferences of a series of
 %designed peptides. Each sequence adopts either a Protein G-like fold ($4 \beta + \alpha$) or forms a

### 2013-01-11 Paper/paper.tex
changed paper 11_5

@@
 
-Calculation of free energy have been successfully attempted by a number of groups \cite{Meirovitch2007}, 
+The calculation of free energy have been successfully attempted by a number of groups \cite{Meirovitch2007}, 
 \cite{Ytreberg2006} - \cite{Zheng2008}. In recent years methods like the Reference system method \cite{Ytreberg2006}, 
@@
 
-\subsection{Confinement method produces correct results in control experiments}
+\subsection{The confinement method produces correct results in control experiments}
 
@@
 \begin{center}
-\caption{Confinement method assigns a more favorable free energy to the experimentally
+\caption{The confinement method assigns a more favorable free energy to the experimentally
 determined structure than to computer-generated predictions. For each target, we examined as many as
@@
 
-\subsection{Confinement method correctly predicts the structural preferences of chameleon
+\subsection{The confinement method correctly predicts the structural preferences of chameleon
 sequences}
@@
 
-To avoid this potential problem, we instead computed the relative free energy of two different
+To avoid this potential problem, we instead computed relative free energy of two different
 structural models for each sequence. One model is based on the GA structure and the other on the GB
@@
 
-\subsection{Per residue free energy calculation can identify the mechanistic detail behind conformational pereference 
+\subsection{The per residue free energy calculation can identify the mechanistic detail behind conformational pereference 
 of a particular residue}
 
-We decomposed the calculated free energy into per residue decomposition as described in the methods
-section. This helps us to
-identify important residues that stabilize a particular conformation. 
+We decomposed the calculated free energy into per residue decomposition as described in the method
+section. This helps us to identify important residues that stabilize a particular conformation. 
+In this direction, we first calculate the per residue free energy for the GA95 and GB95 sequences. 
+The per residue free energy, $\Delta \Delta G (4 \beta + \alpha$ - $3 \alpha)$ is plotted in Figure~\ref{fig:orban_full}(F).
+In this plot, the negative peak indicate the residue will stabilize the $4 \beta + \alpha$ conformation of a particular
+residue and the positive peak will stabilize the $3 \alpha$ structure. As, the differences of amino acid
+residues are only at 3 position, there are some common features of residues in both GA95 and GB95 sequences in both
+$3 \alpha$ and $4 \beta + \alpha$ conformation. Role of some important residues in stabilizing either 
+$4 \beta + \alpha$ or $3 \alpha$ conformation and possible reason for such conformational preferences are discussed in 
+details in Figure~\ref{fig:orban_full} and Table~\ref{table:orban_perresidue}. It is important to note that
+experimental observations classified the protein into two parts: Amino acids 9-51 are fully structured in both
+the folds, where as residues 1-8 and 52-56 are unstructured in $3 \alpha$, but form $\beta$ strand in $4 \beta + \alpha$ structure.
+Most of the amino acid residues in the region 1-9 favor $4 \beta + \alpha$ structure as can be seen from the plot
+Figure~\ref{fig:orban_full}(F). Interestingly, we can explain most of the big peak in this plot. They are presented 
+in Figure~\ref{fig:orban_full} and Table~\ref{table:orban_perresidue}. Here, we explain only one case. The first big peak
+appears at position 7. Here, the hydrophobic Leu-7 stabilizes the $4 \beta + \alpha$ conformation as its sidechain is oriented 
+towards the protein hydrophobic core, whereas in $3 \alpha$ structure it exposed to the solvent (Figure~\ref{fig:orban_full}(A)).
+Thus, if someone want to better design the $3 \alpha$ conformation of this protein, we propose to mutate Leu at this position
+with some hydrophilic residue. Similarly, we found that the free energy differences arise in two conformation for various 
+reasons ranging from hydrogen/salt bridge formation to relative solvent exposure of hydrophobic residues etc. They are 
+listed in Figure~\ref{fig:orban_full} and Table~\ref{table:orban_perresidue}.
+
 \Alberto{This is not Reader Friendly at all. we do not have to describe every little detail. I would
@@
 preference, ditching everything below until the next section. What do you think?}
-In this direction, we first calculate the per residue free energy for the GA95 and GB95 sequences. Similar to the total 
-free energy for each sequence, we calculate the free energy difference between $4 \beta + \alpha$ and $3 \alpha$ structure for each 
-residue. The free energy, $\Delta \Delta G (4 \beta + \alpha$ - $3 \alpha)$ is plotted in Figure~\ref{fig:orban_full}(F). 
-In this plot, the negative peak indicate the residue will stabilize the $4 \beta + \alpha$ conformation of a particular
-residue and the positive peak will stabilize the $3 \alpha$ structure. As, the differences of amino acid 
-residues are only at 3 position, there are some common features of residues in both GA95 and GB95 sequences in both 
-$3 \alpha$ and $4 \beta + \alpha$ conformation. They are reported in Figure~\ref{fig:orban_full} 
-and Table~\ref{table:orban_perresidue}. Before, we further go into that detail, it is important to note that 
-experimental observations classified the protein into two parts: Amino acids 9-51 are fully structured in both
-the folds, where as residues 1-8 and 52-56 are unstructured in $3 \alpha$, but form $\beta$ strand in $4 \beta + \alpha$ structure.
-Most of the amino acid residues in the region 1-9 favor $4 \beta + \alpha$ structure as can be seen from the plot 
-Figure~\ref{fig:orban_full}(F). Here, we try to explain some of the largest peak in that plot. The residues Leu-5 and 
-Leu-7 favor $4 \beta + \alpha$ structure as these hydrophobic residues are oriented towards the protein hydrophobic 
-core, whereas in $3 \alpha$ structure they are exposed to the solvent (Figure~\ref{fig:orban_full}(A)). One can argue that 
-hydrophilic residues in these positions can stabilize the $3 \alpha$ conformation.
-The next big peak is located at position 11. A close observation of the sidechain orientation of residue Gln at position 11 reveal that it
-is exposed to the protein surface in both the conformation, but in $3 \alpha$ structure it has an additional H-bond with Glu-15 which gives
-it more stability compared to the $4 \beta + \alpha$ structure (Figure~\ref{fig:orban_full}(B)). The peak 
-at position 14 is due to the hydrophilic Glu, which is completely
-exposed to the solvent in $4 \beta + \alpha$ conformation as shown in
-(Figure~\ref{fig:orban_full}(E)), but is only partially exposed 
-to the solvent in $3 \alpha$ form. 
-This is the reason why Glu at position 14 stabilize $4 \beta + \alpha$ conformation. Another high peak is at position 26, which indicate
-that the residue in that position stabilizes $4 \beta + \alpha$ conformation. Again, the hydrophobic amino acid residue, Ala at position 26
-is oriented towards the hydrophobic core of the protein in $4 \beta + \alpha$ conformation , whereas it is exposed towards the protein surface 
-in $3 \alpha$ structure (Figure~\ref{fig:orban_full}(C)). Similar situations arise for residues at position 54 and 56. 
-These residues are in the region, where $4 \beta + \alpha$ form is fully structured and $3 \alpha$ form is unstructured. 
-As shown in Figure~\ref{fig:orban_full}(D), the sidechain of hydrophobic Val-54 is oriented towards the hydrophobic core in 
-$4 \beta + \alpha$ conformation and thus have hydrophobic interactions with other residues in this conformation.
-On the other hand, in $3 \alpha$ it is exposed to the solvent. The reverse happens for the residue at position 56.
-Here the hydrophilic residue Glu is partially or not at all exposed to the protein surface in $4 \beta + \alpha$ but fully exposed 
-to the solvent in $3 \alpha$ conformation (Figure~\ref{fig:orban_full}(G)). This is the reason, why one can see a positive peak for position 56. 
-Similarly, hydrophobic Ile-49 in $4 \beta + \alpha$ conformation is exposed to protein surface and solvent, whereas as shown in 
-Figure~\ref{fig:orban_full}(H)$3 \alpha$ conformation 
-it has hydrophobic interaction with other hydrophobic residues in protein core.   
+
+\Arijit{I have rewrite this part.There was a table. I made it better. I want to keep some discussion though if you agree.}
+
 Now, let us discuss the differences of $3 \alpha$ and $4 \beta + \alpha$ fold due to specific difference of mutated amino acid 
@@
 sequence. In $3 \alpha$ fold, Ile-49 has hydrophobic interactions with Leu-20 and Ile-30 with GA sequence, whereas with GB sequence it 
-has such interactions with smaller Ala-20 and Phe-30 residues (Figure~\ref{fig:orban_full}(I). In $3 \alpha$, with GB sequence, Phe-30 
-being a larger group can not be completely accomodated in the hydrophobic core(Figure~\ref{fig:orban_full}(J)) and therefore, 
+has such interactions with smaller Ala-20 (Figure~\ref{fig:orban_full}(I). In $3 \alpha$, with GB sequence, Phe-30 
+being a larger group can not be completely accomodated in the hydrophobic core(Figure~\ref{fig:orban_full}(I)) and therefore, 
 partially exposed to the protein surface. There is a very interesting difference for residue at position 45 as well. In GB95 sequence
@@
 $3 \alpha$ conformation. On the other hand with GA conformation, this H-bond is absent for both the conformation as there is a Leu at position
-45. Although, one can not find much difference between $4 \beta + \alpha$ and $3 \alpha$ conformation
+45 (Figure~\ref{fig:orban_full}(J)). Although, one can not find much difference between $4 \beta + \alpha$ and $3 \alpha$ conformation
 for position 20 for both GA95 and GB95 sequence. There are significant difference for the residues at position 30 and 45.  

### 2013-01-11 Paper/paper.tex
changed paper 11_6

@@
 \newcommand{\Justin}[1]{\color{blue}#1\normalcolor}
-\newcommand{\Arijit}[1]{\color{yellow}#1\normalcolor}
+\newcommand{\Arijit}[1]{\color{magenta}#1\normalcolor}
 \newcommand{\Ken}[1]{\color{red}#1\normalcolor}
@@
 conformations is small, but nonetheless crucial in favoring the native state
- since defects in protein folding may be
-the molecular cause of a range of human genetic disorders. For example a misfolded protein known as
-a
-prion can misfold correctly folded proteins when entering a healthy organism.
-\Alberto{I'm not sure I understand this here} \Arijit{Thus, Free energy can also act as a scale to identify between the misfolded and the native state of
-the protein.} These facts further give free energy a special importance in structural biology.
+ since defects in protein folding may be the molecular cause of a range of human genetic disorders.
+ For example a misfolded protein known as a
+prion can misfold correctly folded proteins when entering a healthy organism. Thus it is important for 
+protein to have its native structure. As protein native structure stays in the global free energy minimum, 
+\Arijit{free energy can be used as a scale for accurate identification of misfolded state from the protein native state.}
+\Alberto{I'm not sure I understand this here} 
+These facts further give free energy a special importance in structural biology.
 
@@
 larger proteins. On that direction, we first test the confinement method for a chamelon sequence and 
-found that it can correctly pick out the prefential structure of the chamelon sequence. Other 
+found that it can correctly pick out the preferential structure of the chamelon sequence. Other 
 applications we explore in the present work is the re-scoring, or
@@
 In this direction, we first calculate the per residue free energy for the GA95 and GB95 sequences. 
-The per residue free energy, $\Delta \Delta G (4 \beta + \alpha$ - $3 \alpha)$ is plotted in Figure~\ref{fig:orban_full}(F).
+The per residue free energy, $\Delta \Delta G ((4 \beta + \alpha$) - ($3 \alpha))$ is plotted in Figure~\ref{fig:orban_full}(F).
 In this plot, the negative peak indicate the residue will stabilize the $4 \beta + \alpha$ conformation of a particular
@@
 GDT\_TS represents the average percentage of residues that are in close proximity in two
-structures optimally superimposed using four different distance cutoffs (1, 2, 4 and 8 Å).
+structures optimally superimposed using four different distance cutoffs (1, 2, 4 and 8 \AA).
 
@@
 techniques perform
-much better compared to those methos which are based on analysis of individual models \cite{Kryshtafovych2011}. 
+much better compared to those methods which are based on analysis of individual models \cite{Kryshtafovych2011}. 
 In this background we try to answer, whether the confinement method can do better quality assessment compare to the best performing
@@
 The method can be also give wrong result if all the models in the decoy are very poor. It may also be difficult to rank structures if 
-all the structures are very very close to the native structure. But that does not matter at all as our main aim is to differentiate
-between the good and the bad structure. In this article we have extensively tested this method using examples from CASP experiment.
+all the structures are very very close to the native structure. But we want to ignore this fact as our main aim is to differentiate
+between the good and the bad structure. In this article we have extensively tested this method using examples from CASP 
 This method can be computationally expensive in actual CASP experiment if one want rank hundreds of submitted models in CASP 

### 2013-01-11 Paper/paper.tex
changed paper 11_6
Reviewed grammar and changed some ideas

--- 2013-01-11 Paper/paper.tex
@@
 
-Calculation of free energy differences is of central importance in the simulation of biochemical
+The calculation of free energy differences is of central importance in the simulation of biochemical
 systems. It is particularly difficult to calculate between pairs of macromolecular
 conformations as well as a computationally expensive task with existing
-methods. In this work, confinement approach is used to calculate absolute
-free energies of biomolecular systems. This method provides two main advantages: it does not require a reaction coordinate or
-transition path and it is fast to compute. Free energy calculated can be decomposed into a per
+methods. In this work, the confinement approach is used to calculate absolute
+free energies of biomolecular systems. The method provides two main advantages: it does not require a reaction coordinate or
+transition path and it is fast to compute. The free energy calculated can be decomposed into a per
 residue contribution in an 
@@
 Free energy also plays an important role in the case of protein
-folding. Ground breaking work of Christian B. Anfinsen and coworkers showed that the native
+folding. The ground breaking work of Christian B. Anfinsen and coworkers showed that the native
 structures of small globular proteins have a unique, thermodynamically stable native structure with
@@
 Methods like umbrella sampling \cite{Torrie1977}, thermodynamic integration \cite{Tironi1994} along with classical MD can suffer from
-overlap problem and requires a larger computational effort. The idea of a reaction coordinate became
+overlap problem and require a large computational effort. The idea of a reaction coordinate became
 even more
-complicated in the case of protein folding.  As free energy landscape for protein folding 
+complicated in the case of protein folding.  As the free energy landscape for protein folding 
 is rough \cite{Dill1997}, \cite{Dill2008}, the molecule 
@@
 
-Calculation of free energy have been successfully attempted by a number of groups \cite{Meirovitch2007}, 
+The calculation of free energy has been successfully attempted by a number of groups \cite{Meirovitch2007}, 
 \cite{Ytreberg2006} - \cite{Zheng2008}. In recent years methods like the Reference system method \cite{Ytreberg2006}, 
@@
 calculate the conformational free energy of a 16 amino acid residue peptide, known as BHP.  We have
-first reproduced that known result. Next we have applied this method to larger proteins for the
+first reproduced that known result. Then we have applied this method to larger proteins for the
 first time. The free energy difference of pairs of conformations in proteins with similar sequence but completely
@@
 
-The results of these calculations are presented in Figure~\ref{fig:orban}. Confinement method
+The results of these calculations are presented in Figure~\ref{fig:orban}. The confinement method
 identifies the correct structure for all five sequences. One of the hypothesis for fold switching
@@
 It is widely believed that if the free energy of the native state and the alternative state is within
- a range of $5 Kcal/Mol$ then it can quickly change fold when the stability of the native state 
+ a range of $-5 Kcal/Mol$ then  it can quickly change fold when the stability of the native state 
 decreases. The stability of the native state can decrease for a number reason ranging from chemical 
@@
 came out from our calculation ranges from
-around 3.5 to 5.0 kcal/mol, which is consistent with the above hypothesis \cite{He2008}, \cite{{Alexander2009}}.
+around 2.9 to 5.7 kcal/mol, which is consistent with the above hypothesis \cite{He2008}, \cite{{Alexander2009}}.
 
@@
 \hline
-    Information from per                                 & What is happening in           & What is happening in                      \\ 
-    residue plot                                         & $3 \alpha$ conformation        & $4 \beta + \alpha$  conformation          \\ \hline
-    Leu-7 favors                                         & hydrophobic residue is         & oriented towards the                      \\ 
-    $4 \beta + \alpha$ (Figure~\ref{fig:orban_full}(A))  & exposed to the solvent         & hydrophobic core                          \\ \hline
-    Gln-11 stabilizes $3 \alpha$                   & Gln-11 has H-bond                         & no such H-bond                           \\ 
-    (Figure~\ref{fig:orban_full}(B))           &  with Glu-15                              &                                           \\ \hline 
+    Observation                                & $3 \alpha$ structure           &    $4 \beta + \alpha$ structure                       \\ \hline
+    Leu-5 and Leu-7 favor                      & hydrophobic residues are       & oriented towards the                                  \\ 
+    $4 \beta + \alpha$                         & exposed to the solvent         & hydrophobic core                                      \\ \hline
+    Gln-11 favors $3 \alpha$                   & Gln-11 has H-bond with Glu-15  & no such H-bond                                        \\ \hline
     Glu-14 favors $4 \beta + \alpha$           & hydrophilic residue partially  & completely exposed                                    \\       
-    (Figure~\ref{fig:orban_full}(E))           & exposed to the solvent         & to the solvent                                        \\ \hline        
-    Ala-26 favors  $4 \beta + \alpha$          & hydrophobic residue            & oriented towards the                                   \\ 
-     (Figure~\ref{fig:orban_full}(C))          & oriented towards surface       & hydrophobic core                                       \\ \hline
-    Ile-49 favors $3 \alpha$                   & hydrophobic interaction        & sidechain exposed                                       \\
-    (Figure~\ref{fig:orban_full}(H) and (I))   & inside the protein             & to the solvent                                          \\ \hline 
-    Ile-49 favors GA more                      & more hydrophobic               &  unchanged                                              \\
-    than GB                                    & interactions with residues     &                                                         \\ 
-    (Figure~\ref{fig:orban_full}(H) and (I))  &  at 20, 30 in GA                &                                                         \\ \hline
-    Val-54 favors $4 \beta + \alpha$           & hydrophobic residue            & hydrophobic interaction                                \\
-    Figure~\ref{fig:orban_full}(D))             & exposed to the solvent        & inside the protein                                      \\ \hline
+                                               & exposed to the solvent         & to the solvent                                        \\ \hline        
+    Ala-26 favors  $4 \beta + \alpha$          & hydrophobic residue oriented   & oriented towards the                                   \\ 
+                                               & towards surface                & hydrophobic core                                       \\ \hline
+    Ile-49 favors $3 \alpha$                   & hydrophobic interaction        & exposed to the                                          \\
+                                               & inside the protein             & solvent                                                \\ \hline 
+    Val-54 favors $4 \beta + \alpha$           & hydrophobic residue exposed    & hydrophobic interaction                                \\
+                                               & to the solvent                 & inside the protein                                      \\ \hline
     Glu-56 favors $3 \alpha$                   & hydrophilic residue partially  & completely exposed                                    \\
-    (Figure~\ref{fig:orban_full}(G))            & exposed to the solvent         & to the solvent                                        \\ \hline
+                                               & exposed to the solvent         & to the solvent                                        \\ \hline
     Phe-30 in GB sequence                      & exposed to the solvent         & oriented towards the                                  \\
-    stabilizes $4 \beta + \alpha$                  & in protein surface             & hydrophobic core                                      \\ \hline   
-    Tyr-45 in GB favors                      & No H-bonding                   & H-bond with Asp-47                                    \\  
-    $4 \beta + \alpha$ (Figure~\ref{fig:orban_full}(K))&                 &                                                       \\ \hline
+    favors $4 \beta + \alpha$                  & in protein surface             & hydrophobic core                                      \\ \hline   
+    Tyr-45 in GB Sequence                      & No H-bonding                   & H-bond with Asp-47                                    \\  
+    favors $4 \beta + \alpha$                  &                                &                                                       \\ \hline
     Asp-47 in GA sequence                      & H-bond with Lys-50             & No H-bond, No Tyr-45                                  \\  
@@
 \label{fig:orban}
-\caption{Confinement method correctly predicts the structural preferences of six chameleon
+\caption{The confinement method correctly predicts the structural preferences of six chameleon
 sequences. (A) The six sequences used in this study. (B) Each sequence adopts either a Protein
@@
 \begin{figure}
-\includegraphics[width=6.4 in,height=5.3 in]{orban_full.pdf}
+\includegraphics[width=6.5 in,height=5.4 in]{orban_full.pdf}
 \label{fig:orban_full}
-\caption{Sidechain orientation of some important residues which stabilize/destabilize either of the $4 \beta + \alpha$ or $3 \alpha$ conformation.
-The details are in Table~\ref{table:orban_perresidue} and the main text.  
-A. Leu-7.
+\caption{Orientation of residues of different sidechains in $4 \beta + \alpha$ and $3 \alpha$ form
+A. Orientation of Hydrophobic residues Leu-5 and Leu-7.
 B. Gln-11 has H-bond with Glu-15 in $3 \alpha$, which is absent in $4 \beta + \alpha$
-C. Ala-26
-D. Val-54.
-E. Glu-14.
-F. plot of free energy difference for GA95 and GB95 sequence. Positive peak stabilize $3 \alpha$ form and negative peak 
+C. Hydrophobic Ala-26
+D. Hydrophobic Val-54.
+E. Hydrophilic Glu-14.
+F. The plot of free energy difference for GA95 and GB95 sequence. Positive peak stabilize $3 \alpha$ form and negative peak 
 stabilize $4 \beta + \alpha$. 
-G. Glu-56 .
-H. sidechain orientation of Hydrophic Ile-49, along with Leu-20 and Ile-30 with GA95 sequence
-I. sidechain orientation of Hydrophic Ile-49, along with Ala-20 and Phe-30 with GB95 sequence
-J. Tyr-45 forms H-bond with Asp-47 with GB95 sequence and $4 \beta + \alpha$ conformation
-K. No H-bond possible between Leu-45 and Asp-47 with GB95 sequence and $4 \beta + \alpha$ conformation .}  
+G. Hydrophilic Glu-56 .
+H. Orientation of Hydrophic Ile-49.
+I. With GA95 sequence, Ile-49 have hydrophic interaction with both Leu-20 and Ile-30.
+J. Larger Phe-30 in GB sequence, move towards the protein surface in $4 \beta + \alpha$.
+K. Tyr-45 form H-bond with Asp-47 in $4 \beta + \alpha$.}  
 \end{figure}
@@
 %\begin{table}
-%\caption{Confinement method correctly reproduces the structural preferences of a series of
+%\caption{The confinement method correctly reproduces the structural preferences of a series of
 %designed peptides. Each sequence adopts either a Protein G-like fold ($4 \beta + \alpha$) or forms a
@@
 The best predictor group for this target was "BAKER-ROSETTASERVER".
-We initlally checked similarity between the models submitted by "BAKER-ROSETTASERVER" and discarded two of them from the 
+We initially checked similarity between the models submitted by "BAKER-ROSETTASERVER" and discarded two of them from the 
 analysis on the basis of being too similar to some of the other models. The GDT\_TS score and rmsd
-values as shown in Figure~\ref{fig:T0559} indicate that the model 1 was predicted correctly, 
-whereas the order of model 3 and 5 was wrong. Main difference
-between the model 3 and the rest of the model is that the orientation of first alpha helix of model 3 is opposite.
-On the other hand, as shown in Figure~\ref{fig:T0559} confine and release method not only can differentiate the native 
-structure from the submitted models , it can also correctly rank all the models like their The GDT\_TS score.
+values as shown in Figure~\ref{fig:T0559} indicate that model 1 was predicted correctly, 
+whereas the order of model 3 and 5 was wrong. The main difference
+between model 3 and the rest of the models is that the orientation of the first alpha helix of model
+3 is \Alberto{opposite??}.
+On the other hand, as shown in Figure~\ref{fig:T0559} the confine and release method not only can differentiate the native 
+structure from the submitted models, the ranking also correlates perfectly with the GDT\_TS score.
 
@@
 
-We mentioned earlier that our predictive abilities are greatly decreased with low model qualities. It is
-better to have at least one good model to pick out the native/native like structures from the rest of the decoy. 
+So far we have seen that the method is good at predicting preferences when the structures are not
+very close to each other, how far from native can we go? In this section we want to explore what
+happens when we use the method to describe low quality models. We use one of the CASP targets (pdb
+id = 2KJX, CASP id = T0531) .
+We mentioned earlier that our predictive abilities are greatly decreased with low model qualities. 
+Comparing low qualitive models between themselves is not very informative,
+    however, comparing all of them to a native/native like structure, allows to rank order this
+structures. 
 In order to test this hypothesis we performed another test with target T0531 of CASP9, in which we compared the crystal 
@@
 energies, meaning that the actual ordering could change a lot \cite{Perez2012}. It is encouraging that at least
-the method can pick out the best model with low GDT score of 44.
+the method can pick out the best model even though it is got a low GDT score: 44.
 
@@
 
-\subsubsection{Can confinement method perform better quality assessment in protein structure prediction ?}
-A part of CASP experiment is dedicated to the assessment of the quality of predicted model \cite{Kryshtafovych2011}. 
-Most of the top performing group in this catagory use consensus approaches for quality assessment \cite{Wang2011}.
+\subsubsection{Can the confinement method perform better quality assessment in protein structure prediction?}
+A part of the CASP experiment is dedicated to the assessment of the quality of predicted models \cite{Kryshtafovych2011}. 
+Most of the top performing groups in this category use consensus approaches for quality assessment \cite{Wang2011}.
 A review of CASP9 quality assessment also pointed out that the methods which are based on clustering
 techniques perform
-much better compared to those methos which are based on analysis of individual models \cite{Kryshtafovych2011}. 
-In this background we try to answer, whether the confinement method can do better quality assessment compare to the best performing
+much better compared to those methods which are based on analysis of individual models \cite{Kryshtafovych2011}. 
+In this background we try to answer, whether the confinement method can do better quality assessment
+compared to the best performing
 group (MUFOLD-WQA) in this catagory. In CASP9, the overall reliability of models in quality assessment (QA) was accepted in a mode known as
 QMODE 1. Here, predictors were asked to score each model on a scale from 0 to 1, with higher values corresponding to better models \cite{Kryshtafovych2011}. 
-We have again picked up the CASP taget T0538 to answer this question. We have chosen the Model 3 submitted by PconsR (GDT\_TS = 96, QMODE 1 = 0.5434) 
+We have again picked up the CASP taget T0538 to answer this question. We have chosen Model 3 submitted by PconsR (GDT\_TS = 96, QMODE 1 = 0.5434) 
 and model 5 from the MULTICOM-NOVEL ( GDT\_TS = 83, QMODE 1 = 0.5865). Both are server predicted models and they were further used for quality assessment 
@@
 
-We have also carried out the similar type of analysis for the case of Target 560 of CASP9 experiment. For that 
+We have also carried out a similar type of analysis for the case of Target 560 from the CASP9 experiment. For that 
 we have chosen the model 1 from the group PconsR (GDT\_TS = 94, QMODE 1 = 0.5178), Bilab\-Enable (GDT\_TS = 87, QMODE 1 = 0.5344) 
-and Multicom\_Refine (GDT\_TS = 74, QMODE 1 = 0.4770). If, one want to rank order with respect to the GDT\_TS values, the model from 
+and Multicom\_Refine (GDT\_TS = 74, QMODE 1 = 0.4770). If, one wants to rank order with respect to the GDT\_TS values, the model from 
 PconsR is the best predicted model, but according to the QA assessment by MUFOLD-WQA, the model from Bilab-Enable was found to be best predicted model. 
@@
 residues, the free energy calculated were $(\Delta \Delta G (Model\_Pconsr - Model\_Bilab\-Enable = 3.3 Kcal/Mol)$ and 
-$(\Delta \Delta G (Model\_Pconsr - Model\_Multicom\_Refine = -30.0 Kcal/Mol)$ which 
+$(\Delta \Delta G (Model\_Pconsr - Model\_Multicom\_Refine = -30.0 Kcal/Mol)$ \Alberto{-30 sounds
+huge!} which 
 support the QA assessment by MUFOLD-WQA. On the other hand, if we calculate the same free energy with the trimmed version of the protein 
@@
 The summary of this whole observation is presented in Table~\ref{table:560QA}.
-There is no doubt that the consensus approach can predict the model quality in a faster manner. But we expect that confinement
-method can predict the model quality in a relatively expensive but much accurate way.
+There is no doubt that the consensus approach can predict the model quality in a faster manner. But
+we expect that the confinement
+method can predict the model quality in a relatively expensive but much more accurate way.
 
@@
 
-In recent past there are number of studies directed towards understanding protein 
+\Alberto{The last 50 years represent our effort to understand what the native state of a protein
+    looks like by using theoretical methods. Experimental evidence coming from NMR and x-ray
+    cristallography remain so far the only real ways to solve the structure of the native state.
+    However, this methods are expensive and slow. Theoretical methods on the other hand are cheaper,
+    but despite improvement still have a long way to go. Often this computational pipelines produce
+    ensembles of diverse structures and identifying which structure is more native like is still a
+    problem. Here we have shown how the application of the confinement method can help identify
+    native like states. The high success rate balances with the computational expense. However, with
+    GPU technology and the rate at which computers improve, this computational expense is becoming
+less of a bottle neck.}
+\Alberto{I don't agree with some of the points here, but I think I covered what you meant in the
+previous paragraph}
+In the recent past there have been a number of studies directed towards understanding protein 
 structure using theoretical tools. Experimentally, it is relatively easy to identify the native structure from the data obtained 
@@
 most populated cluster is picked out as the most probable structure. However, it may happen that the structure may be trapped and 
-spend most of its time in a kinetic trap, which may give a wrong interpretation of the clustering result. On the other hand, some groups
+spend most of its time in a kinetic trap ***kinetic traps in MD,not other methods, local
+minima***, which may give a wrong interpretation of the clustering result. On the other hand, some groups
 also depend on potential energy functions to pick out their best structure, lacking an entropic
 component. In this work we attempted to identify the native/native like protein strcuture from the decoy using free energy as a scale.
-We have used previously developed confinement method to larger systems.
-We found that in majority of the case, the native structure is always the most stable and we can rank the protein structure upto 100 amino 
-acid residues long. This method works very well if atleast one of the strcuture in the decoy is good. Main advantage of this method is 
+We have used the previously developed confinement method, applying it to larger systems.
+
+We found that in majority of cases, the native structure is always the most stable and we can rank
+protein structures upto 100 amino 
+acid residues long. This method works very well if at least one of the strcuture in the decoy set
+has high quality. The main advantage of this method is 
 that it does not require any reaction coordinate. Thus one can calculate the free energy difference between two very different conformations. 
-An interesting application, that we have developed using this method is per residue
-free energy calculation. Although, we have calculated that in an approximate way, but it give important information regarding conformational 
-preference of a particular residue. In future, this may help for protein design. 
-First, we have tested this method for a series of chameleon sequences, which have $88$, $95$ and $98$ percent sequence identity
-but one of them has a $3 \alpha$ structure and another one has $4 \beta + \alpha$ structure. We found that the method can pick the 
-the conformational preferance of a particular sequence, which matches with the experimental results. Per residue free energy calculation 
-identify the residues that stabilize/destabilize a particular conformation and can help for better protein design. 
-Encouraged by that result we have tested this method for different targets of CASP experiments. We found 
-in most cases, it can pick out the native/native like structure in terms of lowest free enrgy value and rank the protein structure
-with terms of free energy. 
+An interesting application from this method that we have exploited is the partition of free energies
+on a per residue basis.
+ Although, it is an approximate partition, it gives important insight regarding the conformational 
+preferences of a particular residue. The use of this decomposition in understanding the effect of a
+small number of aminoacids in our test chamaleon sequences allows us to think that in the future,
+this method may be used for protein design. 
 The method can be also give wrong result if all the models in the decoy are very poor. It may also be difficult to rank structures if 
@@
 The confinement method has been described in details in ref. by Tyka et.al. \cite{Tyka2006} and 
-Cecchini et. al. \cite{Cecchini2009}. The basic approach of confinement approach is same in both
-these paper. However, there are some technical differences. Here we briefly describe what we follow
-to compute free energy $\Delta G_{AB}$ between A and B
-conformation of the protein.
+Cecchini et. al. \cite{Cecchini2009}. The basic approach of the confinement approach is the same in both
+these papers. However, there are some technical differences. Here we briefly describe the procedure
+to compute free energy $\Delta G_{AB}$ between two conformations A and B of a protein. The basic
+idea to compute the free energy between conformations A and B is to perform a thermodynamic cycle:
 
@@
 
-\item  We first minimized both the conformation, A and B. These minimized conformation(A* and B*) 
-       are the reference conformation of that state
+\item  Minimization of A and B. These minimized conformations (A* and B*) 
+       are the reference conformation of that state. During this minimization the backbone is kept
+       restrained to the original position to avoid large conformational changes.
 
-\item Backbone dihedral angle of both the full protein is restrained by a harmonic potential. This is required to keep
-    the configurational state of the protein. This approach is particularly important, so that the any misfolded part 
-    of the protein does not chage its configuration.
-
-\item  The free energy is calculated using a thermodynamic cycle. The first part of this approach is the confinement, where
-       the state A and B are confined to the reference state A* and B*. This is done gradually by applying larger and larger
-       harmonic restraint on all the atoms of the biomolecule. This is done by running 21 molecular dynamics simulation 
-       each of 20 ns long, where the harmonic restraint constant was varied from 0.00005 to 81.92 untill the free and the 
-       restrained state overlap well. The final restrained state was chosen so high so that the rotational contributional 
-       of the protein frozen out and only remaining contribution remain that of vibrational free enrgy. The fluctuation from
-       the reference structure was recorded and the free energy is calculated using a numerical
-       approach developed by Tyka et. al. \cite{Tyka2006}. The confinement free energy calculated in this way recorder as 
+   \item  The free energy of confining the ensemble (A or B) to a microstate (A* or B*) is
+       calculated. This is done by gradually applying larger and larger
+       harmonic restraints on all the atoms of the biomolecule. This is done by running 21 molecular dynamics simulation 
+       (20 ns long each), where the harmonic restraint force constant was scaled from 0.00005
+       \Alberto{units??}
+       (mostly free) to 81.92 (frozen in one microstate). In this final restrained state, the
+       rotational contribution to the free energy is frozen out
+       and the only remaining contribution is the vibrational part. The free energy for this step is
+       estimated from the fluctuations around
+       the reference structure using a numerical
+       approach developed by Tyka et. al. \cite{Tyka2006}. The confinement free energy calculated in
+       this way is recorded as 
        $\Delta G_{A,A*}$ and $\Delta G_{B,B*}$ as shown in Figure~\ref{fig:method}.
@@
 
-In all the calculation amber 11 program is used with ff99SB with GB/SA implicit solvent. Interestingly, we extend the
+ All calculations where performed with the amber 11 suit of programs in combination with ff99SB and GB/SA implicit solvent. Interestingly, we extend the
 method for calculation of per residue free energy in an approximate way. For this purpose, the confinement energy, 
@@
 call this method approximate as we ignore the contribution from the normal mode or quasiharmonic analysis. 
-However, this contrbution to the total free energy is very less (less than $1.0 Kcal/Mol$), which allow us to 
+However, this contrbution to the total free energy is much smaller (less than $1.0
+Kcal/Mol$\Alberto{that sounds huge, i would say what percent of error we stimate, I think it was
+lower than 1kcal/mol}), which allow us to 
 study the mechanistic details of conformational preference of each residue.

### 2013-01-11 Paper/paper.tex

@@
 
-The calculation of free energy has been successfully attempted by a number of groups \cite{Meirovitch2007}, 
+The calculation of free energy have been successfully attempted by a number of groups \cite{Meirovitch2007}, 
 \cite{Ytreberg2006} - \cite{Zheng2008}. In recent years methods like the Reference system method \cite{Ytreberg2006}, 
@@
 came out from our calculation ranges from
-around 2.9 to 5.7 kcal/mol, which is consistent with the above hypothesis \cite{He2008}, \cite{{Alexander2009}}.
+around 3.5 to 5.0 kcal/mol, which is consistent with the above hypothesis \cite{He2008}, \cite{{Alexander2009}}.
 
-\subsection{Per residue free energy calculation can identify the mechanistic detail behind conformational pereference 
+\subsection{The per residue free energy calculation can identify the mechanistic detail behind conformational pereference 
 of a particular residue}
 
-We decomposed the calculated free energy into per residue decomposition as described in the methods
-section. This helps us to
-identify important residues that stabilize a particular conformation. 
+We decomposed the calculated free energy into per residue decomposition as described in the method
+section. This helps us to identify important residues that stabilize a particular conformation. 
+In this direction, we first calculate the per residue free energy for the GA95 and GB95 sequences. 
+The per residue free energy, $\Delta \Delta G ((4 \beta + \alpha$) - ($3 \alpha))$ is plotted in Figure~\ref{fig:orban_full}(F).
+In this plot, the negative peak indicate the residue will stabilize the $4 \beta + \alpha$ conformation of a particular
+residue and the positive peak will stabilize the $3 \alpha$ structure. As, the differences of amino acid
+residues are only at 3 position, there are some common features of residues in both GA95 and GB95 sequences in both
+$3 \alpha$ and $4 \beta + \alpha$ conformation. Role of some important residues in stabilizing either 
+$4 \beta + \alpha$ or $3 \alpha$ conformation and possible reason for such conformational preferences are discussed in 
+details in Figure~\ref{fig:orban_full} and Table~\ref{table:orban_perresidue}. It is important to note that
+experimental observations classified the protein into two parts: Amino acids 9-51 are fully structured in both
+the folds, where as residues 1-8 and 52-56 are unstructured in $3 \alpha$, but form $\beta$ strand in $4 \beta + \alpha$ structure.
+Most of the amino acid residues in the region 1-9 favor $4 \beta + \alpha$ structure as can be seen from the plot
+Figure~\ref{fig:orban_full}(F). Interestingly, we can explain most of the big peak in this plot. They are presented 
+in Figure~\ref{fig:orban_full} and Table~\ref{table:orban_perresidue}. Here, we explain only one case. The first big peak
+appears at position 7. Here, the hydrophobic Leu-7 stabilizes the $4 \beta + \alpha$ conformation as its sidechain is oriented 
+towards the protein hydrophobic core, whereas in $3 \alpha$ structure it exposed to the solvent (Figure~\ref{fig:orban_full}(A)).
+Thus, if someone want to better design the $3 \alpha$ conformation of this protein, we propose to mutate Leu at this position
+with some hydrophilic residue. Similarly, we found that the free energy differences arise in two conformation for various 
+reasons ranging from hydrogen/salt bridge formation to relative solvent exposure of hydrophobic residues etc. They are 
+listed in Figure~\ref{fig:orban_full} and Table~\ref{table:orban_perresidue}.
+
 \Alberto{This is not Reader Friendly at all. we do not have to describe every little detail. I would
@@
 This method can be further used for calculating free energy due to very large conformational change. Another application may be
-for the case of ligand binding. It can also be used to distinguish the most stable structure during protein design. Decompition of 
+the case of ligand binding. It can also be used to distinguish the most stable structure during protein design. Decompition of 
 total free energy into the per residue component help identifying residues that give stability to a particular conformation. 

### 2013-01-13 Paper/paper.tex
Cleaned up intro and first section of results.

@@
 \author{Arijit~Roy}
-\author{Justin~L.~MacCallum}
 \author{Alberto~Perez}
 \author{Ken~A.~Dill}
+\author{Justin~L.~MacCallum}
 \affil{Laufer Center for Physical and Quantitative Biology\\
@@
 
-\title{Physics-based protein structure prediction and design using the confinement method}
+\title{Predicting the conformational preferences of proteins using a physics-based free energy
+method}
 
@@
 
-In computational structural biology there are numerous cases where the free energy difference
-between two well defined 
-states plays an important role. Examples range from small to large conformational changes of proteins due to
-ligand binding, change of pH or other conditions \cite{Meirovitch2007}, \cite{Chipot2007}, \cite{Jorgensen2004}. 
-Free energy also plays an important role in the case of protein
-folding. Ground breaking work of Christian B. Anfinsen and coworkers showed that the native
-structures of small globular proteins have a unique, thermodynamically stable native structure with
-their conformation at the global free energy minimum \cite{Anfinsen1973}. Regardless of the starting
-point or even after unfolding 
-it by changing different condition most proteins will go back to the native state after
-reinstituting native like conditions. Often, the changes in free energy between different
-conformations is small, but nonetheless crucial in favoring the native state
- since defects in protein folding may be the molecular cause of a range of human genetic disorders.
- For example a misfolded protein known as a
-prion can misfold correctly folded proteins when entering a healthy organism. Thus it is important for 
-protein to have its native structure. As protein native structure stays in the global free energy minimum, 
-\Arijit{free energy can be used as a scale for accurate identification of misfolded state from the protein native state.}
-\Alberto{I'm not sure I understand this here} 
-These facts further give free energy a special importance in structural biology.
+Protein conformation and associated conformational changes are important in biology. The mechanism
+of enzymes, the function of structural proteins, and the binding of proteins to DNA or other
+proteins all depend on protein conformations and the relative free energies of different
+conformational states. Proteins can undergo conformational changes for a variety of reasons,
+including ligand binding, change of pH or other conditions \cite{Meirovitch2007}, \cite{Chipot2007},
+\cite{Jorgensen2004}. The process of protein folding also depends on the free energies of different
+conformational states. The native structure is at a free energy minimum \cite{Anfinsen1973}, while
+the kinetics are determined by the free energies of various on- and off-pathway intermediates.
 
-However, the theoretical calculation of conformational free energy change becomes difficult if the
-states of interest are very different \cite{Meirovitch2007}. In such scenario timescales involved in
-such conformational transitions may
-be beyond the sampling ability of classical molecular dynamics simulation.  Generally, in order to
-calculate the free energy a pathway or reaction coordinate is needed with several intermediate structures.
-Methods like umbrella sampling \cite{Torrie1977}, thermodynamic integration \cite{Tironi1994} along with classical MD can suffer from
-overlap problem and requires a larger computational effort. The idea of a reaction coordinate became
-even more
-complicated in the case of protein folding.  As free energy landscape for protein folding 
-is rough \cite{Dill1997}, \cite{Dill2008}, the molecule 
-can get trapped in an energy well during conformational sampling. Even with 
-advanced methods like replica exchange molecular dynamics, the three dimensional structure may be
-trapped in a local minima for a considerable time, giving rise to nonconverged distributions where
-misfolded states might be selected as native like
-% (give the example of human pin1ww domain from Benout Roux and Klaus Schultan).
+However, the theoretical calculation of conformational free energy changes is difficult when the
+states of interest are very different \cite{Meirovitch2007}. In these cases, the timescales involved
+in such conformational transitions may be too long to simulate directly. For example, in protein
+folding the free energy landscape is rough \cite{Dill1997}, \cite{Dill2008}, and the molecule can
+become trapped in an energy well during conformational sampling. Even with advanced methods like
+replica exchange molecular dynamics, the three dimensional structure may be trapped in a local
+minima for a considerable time. Sometimes this problem can be circumvented by defining a pathway or
+reaction coordinate connecting the states of interest. The free energy along this reaction
+coordinate can then be determined using methods such as umbrella sampling \cite{Torrie1977},
+targeted MD \cite{}, or thermodynamic integration \cite{Tironi1994}. However, accurately sampling
+this reaction coordinate may still be difficult and require a large computationl effort. In the case
+of protein folding, it is often difficult to even devise a meaningful reaction coordinate that could
+connect the unfolded and folded states efficiently \cite{}.
 
-The calculation of free energy have been successfully attempted by a number of groups \cite{Meirovitch2007}, 
-\cite{Ytreberg2006} - \cite{Zheng2008}. In recent years methods like the Reference system method \cite{Ytreberg2006}, 
-Decativated Morphing \cite{Park2008}, Orthogonal Space Random Work \cite{Zheng2008} or the Confinement 
-Method \cite{Tyka2006}, \cite{Cecchini2009} to name a few have been used to calculate free energy
-differences.
-In this work we have applied the confinement method which was originally devoled by Tyka et. al. \cite{Tyka2006} and
-Cecchini et. al. \cite{Cecchini2009}. This method relies on the fact that the free energy difference is a state function and 
-thus it is independent of the sampling path. This approach uses a thermodynamic cycle where first, non-harmonic degrees of
-freedom are removed by applying a series of restraints to the system. Then the free energy of the
-remaining harmonic system is obtained using a normal mode calculation and combined with the
-non-harmonic part to give the total absolute free energy.  Previously, this method was applied to
-calculate the conformational free energy of a 16 amino acid residue peptide, known as BHP.  We have
-first reproduced that known result. Next we have applied this method to larger proteins for the
-first time. The free energy difference of pairs of conformations in proteins with similar sequence but completely
-different fold was calculated. \Alberto{i made this part too long, have to summarize here and define
-better inside text?-->} Encouraged by the result, we applied the method to help tackle the
-problem of protein conformation ranking. We did this by selecting different targets from the
-Critical Assessment of Structure Prediction (CASP) event. In this event, different research groups
-try to find the 3D structure of a protein given a sequence in a blind way using their methodologies. Upon comparing
-with the real structure it is often seen that several groups are able to generate better structures
-than the ones they submit, but are not able to identify them. This is known as the ranking problem,
-and this method is able to correctly rank structures in the majority of cases we have tried.
- Finally, we have been exploring the ability of working on a per residue basis to identify which
- residues are the major players in the stabilization/destabilization of a pair of conformations. Our
- initial experience makes us thinkg that this method might be helpful in identifying key residues
- for protein design.
+The calculation of protein conformational free energy has been successfully attempted by a number of
+groups \cite{Meirovitch2007,Ytreberg2006,Zheng2008}. In recent years, methods such as the reference
+system method \cite{Ytreberg2006}, decativated morphing \cite{Park2008}, orthogonal space random
+walk \cite{Zheng2008}, and the confinement method \cite{Tyka2006,Cecchini2009} have been used to
+calculate free energy differences.
 
-%of protein can be obtained to atomic resolution by X-ray crystallography or NMR. Unfortunately, the
-%sheer amount of time and economic investment for structure determination cannot keep up with the
-%rate at which proteins of unknown structure are sequenced. Therefore, computational models can be a
-%good way to bridge the increasing gap between sequence and structure. In this respect CASP
-%(Critical Assessment of Structure Prediction) has played a crucial role which is a community wide
-%experiments for the protein structure prediction.
+In this work we explore confinement method\cite{Tyka2006,Cecchini2009}, which relies on the fact
+that the free energy difference is a state function and thus is independent of the sampling path.
+This approach uses a thermodynamic cycle \ref{fig:method} where first protein conformational entropy
+of the two end-states is frozen out by applying a series of restraints to the system. Next, the free
+energy difference between the two highly-restrained harmocic systems is obtained using a normal mode
+calculation.
+
+Previously, this method was applied to calculate the conformational free energy of a 16 amino acid
+residue peptide, known as BHP.  We have first reproduced that known result. Next, we have applied
+this method to larger proteins in two different case studies. In the first, we attempt to pick the
+most native like structures from a set of models produced during the Critical Assessment of
+Structure Prediction (CASP) \cite{} experiment. In the second case study, we examine the
+conformational preferences of a series of chameleon proteins designed by Orban and co-workers
+\cite{} that can switch between two completely different folds with only small changes in amino acid
+sequence.
 
@@
 
-\subsection{The confinement method produces correct results in control experiments}
+\subsection{Our implementation of the confiment method reproduces results from the literature}
 
-\Alberto{ This is benchmarking, good internal, but not relevant. Should go to SI,out, or summarized
-    in one sentence:  "After reproducing the results of Cecchini and coworkers we were interested in
-the ability to apply this method to larger systems"}
-
-\Alberto{As a first step, we performed several control experiments to verify that our implementation of the
+As a first step, we performed several control experiments to verify that our implementation of the
 confinement method produces results compatible with previous calculations reported in the
@@
 
-The method has previously been applied to a 16 amino acid residue $\beta$-hairpin from
-protein G, known as BHP \cite{Cecchini2009}. We calculated the free energy difference between two different
+The method has previously been applied to a 16 amino acid residue $\beta$-hairpin from protein G,
+known as BHP \cite{Cecchini2009}. We calculated the free energy difference between two different
 conformations of the peptide: (1) the native conformation, called bhp1, with a two-stranded
 $\beta$-sheet; and (2) a conformation, called bhp3, which has a three-stranded $\beta$-sheet.
-Analysis of long (4 $\mu$s) equilibrium simulations \cite{Cecchini2009},\cite{Krivov2004}  shows that bhp1 is the more favorable
-configuration by 1.8 kcal/mol. Using the confinement method, we obtain a value 1.7 kcal/mol, which
-is in good agreement with the equilibrium simulations and with previous calculations using the
-confinement method \cite{Cecchini2009} (see Supporting Information for further details).}
+Analysis of long (4 $\mu$s) equilibrium simulations \cite{Cecchini2009,Krivov2004}  shows that bhp1
+is the more favorable configuration by 1.8 kcal/mol. Using the confinement method, we obtain a value
+1.7 kcal/mol, which is in good agreement with the equilibrium simulations and with previous
+calculations using the confinement method \cite{Cecchini2009}.
 
-Previous applications of the confiment method have focused on relatively short peptides, up to $17$
-residues in length. In this work we apply the confinement method to
-larger proteins. On that direction, we first test the confinement method for a chamelon sequence and 
-found that it can correctly pick out the preferential structure of the chamelon sequence. Other 
-applications we explore in the present work is the re-scoring, or
-metaprediction, of structures submitted during the Critical Assessment of Structure Prediction
-(CASP) experiment (described in detail later). We computed the relative free energies of predictions
-submitted during CASP9, with the expectation that the most native-like prediction will have
-the lowest free energy. As a control, for several targets we also calculated the relative free
-energy of the experimentally determined structure. \Alberto{should not start with results yet, first
-we are describing what we are going to show them} As
-Table~\ref{table:casp_control} shows, in most cases, the confinement method correctly assigns a
-lower free energy to the experimentally determined structure than to any of the decoys. 
-Most interestingly, we decomposed the free energy into its per residue component, which help 
-us to identify the residues which stabilize or destabilize a particular conformation of the protein.
-%Although differentiating between experimentally determined structures and computer generated predictions is
-%not a stringent test of a scoring method---for example, examining the residue-residue packing can
-%easily distinguish between the two [ref Rosetta Holes]---the results none the less serve as a useful
-%"sanity check".
+\subsection{The confiment method predicts that experimental structures have lower free energies than
+computer-generated models}
+
+Previous applications of the confiment method have focused on relatively short peptides, up to 17
+residues in length. In this work we apply the confinement method to larger proteins. As a first
+test, we calculated the free energies of models submitted by predictors during the CASP9 experiment
+(often referred to as decoys) and compared them with the calculated free energy of the corresponding
+experimentally determined structure from the Protein Data Bank (PDB). As
+Table~\ref{table:casp_control} shows, in 5 out of 6 cases, the confinement method assigns a lower
+free energy to the experimentally determined structure than to any of the decoys---as would be
+expected. Although differentiating between experimentally determined structures and computer
+generated predictions is not a stringent test of a scoring method---for example, examining the
+residue-residue packing can easily distinguish between the two [ref Rosetta Holes]---the results
+none the less serve as a useful "sanity check".
 
@@
 \caption{The confinement method assigns a more favorable free energy to the experimentally
-determined structure than to computer-generated predictions. For each target, we examined as many as
-five predictions submitted by CASP participants. Positive $\Delta\Delta G$ values indicate that the
-experimental structure is predicted to be more favorable than any of the decoys.}
+    determined structure than to computer-generated predictions. For each target, we examined as
+    many as five predictions submitted by CASP participants. We report the free energy difference
+    between the most favorable decoy and the experimentally determined structure. Positive
+    $\Delta\Delta G$ values indicate that the experimental structure is predicted to be more
+favorable than any of the decoys.}
 \label{table:casp_control}
@@
 \end{center}
-
 \end{table}

### 2013-01-14 Paper/paper.tex
Edited chameleon peptide section.

@@
 is marginally stable in one of two possible folds. By mutating key residues in this sequence they
-are able to stabilize one fold or the other (see Figure~\ref{fig:orban}). Sequences that adopt a mixed alpha/beta structure similar to Protein G
-are denoted as ``GB'', while sequences that form a three-helix bundle are denoted as ``GA''. One
-pair of sequences (GA88/GB88) are 88 percent identical in sequence and differ in seven positions. Another pair
-(GA95/GB95) are 95 percent identical and differ in three positions. The last pair (GA98/GB98) differ
-only in a single tyrosine to alanine mutation.
-
+are able to stabilize one fold or the other (see Figure~\ref{fig:orban}). Sequences that adopt a
+mixed alpha/beta structure similar to Protein G are denoted as ``GB'', while sequences that form a
+three-helix bundle are denoted as ``GA''. We label the two folds as $3 \alpha$ and $4 \beta +
+\alpha$. The GA sequences are observed to fold into the $3 \alpha$ fold and the GB sequences fold
+into the $4 \beta + \alpha$ fold. One pair of sequences (GA88/GB88) are 88 percent identical in
+sequence and differ in seven positions. Another pair (GA95/GB95) are 95 percent identical and differ
+in three positions. The last pair (GA98/GB98) differ only in a single tyrosine to alanine mutation.
 The fact such small changes in sequence can lead to such dramatic changes in structure is rather
@@
 the GA88 sequence and the GB88 structure). The confinement method was able to predict the
-conformational preferences correctly for all six sequences (data not shown). This is however not
+conformational preferences correctly for all six sequences (data not shown). This is, however, not
 surprising. It is well known [CITE] that it is easy to distinguish computational models from native
-structures. Despite the fact that this models are built by having a huge amount of structural
-information, it might be possible that we were able to make correct predictions simply
-because artefacts of our modeling procedure always lead to the model having a higher free energy
-than the experimental structure.
+structures. To avoid this potential problem, we instead computed relative free energy of two
+different computer generated models for each sequence. One model is based on the GA structure and
+the other on the GB structure (see Supporting Information for details on the modeling procedure).
+This is a much more realistic test of the confinement method's ability to accurately calculate
+relative free energies.
 
-To avoid this potential problem, we instead computed relative free energy of two different
-structural models for each sequence. One model is based on the GA structure and the other on the GB
-structure (see Supporting Information for details on the modeling procedure). This is a much more
-realistic test of the confinement method's ability to accurately calculate relative free energies.
+The results of these calculations are presented in Figure~\ref{fig:orban}. The confinement method
+identifies the correct structure for all five sequences. One hypothesis for fold switching of
+chameleon sequences is that the structural transitions require states with diminished stability
+[ref]. It is believed that if the free energy of the native state and the alternative state are
+within a range of 5 Kcal/mol, then it is possible for the native state to be destabilized relative
+to the alternative state with only small changes in sequence. The stability of the native state can
+decrease for a number reason ranging from chemical modification, breaking of disulphide bonds or
+mutations---as is the case here. The calculated free energy differences range from around 3.5 to 5.0
+kcal/mol, which is consistent with the above hypothesis \cite{He2008}, \cite{{Alexander2009}}.
 
-The results of these calculations are presented in Figure~\ref{fig:orban}. Confinement method
-identifies the correct structure for all five sequences. One of the hypothesis for fold switching
-of chamelon sequence is that the structural transitions require states with diminished stability.
-It is widely believed that if the free energy of the native state and the alternative state is within
- a range of $5 Kcal/Mol$ then it can quickly change fold when the stability of the native state 
-decreases. The stability of the native state can decrease for a number reason ranging from chemical 
-modification, breaking of disulphide bonds or mutations as in this case. The free energy differences, that
-came out from our calculation ranges from
-around 3.5 to 5.0 kcal/mol, which is consistent with the above hypothesis \cite{He2008}, \cite{{Alexander2009}}.
+\subsection{Per-residue free energy calculation can identify the mechanistic detail behind
+conformational preferences}
 
-\subsection{The per residue free energy calculation can identify the mechanistic detail behind conformational pereference 
-of a particular residue}
+To understand the mechanism behind the conformational preferences of the different protein
+sequences, we decomposed the calculated free energy into per-residue contributions as described in
+the method section. This helps us to identify important residues that stabilize a particular
+conformation. The per residue free energy, $\Delta \Delta G ((4 \beta + \alpha$) - ($3 \alpha))$ is
+shown in Figure~\ref{fig:orban_full}(F). In this plot, a negative peak indicates that a residue
+favors the $4 \beta + \alpha$ conformation and a positive peak indicates that the residue favors the
+$3 \alpha$ structure.
 
-We decomposed the calculated free energy into per residue decomposition as described in the method
-section. This helps us to identify important residues that stabilize a particular conformation. 
-In this direction, we first calculate the per residue free energy for the GA95 and GB95 sequences. 
-The per residue free energy, $\Delta \Delta G ((4 \beta + \alpha$) - ($3 \alpha))$ is plotted in Figure~\ref{fig:orban_full}(F).
-In this plot, the negative peak indicate the residue will stabilize the $4 \beta + \alpha$ conformation of a particular
-residue and the positive peak will stabilize the $3 \alpha$ structure. As, the differences of amino acid
-residues are only at 3 position, there are some common features of residues in both GA95 and GB95 sequences in both
-$3 \alpha$ and $4 \beta + \alpha$ conformation. Role of some important residues in stabilizing either 
-$4 \beta + \alpha$ or $3 \alpha$ conformation and possible reason for such conformational preferences are discussed in 
-details in Figure~\ref{fig:orban_full} and Table~\ref{table:orban_perresidue}. It is important to note that
-experimental observations classified the protein into two parts: Amino acids 9-51 are fully structured in both
-the folds, where as residues 1-8 and 52-56 are unstructured in $3 \alpha$, but form $\beta$ strand in $4 \beta + \alpha$ structure.
-Most of the amino acid residues in the region 1-9 favor $4 \beta + \alpha$ structure as can be seen from the plot
-Figure~\ref{fig:orban_full}(F). Interestingly, we can explain most of the big peak in this plot. They are presented 
-in Figure~\ref{fig:orban_full} and Table~\ref{table:orban_perresidue}. Here, we explain only one case. The first big peak
-appears at position 7. Here, the hydrophobic Leu-7 stabilizes the $4 \beta + \alpha$ conformation as its sidechain is oriented 
-towards the protein hydrophobic core, whereas in $3 \alpha$ structure it exposed to the solvent (Figure~\ref{fig:orban_full}(A)).
-Thus, if someone want to better design the $3 \alpha$ conformation of this protein, we propose to mutate Leu at this position
-with some hydrophilic residue. Similarly, we found that the free energy differences arise in two conformation for various 
-reasons ranging from hydrogen/salt bridge formation to relative solvent exposure of hydrophobic residues etc. They are 
-listed in Figure~\ref{fig:orban_full} and Table~\ref{table:orban_perresidue}.
+The GA and GB sequences are nearly identical, so there are some common features observed for all
+sequences. Roles of some important residues in stabilizing either $4 \beta + \alpha$ or $3 \alpha$
+conformation and possible reason for such conformational preferences are discussed in details in
+Figure~\ref{fig:orban_full} and Table~\ref{table:orban_perresidue}. The experimental observations
+classified the protein into two parts: Amino acids 9-51 are fully structured in both the folds,
+where as residues 1-8 and 52-56 are unstructured in $3 \alpha$, but form $\beta$ strand in $4 \beta
++ \alpha$ structure. Most of the amino acid residues in the region 1-9 favor $4 \beta + \alpha$
+structure as can be seen from the plot Figure~\ref{fig:orban_full}(F). We can explain most of the
+major peaks in this plot. For example,  first major peak appears at residue 7. Here, the hydrophobic
+Leu-7 stabilizes the $4 \beta + \alpha$ conformation as its sidechain is oriented towards the
+protein hydrophobic core, whereas in $3 \alpha$ structure it exposed to the solvent
+(Figure~\ref{fig:orban_full}(A)). Thus, we hypothesise that a mutation of Leu-7 to a hydrophilic
+residue would stabilize the $3 \alpha$ fold. The conformational preferences of most of the other
+large peaks can be similarly explained by hydrogen bonding, salt bridge formation, and solvent
+exposure of hydrophobic residues. They are listed in Figure~\ref{fig:orban_full} and
+Table~\ref{table:orban_perresidue}.
 
-\Alberto{This is not Reader Friendly at all. we do not have to describe every little detail. I would
-    rather have a table where we have the pairing and the reason for favoring one structure or the
-    other. In the main text I would say that the main reason for preferences of one fold or another
-    has to do with the relative solvent exposure of the residue in one conformation or another, or
-    the possibility of stablishing hydrogen/salt bridges in one conformation. This would be my
-preference, ditching everything below until the next section. What do you think?}
-
-\Arijit{I have rewrite this part.There was a table. I made it better. I want to keep some discussion though if you agree.}
-
-Now, let us discuss the differences of $3 \alpha$ and $4 \beta + \alpha$ fold due to specific difference of mutated amino acid 
-residues in GA and GB sequence. The differences are in positions 20, 30 and 45. While GA95 has Leu-20-Ile-30-Leu-45, GB95 sequence 
-has Ala-20-Phe-30-Tyr-45 at those position. One can find that the peak at position 49 stabilizes the GA sequence more than that of GB
-sequence. In $3 \alpha$ fold, Ile-49 has hydrophobic interactions with Leu-20 and Ile-30 with GA sequence, whereas with GB sequence it 
-has such interactions with smaller Ala-20 (Figure~\ref{fig:orban_full}(I). In $3 \alpha$, with GB sequence, Phe-30 
-being a larger group can not be completely accomodated in the hydrophobic core(Figure~\ref{fig:orban_full}(I)) and therefore, 
-partially exposed to the protein surface. There is a very interesting difference for residue at position 45 as well. In GB95 sequence
-and $4 \beta + \alpha$ fold there is H-bonding between Tyr-45 and Asp-47 (Figure~\ref{fig:orban_full}(K)), which is absent in the
-$3 \alpha$ conformation. On the other hand with GA conformation, this H-bond is absent for both the conformation as there is a Leu at position
-45 (Figure~\ref{fig:orban_full}(J)). Although, one can not find much difference between $4 \beta + \alpha$ and $3 \alpha$ conformation
-for position 20 for both GA95 and GB95 sequence. There are significant difference for the residues at position 30 and 45.  
-These residues have some long term effects as well. As already explained, residue Ile-49 has much more hydrophic interactions with residues
-at position 20 and 30 in GA sequence, which is less in GB sequence. There are difference for Asp-47 as well. With $3 \alpha$ conformation 
-and GA95 sequence, Asp-47 has H-bonding with Lys-50, which is absent in the other conformation.
+We will now discuss the differences between the $3 \alpha$ and $4 \beta + \alpha$ folds due to the
+specific differences between the GA and GB sequences. The differences are in positions 20, 30, and
+45. GA95 has Leu-20, Ile-30, and Leu-45; GB95 has Ala-20, Phe-30, and Tyr-45. In the $3 \alpha$
+fold, residues 20 and 30 are involved in interactions with residue 49. The peak at position 49
+stabilizes the GA sequence more than that of GB sequence. In the $3 \alpha$ fold, Ile-49 has
+hydrophobic interactions with Leu-20 and Ile-30 with GA sequence, whereas with GB sequence it has
+such interactions with smaller Ala-20 (Figure~\ref{fig:orban_full}(I). Also, in the $3 \alpha$ fold
+with the GB sequence, Phe-30 being a larger group can not be completely accommodated in the
+hydrophobic core (Figure~\ref{fig:orban_full}(I)) and therefore, partially exposed to the protein
+surface. There is an interesting difference for residue at position 45 as well. In GB95 sequence and
+$4 \beta + \alpha$ fold there is H-bonding between Tyr-45 and Asp-47
+(Figure~\ref{fig:orban_full}(K)), which is absent in the $3 \alpha$ conformation. On the other hand
+with GA sequence, this H-bond is absent for both the conformation as there is a Leu at position 45
+(Figure~\ref{fig:orban_full}(J)).
 
@@
 \begin{center}
-\caption{Analysis of per residue free energy decomposition of GA95 and GB95 sequence reveal that certain residues prefer
+\caption{Analysis of the per-residue free energy decomposition of GA95 and GB95 sequence reveal that certain residues prefer
 either $3 \alpha$  or $4 \beta + \alpha$ structure. These observations and the reason behind such conformational preferences are listed in this
@@
 
-%\begin{table}
-%\caption{Confinement method correctly reproduces the structural preferences of a series of
-%designed peptides. Each sequence adopts either a Protein G-like fold ($4 \beta + \alpha$) or forms a
-%helix bundle ($3 \alpha$).}
-%\label{table:orban}
-%\begin{center}
-%\begin{tabular}{l l l}\hline
-%Changes in & $\Delta G$ (kcal/mol)    & $\Delta G$ (kcal/mol)     \\
- %Sequences & ($3 \alpha$ fold)        & ($4 \beta + \alpha$ fold) \\ \hline
-%GA95       & $0.000$                  & $2.901 \pm 0.475$         \\
-%GB95       & $4.365 \pm 0.460$        & $0.000$                   \\
-%GA88       & $0.000$                  & $ 3.941 \pm 0.517$        \\
-%GB88       & $5.576 \pm 0.464$        & $0.000$                   \\
-%GA98       &  &                           \\ \hline
-%GB98       &                          &                           \\ \hline
-%\end{tabular}
-%\end{center}
-%\end{table}
-
 \subsection{The confinement method correctly identifies the most native-like predictions from a
 subset of CASP predictions}
-
-%\Alberto{Arijit: we need a global table of success, rather than individual tables. The simplest form
-%    would have 3 columns: name of the PDB, number of structures tested, percent of success of
-%    picking 1st model}
 

### 2013-02-16 Paper/paper.tex
summarize conclusions

--- 2013-01-11 Paper/paper.tex
@@
 
-\Alberto{The last 50 years represent our effort to understand what the native state of a protein
+The last 50 years represent our effort to understand what the native state of a protein
     looks like by using theoretical methods. Experimental evidence coming from NMR and x-ray
@@
     GPU technology and the rate at which computers improve, this computational expense is becoming
-less of a bottle neck.}
-\Alberto{I don't agree with some of the points here, but I think I covered what you meant in the
-previous paragraph}
-In the recent past there have been a number of studies directed towards understanding protein 
-structure using theoretical tools. Experimentally, it is relatively easy to identify the native structure from the data obtained 
-from x-ray crystallography or NMR. On the other hand using theoretical tools it is often difficult to identify the native structure 
-using both physics based methods and knowledge based method. In order to rank structure from the available data, some groups use 
-clustering algorithms and the structure of the
-most populated cluster is picked out as the most probable structure. However, it may happen that the structure may be trapped and 
-spend most of its time in a kinetic trap ***kinetic traps in MD,not other methods, local
-minima***, which may give a wrong interpretation of the clustering result. On the other hand, some groups
-also depend on potential energy functions to pick out their best structure, lacking an entropic
-component. In this work we attempted to identify the native/native like protein strcuture from the decoy using free energy as a scale.
-We have used the previously developed confinement method, applying it to larger systems.
-
-We found that in majority of cases, the native structure is always the most stable and we can rank
-protein structures upto 100 amino 
-acid residues long. This method works very well if at least one of the strcuture in the decoy set
-has high quality. The main advantage of this method is 
-that it does not require any reaction coordinate. Thus one can calculate the free energy difference between two very different conformations. 
-An interesting application from this method that we have exploited is the partition of free energies
-on a per residue basis.
- Although, it is an approximate partition, it gives important insight regarding the conformational 
-preferences of a particular residue. The use of this decomposition in understanding the effect of a
-small number of aminoacids in our test chamaleon sequences allows us to think that in the future,
-this method may be used for protein design. 
-The method can be also give wrong result if all the models in the decoy are very poor. It may also be difficult to rank structures if 
-all the structures are very very close to the native structure. But we want to ignore this fact as our main aim is to differentiate
-between the good and the bad structure. In this article we have extensively tested this method using examples from CASP 
-This method can be computationally expensive in actual CASP experiment if one want rank hundreds of submitted models in CASP 
-experiments. However, in real life experiment with a particular target, we expect this method can give accurate result. The important 
-fact about confinement method is that all the part of the calculation can be done independently. 
-There is another big bottlenecks in this calculation. This has to do with the amount of
-simulations that have to be performed to confine the ensemble of conformations close to a microstate
-into a single microstate. We have tackled this issue by using GPU (Graphical Processing Unit) technology as opposed to the classical
-CPUs, which gives us a two order of magnitude boost in computational efficiency.
-For example, we have mostly carried out our calculation with Amber program \cite{Case2005} in GPU computer. With a avarage 56 residue 
-we found that it take only 4 hour to complete 20 ns of the confinement step. So, with available computer power this method will 
-be easy to use for real problems. 
-%There is another issue, which is normal mode
-%analysis calculations are Oh(N**3) and require large amounts of memory, effectively limiting the
-%size of the systems under study. In most cases quasiharmonic analysis.
-
-This method can be further used for calculating free energy due to very large conformational change. Another application may be
-the case of ligand binding. It can also be used to distinguish the most stable structure during protein design. Decompition of 
-total free energy into the per residue component help identifying residues that give stability to a particular conformation. 
-Another, important application may be to study the protein folding kinetics, specially to study the phi value analysis.
+less of a bottle neck.
+We foresee great potential for this methodology for two reasons: independence from a reaction
+coordinate (just need start and end state) and avility to decompose the free energies on a per
+residues basis. This last point can be very useful in protein design. Finally, we have presented
+data for structure comparison of proteins, but we are currently exploring the possibility of
+using this methodology for getting free energies of ligand-protein binding. 
+\Alberto{Do we mention this in the text? too technical for conclusions:
+There is another issue, which is normal mode
+analysis calculations are Oh(N**3) and require large amounts of memory, effectively limiting the
+size of the systems under study. In most cases quasiharmonic analysis.   
+}
 

### 2013-03-06 Paper/paper.tex
Started changes to text

--- 2013-01-14 Paper/paper.tex
@@
 Calculation of free energy differences is of central importance in the simulation of biochemical
-systems. It is particularly difficult to calculate between pairs of macromolecular
-conformations as well as a computationally expensive task with existing
-methods. In this work, confinement approach is used to calculate absolute
-free energies of biomolecular systems. This method provides two main advantages: it does not require a reaction coordinate or
-transition path and it is fast to compute. Free energy calculated can be decomposed into a per
-residue contribution in an 
-approximate way. Per residue free energy allows us
-to identify the reason behind conformational preferences in biomolecules. Through out the article we show its use in different
-challenging modeling problems. In particular, we show its use in predicting the conformational
-preference of chamaleon sequences (sequences with high sequence identity and different
-folds). This sequence dependent conformational preferences and per residue free energy decomposition set the stage for the use of this 
-method in protein design.
+systems. It is particularly difficult to calculate between pairs of macromolecular conformations as
+well as a computationally expensive task with existing methods. In this work, confinement approach
+is used to calculate absolute free energies of biomolecular systems. This method provides two main
+advantages: it does not require a reaction coordinate or transition path and it is fast to compute.
+Free energy calculated can be decomposed into a per residue contribution in an approximate way. Per
+residue free energy allows us to identify the reason behind conformational preferences in
+biomolecules. Through out the article we show its use in different challenging modeling problems. In
+particular, we show its use in predicting the conformational preference of chamaleon sequences
+(sequences with high sequence identity and different folds). This sequence dependent conformational
+preferences and per residue free energy decomposition set the stage for the use of this method in
+protein design.
 
@@
 
-\subsection{The confinement method correctly identifies the most native-like predictions from a
-subset of CASP predictions}
+\subsection{We have applied the confinement method to structure prediction}
 
-This method is particularly well suited to pick out the native/native like structure from misfolded or decoys. We have
-tested this using different models from CASP (critical assessment of structure
-prediction). CASP is a blind test in which different groups world wide apply their
-methods to predict the 3D structure of proteins given their sequence. This is done with a strict 3
-week limit on each target (3 day for servers) and each group is allowed to submit 5 possible
-structures (ranked from best to worst). In our role as assessors during the CASP8 and CASP9 \cite{MacCallum2011}
-competition we observed no correlation at all between the ordering of structures submited by the
-groups and the real ranking compared to an experimental model \cite{Kryshtafovych2011}. The
-consequences of this goes beyond
-those five structures; the deeper meaning is that groups producing ensembles of structures are
-generating structures that are better than the ones they submit, but they do not know about it. In
-fact, when querrying different groups after the results were known, most groups agree that this is
-the case. Beyond the CASP problem, this reflects on the modelers ability to correctly rank order
-models in many different environments, such as structures for drug design leads or designing more
-stable proteins or peptide mimetics such as peptoids. One of the main culprits of this lack of
-accuracy is the fact that ranking is often done via a potential energy function, which in many cases
-lacks an entropy component. Other initiatives including knowledge based potentials do use some sort
-of free energy to rank order, but its accuracy is not enough. Our method provides a physics based
-solution to this problem. We have tried two experiments centered on CASP. First we tried to rank
-order some structures from the previous CASP9 experiment. We have tested different options for that purpose.
-First, we rank order submitions from same group and then from different groups. Our second test was 
-to check whether this method can do better quality assessment/metaprediction than the best group in that catagory in CASP9. 
-In both methods, our ranking is determined as a free energy measure,
-and it is compared with the ranking given by a geometrical comparison between native and submitted
-models called GDT\_TS (Global distance test score \cite{Zemla2003}). 
-GDT\_TS represents the average percentage of residues that are in close proximity in two
-structures optimally superimposed using four different distance cutoffs (1, 2, 4 and 8 \AA).
+We have applied the confinement method as a ``meta-predictor'' for structure prediction. Here, the
+task is to correctly identify the most correct models out of a set of ``decoys'' generated by
+different methods during the Critical Assessment of Structure Prediction (CASP) experiment. CASP is
+a blind test in which different groups apply methods to predict the 3-dimensional structure of
+proteins from their sequences. This is done with a either a three week (for humans) or three day
+(for fully automated servers) time limit on each target. Each group is allowed to submit five
+possible structures, which they are supposed to rank from best to worst. We have performed two
+experiments centered on CASP.
 
-We have chosen the inital models on which to test the methods based on server groups that have
-traditionally done good in past CASP events. There is a high correlation between the structures that
-are chosen for this method and its predictive accuracy. In particular, when GDT\_TS scores are below
-50, it is difficult for us to say anything about the models. Our approach is simple, for every
-subset of structures we choose, we use the confinement method in a pairwise procedure, and then rank
-the structures. In all cases, we have also calculated the free energy with respect to the native
-structure to check the stability of the models.
-However, this step is not needed if we just want to rank order.
+In the first experiment, we examined several targets and tried to rank-order the predictions
+produced by other groups. In some cases the structures all came from the same group, in others the
+structures were drawn from several different groups. The goal of this experiment is to determine if
+the physics-based confinement method can correctly identify more native-like structures as having
+low free energies.
 
-\subsubsection{The confinement method can rank different models generated by the same methodology and distinguish the native structure}
+Our second experiment was to see if the confinement method can identify structures that are missed
+by other meta-prediction servers. Most successful meta-prediction servers are based on the idea of
+consensus: if many different prediction methods produce similar results, then that is probably a
+correct prediction. This is often a powerful heuristic, but it can miss cases where there is a very
+good result that is only predicted by one method. We chose several cases where such structures were
+missed and assessed if the confinement method can correctly identify these accurate models.
+
+As is common in the CASP experiment, we assess our results in terms of Global Distance Test Total
+Score (GDT-TS) \cite{Zemla2003}, which is a C$\alpha$ based measure of structural accuracy. It can
+be understood roughly as the percentage of residues that are correctly positioned in the model
+(higher is better, range 0 to 100).
+
+We have chosen the initial models on which to test the methods based on server groups that have
+traditionally done well in past CASP events.
+
+\subsubsection{The confinement method can correctly rank different models generated by the same methodology
+and distinguish the native structure}
 
 The question that we want to answer here is whether we can rank models that have been generated with
-the same methodology. In all cases, we have calculated the free energy
-difference between the native structure and the submitted models as well. 
+the same methodology. In all cases, we have calculated the free energy difference between the native
+structure and the submitted models as well. 
 %Identifying the native state from a set of models is a very easy test, that most metapredictors
@@
 2L01 and T0559 respectively. The native structure of this 69 residue protein was solved using NMR.
-The best predictor group for this target was "BAKER-ROSETTASERVER".
-We initially checked similarity between the models submitted by "BAKER-ROSETTASERVER" and discarded two of them from the 
-analysis on the basis of being too similar to some of the other models. The GDT\_TS score and rmsd
-values as shown in Figure~\ref{fig:T0559} indicate that model 1 was predicted correctly, 
-whereas the order of model 3 and 5 was wrong. The main difference
-between model 3 and the rest of the models is that the orientation of the first alpha helix of model
-3 is \Alberto{opposite??}.
-On the other hand, as shown in Figure~\ref{fig:T0559} the confine and release method not only can differentiate the native 
+The best predictor group for this target was "BAKER-ROSETTASERVER". We initially checked similarity
+between the models submitted by "BAKER-ROSETTASERVER" and discarded two of them from the analysis on
+the basis of being too similar to some of the other models. The GDT\_TS score and rmsd values as
+shown in Figure~\ref{fig:T0559} indicate that model 1 was predicted correctly, whereas the order of
+model 3 and 5 was wrong. The main difference between model 3 and the rest of the models is that the
+orientation of the first alpha helix of model 3 is \Alberto{opposite??}. On the other hand, as shown
+in Figure~\ref{fig:T0559} the confine and release method not only can differentiate the native
 structure from the submitted models, the ranking also correlates perfectly with the GDT\_TS score.
 
-To further test the method we have taken the example of protein BT2368 from Bacteroides thetaiotaomicron. The PDB id of this 
-74 amino acid residue protein is 2L02 and CASP target code is T0560. We have compared the free energy difference between 
-the native structure and the two of the five submitted models from the group "Splicer". The remaining three models are discarded as 
-again they are similar to the rest of the models. The comparison of GDT\_TS score as shown in 
-Figure~\ref{fig:T0560} and published in the final CASP result is only for models with residues 3 to 66. In order to keep
-consistency, we have also done our analysis with models and crystal structure consisting residues 3 to 66. As expected,
-the native state is identified correctly, and the two other models are ranked in the correct order in accordance with the GDT\_TS score.
+To further test the method we have taken the example of protein BT2368 from Bacteroides
+thetaiotaomicron. The PDB id of this 74 amino acid residue protein is 2L02 and CASP target code is
+T0560. We have compared the free energy difference between the native structure and the two of the
+five submitted models from the group "Splicer". The remaining three models are discarded as again
+they are similar to the rest of the models. The comparison of GDT\_TS score as shown in
+Figure~\ref{fig:T0560} and published in the final CASP result is only for models with residues 3 to
+66. In order to keep consistency, we have also done our analysis with models and crystal structure
+consisting residues 3 to 66. As expected, the native state is identified correctly, and the two
+other models are ranked in the correct order in accordance with the GDT\_TS score.
 
@@
 Our next test was between models that are produced with different methodologies. We do this by
-looking at some of the best CASP submissions from different research groups. 
-For this purpose we have chosen the x-ray crystallographic structure of 
-fas apoptosis inhibitory protein molecule whose pdb id and CASP target codes are 3MX7 and T0540 respectively.
-This protein contain 90 amino acid residues with 8 beta strands. We have chosen best models from groups
-"LTB" and "Mufold" for our analysis, which are labelled as Model 1 and Model 2 in Figure~\ref{fig:T0540}.
-As presented in  Figure~\ref{fig:T0540} we found that this method is able to correctly order the models which matches
+looking at some of the best CASP submissions from different research groups. For this purpose we
+have chosen the x-ray crystallographic structure of fas apoptosis inhibitory protein molecule whose
+pdb id and CASP target codes are 3MX7 and T0540 respectively. This protein contain 90 amino acid
+residues with 8 beta strands. We have chosen best models from groups "LTB" and "Mufold" for our
+analysis, which are labelled as Model 1 and Model 2 in Figure~\ref{fig:T0540}. As presented in
+Figure~\ref{fig:T0540} we found that this method is able to correctly order the models which matches
 with the GDT\_TS score.
@@
 between this NMR structure and the best CASP9 predicted model (from the 'Mufold' group, GDT\_TS=78).
-Visualization upon superimposition of two structures reveals that
-the difference between the two structure is at the region consist of residues 48 to 65, where model1 is much more disordered compare to the
-native structure(see Figure~\ref{fig:T0569}). Other regions look similar for both the cases with of model1 is within 2.6 \AA backbone
-rmsd value of the native structure. We found that, the confinement method can still differentiate between the native structure and 
-the generated model. We have also calculated the per residue free energy with a aim that this can identify the residue which are responsible
-for such free enerfy difference. The per residue free energy decomposition is shown in  Figure~\ref{fig:T0569_per_residue}(A). We indeed 
-found that two hydrophobic residues Val-59 and Ile-61 are oriented towards the solvent in the protein 
-surface in the generated model (Figure~\ref{fig:T0569_per_residue}(B). This the region, where there is a beta sheet in the original crystal structure, which is absent in the 
-model. Interestinlgy, some regions are more stable in the generated model. A close inspection
-reveals that 
-the hydrophilic Glu in position 67 is completely exposed to the protein surface and solvent in the best model, whereas it is partially exposed in the 
-crystal (Figure~\ref{fig:T0569_per_residue}(C)). The peak at position 76 is due to the H-bond between Lys-76 and Asp-11, 
-which is absent in the original crystalographic structure (Figure~\ref{fig:T0569_per_residue}(D)).
+Visualization upon superimposition of two structures reveals that the difference between the two
+structure is at the region consist of residues 48 to 65, where model1 is much more disordered
+compare to the native structure(see Figure~\ref{fig:T0569}). Other regions look similar for both the
+cases with of model1 is within 2.6 \AA backbone rmsd value of the native structure. We found that,
+the confinement method can still differentiate between the native structure and the generated model.
+We have also calculated the per residue free energy with a aim that this can identify the residue
+which are responsible for such free enerfy difference. The per residue free energy decomposition is
+shown in  Figure~\ref{fig:T0569_per_residue}(A). We indeed found that two hydrophobic residues
+Val-59 and Ile-61 are oriented towards the solvent in the protein surface in the generated model
+(Figure~\ref{fig:T0569_per_residue}(B). This the region, where there is a beta sheet in the original
+crystal structure, which is absent in the model. Interestinlgy, some regions are more stable in the
+generated model. A close inspection reveals that the hydrophilic Glu in position 67 is completely
+exposed to the protein surface and solvent in the best model, whereas it is partially exposed in the
+crystal (Figure~\ref{fig:T0569_per_residue}(C)). The peak at position 76 is due to the H-bond
+between Lys-76 and Asp-11, which is absent in the original crystalographic structure
+(Figure~\ref{fig:T0569_per_residue}(D)).
 
@@
 Despite the great success in most of the studied systems, we found a few failures, specially when
-the GDT scores of the compared structures are
-very close. We have found this for an engineered protein from Asr4154 protein (PDB ID: 2L09 and CASP Target T0538).
-This protein contains 54 amino acid residues with three alpha helixes and two very short beta
-strands. 
-The model that is closest to the native structure was generated by the group PconsR with a GDT\_TS value of 96.23.
-We have also considered one model from group "Shell" (GDT\_TS = 90.09) and "FOLDIT" (GDT\_TS = 86.32) for our analysis along with
-the native crystal struture (RMSD values between 1.6 and 2.0\AA \Alberto{from native, right?all
-atom? CA?}). The models from PconsR, Shell and FOLDIT are labelled as model 1, model 2 and model 3 
-respectively in Figure~\ref{fig:T0538}. As seen in Figure~\ref{fig:T0538}, one can find that the 
-model from PconsR has lowest free energy value, making it a more stable structure. Although ranking of the other models
-remains consistent, model 1 is predicted to be more stable than the
-crystal structure. To rationalize this, we have decomposed the total free energy into the per
+the GDT scores of the compared structures are very close. We have found this for an engineered
+protein from Asr4154 protein (PDB ID: 2L09 and CASP Target T0538). This protein contains 54 amino
+acid residues with three alpha helixes and two very short beta strands. The model that is closest to
+the native structure was generated by the group PconsR with a GDT\_TS value of 96.23. We have also
+considered one model from group "Shell" (GDT\_TS = 90.09) and "FOLDIT" (GDT\_TS = 86.32) for our
+analysis along with the native crystal struture (RMSD values between 1.6 and 2.0\AA \Alberto{from
+native, right?all atom? CA?}). The models from PconsR, Shell and FOLDIT are labelled as model 1,
+model 2 and model 3 respectively in Figure~\ref{fig:T0538}. As seen in Figure~\ref{fig:T0538}, one
+can find that the model from PconsR has lowest free energy value, making it a more stable structure.
+Although ranking of the other models remains consistent, model 1 is predicted to be more stable than
+the crystal structure. To rationalize this, we have decomposed the total free energy into the per
 residue components. The results show us that despite the small variation at the backbone level (as
 shown by high GDT\_TS scores and low RMSDs), the sidechains are oriented in very different ways,
-giving raise to large differences in the stabilization of certain residues.
-In particular, some of the differences arise from different salt bridge patterns (Arg-32 with Glu-35
-and Glu-28 with Lys-24
+giving raise to large differences in the stabilization of certain residues. In particular, some of
+the differences arise from different salt bridge patterns (Arg-32 with Glu-35 and Glu-28 with Lys-24
 in crystal versus Arg-32 with Glu-28 in model 1) and certain flexible polar residues exposed to the
@@
 Similar patterns arise in other places of the protein (see Figure ~\ref{fig:T0538compare}(B)), where
-the salt bridge between Arg-26 and Glu-50 is absent in model 1 and is further \Alberto{stabilized?} by Phe-51
-having a very different conformation that in the crystal structure.
-All in all, this unexpected result just shows us that this method is sensitive to local interactions
-such as those happening from side chain reorientation. At the same time, the results explained in
-past sections showed that for larger changes the accuracy of the method is very high.
+the salt bridge between Arg-26 and Glu-50 is absent in model 1 and is further \Alberto{stabilized?}
+by Phe-51 having a very different conformation that in the crystal structure. All in all, this
+unexpected result just shows us that this method is sensitive to local interactions such as those
+happening from side chain reorientation. At the same time, the results explained in past sections
+showed that for larger changes the accuracy of the method is very high.
 
@@
 happens when we use the method to describe low quality models. We use one of the CASP targets (pdb
-id = 2KJX, CASP id = T0531) .
-We mentioned earlier that our predictive abilities are greatly decreased with low model qualities. 
-Comparing low qualitive models between themselves is not very informative,
-    however, comparing all of them to a native/native like structure, allows to rank order this
-structures. 
-In order to test this hypothesis we performed another test with target T0531 of CASP9, in which we compared the crystal 
-structure to five models from the MUFOLD server. Pdb id of this extracellular domain of the jumping translocation 
-breakpoint protein is 2KJX and CASP target code is T0531.
-Looking at Figure~\ref{fig:T0531} shows: 1.) native is correctly
-identified as expected and 2.) Surprisingly there is a high level of correlation between the GDT
-ranking and the free energy ranking for model 1 and model 3, the rest three structures with GDT\_TS score less than  35  are ordered
-incorrectly. It is worth to note that models 2 and 3 have the same GDT and very different free
-energies, meaning that the actual ordering could change a lot \cite{Perez2012}. It is encouraging that at least
-the method can pick out the best model even though it is got a low GDT score: 44.
-
-%\subsubsection{Difficulty in CASP Experiments}
-
-%It is often difficult to model CASP targets using the confinement and release method. There are many cases,
-%where the sequence of a particular target is trimmed during actual result presentation. This is illustrated
-%with the crystal structure of the leucine zipper domain of cGMP dependent protein kinase I (pdb id: 3NMD
-%and CASP9 code: T0605) as shown in Figure~\ref{fig:T0605}. The actual CASP 9 target consisted of 72 amino acid residues, but
-%as only 18-66 amino acid residues could be detected in actual crystal structure of this protein, the final result was trimmed 
-%with respect to that crystal structure with 49 amino acid residues which consist of a single alpha helix. The amino acid residues 18 and 66 are shown in
-%three models submitted by the group "Baker". It is important to note that blind prediction of model ranking on the basis of free energy can be hugely 
-%affected in such scenario as the only major difference between those models are the trimmed parts, specially the region consisting residues 1-18.
+id = 2KJX, CASP id = T0531) . We mentioned earlier that our predictive abilities are greatly
+decreased with low model qualities. Comparing low qualitive models between themselves is not very
+informative, however, comparing all of them to a native/native like structure, allows to rank order
+this structures. In order to test this hypothesis we performed another test with target T0531 of
+CASP9, in which we compared the crystal structure to five models from the MUFOLD server. Pdb id of
+this extracellular domain of the jumping translocation breakpoint protein is 2KJX and CASP target
+code is T0531. Looking at Figure~\ref{fig:T0531} shows: 1.) native is correctly identified as
+expected and 2.) Surprisingly there is a high level of correlation between the GDT ranking and the
+free energy ranking for model 1 and model 3, the rest three structures with GDT\_TS score less than
+35  are ordered incorrectly. It is worth to note that models 2 and 3 have the same GDT and very
+different free energies, meaning that the actual ordering could change a lot \cite{Perez2012}. It is
+encouraging that at least the method can pick out the best model even though it is got a low GDT
+score: 44.
 
 \subsubsection{Can the confinement method perform better quality assessment in protein structure prediction?}
-A part of the CASP experiment is dedicated to the assessment of the quality of predicted models \cite{Kryshtafovych2011}. 
-Most of the top performing groups in this category use consensus approaches for quality assessment \cite{Wang2011}.
-A review of CASP9 quality assessment also pointed out that the methods which are based on clustering
-techniques perform
-much better compared to those methods which are based on analysis of individual models \cite{Kryshtafovych2011}. 
-In this background we try to answer, whether the confinement method can do better quality assessment
-compared to the best performing
-group (MUFOLD-WQA) in this catagory. In CASP9, the overall reliability of models in quality assessment (QA) was accepted in a mode known as
-QMODE 1. Here, predictors were asked to score each model on a scale from 0 to 1, with higher values corresponding to better models \cite{Kryshtafovych2011}. 
-We have again picked up the CASP taget T0538 to answer this question. We have chosen Model 3 submitted by PconsR (GDT\_TS = 96, QMODE 1 = 0.5434) 
-and model 5 from the MULTICOM-NOVEL ( GDT\_TS = 83, QMODE 1 = 0.5865). Both are server predicted models and they were further used for quality assessment 
-by the group MUFOLD-WQA \cite{Wang2011}. It is important to note that the model from PconsR was the most accurate predicted model for this 
-target and the model from MULTICOM-NOVEL was predicted as best by the quality assessment experiments of MUFOLD-WQA. Also, 
-the QMODE 1 value presented here is from the group MUFOLD-WQA. 
-Our free energy analysis predict that the model from Pconsr is indeed better than the model from 
-MULTICOM-NOVEL $(\Delta \Delta G (Model\_Pconsr - Model\_MULTICOM-NOVEL = -3.9 Kcal/Mol)$, which   
-matches well with the GDT\_TS score. On the other hand the MUFOLD-WQA predicted the ranking in the reverse order.
 
-We have also carried out a similar type of analysis for the case of Target 560 from the CASP9 experiment. For that 
-we have chosen the model 1 from the group PconsR (GDT\_TS = 94, QMODE 1 = 0.5178), Bilab\-Enable (GDT\_TS = 87, QMODE 1 = 0.5344) 
-and Multicom\_Refine (GDT\_TS = 74, QMODE 1 = 0.4770). If, one wants to rank order with respect to the GDT\_TS values, the model from 
-PconsR is the best predicted model, but according to the QA assessment by MUFOLD-WQA, the model from Bilab-Enable was found to be best predicted model. 
-All the QMODE 1 values are from MUFOLD\-WQA. In CASP 9, the sequence of 74 residues were given to predict the structure and for 
-quality assessment, whereas during GDT\_TS analysis only residues 3-66 were considered. When we considered the full 74 amino acid 
-residues, the free energy calculated were $(\Delta \Delta G (Model\_Pconsr - Model\_Bilab\-Enable = 3.3 Kcal/Mol)$ and 
-$(\Delta \Delta G (Model\_Pconsr - Model\_Multicom\_Refine = -30.0 Kcal/Mol)$ \Alberto{-30 sounds
-huge!} which 
-support the QA assessment by MUFOLD-WQA. On the other hand, if we calculate the same free energy with the trimmed version of the protein 
-(with residues 3-66), the values of the free energy difference become -1.7 and -31.1 Kcal/Mol respectively, which support the GDT\_TS score. 
-The summary of this whole observation is presented in Table~\ref{table:560QA}.
-There is no doubt that the consensus approach can predict the model quality in a faster manner. But
-we expect that the confinement
-method can predict the model quality in a relatively expensive but much more accurate way.
+A part of the CASP experiment is dedicated to the assessment of the quality of predicted models
+\cite{Kryshtafovych2011}. Most of the top performing groups in this category use consensus
+approaches for quality assessment \cite{Wang2011}. A review of CASP9 quality assessment also pointed
+out that the methods which are based on clustering techniques perform much better compared to those
+methods which are based on analysis of individual models \cite{Kryshtafovych2011}. In this
+background we try to answer, whether the confinement method can do better quality assessment
+compared to the best performing group (MUFOLD-WQA) in this catagory. In CASP9, the overall
+reliability of models in quality assessment (QA) was accepted in a mode known as QMODE 1. Here,
+predictors were asked to score each model on a scale from 0 to 1, with higher values corresponding
+to better models \cite{Kryshtafovych2011}. We have again picked up the CASP taget T0538 to answer
+this question. We have chosen Model 3 submitted by PconsR (GDT\_TS = 96, QMODE 1 = 0.5434) and model
+5 from the MULTICOM-NOVEL ( GDT\_TS = 83, QMODE 1 = 0.5865). Both are server predicted models and
+they were further used for quality assessment by the group MUFOLD-WQA \cite{Wang2011}. It is
+important to note that the model from PconsR was the most accurate predicted model for this target
+and the model from MULTICOM-NOVEL was predicted as best by the quality assessment experiments of
+MUFOLD-WQA. Also, the QMODE 1 value presented here is from the group MUFOLD-WQA. Our free energy
+analysis predict that the model from Pconsr is indeed better than the model from MULTICOM-NOVEL
+$(\Delta \Delta G (Model\_Pconsr - Model\_MULTICOM-NOVEL = -3.9 Kcal/Mol)$, which   matches well
+with the GDT\_TS score. On the other hand the MUFOLD-WQA predicted the ranking in the reverse order.
+
+We have also carried out a similar type of analysis for the case of Target 560 from the CASP9
+experiment. For that we have chosen the model 1 from the group PconsR (GDT\_TS = 94, QMODE 1 =
+0.5178), Bilab\-Enable (GDT\_TS = 87, QMODE 1 = 0.5344) and Multicom\_Refine (GDT\_TS = 74, QMODE 1
+= 0.4770). If, one wants to rank order with respect to the GDT\_TS values, the model from PconsR is
+the best predicted model, but according to the QA assessment by MUFOLD-WQA, the model from
+Bilab-Enable was found to be best predicted model. All the QMODE 1 values are from MUFOLD\-WQA. In
+CASP 9, the sequence of 74 residues were given to predict the structure and for quality assessment,
+whereas during GDT\_TS analysis only residues 3-66 were considered. When we considered the full 74
+amino acid residues, the free energy calculated were $(\Delta \Delta G (Model\_Pconsr -
+Model\_Bilab\-Enable = 3.3 Kcal/Mol)$ and $(\Delta \Delta G (Model\_Pconsr - Model\_Multicom\_Refine
+= -30.0 Kcal/Mol)$ \Alberto{-30 sounds huge!} which support the QA assessment by MUFOLD-WQA. On the
+other hand, if we calculate the same free energy with the trimmed version of the protein (with
+residues 3-66), the values of the free energy difference become -1.7 and -31.1 Kcal/Mol
+respectively, which support the GDT\_TS score. The summary of this whole observation is presented in
+Table~\ref{table:560QA}. There is no doubt that the consensus approach can predict the model quality
+in a faster manner. But we expect that the confinement method can predict the model quality in a
+relatively expensive but much more accurate way.
 
@@
 previous paragraph}
+
 In the recent past there have been a number of studies directed towards understanding protein
-structure using theoretical tools. Experimentally, it is relatively easy to identify the native structure from the data obtained 
-from x-ray crystallography or NMR. On the other hand using theoretical tools it is often difficult to identify the native structure 
-using both physics based methods and knowledge based method. In order to rank structure from the available data, some groups use 
-clustering algorithms and the structure of the
-most populated cluster is picked out as the most probable structure. However, it may happen that the structure may be trapped and 
-spend most of its time in a kinetic trap ***kinetic traps in MD,not other methods, local
-minima***, which may give a wrong interpretation of the clustering result. On the other hand, some groups
-also depend on potential energy functions to pick out their best structure, lacking an entropic
-component. In this work we attempted to identify the native/native like protein strcuture from the decoy using free energy as a scale.
-We have used the previously developed confinement method, applying it to larger systems.
+structure using theoretical tools. Experimentally, it is relatively easy to identify the native
+structure from the data obtained from x-ray crystallography or NMR. On the other hand using
+theoretical tools it is often difficult to identify the native structure using both physics based
+methods and knowledge based method. In order to rank structure from the available data, some groups
+use clustering algorithms and the structure of the most populated cluster is picked out as the most
+probable structure. However, it may happen that the structure may be trapped and spend most of its
+time in a kinetic trap ***kinetic traps in MD,not other methods, local minima***, which may give a
+wrong interpretation of the clustering result. On the other hand, some groups also depend on
+potential energy functions to pick out their best structure, lacking an entropic component. In this
+work we attempted to identify the native/native like protein strcuture from the decoy using free
+energy as a scale. We have used the previously developed confinement method, applying it to larger
+systems.
 
 We found that in majority of cases, the native structure is always the most stable and we can rank
-protein structures upto 100 amino 
-acid residues long. This method works very well if at least one of the strcuture in the decoy set
-has high quality. The main advantage of this method is 
-that it does not require any reaction coordinate. Thus one can calculate the free energy difference between two very different conformations. 
-An interesting application from this method that we have exploited is the partition of free energies
-on a per residue basis.
- Although, it is an approximate partition, it gives important insight regarding the conformational 
-preferences of a particular residue. The use of this decomposition in understanding the effect of a
-small number of aminoacids in our test chamaleon sequences allows us to think that in the future,
-this method may be used for protein design. 
-The method can be also give wrong result if all the models in the decoy are very poor. It may also be difficult to rank structures if 
-all the structures are very very close to the native structure. But we want to ignore this fact as our main aim is to differentiate
-between the good and the bad structure. In this article we have extensively tested this method using examples from CASP 
-This method can be computationally expensive in actual CASP experiment if one want rank hundreds of submitted models in CASP 
-experiments. However, in real life experiment with a particular target, we expect this method can give accurate result. The important 
-fact about confinement method is that all the part of the calculation can be done independently. 
-There is another big bottlenecks in this calculation. This has to do with the amount of
-simulations that have to be performed to confine the ensemble of conformations close to a microstate
-into a single microstate. We have tackled this issue by using GPU (Graphical Processing Unit) technology as opposed to the classical
-CPUs, which gives us a two order of magnitude boost in computational efficiency.
-For example, we have mostly carried out our calculation with Amber program \cite{Case2005} in GPU computer. With a avarage 56 residue 
-we found that it take only 4 hour to complete 20 ns of the confinement step. So, with available computer power this method will 
-be easy to use for real problems. 
-%There is another issue, which is normal mode
-%analysis calculations are Oh(N**3) and require large amounts of memory, effectively limiting the
-%size of the systems under study. In most cases quasiharmonic analysis.
+protein structures upto 100 amino acid residues long. This method works very well if at least one of
+the strcuture in the decoy set has high quality. The main advantage of this method is that it does
+not require any reaction coordinate. Thus one can calculate the free energy difference between two
+very different conformations. An interesting application from this method that we have exploited is
+the partition of free energies on a per residue basis. Although, it is an approximate partition, it
+gives important insight regarding the conformational preferences of a particular residue. The use of
+this decomposition in understanding the effect of a small number of aminoacids in our test chamaleon
+sequences allows us to think that in the future, this method may be used for protein design. The
+method can be also give wrong result if all the models in the decoy are very poor. It may also be
+difficult to rank structures if all the structures are very very close to the native structure. But
+we want to ignore this fact as our main aim is to differentiate between the good and the bad
+structure. In this article we have extensively tested this method using examples from CASP This
+method can be computationally expensive in actual CASP experiment if one want rank hundreds of
+submitted models in CASP experiments. However, in real life experiment with a particular target, we
+expect this method can give accurate result. The important fact about confinement method is that all
+the part of the calculation can be done independently. There is another big bottlenecks in this
+calculation. This has to do with the amount of simulations that have to be performed to confine the
+ensemble of conformations close to a microstate into a single microstate. We have tackled this issue
+by using GPU (Graphical Processing Unit) technology as opposed to the classical CPUs, which gives us
+a two order of magnitude boost in computational efficiency. For example, we have mostly carried out
+our calculation with Amber program \cite{Case2005} in GPU computer. With a avarage 56 residue we
+found that it take only 4 hour to complete 20 ns of the confinement step. So, with available
+computer power this method will be easy to use for real problems. 
+%There is another issue, which is normal mode analysis calculations are Oh(N**3) and require large
+%amounts of memory, effectively limiting the size of the systems under study. In most cases
+%quasiharmonic analysis.
 
-This method can be further used for calculating free energy due to very large conformational change. Another application may be
-the case of ligand binding. It can also be used to distinguish the most stable structure during protein design. Decompition of 
-total free energy into the per residue component help identifying residues that give stability to a particular conformation. 
-Another, important application may be to study the protein folding kinetics, specially to study the phi value analysis.
+This method can be further used for calculating free energy due to very large conformational change.
+Another application may be the case of ligand binding. It can also be used to distinguish the most
+stable structure during protein design. Decompition of total free energy into the per residue
+component help identifying residues that give stability to a particular conformation. Another,
+important application may be to study the protein folding kinetics, specially to study the phi value
+analysis.
 
@@
 The confinement method has been described in details in ref. by Tyka et.al. \cite{Tyka2006} and
-Cecchini et. al. \cite{Cecchini2009}. The basic approach of the confinement approach is the same in both
-these papers. However, there are some technical differences. Here we briefly describe the procedure
-to compute free energy $\Delta G_{AB}$ between two conformations A and B of a protein. The basic
-idea to compute the free energy between conformations A and B is to perform a thermodynamic cycle:
+Cecchini et. al. \cite{Cecchini2009}. The basic approach of the confinement approach is the same in
+both these papers. However, there are some technical differences. Here we briefly describe the
+procedure to compute free energy $\Delta G_{AB}$ between two conformations A and B of a protein. The
+basic idea to compute the free energy between conformations A and B is to perform a thermodynamic
+cycle:
 
@@
 
- All calculations where performed with the amber 11 suit of programs in combination with ff99SB and GB/SA implicit solvent. Interestingly, we extend the
-method for calculation of per residue free energy in an approximate way. For this purpose, the confinement energy, 
-$\Delta G_{A,A*}$ and $\Delta G_{B,B*}$ of each residue is calculated in the usual numerical way as described above. The
-internal energy of each residue is calculated using the decomp module of amber from the final restrained trajectory. We 
-call this method approximate as we ignore the contribution from the normal mode or quasiharmonic analysis. 
-However, this contrbution to the total free energy is much smaller (less than $1.0
-Kcal/Mol$\Alberto{that sounds huge, i would say what percent of error we stimate, I think it was
-lower than 1kcal/mol}), which allow us to 
-study the mechanistic details of conformational preference of each residue.
+All calculations where performed with the amber 11 suit of programs in combination with ff99SB and
+GB/SA implicit solvent. Interestingly, we extend the method for calculation of per residue free
+energy in an approximate way. For this purpose, the confinement energy, $\Delta G_{A,A*}$ and
+$\Delta G_{B,B*}$ of each residue is calculated in the usual numerical way as described above. The
+internal energy of each residue is calculated using the decomp module of amber from the final
+restrained trajectory. We call this method approximate as we ignore the contribution from the normal
+mode or quasiharmonic analysis. However, this contrbution to the total free energy is much smaller
+(less than $1.0 Kcal/Mol$\Alberto{that sounds huge, i would say what percent of error we stimate, I
+think it was lower than 1kcal/mol}), which allow us to study the mechanistic details of
+conformational preference of each residue.
 
 %There are two big bottlenecks in this calculation. The first has to do with the amount of
-%simulations that have to be performed to confine the ensemble of conformations close to a microstate
-%into a single microstate, the second bottleneck has to do with the size of the system in the normal
-%mode calculation. We have tackled the first issue by using GPU (Graphical Processing Unit) technology as opposed to the classical
-%CPUs, which gives us a two order of magnitude boost in computational efficiency. Normal mode
-%analysis calculations are Oh(N**3) and require large amounts of memory, effectively limiting the
-%size of the systems under study. Where as this can be overcome by larger memory machines, these are
-%often not available, and so we propose an alternative by using quasiharmonic analysis. This is still
-%an Oh(N**3) problem, but we now the motions of atoms are already determined in trajectories, and so
-%we can choose to do quasiharmonic analysis excluding hydrogens, which will reduce the number of
-%atoms N to almost N/2. This means that we can use this method for systems twice as large.
+%simulations that have to be performed to confine the ensemble of conformations close to a
+%microstate into a single microstate, the second bottleneck has to do with the size of the system in
+%the normal mode calculation. We have tackled the first issue by using GPU (Graphical Processing
+%Unit) technology as opposed to the classical CPUs, which gives us a two order of magnitude boost in
+%computational efficiency. Normal mode analysis calculations are Oh(N**3) and require large amounts
+%of memory, effectively limiting the size of the systems under study. Where as this can be overcome
+%by larger memory machines, these are often not available, and so we propose an alternative by using
+%quasiharmonic analysis. This is still an Oh(N**3) problem, but we now the motions of atoms are
+%already determined in trajectories, and so we can choose to do quasiharmonic analysis excluding
+%hydrogens, which will reduce the number of atoms N to almost N/2. This means that we can use this
+%method for systems twice as large.
 
@@
 \end{figure}
-
-%\begin{figure}
-%\begin{center}
-%\includegraphics[width=12cm,height=10cm]{T0605.pdf}
-%\end{center}
-%\caption{ Three models of the leucine zipper domain of cGMP dependent protein
-%kinase I (original pdb id: 3NMD and CASP target: T0605). The original target sequence of this protein was 72 residues.
-%However only residues 3 to 66 is only kept during result announcement as that part can be found from crystallographic analysis. The
-%region 3 to 66 are almost identical in all the 3 models  }
-%\label{fig:T0605}
-%\end{figure}
 

### 2013-03-06 Paper/paper.tex

@@
     GPU technology and the rate at which computers improve, this computational expense is becoming
-less of a bottle neck.}
+
+less of a bottle neck.
+
 \Alberto{I don't agree with some of the points here, but I think I covered what you meant in the
@@
 analysis.
+
+less of a bottle neck.
+We foresee great potential for this methodology for two reasons: independence from a reaction
+coordinate (just need start and end state) and avility to decompose the free energies on a per
+residues basis. This last point can be very useful in protein design. Finally, we have presented
+data for structure comparison of proteins, but we are currently exploring the possibility of
+using this methodology for getting free energies of ligand-protein binding.
+
+\Alberto{Do we mention this in the text? too technical for conclusions:
+There is another issue, which is normal mode
+analysis calculations are Oh(N**3) and require large amounts of memory, effectively limiting the
+size of the systems under study. In most cases quasiharmonic analysis.
+}
 

### 2013-03-06 Paper/paper.tex
Made changes to introduction

@@
 
-Protein conformation and associated conformational changes are important in biology. The mechanism
-of enzymes, the function of structural proteins, and the binding of proteins to DNA or other
-proteins all depend on protein conformations and the relative free energies of different
-conformational states. Proteins can undergo conformational changes for a variety of reasons,
-including ligand binding, change of pH or other conditions \cite{Meirovitch2007}, \cite{Chipot2007},
-\cite{Jorgensen2004}. The process of protein folding also depends on the free energies of different
-conformational states. The native structure is at a free energy minimum \cite{Anfinsen1973}, while
-the kinetics are determined by the free energies of various on- and off-pathway intermediates.
+Protein structure and conformational changes are of central importance in biochemistry. The
+populations of different conformational states are determined by their free energies. The ability to
+compute these free energies would allow us to understand and make predictions about a wide variety
+of biochemical processes---including the mechanism of enzymes, the function of structural proteins,
+binding of proteins to other proteins or DNA, and protein folding.
 
-However, the theoretical calculation of conformational free energy changes is difficult when the
-states of interest are very different \cite{Meirovitch2007}. In these cases, the timescales involved
-in such conformational transitions may be too long to simulate directly. For example, in protein
-folding the free energy landscape is rough \cite{Dill1997}, \cite{Dill2008}, and the molecule can
-become trapped in an energy well during conformational sampling. Even with advanced methods like
-replica exchange molecular dynamics, the three dimensional structure may be trapped in a local
-minima for a considerable time. Sometimes this problem can be circumvented by defining a pathway or
-reaction coordinate connecting the states of interest. The free energy along this reaction
+Computer simulations have proven to be a powerful tool for studying biomolecular systems at atomic
+detail [ref]. However, calculating the free energies for conformational changes remains a
+challenging problem \cite{Meirovitch2007}, primarily because the timescales for such conformational
+transitions may be too long to simulate directly. For example, in protein folding the free energy
+landscape is very rough \cite{Dill1997,Dill2008}, which limits conformational sampling and leads to
+large uncertainties in the relative populations of different states.
+
+Even with advanced methods like replica exchange molecular dynamics [ref], the molecule may be
+become trapped in a local minima. Sometimes this problem may be circumvented by defining a pathway
+or reaction coordinate connecting the states of interest. The free energy along this reaction
 coordinate can then be determined using methods such as umbrella sampling \cite{Torrie1977},
 targeted MD \cite{}, or thermodynamic integration \cite{Tironi1994}. However, accurately sampling
-this reaction coordinate may still be difficult and require a large computationl effort. In the case
-of protein folding, it is often difficult to even devise a meaningful reaction coordinate that could
-connect the unfolded and folded states efficiently \cite{}.
+this reaction coordinate may still be difficult and require a large computational effort. In the
+case of protein folding, it is often difficult to even devise a meaningful reaction coordinate that
+could connect the unfolded and folded states efficiently \cite{}.
 
@@
 groups \cite{Meirovitch2007,Ytreberg2006,Zheng2008}. In recent years, methods such as the reference
-system method \cite{Ytreberg2006}, decativated morphing \cite{Park2008}, orthogonal space random
+system method \cite{Ytreberg2006}, deactivated morphing \cite{Park2008}, orthogonal space random
 walk \cite{Zheng2008}, and the confinement method \cite{Tyka2006,Cecchini2009} have been used to
@@
 
-In this work we explore confinement method\cite{Tyka2006,Cecchini2009}, which relies on the fact
+In this work we explore the confinement method\cite{Tyka2006,Cecchini2009}, which relies on the fact
 that the free energy difference is a state function and thus is independent of the sampling path.
-This approach uses a thermodynamic cycle \ref{fig:method} where first protein conformational entropy
-of the two end-states is frozen out by applying a series of restraints to the system. Next, the free
-energy difference between the two highly-restrained harmocic systems is obtained using a normal mode
-calculation.
+This allows us to use the alchemical thermodynamic cycle shown in Figure \ref{fig:method}. First,
+the two end states are slowly ``frozen'' into a very tight, harmonically restrained ensemble and the
+free energies for these two processes are calculated. Next, the free energy difference between the
+two restrained ensembles is calculated using normal mode analysis. This allows us to calculate the
+free energy difference between the two end states without needing to define a physical path or
+reaction coordinate connecting the two states.
 
-Previously, this method was applied to calculate the conformational free energy of a 16 amino acid
-residue peptide, known as BHP.  We have first reproduced that known result. Next, we have applied
-this method to larger proteins in two different case studies. In the first, we attempt to pick the
-most native like structures from a set of models produced during the Critical Assessment of
-Structure Prediction (CASP) \cite{} experiment. In the second case study, we examine the
-conformational preferences of a series of chameleon proteins designed by Orban and co-workers
+The confinement has previously been applied to several small proteins [refs]. Here, we use the power
+of GPU computing (using graphic cards to run calculations [cite amber gpu]), which allows us to
+extend the method to larger systems while maintaining reasonable computation times. We first
+demonstrate that our implementation reproduces the previously reported results for small peptides.
+Next, we have applied this method to larger proteins in two different case studies. In the first, we
+attempt to pick the most native-like structures from a set of models produced during the Critical
+Assessment of Structure Prediction (CASP) \cite{} experiment. In the second case study, we examine
+the conformational preferences of a series of chameleon proteins designed by Orban and co-workers
 \cite{} that can switch between two completely different folds with only small changes in amino acid

### 2013-04-24 Paper/paper.tex
correction_24thApril

--- 2013-03-06 Paper/paper.tex
@@
 \affil{Laufer Center for Physical and Quantitative Biology\\
+    and Departments of Physics and Chemistry\\
     Stony Brook University\\
@@
 
+\Ken{[Guys - I recommend we get a better name than confinement.  It's totally uninformative.  In my view, better would be `confine-and-configure' or something.  
+Please give it some thought.]}
+
 \section{Introduction}
 
-Protein conformation and associated conformational changes are important in biology. The mechanism
-of enzymes, the function of structural proteins, and the binding of proteins to DNA or other
-proteins all depend on protein conformations and the relative free energies of different
-conformational states. Proteins can undergo conformational changes for a variety of reasons,
-including ligand binding, change of pH or other conditions \cite{Meirovitch2007}, \cite{Chipot2007},
-\cite{Jorgensen2004}. The process of protein folding also depends on the free energies of different
-conformational states. The native structure is at a free energy minimum \cite{Anfinsen1973}, while
-the kinetics are determined by the free energies of various on- and off-pathway intermediates.
+In some problems of protein science, you want to know the relative stability of a protein's conformation $A$ compared to it's 
+conformation $B$.  We call this the \emph{difference free energy} (DFE). For example, in allosteric mechanisms, a protein often 
+adopts one conformation when a ligand is bound and another conformation when no ligand is bound \cite{Elber2007}. To understand the mechanism requires knowledge 
+of the relative free energies of the protein conformations.  Another example is in protein structure prediction, for example as practiced in 
+the CASP competition \cite{Moult2011}.  If you are using some computational model and have predicted two putative native structures, $A$ 
+and $B$, you want to compute which has the lower free energy, in order to know which is the more native-like.  In a third example, you may 
+want to know whether mutating a few amino acids in a protein could cause the protein to switch from one stable conformation to another, because 
+that could have important consequences for biological mechanisms and disease.  And fourth, sometimes binding a ligand induces a protein from 
+one conformation to another.  A quantitative understanding of such \emph{induced-fit} situations requires knowledge of the difference free 
+energy of the protein in the two states. In all these cases, it would be useful to have a computational method that can be used efficiently 
+with atomically detailed physical forcefield models to compute the free energy differences between different given protein conformations.
 
-However, the theoretical calculation of conformational free energy changes is difficult when the
-states of interest are very different \cite{Meirovitch2007}. In these cases, the timescales involved
-in such conformational transitions may be too long to simulate directly. For example, in protein
-folding the free energy landscape is rough \cite{Dill1997}, \cite{Dill2008}, and the molecule can
-become trapped in an energy well during conformational sampling. Even with advanced methods like
-replica exchange molecular dynamics, the three dimensional structure may be trapped in a local
-minima for a considerable time. Sometimes this problem can be circumvented by defining a pathway or
-reaction coordinate connecting the states of interest. The free energy along this reaction
-coordinate can then be determined using methods such as umbrella sampling \cite{Torrie1977},
-targeted MD \cite{}, or thermodynamic integration \cite{Tironi1994}. However, accurately sampling
-this reaction coordinate may still be difficult and require a large computationl effort. In the case
-of protein folding, it is often difficult to even devise a meaningful reaction coordinate that could
-connect the unfolded and folded states efficiently \cite{}.
+Methods like thermodynamic integration \cite{Tironi1994},\cite{Meirovitch2007} which are successful in alchemical free energy calculation
+may not be as useful for calculation of conformational free energy.
+One widely explored strategy is to use molecular dynamics simulations along some putative reaction coordinate pathway from conformation $A$ 
+to $B$ \cite{Elber2007,West2007,Haas2007,Jonsson1998,E2007,Dellago2002,Cheng2006}. \Ken{we need 
+many refs here: cite Elber's milestoning, Jhih-Wei Chu, and the many other path-sampling approaches.]} The free energy along this reaction 
+coordinate can then be determined using methods such as umbrella sampling \cite{Torrie1977}. Such approaches have several limitations.  
+First, it is necessary to know a reaction pathway from $A$ to $B$. If conformations $A$ and $B$ are quite different, then it can be 
+challenging to find such paths. Second, these methods 
+are computationally slow. To get an accurate estimate of the total free energy difference $\Delta G = G_B - G_A$ requires accurate determinations 
+of the many small free energy differences $A \rightarrow 1 \rightarrow 2 \ldots \rightarrow B$, so each microstep requires substantial amounts 
+of sampling.  And third, these methods are prone to large errors, because the pathway error is a sum of many errors among the many microsteps.
 
+\Arijit{Will it be useful to add methods like PMF or replica exchange based free energy methods ?}
+
+\Ken{[Are all these methods below ones that don't involve paths?  Also, are there other key methods we should cite?  We need a good scholarly 
+list of all the relevant works.]} 
 The calculation of protein conformational free energy has been successfully attempted by a number of
-groups \cite{Meirovitch2007,Ytreberg2006,Zheng2008}. In recent years, methods such as the reference
-system method \cite{Ytreberg2006}, decativated morphing \cite{Park2008}, orthogonal space random
-walk \cite{Zheng2008}, and the confinement method \cite{Tyka2006,Cecchini2009} have been used to
-calculate free energy differences.
+groups \cite{Shell2010,Ytreberg2006,Zheng2008,Spichty2010,Strajbl2000,Park2008,Tyka2006,Cecchini2009,Ovchinnikov2013}. 
+Some of these methods like reference system method \cite{Ytreberg2006}, deactivated morphing \cite{Park2008}, confinement 
+method \cite{Tyka2006,Cecchini2009,Ovchinnikov2013} take an alternative strategy for computing difference free energies 
+that does not require knowing a pathway from $A$ to $B$.  
+\Ken{[please add the several new Karplus papers from 2012-2013.]}.
 
-In this work we explore confinement method\cite{Tyka2006,Cecchini2009}, which relies on the fact
-that the free energy difference is a state function and thus is independent of the sampling path.
-This approach uses a thermodynamic cycle \ref{fig:method} where first protein conformational entropy
-of the two end-states is frozen out by applying a series of restraints to the system. Next, the free
-energy difference between the two highly-restrained harmocic systems is obtained using a normal mode
-calculation.
+Here, we adopt the confinement method of Tyka et al \cite{Tyka2006} and Cecchini et al.\cite{Cecchini2009}. This method 
+also resembles the `confine-and-release' method \cite{Mobley2007}, \cite{Mobley2012}, \cite{Mobley2006}.  
+\Ken{[cite mobley's jctc 07 (ref 216 on our labsite), and mobley's recent j chem phys review, 
+and mobley (ref 199) j chem phys, and others that are relevant.]}. The method is based on the thermodynamic cycle showns in 
+Figure \ref{fig:method}.  To start, conformation $A$ is an ensemble.  We reduce ensemble $A$ to one particular microstate $A^\ast$ 
+in that ensemble by applying restraints in an MD simulation.  In the same way, we reduce conformational ensemble $B$ to one of its 
+microstates, $B^\ast$.  Then, we simply compute the free energy difference between the microstates 
+as $\Delta G_{A^\ast, B^\ast} = \Delta H_{A^\ast, B^\ast}$ (which is just an enthalpy difference since there is no conformational 
+entropy difference between the microstates).  $\Delta H_{A^\ast, B^\ast}$ is readily calculated using normal mode analysis. In 
+this way, we can calculate the free energy difference between the two end states without needing to define a physical path or reaction 
+coordinate connecting the two states.
 
-Previously, this method was applied to calculate the conformational free energy of a 16 amino acid
-residue peptide, known as BHP.  We have first reproduced that known result. Next, we have applied
-this method to larger proteins in two different case studies. In the first, we attempt to pick the
-most native like structures from a set of models produced during the Critical Assessment of
-Structure Prediction (CASP) \cite{} experiment. In the second case study, we examine the
-conformational preferences of a series of chameleon proteins designed by Orban and co-workers
-\cite{} that can switch between two completely different folds with only small changes in amino acid
-sequence.
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.5 in,height=3.5 in]{method.pdf}
+\end{center}
+\caption{Graphical representation of the thermodynamic cycle involving confinement method.}
+\label{fig:method}
+\end{figure}
+
+\Ken{For this fig, let's make A and A* green and B and B* red or something, to distinguish them visually.
+Let's also blur the ensemble more relative to the microstate, so it's clearer the distinction.]}
+
+Previously, the confinement method has been validated on small model peptides \cite{Tyka2006,Cecchini2009}. Here, we show that the confinement 
+method can also be applied successfully to larger systems, such as the series of chameleon proteins designed by Orban and co-workers
+\cite{Alexander2007,He2008,Alexander2009,He2012}, which can switch between two completely different folds with only few-amino-acid changes 
+in sequence.  We also show that our computed difference free energies are useful in evaluating the quality of CASP target protein predictions.  
+This could be of significant value for ultimately improving the energetics in protein-structure models, not just the structures.  And, 
+finally, we show that we can approximately decompose difference free energies into individual amino acid components.  This offers the 
+opportunity for diagnosing the structural basis for difference free energies, which can be useful for interpreting biological mechanisms.
 
@@
 
-\subsection{Our implementation of the confiment method reproduces results from the literature}
+\subsection{The confinement method succeeds at some basic consistency checks}
 
-As a first step, we performed several control experiments to verify that our implementation of the
-confinement method produces results compatible with previous calculations reported in the
-literature.
+First, we validated that our implementation of the confinement method produces results compatible with previous calculations reported in the 
+literature.  The method has previously been applied to a 16 amino acid residue $\beta$-hairpin from protein G, known as 
+BHP \cite{Cecchini2009}. We calculated the free energy difference between the native conformation (called bhp1), which has 
+a two-stranded $\beta$-sheet and a three-stranded $\beta$-sheet called bhp3.  Our confinement calculation shows that bhp1 
+is more stable by 1.7 kcal/mol, consistent with 4 $\mu$s equilibrium simulations showing that bhp1 is more favorable configuration 
+by 1.8 kcal/mol, and in agreement with previous calculations \cite{Cecchini2009}.
 
-The method has previously been applied to a 16 amino acid residue $\beta$-hairpin from protein G,
-known as BHP \cite{Cecchini2009}. We calculated the free energy difference between two different
-conformations of the peptide: (1) the native conformation, called bhp1, with a two-stranded
-$\beta$-sheet; and (2) a conformation, called bhp3, which has a three-stranded $\beta$-sheet.
-Analysis of long (4 $\mu$s) equilibrium simulations \cite{Cecchini2009,Krivov2004}  shows that bhp1
-is the more favorable configuration by 1.8 kcal/mol. Using the confinement method, we obtain a value
-1.7 kcal/mol, which is in good agreement with the equilibrium simulations and with previous
-calculations using the confinement method \cite{Cecchini2009}.
-
-\subsection{The confiment method predicts that experimental structures have lower free energies than
-computer-generated models}
-
-Previous applications of the confiment method have focused on relatively short peptides, up to 17
-residues in length. In this work we apply the confinement method to larger proteins. As a first
-test, we calculated the free energies of models submitted by predictors during the CASP9 experiment
-(often referred to as decoys) and compared them with the calculated free energy of the corresponding
-experimentally determined structure from the Protein Data Bank (PDB). As
-Table~\ref{table:casp_control} shows, in 5 out of 6 cases, the confinement method assigns a lower
-free energy to the experimentally determined structure than to any of the decoys---as would be
-expected. Although differentiating between experimentally determined structures and computer
-generated predictions is not a stringent test of a scoring method---for example, examining the
-residue-residue packing can easily distinguish between the two [ref Rosetta Holes]---the results
-none the less serve as a useful "sanity check".
+Second, we looked at 6 target proteins from the CASP9 experiment.  For each target, we examined up to 5 submitted models.  We 
+computed the difference free energy between the true native and the best model.  Table~\ref{table:casp_control} shows that, in 5 
+out of 6 cases, the confinement method assigns a lower free energy to the experimentally determined structure than to any of the 
+decoys.  Other well-known discriminators  can also successfully make this recognition \cite{Sheffler2009}; 
+ it is just an independent useful validation that the confinement calculations make sense.
 
@@
 \begin{tabular}{l l l}\hline
-    CASP Target  & PDB Identifier & $\Delta\Delta G_{native \to best decoy}$ (kcal/mol) \\ \hline
+    CASP Target  & PDB Identifier & $\Delta \Delta G = G_{best~decoy} - G_{native}$ (kcal/mol) \\ \hline
      T0531       &    2KJX        &          $11.15 \pm 0.70$ \\ \hline
@@
 
-\subsection{The confinement method correctly predicts the structural preferences of chameleon
-sequences}
+\Ken{[ I think a better notation is $\Delta \Delta G = G_{best~decoy} - G_{native}$, rather than having the arrow.]}
 
-In general, proteins with similar sequences tend to have similar structures. This idea is the basis
-of comparative modeling and fold recognition in protein structure prediction. There are, however,
-examples---often referred to as chameleon sequences---of proteins with similar sequences that have
-remarkably different structures. Orban and co-workers have designed a sequence of 56-residues that
+\subsection{The confinement method correctly predicts the structures of chameleon sequences}
+
+We also tested the confinement method predictions for difference free energies on the chameleon sequences of Orban 
+et al. \cite{Alexander2007,He2008,Alexander2009,Bryan2010,He2012,Shortle2009}; these 
+are instances in which two highly similar sequences fold into remarkably different structures. Orban and co-workers have designed a 
+protein-G-like sequence of 56-residues that
 is marginally stable in one of two possible folds. By mutating key residues in this sequence they
-are able to stabilize one fold or the other (see Figure~\ref{fig:orban}). Sequences that adopt a
-mixed alpha/beta structure similar to Protein G are denoted as ``GB'', while sequences that form a
-three-helix bundle are denoted as ``GA''. We label the two folds as $3 \alpha$ and $4 \beta +
-\alpha$. The GA sequences are observed to fold into the $3 \alpha$ fold and the GB sequences fold
-into the $4 \beta + \alpha$ fold. One pair of sequences (GA88/GB88) are 88 percent identical in
-sequence and differ in seven positions. Another pair (GA95/GB95) are 95 percent identical and differ
-in three positions. The last pair (GA98/GB98) differ only in a single tyrosine to alanine mutation.
-The fact such small changes in sequence can lead to such dramatic changes in structure is rather
-remarkable. Accurately predicting the structural preferences of these structures presents a serious
-challenge for computational methods \cite{Alexander2007}$-$\cite{Shortle20009}.
+are able to stabilize one fold or the other (see Figure~\ref{fig:orban}). We call sequences ``$\beta$'' that fold into 
+the protein-G-like that fold into a $4 \beta + \alpha$ fold, and we call sequences ``$\alpha$'' that fold into $3 \alpha$ helical folds.  One pair 
+of sequences (GA88/GB88) is 88 percent identical in sequence, differing in seven positions. Another pair (GA95/GB95) is 95 percent identical, 
+differing in three positions. The third pair (GA98/GB98) differs in only a single tyrosine-to-alanine mutation. Accurately predicting the 
+structural preferences of these structures presents a serious challenge for computational methods 
+\cite{Alexander2007,He2008,Alexander2009,Bryan2010,He2012,Shortle2009}.
+
+\begin{figure}
+\includegraphics[width=5.0in]{orban.pdf}
+\label{fig:orban}
+\caption{Confinement method correctly predicts the structural preferences of four chameleon
+sequences. The top part of the figure represent four sequences used in this study along with the
+protein data bank id. Each sequence adopts either a three-helix bundle fold (denoted $\alpha$) or
+$4 \beta + \alpha$ fold (denoted $\beta$). The relative free energies of
+the two folds are reported for each sequence.}
+\end{figure}
 
@@
 the GA88 sequence and the GB88 structure). The confinement method was able to predict the
-conformational preferences correctly for all six sequences (data not shown). This is, however, not
-surprising. It is well known [CITE] that it is easy to distinguish computational models from native
-structures. To avoid this potential problem, we instead computed relative free energy of two
-different computer generated models for each sequence. One model is based on the GA structure and
-the other on the GB structure (see Supporting Information for details on the modeling procedure).
+conformational preferences correctly for all four sequences (data not shown). This is, however, not
+surprising. It is well known [CITE]\Arijit{still searching this citation} that it is easy to distinguish 
+computational models from native structures. To avoid this potential problem, we instead computed relative free energy of two
+different computer generated models for each sequence. One model is based on the $\alpha$ structure and
+the other on the $\beta$ structure (see Supporting Information for details on the modeling procedure).
 This is a much more realistic test of the confinement method's ability to accurately calculate
@@
 
-The results of these calculations are presented in Figure~\ref{fig:orban}. The confinement method
-identifies the correct structure for all five sequences. One hypothesis for fold switching of
+\Ken{[ It's not clear what this test is and why it's better, and why there are only 5 sequences.  Please clarify.]}
+\Arijit{ Now there are 4 sequences as they have the experimentally available structure. The fifth sequence is discussed 
+at the end of the paragraph.}
+
+Figure~\ref{fig:orban} shows that the confinement method identifies the correct structure for all four sequences. One hypothesis for fold switching of
 chameleon sequences is that the structural transitions require states with diminished stability
-[ref]. It is believed that if the free energy of the native state and the alternative state are
-within a range of 5 Kcal/mol, then it is possible for the native state to be destabilized relative
+\cite{Bryan2010}. It is believed that if the free energy of the native state and the alternative state are
+within a range of 5 kcal/mol, then it is possible for the native state to be destabilized relative
 to the alternative state with only small changes in sequence. The stability of the native state can
@@
 mutations---as is the case here. The calculated free energy differences range from around 3.5 to 5.0
-kcal/mol, which is consistent with the above hypothesis \cite{He2008}, \cite{{Alexander2009}}.
+kcal/mol, which is consistent with the above hypothesis \cite{He2008,Alexander2009,Bryan2010}. 
+In a more recent study\cite{He2012}, the amino acid residue at position 45 (Tyr for $\beta$ and Leu for $\alpha$) was
+found to be important for switching between  $\alpha$ and $\beta$ conformation. This inspired us to 
+introduce another mutation at 45 position. For our future reference this Tyr-45-Ala mutation of GB95 sequence will
+be known as GB98. Interestingly, with this mutation, the equilibrium shifts to $\alpha$ conformation as it gains 3.84 Kcal/Mol 
+stability compare to $\beta$ conformation and indicate the importance of Tyr-45 residue. It will be interesting to 
+know the actutal experimental structure which can validate our result.
 
 \subsection{Per-residue free energy calculation can identify the mechanistic detail behind
-conformational preferences}
+conformational preferences} \label{sec:orban_per_residue}
 
 To understand the mechanism behind the conformational preferences of the different protein
-sequences, we decomposed the calculated free energy into per-residue contributions as described in
-the method section. This helps us to identify important residues that stabilize a particular
-conformation. The per residue free energy, $\Delta \Delta G ((4 \beta + \alpha$) - ($3 \alpha))$ is
-shown in Figure~\ref{fig:orban_full}(F). In this plot, a negative peak indicates that a residue
-favors the $4 \beta + \alpha$ conformation and a positive peak indicates that the residue favors the
-$3 \alpha$ structure.
+sequences, we decomposed the calculated free energy into per-residue contributions in
+an approximate way \cite{Tyka2006}. For this purpose, the confinement free energy, $\Delta G_{A,A*}$ and
+$\Delta G_{B,B*}$ of each residue is calculated in the usual numerical way as described
+the method section. We call this method approximate as we ignore the contribution from the normal 
+mode or quasiharmonic analysis. Instead, decomp module of amber is used to calculate the internal energy 
+of each residue from the final restrained trajectory.
 
-The GA and GB sequences are nearly identical, so there are some common features observed for all
-sequences. Roles of some important residues in stabilizing either $4 \beta + \alpha$ or $3 \alpha$
-conformation and possible reason for such conformational preferences are discussed in details in
-Figure~\ref{fig:orban_full} and Table~\ref{table:orban_perresidue}. The experimental observations
-classified the protein into two parts: Amino acids 9-51 are fully structured in both the folds,
-where as residues 1-8 and 52-56 are unstructured in $3 \alpha$, but form $\beta$ strand in $4 \beta
-+ \alpha$ structure. Most of the amino acid residues in the region 1-9 favor $4 \beta + \alpha$
-structure as can be seen from the plot Figure~\ref{fig:orban_full}(F). We can explain most of the
-major peaks in this plot. For example,  first major peak appears at residue 7. Here, the hydrophobic
-Leu-7 stabilizes the $4 \beta + \alpha$ conformation as its sidechain is oriented towards the
-protein hydrophobic core, whereas in $3 \alpha$ structure it exposed to the solvent
-(Figure~\ref{fig:orban_full}(A)). Thus, we hypothesise that a mutation of Leu-7 to a hydrophilic
-residue would stabilize the $3 \alpha$ fold. The conformational preferences of most of the other
-large peaks can be similarly explained by hydrogen bonding, salt bridge formation, and solvent
-exposure of hydrophobic residues. They are listed in Figure~\ref{fig:orban_full} and
-Table~\ref{table:orban_perresidue}.
-
-We will now discuss the differences between the $3 \alpha$ and $4 \beta + \alpha$ folds due to the
-specific differences between the GA and GB sequences. The differences are in positions 20, 30, and
-45. GA95 has Leu-20, Ile-30, and Leu-45; GB95 has Ala-20, Phe-30, and Tyr-45. In the $3 \alpha$
-fold, residues 20 and 30 are involved in interactions with residue 49. The peak at position 49
-stabilizes the GA sequence more than that of GB sequence. In the $3 \alpha$ fold, Ile-49 has
-hydrophobic interactions with Leu-20 and Ile-30 with GA sequence, whereas with GB sequence it has
-such interactions with smaller Ala-20 (Figure~\ref{fig:orban_full}(I). Also, in the $3 \alpha$ fold
-with the GB sequence, Phe-30 being a larger group can not be completely accommodated in the
-hydrophobic core (Figure~\ref{fig:orban_full}(I)) and therefore, partially exposed to the protein
-surface. There is an interesting difference for residue at position 45 as well. In GB95 sequence and
-$4 \beta + \alpha$ fold there is H-bonding between Tyr-45 and Asp-47
-(Figure~\ref{fig:orban_full}(K)), which is absent in the $3 \alpha$ conformation. On the other hand
-with GA sequence, this H-bond is absent for both the conformation as there is a Leu at position 45
-(Figure~\ref{fig:orban_full}(J)).
-
-\begin{table}
-\begin{center}
-\caption{Analysis of the per-residue free energy decomposition of GA95 and GB95 sequence reveal that certain residues prefer
-either $3 \alpha$  or $4 \beta + \alpha$ structure. These observations and the reason behind such conformational preferences are listed in this
-table.}
-\label{table:orban_perresidue}
-\begin{tabular}{| l | l | l |}
-\hline
-    Information from per                                 & What is happening in           & What is happening in                      \\
-    residue plot                                         & $3 \alpha$ conformation        & $4 \beta + \alpha$  conformation          \\ \hline
-    Leu-7 favors                                         & hydrophobic residue is         & oriented towards the                      \\
-    $4 \beta + \alpha$ (Figure~\ref{fig:orban_full}(A))  & exposed to the solvent         & hydrophobic core                          \\ \hline
-    Gln-11 stabilizes $3 \alpha$                   & Gln-11 has H-bond                         & no such H-bond                           \\
-    (Figure~\ref{fig:orban_full}(B))           &  with Glu-15                              &                                           \\ \hline
-    Glu-14 favors $4 \beta + \alpha$           & hydrophilic residue partially  & completely exposed                                    \\
-    (Figure~\ref{fig:orban_full}(E))           & exposed to the solvent         & to the solvent                                        \\ \hline
-    Ala-26 favors  $4 \beta + \alpha$          & hydrophobic residue            & oriented towards the                                   \\
-     (Figure~\ref{fig:orban_full}(C))          & oriented towards surface       & hydrophobic core                                       \\ \hline
-    Ile-49 favors $3 \alpha$                   & hydrophobic interaction        & sidechain exposed                                       \\
-    (Figure~\ref{fig:orban_full}(H) and (I))   & inside the protein             & to the solvent                                          \\ \hline
-    Ile-49 favors GA more                      & more hydrophobic               &  unchanged                                              \\
-    than GB                                    & interactions with residues     &                                                         \\
-    (Figure~\ref{fig:orban_full}(H) and (I))  &  at 20, 30 in GA                &                                                         \\ \hline
-    Val-54 favors $4 \beta + \alpha$           & hydrophobic residue            & hydrophobic interaction                                \\
-    Figure~\ref{fig:orban_full}(D))             & exposed to the solvent        & inside the protein                                      \\ \hline
-    Glu-56 favors $3 \alpha$                   & hydrophilic residue partially  & completely exposed                                    \\
-    (Figure~\ref{fig:orban_full}(G))            & exposed to the solvent         & to the solvent                                        \\ \hline
-    Phe-30 in GB sequence                      & exposed to the solvent         & oriented towards the                                  \\
-    stabilizes $4 \beta + \alpha$                  & in protein surface             & hydrophobic core                                      \\ \hline
-    Tyr-45 in GB favors                      & No H-bonding                   & H-bond with Asp-47                                    \\
-    $4 \beta + \alpha$ (Figure~\ref{fig:orban_full}(K))&                 &                                                       \\ \hline
-    Asp-47 in GA sequence                      & H-bond with Lys-50             & No H-bond, No Tyr-45                                  \\
-    favors $3 \alpha$                          &                         & in GA sequence                                         \\ \hline
-\end{tabular}
-\end{center}
-\end{table}
+\Ken{[Arijit -- I didn't find this methods calculation.  We should at least put in a sentence or two here summarizing it.]}  
+This helps us to identify important residues that stabilize a particular
+conformation. The per residue free energy, $\Delta \Delta G ((\beta) - (\alpha))$ is
+shown in Figure~\ref{fig:perresidue_orban}. 
+In this plot, residues colored in blue favor the $\alpha$ structure, residues in red favor the $\beta$ structure, and white residues
+have no preference.
 
 \begin{figure}
-\includegraphics[width=3.5 in,height=4 in]{orban.pdf}
-\label{fig:orban}
-\caption{Confinement method correctly predicts the structural preferences of six chameleon
-sequences. (A) The six sequences used in this study. (B) Each sequence adopts either a Protein
-G-like fold (denoted GB) or a three-helix bundle fold (denoted GA). The relative free energies of
-the two folds are reported for each sequence.}
+\begin{center}
+\includegraphics[width=4.5in]{delta_g.pdf}
+\end{center}
+\caption{The $\alpha$ and $\beta$ conformation associated with GA95 and GB95 sequences are colored according to the relative per residue
+free energy value. In this figure, residues colored in blue favor the $\alpha$ structure, residues in red favor the $\beta$ structure, and 
+white residues have no preference. Some of the mutated residues are also labelled.}
+\label{fig:perresidue_orban}
 \end{figure}
 
-\begin{figure}
-\includegraphics[width=6.4 in,height=5.3 in]{orban_full.pdf}
-\label{fig:orban_full}
-\caption{Sidechain orientation of some important residues which stabilize/destabilize either of the $4 \beta + \alpha$ or $3 \alpha$ conformation.
-The details are in Table~\ref{table:orban_perresidue} and the main text.
-A. Leu-7.
-B. Gln-11 has H-bond with Glu-15 in $3 \alpha$, which is absent in $4 \beta + \alpha$
-C. Ala-26
-D. Val-54.
-E. Glu-14.
-F. plot of free energy difference for GA95 and GB95 sequence. Positive peak stabilize $3 \alpha$ form and negative peak
-stabilize $4 \beta + \alpha$.
-G. Glu-56 .
-H. sidechain orientation of Hydrophic Ile-49, along with Leu-20 and Ile-30 with GA95 sequence
-I. sidechain orientation of Hydrophic Ile-49, along with Ala-20 and Phe-30 with GB95 sequence
-J. Tyr-45 forms H-bond with Asp-47 with GB95 sequence and $4 \beta + \alpha$ conformation
-K. No H-bond possible between Leu-45 and Asp-47 with GB95 sequence and $4 \beta + \alpha$ conformation .}
-\end{figure}
+Although the overall free energy difference between the two structures is small (within 3-5 Kcal/Mol), the individual residues 
+can show marked preferences for being in either $\alpha$ or $\beta$
+conformation. Such differences can be undestood by looking in detail at the local environemnt for
+those resiudes. For example, the region around residue 7 forms a random coil in the $\alpha$
+structure, whereas it is forming a beta sheet structure in the $\beta$ structure. These residues
+strongly favour the locally well packed and hydrogen bonded environment found in the $\beta$ sheet.
+Overall, favoring $\alpha$ or $\beta$ structure is a delicate balance, were the relative global free
+energy difference is small and the contributions of different per residue tendencies balance out. It
+is therefore very likely that by changing key resiude preferences the global fold preferences can be
+changed.
 
-\subsection{We have applied the confinement method to structure prediction}
+Three residues are mutated between GA95 and GB95, at positions 20, 30, and 45. In GA95, these
+residues (L20, I30 and L45) stabilize the $\alpha$ structure. In GB95, two of these residues (F30
+and Y45) favor the $\beta$ structure, because they have large solvent exposed surface areas in the
+$\alpha$ structure, but are more buried in the $\beta$ structure. Additionally Y45 forms a hydrogen
+bond with D47 in the $\beta$ structure. On the other hand, residue A20 from GB95 still favors the
+$\alpha$ structure, although not as strongly as L20 in GA95.
 
-We have applied the confinement method as a ``meta-predictor'' for structure prediction. Here, the
-task is to correctly identify the most correct models out of a set of ``decoys'' generated by
+The $\alpha$ and $\beta$ sequences are nearly identical, so there are some common features observed for all
+sequences. Roles of some important residues in stabilizing either the $\alpha$ or $\beta$
+conformation and possible reasons for such conformational preferences are summerized in 
+supporting information Figure S1. The experimental observations
+classified the protein into two parts: Amino acids 9-51 are fully structured in both folds, whereas
+residues 1-8 and 52-56 are unstructured in the $\alpha$ fold, but form a $\beta$-strand in the
+$\beta$ fold. Most of the amino acid residues in the region 1-9 have negative per-residue free
+energies, which means that these residues favor the $\beta$ structure.
+
+In addition to the direct effects of the mutation, there are also indirect effects due to small
+perturbations due to the mutations. For example, the L20A mutation causes a slight repacking around
+residue 20. This causes large changes in the per-residue free energies of nearby residues T25 and
+A26. However, these differences largely cancel and the overall effect of the L20A mutation is weaker
+than either the I30F or L45Y mutations.
+
+Such per-residue free energy decompositions allow us to understand the driving forces behind protein
+folding and conformational change and may be useful for designing proteins with specific structures
+and functions.
+
+\subsection{Applcation of confinement method for structure prediction}
+
+We applied the confinement method as a ``meta-predictor'' for structure prediction. Here, the
+task is to correctly identify the most accurate models out of a set of ``decoys'' generated by
 different methods during the Critical Assessment of Structure Prediction (CASP) experiment. CASP is
 a blind test in which different groups apply methods to predict the 3-dimensional structure of
-proteins from their sequences. This is done with a either a three week (for humans) or three day
-(for fully automated servers) time limit on each target. Each group is allowed to submit five
+proteins from their sequences. Each group is allowed to submit five
 possible structures, which they are supposed to rank from best to worst. We have performed two
@@
 In the first experiment, we examined several targets and tried to rank-order the predictions
-produced by other groups. In some cases the structures all came from the same group, in others the
-structures were drawn from several different groups. The goal of this experiment is to determine if
+made by the same group (same methodology used to generate the models), in others the
+structures were drawn from several different groups (different methodology used). The goal of this 
+experiment is to determine whether 
 the physics-based confinement method can correctly identify more native-like structures as having
@@
 be understood roughly as the percentage of residues that are correctly positioned in the model
-(higher is better, range 0 to 100).
-
-We have chosen the initial models on which to test the methods based on server groups that have
-traditionally done well in past CASP events.
+(range 0 to 100, higher is better). The initial models for the test was chosen from different server 
+groups those have traditionally done well in past CASP events.
 
@@
 
-The question that we want to answer here is whether we can rank models that have been generated with
-the same methodology. In all cases, we have calculated the free energy difference between the native
-structure and the submitted models as well. 
-%Identifying the native state from a set of models is a very easy test, that most metapredictors
-%already handle correctly, but it is also a first proof of principle. 
-First we choose a protein BVU3908 from Bacteroides vulgatus whose PDB id and CASP target code are
-2L01 and T0559 respectively. The native structure of this 69 residue protein was solved using NMR.
-The best predictor group for this target was "BAKER-ROSETTASERVER". We initially checked similarity
-between the models submitted by "BAKER-ROSETTASERVER" and discarded two of them from the analysis on
+Here, our task is to rank different submitted models of CASP9 targets that have been generated with
+the same methodology. In absence of any other reliable alternative, we compare all our free energy based 
+ranking with the GDT-TS based ranking which are available in CASP website.
+
+The first test case is a protein BVU3908 from Bacteroides vulgatus whose PDB id and CASP target 
+code are 2L01 and T0559 respectively.
+The best predictor group for this 69 amino acid long target was "BAKER-ROSETTASERVER". We initially checked similarity
+between the models submitted by them and discarded two from the analysis on
 the basis of being too similar to some of the other models. The GDT\_TS score and rmsd values as
 shown in Figure~\ref{fig:T0559} indicate that model 1 was predicted correctly, whereas the order of
-model 3 and 5 was wrong. The main difference between model 3 and the rest of the models is that the
-orientation of the first alpha helix of model 3 is \Alberto{opposite??}. On the other hand, as shown
-in Figure~\ref{fig:T0559} the confine and release method not only can differentiate the native
+model 3 and 5 was wrong. The main difference between model 3 and rest of the models was the  
+orientation of the first alpha helix. On the other hand, as shown
+in Figure~\ref{fig:T0559} the confinement method not only can differentiate the native
 structure from the submitted models, the ranking also correlates perfectly with the GDT\_TS score.
 
-To further test the method we have taken the example of protein BT2368 from Bacteroides
+\Ken{Arijit- we could probably make a better visualization of this figure above by putting it onto a toy landscape: x-axis is rmsd, y-axis is DDG, and above each of the 4 points on this graph will be the ribbon diagram and the GDT-TS.
+I suggest we do that with the fig below too.
+Let's don't use 4 different colors for the fig below.
+Let's just show differences in structures in some different color, or show them all as green.]}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.8 in,height=3.2 in]{T0559.pdf}
+\end{center}
+\caption{The native and three submitted model structure, along with their GDT\_TS, RMSD and relative Free energy values
+of protein BVU3908 from Bacteroides vulgatus (PDB id: 2L01 and CASP code: T0559).}
+\label{fig:T0559}
+\end{figure}
+
+To further test the method we continued with the example of protein BT2368 from Bacteroides
 thetaiotaomicron. The PDB id of this 74 amino acid residue protein is 2L02 and CASP target code is
-T0560. We have compared the free energy difference between the native structure and the two of the
-five submitted models from the group "Splicer". The remaining three models are discarded as again
-they are similar to the rest of the models. The comparison of GDT\_TS score as shown in
-Figure~\ref{fig:T0560} and published in the final CASP result is only for models with residues 3 to
-66. In order to keep consistency, we have also done our analysis with models and crystal structure
-consisting residues 3 to 66. As expected, the native state is identified correctly, and the two
-other models are ranked in the correct order in accordance with the GDT\_TS score.
+T0560. We compared the free energy difference between the native structure and the two of the
+five submitted models from the group "Splicer". The remaining three models were discarded as 
+they were too similar to the rest of the models. As shown in Figure~\ref{fig:T0560}, we can identify 
+the native state correctly and our calculated free energy based ranking matches well with the
+GDT\_TS score.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=4.0 in,height=3.0 in]{T0560.pdf}
+\end{center}
+\caption{Native and two model structure of protein BT2368 from Bacteroides thetaiotaomicron (pdb id: 2L02 and CASP code: T0560). The two models
+were from the group "Splicer".}
+\label{fig:T0560}
+\end{figure}
 
@@
 
-Our next test was between models that are produced with different methodologies. We do this by
-looking at some of the best CASP submissions from different research groups. For this purpose we
-have chosen the x-ray crystallographic structure of fas apoptosis inhibitory protein molecule whose
-pdb id and CASP target codes are 3MX7 and T0540 respectively. This protein contain 90 amino acid
-residues with 8 beta strands. We have chosen best models from groups "LTB" and "Mufold" for our
-analysis, which are labelled as Model 1 and Model 2 in Figure~\ref{fig:T0540}. As presented in
-Figure~\ref{fig:T0540} we found that this method is able to correctly order the models which matches
-with the GDT\_TS score.
+Our next test was between models that are produced with different methodologies. 
+Here, our model system was a fas apoptosis inhibitory protein molecule whose
+pdb id and CASP target codes are 3MX7 and T0540 respectively. 
+This protein contain 8 beta strands and 90 amino acid residues. 
+The top models from groups "LTB" and "Mufold" was chosen for analysis.
+We will call this as Model 1 and Model 2 for future references. As summerized in
+Figure~\ref{fig:T0540}, the free energy based ranking once again correlates well with the 
+GDT\_TS based score.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=4.0 in,height=3.4 in]{T0540.pdf}
+\end{center}
+\caption{X-ray crystallographic structure and two submitted models of fas apoptosis inhibitory protein (pdb id: 3MX7 and CASP code: T0540).
+Model 1 and Model 2 in this analysis were submitted by the group LTB and MUFOLD respectively.}
+\label{fig:T0540}
+\end{figure}
 
@@
 
-We used one of the CASP targets (pdb id=2KWY, casp id=T0569) to study the per residue contribution
-between this NMR structure and the best CASP9 predicted model (from the 'Mufold' group, GDT\_TS=78).
-Visualization upon superimposition of two structures reveals that the difference between the two
-structure is at the region consist of residues 48 to 65, where model1 is much more disordered
-compare to the native structure(see Figure~\ref{fig:T0569}). Other regions look similar for both the
-cases with of model1 is within 2.6 \AA backbone rmsd value of the native structure. We found that,
-the confinement method can still differentiate between the native structure and the generated model.
-We have also calculated the per residue free energy with a aim that this can identify the residue
-which are responsible for such free enerfy difference. The per residue free energy decomposition is
-shown in  Figure~\ref{fig:T0569_per_residue}(A). We indeed found that two hydrophobic residues
-Val-59 and Ile-61 are oriented towards the solvent in the protein surface in the generated model
-(Figure~\ref{fig:T0569_per_residue}(B). This the region, where there is a beta sheet in the original
-crystal structure, which is absent in the model. Interestinlgy, some regions are more stable in the
-generated model. A close inspection reveals that the hydrophilic Glu in position 67 is completely
-exposed to the protein surface and solvent in the best model, whereas it is partially exposed in the
-crystal (Figure~\ref{fig:T0569_per_residue}(C)). The peak at position 76 is due to the H-bond
-between Lys-76 and Asp-11, which is absent in the original crystalographic structure
-(Figure~\ref{fig:T0569_per_residue}(D)).
+\Ken{[Arijit - This and the following sections need to be shortened and made more concise.]}
+
+In section~\protect\ref{sec:orban_per_residue} we discussed how per residue free energy calculation
+can identify the mechanistic detail behind conformational preference of a chameleon sequence. In this section, our aim 
+is to apply the same method and try to understand whether per residue free energy can help us identify
+residues which stabilizes/destabilizes a particular region of a protein. The difference  
+here is that the protein has similar fold with only mismatch in some region. 
+To investigate this, we choose a domain of adhesion exoprotein from Pediococcus pentosaceus (pdb id=2KYW) from protein data bank 
+and its best predicted model of CASP9 (casp id=T0569, submitted by 'Mufold' group which had a GDT\_TS of 78).
+We first calculated the free energy of the whole protein and found that the native 
+structure is stabilized by 20 Kcal/Mol (see suppurting information Figure S2).
+
+Next we proceed with the per residue free energy calculation. Our effort clearly identifies two hydrophobic residues 
+Val-59 and Ile-61, which destabilizes the submitted model with respect to the crystallograohic structure.
+The sidechains of these hydrophobic residues are oriented towards the protein hydrophobic core in the
+native NMR structure but oriented outside protein and solvent exposed in the model. These residues are the primary reason
+why two of the beta sheets in the native structure are disordered in the model structure ( see Figure~\ref{fig:T0569_per_residue}
+and supporting information Figure S3). This method also identifies some other residues which stabilize/destabilize either 
+of the conformation. For
+example, Lys-76 stabilizes the model as it has a H-bond with Asp-11 which is missing in the native structure
+(Figure~\ref{fig:T0569_per_residue} and supporting information figue S3).
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=4.3 in,height=4 in]{T0569_perres1.pdf}
+\end{center}
+\caption{To better understand the reason behind difference in conformation, structure of CASP Target 569 (pdb id: 2KWY) and a 
+generated model was colored according to the per residue free energy value. The amino acid residues that are colored in deep red 
+and deep blue stabilizes the crystal and the model respectively, while the residues with light blue color does not have a preferance.
+ Some of the important residues are also labelled.}   
+\label{fig:T0569_per_residue}
+\end{figure}
 
@@
 
-Despite the great success in most of the studied systems, we found a few failures, specially when
-the GDT scores of the compared structures are very close. We have found this for an engineered
-protein from Asr4154 protein (PDB ID: 2L09 and CASP Target T0538). This protein contains 54 amino
-acid residues with three alpha helixes and two very short beta strands. The model that is closest to
-the native structure was generated by the group PconsR with a GDT\_TS value of 96.23. We have also
-considered one model from group "Shell" (GDT\_TS = 90.09) and "FOLDIT" (GDT\_TS = 86.32) for our
-analysis along with the native crystal struture (RMSD values between 1.6 and 2.0\AA \Alberto{from
-native, right?all atom? CA?}). The models from PconsR, Shell and FOLDIT are labelled as model 1,
-model 2 and model 3 respectively in Figure~\ref{fig:T0538}. As seen in Figure~\ref{fig:T0538}, one
-can find that the model from PconsR has lowest free energy value, making it a more stable structure.
-Although ranking of the other models remains consistent, model 1 is predicted to be more stable than
-the crystal structure. To rationalize this, we have decomposed the total free energy into the per
-residue components. The results show us that despite the small variation at the backbone level (as
+Despite the great success in most of the studied systems, there are few failures, specially when
+the GDT scores of the compared structures are very close. We observed this for a $54$ amino acid residue long
+ engineered protein from Asr4154 protein (pdb id: 2L09 and CASP code T0538).
+Like previous cases, we kept the native structure and include models from the group "PconsR" (GDT\_TS=96.23, model1),
+"Shell" (GDT\_TS = 90.09, model2) and "FOLDIT" (GDT\_TS = 86.32, model3) for analysis. 
+Contrary to our expectation, model1 was found to be more stable than the crystal structure 
+(see Figure~\ref{fig:T0538}). In order to rationalize such obseravtion we calculates per residue free energy
+between crystal structure and model1. The results show us that despite the small variation at the backbone level (as
 shown by high GDT\_TS scores and low RMSDs), the sidechains are oriented in very different ways,
-giving raise to large differences in the stabilization of certain residues. In particular, some of
-the differences arise from different salt bridge patterns (Arg-32 with Glu-35 and Glu-28 with Lys-24
+giving rise to large differences in the stabilization of certain residues. In particular, some of
+the differences arise from different salt bridge patterns (Arg-32 with Glu-35, Glu-28 with Lys-24 and Arg-26 with Glu-50
 in crystal versus Arg-32 with Glu-28 in model 1) and certain flexible polar residues exposed to the
-surface (Lys-24 in model 1, which has an entropic gain from not forming a salt bridge and is
-stabilized by interactions with the solvent). This is summarized in Figure~\ref{fig:T0538compare}(A)
-Similar patterns arise in other places of the protein (see Figure ~\ref{fig:T0538compare}(B)), where
-the salt bridge between Arg-26 and Glu-50 is absent in model 1 and is further \Alberto{stabilized?}
-by Phe-51 having a very different conformation that in the crystal structure. All in all, this
+surface (Lys-24 in model 1, which has an entropic gain from absence of salt bridge and is
+stabilized by interactions with the solvent) (See supporting information Figure S4). All in all, this
 unexpected result just shows us that this method is sensitive to local interactions such as those
-happening from side chain reorientation. At the same time, the results explained in past sections
-showed that for larger changes the accuracy of the method is very high.
+happening from side chain reorientation and also indicate the limitations of GDT\_TS based ranking.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.6 in,height=3.4 in]{T0538.pdf}
+\end{center}
+\caption{The native and three model structure of engineered protein from Asr4154 protein (PDB ID: 2L09 and CASP code:T0538). The model 1,2 and 3 are
+from the group PconsR, Shell and FOLDIT respectively.}
+\label{fig:T0538}
+\end{figure}
 
@@
 So far we have seen that the method is good at predicting preferences when the structures are not
-very close to each other, how far from native can we go? In this section we want to explore what
-happens when we use the method to describe low quality models. We use one of the CASP targets (pdb
-id = 2KJX, CASP id = T0531) . We mentioned earlier that our predictive abilities are greatly
-decreased with low model qualities. Comparing low qualitive models between themselves is not very
-informative, however, comparing all of them to a native/native like structure, allows to rank order
-this structures. In order to test this hypothesis we performed another test with target T0531 of
-CASP9, in which we compared the crystal structure to five models from the MUFOLD server. Pdb id of
-this extracellular domain of the jumping translocation breakpoint protein is 2KJX and CASP target
-code is T0531. Looking at Figure~\ref{fig:T0531} shows: 1.) native is correctly identified as
+very far from the native. But the question remains how far from native can we go and still see 
+that the method produces correct result. In this section we explored this question with models from  
+extracellular domain of the jumping translocation breakpoint protein (pdb id: 2KJX). Most of the
+group could only generate low resolution models for this CASP9 target (id: T0531) . In our comparison, we choose five
+models by the group MUFOLD-MD, which was the best performing group for this target with their best model 
+had a GDT\_TS value of 44. The result presented in Figure~\ref{fig:T0531} shows: 1.) native is correctly identified as
 expected and 2.) Surprisingly there is a high level of correlation between the GDT ranking and the
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.5 in,height=3.5 in]{T0531.pdf}
+\end{center}
+\caption{The native structure and 5 models of extracellular domain of the jumping translocation
+breakpoint protein (pdb id: 2KJX and the CASP code: T0531).}
+\label{fig:T0531}
+\end{figure}
+
 \subsubsection{Can the confinement method perform better quality assessment in protein structure prediction?}
 
-A part of the CASP experiment is dedicated to the assessment of the quality of predicted models
-\cite{Kryshtafovych2011}. Most of the top performing groups in this category use consensus
-approaches for quality assessment \cite{Wang2011}. A review of CASP9 quality assessment also pointed
-out that the methods which are based on clustering techniques perform much better compared to those
-methods which are based on analysis of individual models \cite{Kryshtafovych2011}. In this
-background we try to answer, whether the confinement method can do better quality assessment
-compared to the best performing group (MUFOLD-WQA) in this catagory. In CASP9, the overall
-reliability of models in quality assessment (QA) was accepted in a mode known as QMODE 1. Here,
-predictors were asked to score each model on a scale from 0 to 1, with higher values corresponding
-to better models \cite{Kryshtafovych2011}. We have again picked up the CASP taget T0538 to answer
-this question. We have chosen Model 3 submitted by PconsR (GDT\_TS = 96, QMODE 1 = 0.5434) and model
-5 from the MULTICOM-NOVEL ( GDT\_TS = 83, QMODE 1 = 0.5865). Both are server predicted models and
-they were further used for quality assessment by the group MUFOLD-WQA \cite{Wang2011}. It is
-important to note that the model from PconsR was the most accurate predicted model for this target
-and the model from MULTICOM-NOVEL was predicted as best by the quality assessment experiments of
-MUFOLD-WQA. Also, the QMODE 1 value presented here is from the group MUFOLD-WQA. Our free energy
-analysis predict that the model from Pconsr is indeed better than the model from MULTICOM-NOVEL
-$(\Delta \Delta G (Model\_Pconsr - Model\_MULTICOM-NOVEL = -3.9 Kcal/Mol)$, which   matches well
-with the GDT\_TS score. On the other hand the MUFOLD-WQA predicted the ranking in the reverse order.
+\Ken{[Arijit - let's make the following section more concise and focused.]}
 
-We have also carried out a similar type of analysis for the case of Target 560 from the CASP9
-experiment. For that we have chosen the model 1 from the group PconsR (GDT\_TS = 94, QMODE 1 =
-0.5178), Bilab\-Enable (GDT\_TS = 87, QMODE 1 = 0.5344) and Multicom\_Refine (GDT\_TS = 74, QMODE 1
-= 0.4770). If, one wants to rank order with respect to the GDT\_TS values, the model from PconsR is
-the best predicted model, but according to the QA assessment by MUFOLD-WQA, the model from
-Bilab-Enable was found to be best predicted model. All the QMODE 1 values are from MUFOLD\-WQA. In
-CASP 9, the sequence of 74 residues were given to predict the structure and for quality assessment,
-whereas during GDT\_TS analysis only residues 3-66 were considered. When we considered the full 74
-amino acid residues, the free energy calculated were $(\Delta \Delta G (Model\_Pconsr -
-Model\_Bilab\-Enable = 3.3 Kcal/Mol)$ and $(\Delta \Delta G (Model\_Pconsr - Model\_Multicom\_Refine
-= -30.0 Kcal/Mol)$ \Alberto{-30 sounds huge!} which support the QA assessment by MUFOLD-WQA. On the
-other hand, if we calculate the same free energy with the trimmed version of the protein (with
-residues 3-66), the values of the free energy difference become -1.7 and -31.1 Kcal/Mol
-respectively, which support the GDT\_TS score. The summary of this whole observation is presented in
-Table~\ref{table:560QA}. There is no doubt that the consensus approach can predict the model quality
-in a faster manner. But we expect that the confinement method can predict the model quality in a
+A part of the CASP experiment is dedicated to the quality assessment (QA) of predicted models
+\cite{Kryshtafovych2011}. Here, predictors were asked to score each model on a scale 
+(known as qmode) from 0 to 1, with higher values corresponding to better models \cite{Kryshtafovych2011}. 
+It will be interesting to know how confinement method perform in quality assessment compare to 
+the other groups in CASP9. We investigated this using couple of CASP Targets. Here, we present a 
+case, where confinement method perform well than the top performing group MUFOLD-WQA \cite{Wang2011} 
+of CASP9. We choose  
+two models from CASP taget T0538. They were model 3 submitted by PconsR (GDT\_TS = 96, qmode 1 = 0.5434) 
+and model 5 from the MULTICOM-NOVEL ( GDT\_TS = 83, qmode 1 = 0.5865). Both are server predicted models and
+the qmode value presented are from MUFOLD-WQA \cite{Wang2011}. We choose these two models as
+the model by PconsR was the most accurate predicted model for this target, whereas the other model 
+was predicted best by QA test of MUFOLD-WQA. The calculated free energy using confinement method indicate
+that the model by PconsR is more stable by $3.9 Kcal/Mol$ which support the GDT\_TS trend.
+There is no doubt that the consensus approach can predict the model quality
+in a faster manner. But we expect confinement method can predict the model quality in a
 relatively expensive but much more accurate way.
-
-\begin{table}
-\label{table:560QA}
-\begin{center}
-\begin{tabular}{| l | l | l | l | l |}
-\hline
-Model            &  $\Delta \Delta G$ (1-74)  & $\Delta \Delta G$ (3-66) &    GDT\_TS    & QMODE 1  \\ \hline
-PconsR           &         0.0                &      0.0                 &    94         &  0.5178   \\ \hline
-Bilab\-Enable     &        -3.3                &      1.7                 &    87         &  0.5344   \\ \hline
-Multicom\_Refine &        30.0                &     31.1                 &    74         &  0.4770   \\ \hline
-\end{tabular}
-\end{center}
-\caption{Comparison of quality assessment by the group MUFOLD\-WQA and the confinement method of the CASP9 target 560 (pdb id: 2L02). 
-The GDT\_TS score obtained from CASP9 website, was based on the trimmed model containing residue 3-66, whereas, the QMODE 1 values are from
-quality assessment by the group MUFOLD\-WQA and based on residues 1-74.}
-\end{table}
 
@@
 
-\Alberto{The last 50 years represent our effort to understand what the native state of a protein
-    looks like by using theoretical methods. Experimental evidence coming from NMR and x-ray
-    cristallography remain so far the only real ways to solve the structure of the native state.
-    However, this methods are expensive and slow. Theoretical methods on the other hand are cheaper,
-    but despite improvement still have a long way to go. Often this computational pipelines produce
-    ensembles of diverse structures and identifying which structure is more native like is still a
-    problem. Here we have shown how the application of the confinement method can help identify
-    native like states. The high success rate balances with the computational expense. However, with
-    GPU technology and the rate at which computers improve, this computational expense is becoming
-less of a bottle neck.}
-\Alberto{I don't agree with some of the points here, but I think I covered what you meant in the
-previous paragraph}
+We have described a computational method called confinement for computing the difference free energy from a protein conformation $A$ to $B$.  We 
+show that it can give accurate values, even for relatively large conformational changes.  We have demonstrated that it can discriminate 
+which Orban chameleon peptides fold into a 3-helix bundle vs. a 4$\beta$+1 helix structure. We show that it can discriminate between good and bad CASP 
+models.  Perhaps most importantly, we have shown that it can be used to give residue-level insights into what are the dominant structural factors in a 
+protein that are responsible for the difference free energies.  A key advantage is that this method does not require any reaction coordinate, or sampling 
+a pathway from conformation $A$ to $B$. We have tested the method for structures of proteins having up to 100 amino acids.  The computational cost 
+for a 56-residue protein is about 4 hours [on 1 gpu] for 20 ns of confinement [Is 20ns all you need?].
 
-In the recent past there have been a number of studies directed towards understanding protein
-structure using theoretical tools. Experimentally, it is relatively easy to identify the native
-structure from the data obtained from x-ray crystallography or NMR. On the other hand using
-theoretical tools it is often difficult to identify the native structure using both physics based
-methods and knowledge based method. In order to rank structure from the available data, some groups
-use clustering algorithms and the structure of the most populated cluster is picked out as the most
-probable structure. However, it may happen that the structure may be trapped and spend most of its
-time in a kinetic trap ***kinetic traps in MD,not other methods, local minima***, which may give a
-wrong interpretation of the clustering result. On the other hand, some groups also depend on
-potential energy functions to pick out their best structure, lacking an entropic component. In this
-work we attempted to identify the native/native like protein strcuture from the decoy using free
-energy as a scale. We have used the previously developed confinement method, applying it to larger
-systems.
-
-We found that in majority of cases, the native structure is always the most stable and we can rank
-protein structures upto 100 amino acid residues long. This method works very well if at least one of
-the strcuture in the decoy set has high quality. The main advantage of this method is that it does
-not require any reaction coordinate. Thus one can calculate the free energy difference between two
-very different conformations. An interesting application from this method that we have exploited is
-the partition of free energies on a per residue basis. Although, it is an approximate partition, it
-gives important insight regarding the conformational preferences of a particular residue. The use of
-this decomposition in understanding the effect of a small number of aminoacids in our test chamaleon
-sequences allows us to think that in the future, this method may be used for protein design. The
-method can be also give wrong result if all the models in the decoy are very poor. It may also be
-difficult to rank structures if all the structures are very very close to the native structure. But
-we want to ignore this fact as our main aim is to differentiate between the good and the bad
-structure. In this article we have extensively tested this method using examples from CASP This
-method can be computationally expensive in actual CASP experiment if one want rank hundreds of
-submitted models in CASP experiments. However, in real life experiment with a particular target, we
-expect this method can give accurate result. The important fact about confinement method is that all
-the part of the calculation can be done independently. There is another big bottlenecks in this
-calculation. This has to do with the amount of simulations that have to be performed to confine the
-ensemble of conformations close to a microstate into a single microstate. We have tackled this issue
-by using GPU (Graphical Processing Unit) technology as opposed to the classical CPUs, which gives us
-a two order of magnitude boost in computational efficiency. For example, we have mostly carried out
-our calculation with Amber program \cite{Case2005} in GPU computer. With a avarage 56 residue we
-found that it take only 4 hour to complete 20 ns of the confinement step. So, with available
-computer power this method will be easy to use for real problems. 
+\Arijit{I need 21 (confinement run) x 20 ns for single confinement}
 %There is another issue, which is normal mode analysis calculations are Oh(N**3) and require large
@@
 
-This method can be further used for calculating free energy due to very large conformational change.
-Another application may be the case of ligand binding. It can also be used to distinguish the most
-stable structure during protein design. Decompition of total free energy into the per residue
-component help identifying residues that give stability to a particular conformation. Another,
-important application may be to study the protein folding kinetics, specially to study the phi value
-analysis.
+\Ken{[If this is a relevant limitation, let's say something about it.]}
+\Arijit{I think Ken is pointing towards the normal mode issue and number of residues that we can handle}
 
@@
 
-The confinement method has been described in details in ref. by Tyka et.al. \cite{Tyka2006} and
-Cecchini et. al. \cite{Cecchini2009}. The basic approach of the confinement approach is the same in
+The confinement method has been described in details in ref. by Tyka et al. \cite{Tyka2006} and
+Cecchini et  al. \cite{Cecchini2009}. The basic approach of the confinement approach is the same in
 both these papers. However, there are some technical differences. Here we briefly describe the
@@
 restrained trajectory. We call this method approximate as we ignore the contribution from the normal
-mode or quasiharmonic analysis. However, this contrbution to the total free energy is much smaller
-(less than $1.0 Kcal/Mol$\Alberto{that sounds huge, i would say what percent of error we stimate, I
-think it was lower than 1kcal/mol}), which allow us to study the mechanistic details of
-conformational preference of each residue.
+mode or quasiharmonic analysis. However, this contrbution to the total free energy is much smaller, 
+which allow us to study the mechanistic details of conformational preference of each residue.
 
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.5 in,height=3.5 in]{method.pdf}
-\end{center}
-\caption{Graphical representation of the thermodynamic cycle involving confinement method.}
-\label{fig:method}
-\end{figure}
+\begin{thebibliography}{99}
 
-%\begin{figure}
-%\begin{center}
-%\includegraphics[width=3.5 in,height=3.0 in]{G95.pdf}
-%\end{center}
-%\caption{Two protein with 95 \% similar sequence but different folds}
-%\label{fig:G95}
-%\end{figure}
+\bibitem{Elber2007}
+Elber, R. A Milestoning Study of the Kinetics of an Allosteric Transition: Atomically Detailed Simulations of Deoxy Scapharca
+Hemoglobin. Biophysical J., 2007, 92, 85-87.
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.8 in,height=3.0 in]{T0559.pdf}
-\end{center}
-\caption{The native and three submitted model structure, along with their GDT\_TS, RMSD and relative Free energy values 
-of protein BVU3908 from Bacteroides vulgatus (PDB id: 2L01 and CASP code: T0559).}
-\label{fig:T0559}
-\end{figure}
+\bibitem{Moult2011}
+Moult, J.; Fidelis, K.; Kryshtafovych, A.; Tramontano, A. Critical assessment of methods of protein structure prediction (CASP)-round IX.
+Proteins, 2011, 79, 1-5.
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.8 in,height=3.0 in]{T0538.pdf}
-\end{center}
-\caption{The native and three model structure of engineered protein from Asr4154 protein (PDB ID: 2L09 and CASP code:T0538). The model 1,2 and 3 are
-from the group PconsR, Shell and FOLDIT respectively.}
-\label{fig:T0538}
-\end{figure}
+\bibitem{West2007}
+West, A.M.; Elber, R.; Shalloway, D. Extending molecular dynamics time scales with milestoning: example of complex kinetics
+in a solvated peptide. J Chem Phys. 2007, 126, 145104-1451014.
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=4.9 in,height=4.0 in]{Target_538_compare.pdf}
-\end{center}
-\caption{Difference of the sidechain orientation between model 1 and crystallographic structure in Target-538.
-A. The salt bridge between Glu-35 and Arg-32 in crystal structure. In model 1, this is compensated by H-bonding between 
-Arg-32 and Glu-28. There is another salt bridge between Glu-28 and Lys-24.
-B. Salt bridge between Arg-26 and Glu-50 in crystal structure. Orientation of Phe-51 is different in crystal than in model1.}
-\label{fig:T0538compare}
-\end{figure}
+\bibitem{Haas2007}
+Haas, K.; Chu, J.W. Decomposition of energy and free energy changes by following the flow of work along reaction path.
+J. Chem. Phys. 2009, 131, 144105-144111.
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.5 in,height=3.0 in]{T0560.pdf}
-\end{center}
-\caption{Native and two model structure of protein BT2368 from Bacteroides thetaiotaomicron (pdb id: 2L02 and CASP code: T0560). The two models
-were from the group "Splicer".}
-\label{fig:T0560}
-\end{figure}
+\bibitem{Jonsson1998}
+Jónsson, H.; Mills, G.; Jacobsen, K.W. Nudged Elastic Band Method for Finding Minimum Energy Paths of Transitions,
+in Classical and Quantum Dynamics in Condensed Phase Simulations, Ed. B. J. Berne, G. Ciccotti and D. F.
+Coker, 385 (World Scientific, 1998).
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.5 in,height=3.0 in]{T0540.pdf}
-\end{center}
-\caption{X-ray crystallographic structure and two submitted models of fas apoptosis inhibitory protein (pdb id: 3MX7 and CASP code: T0540). 
-Model 1 and Model 2 in this analysis were submitted by the group LTB and MUFOLD respectively.}
-\label{fig:T0540}
-\end{figure}
+\bibitem{E2007}
+E, W.; Ren, W.; Vanden-Eijnden, E. Simplified and improved string method for computing the minimum energy paths in
+barrier-crossing eventsi. J. Chem. Phys. 2007, 126, 164103.
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.5 in,height=5.7 in]{T0531.pdf}
-\end{center}
-\caption{The native structure and 5 models of extracellular domain of the jumping translocation
-breakpoint protein (pdb id: 2KJX and the CASP code: T0531).}
-\label{fig:T0531}
-\end{figure}
+\bibitem{Dellago2002}
+Dellago, C.; Bolhuis, P.G.; Geissler, P.L. Transition Path Sampling, Adv. Chem. Phys. 2002, 123, 1-84.
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.2 in,height=3.8 in]{T0569.pdf}
-\end{center}
-\caption{Native and best model structure of a domain of adhesion exoprotein from Pediococcus pentosaceus (pdb id: 2KWY and CASP code: T0569).
-The best model was from the group "Mufold".}
-\label{fig:T0569}
-\end{figure}
+\bibitem{Cheng2006}
+Cheng, X.; Wang, H.; Grant, B.; Sine, S.M.; McCammon, J.A. Targeted Molecular Dynamics Study of C-Loop Closure
+and Channel Gating in Nicotinic Receptors. 2006, 9, 134.
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=5 in,height=4 in]{Target_569_per_residue.pdf}
-\end{center}
-\caption{A. Plot of per residue free energy difference between the model and the crystal structure of Traget 569 (pdb id: 2KWY) of CASP9.
-B. In model 1, the beta sheet containing Val-59 and Ile-61 is disordered. These hydrophobic residues are exposed to the solvent.
-C. Glu-67 is partially exposed to the solvent in crystal and it is fully exposed in model 1.
-D. There is a H-bond between Lys-76 and Asp-11 in model 1, which is absent in the crystallaographic structure.}
-\label{fig:T0569_per_residue}
-\end{figure}
+\bibitem{Elber2005}
+Elber, R. Long-timescale simulation methods. Cur. Opi. in Str. Biol. 2005, 15, 151-156.
 
-\begin{thebibliography}{99}
+\bibitem{Torrie1977}
+Torrie, G. M.; Valleau, J. P. Nonphysical sampling distributions in Monte Carlo free-energy estimation: Umbrella sampling
+(1977) J. Comput. Phys. 23, 187
+
+\bibitem{Tironi1994}
+Tironi, I.G.; van Gunsteren, W.F. A molecular-dynamics simulation study of chloroform. Mol. Phys. 1994, 83, 381-403.
 
@@
 Current Opinion in Structural Biology, 2007, 17, 181-186.
+
+\bibitem{Mobley2007}
+Mobley, D.L.; Chodera, J.D.; Dill, K.A. The combining and release method: obtaining correct binding free energies in
+the presence of protein conformational change. Journal of Chemical Theory and Computation 2007, 3, 1231-1235.
+
+\bibitem{Mobley2012}
+Mobley, D.L.; Klimovich, P.V. Perspective: Alchemical free energy calculations for drug discovery.
+J. Chem. Phys. 2012, 137, 230901-12.
+
+\bibitem{Mobley2006}
+Mobley, D.L.; Chodera, J.D.; Dill, K.A. On the use of orientational restraints and symmetry corrections in alchemical
+free energy calculations. J. Chem. Phy. 2006, 125, 084902.
 
@@
 Gilson, M.K.; Zhou, H.X. Calculation of protein-ligand binding affinities. Annu Rev Biophys Biomol Struct. (2007) 36, 21-42.
-
-\bibitem{Torrie1977}
-Torrie, G. M.; Valleau, J. P. iNonphysical sampling distributions in Monte Carlo free-energy estimation: Umbrella sampling 
-(1977) J. Comput. Phys. 23, 187
-
-\bibitem{Tironi1994}
-Tironi, I.G.; van Gunsteren, W.F. A molecular-dynamics simulation study of chloroform. Mol. Phys. (1994) 83, 381-403.
 
@@
 \bibitem{Zheng2008}
-Zheng, L.; Chen, M.; Yang, W. Random walk in orthogonal space to achieve efficient free-energy simulation of complex systems, Proc. Natl. Acad. Sci. 2008, 105 (51), 20227.
+Zheng, L.; Chen, M.; Yang, W. Random walk in orthogonal space to achieve efficient free-energy simulation of complex systems, Proc. Natl. Acad. Sci. 200
+8, 105 (51), 20227.
 
-\bibitem{Tyka2006}
-Tyka, M.; Clarke, A.; Sessions, R. An Efficient, Path-Independent Method for Free-Energy Calculations. J.Phys.Chem. B 2006, 110, 17212-17220.
+\bibitem{Shell2010}
+Shell, S.M. A replica-exchange approach to computing peptide conformational free energies. Mol. Sim. 2010, 7, 505-515.
+
+\bibitem{Tyka2006}Tyka, M.; Clarke, A.; Sessions, R. An Efficient, Path-Independent Method for Free-Energy Calculations. J.Phys.Chem. B 2006, 110, 17212-17220.
 
 \bibitem{Cecchini2009}
-Cecchini, M., Krivov, S.V., Spichty, M., Karplus, M. Calculation of free-energy differences by confinement simulations. Application to peptide conformers. 
-J. Phys. Chem. B 113, p. 9728-9740 (2009).
+Cecchini, M., Krivov, S.V., Spichty, M., Karplus, M. Calculation of free-energy differences by confinement simulations. Application to peptide conformers. J. Phys. Chem. B. 2009, 113, 9728-9740.
+
+\bibitem{Ovchinnikov2013}
+Ovchinnikov, V.; Cecchini, M.; Karplus, M. A Simplified Confinement Method for Calculating Absolute Free Energies
+and Free Energy and Entropy Differences. J. Phys. Chem. B. 2013, 117, 750-762.
+
+\bibitem{Spichty2010}Spichty, M.; Cecchini, M.; Karplus, M. Conformational Free-Energy Difference of a Miniprotein from Nonequilibrium
+Simulations. J. Phys. Chem. Lett., 2010, 1, 1922-1926.
 
@@
 
-\bibitem{Alexander2007}
-Alexander, P.A.; He, Y.; Chen, Y.; Orban, J. Bryan, P. The design and characterization of two proteins with $88 \%$ sequence identity but different 
-structure and function. Proc. Natl. Acad. Sci. 2007, 104 (29), 11963-11968.
+\bibitem{Alexander2007}Alexander, P.A.; He, Y.; Chen, Y.; Orban, J.
+Bryan, P.
+The design and characterization of two proteins with $88 \%$ sequence identity but differentstructure and function.
+Proc.
+Natl.
+Acad.
+Sci.
+2007, 104 (29), 11963-11968.
 
@@
 \bibitem{Alexander2009}
-Alexander, P.A.; He, Y.; Chen, Y.; Orban, J. Bryan, P. A minimal sequence code for switching protein structure and function. Proc. Natl. Acad. Sci. 2009, 106(50), 21149-21154.
+Alexander, P.A.; He, Y.; Chen, Y.; Orban, J. Bryan, P. A minimal sequence code for switching protein structure and function. Proc. Natl. Acad. Sci. 2009
+, 106(50), 21149-21154.
 
-\bibitem{Shortle20009}
+\bibitem{Bryan2010}
+Bryan, P.N.; Orban, J. Proteins that switch folds. Curr Opin Struct Biol. 2010, 20(4), 482-488.
+
+\bibitem{He2012}
+He, Y.; Chen, Y.; Alexander, P.A.; Bryan, P.N.; Orban, J. Mutational tipping points for switching protein folds and functions. Structure. 2012, 20(2), 2
+83-91.
+
+\bibitem{Shortle2009}
 Shortle, D. One sequence plus one mutation equals two folds. Proc. Natl. Acad. Sci. 2009, 106(50), 21011-21012.
@@
 \bibitem{Sheffler2009}
-Sheffler, W.; Baker, D. RosettaHoles: Rapid assessment of protein core packing for structure prediction, refinement, design, and validation. Protein Science. 2009, 18(1), 229-239.
+Sheffler, W.; Baker, D. RosettaHoles: Rapid assessment of protein core packing for structure prediction, refinement, design, and validation. Protein Sci
+ence. 2009, 18(1), 229-239.
+
+\bibitem{Case2012}
+D.A.
+Case, T.A.
+Darden, T.E.
+Cheatham, III, C.L.
+Simmerling, J.
+Wang, R.E.
+Duke, R.
+Luo, R.C.
+Walker, W.
+Zhang, K.M.
+Merz, B.
+Roberts, S.
+Hayik, A.
+Roitberg, G.
+Seabra, J.
+Swails, A.W.
+Goetz, I.
+Kolossváry, K.F.
+Wong, F.
+Paesani, J.
+Vanicek, R.M.
+Wolf, J.
+Liu, X.
+Wu, S.R.
+Brozell, T.
+Steinbrecher, H.
+Gohlke, Q.
+Cai, X.
+Ye, J.
+Wang, M.-J.
+Hsieh, G.
+Cui, D.R.
+Roe, D.H.
+Mathews, M.G.
+Seetin, R.
+Salomon-Ferrer, C.
+Sagui, V.
+Babin, T.
+Luchko, S.
+Gusarov, A.
+Kovalenko, and P.A.
+Kollman (2012), AMBER 12, University of California, San Francisco.
+
+\bibitem{Goetz2012}
+Goetz, A.W.; Williamson, M.J.; Xu, D.; Poole, D.; Le Grand, S.; Walker, R.C. Routine microsecond molecular dynamics simulations with AMBER - Part I: Generalized
+Born. J. Chem. Theory Comput. 2012, 8(5) 1542.
 

### 2013-04-24 Paper/supporting_information.tex
correction_24thApril

% vim:set textwidth=100:
% vim:set fo+=t:

\documentclass[12pt]{article}

\usepackage{amsmath}
\usepackage[nofiglist,notablist]{endfloat}
\usepackage[usenames,dvipsnames]{color}
\usepackage{color}
\usepackage{authblk}
\usepackage{graphicx}
\usepackage{palatino}
\usepackage[activate={true,nocompatibility},final]{microtype}
\usepackage[super,sort&compress]{natbib}
\pagenumbering{arabic}
\parskip = 0.08in \parindent = 0.0in

% Custom macros for author comments
\newcommand{\Alberto}[1]{\color{ForestGreen}#1\normalcolor }
\newcommand{\Justin}[1]{\color{blue}#1\normalcolor}
\newcommand{\Arijit}[1]{\color{magenta}#1\normalcolor}
\newcommand{\Ken}[1]{\color{red}#1\normalcolor}

\author{Arijit~Roy}
\author{Alberto~Perez}
\author{Ken~A.~Dill}
\author{Justin~L.~MacCallum}
\affil{Laufer Center for Physical and Quantitative Biology\\
    and Departments of Physics and Chemistry\\
    Stony Brook University\\
    Stony Brook, NY 11794-5252.}

\title{Predicting the conformational preferences of proteins using a physics-based free energy
method}

\begin{document}

\maketitle

\section{Preparation of input model for chamelon sequences}

We prepared computer generated models in both $\alpha$ and $\beta$ conformation for all five chamelon sequences mentioned 
in the main text. Crystallographic structure of GA95 and GB95 sequence have $\alpha$ and $\beta$ fold respectively. When,
we attempt to prepare the $\alpha$ and $\beta$ fold with GA95 sequence we start with the crystallographic 
structure of GA95 and GB95 with GA95 sequence. However, we remove the sidechains at three mutated positions in both the fold 
and call the program SCWRL4 \cite{Krivov2009} to generate the sidechain conformation of the mutated residues and its neighbors. Followed by 
this step, we do long molecular dynamics
minimization to remove any possible bad contact. This procedure was followed to generate all $\alpha$ and 
$\beta$ conformations of five sequences that are used for free energy calculation.

\begin{figure}
\includegraphics[width=5.6 in,height=5.3 in]{orban_perres_SI.pdf}
\label{fig:orban_full}
\caption{Per residue free energy calculation indicate significant differences of sidechain orientation of key amino acid residues. These amino acid
residues control the free energy equilibrium to either $\beta$ or
$\alpha$ conformation. All figures in the left are in $\beta$ conformation, whereas the right ones are in $\alpha$ conformation except in Figure (F).
Sidechains of hydrophobic residues Leu-7 in (A) and Ala-26 in (B) are oriented towards the protein interior in $\beta$ 
conformation. In such conformation they have hydrophobic interactions with other hydrophobic residues colored in white. In $\alpha$ conformation these residues are 
exposed to the protein surface. 
(C) Tyr-45 forms H-bond with Asp-47 with GB95 sequence and $\beta$ conformation. (D) Gln-11 and Glu-15 form a H-bond in $\beta$ conformation which is absent
in the $\alpha$ conformation. (E) Hydrophobic Ile-49 is exposed to the protein surface in $\beta$ conformation, whereas it is inside protein hydrophobic core in
$\alpha$ conformation. (F) Sidechain of Ile-30 is inside the protein hydrophobic core in as it has a smaller sidechain with GA95 sequence in $\alpha$ conformation. 
In contrast, Phe-30 with larger sidechain is exposed to the protein surface with GB95 sequence.}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=3.6 in,height=3.4 in]{T0569.pdf}
\end{center}
\caption{Native and best model structure of a domain of adhesion exoprotein from Pediococcus pentosaceus (pdb id: 2KWY and CASP code: T0569).
The best model was from the group "Mufold".}
\label{fig:T0569}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=5 in,height=4 in]{T0569_perres2.pdf}
\end{center}
\caption{Key differences between native crystal structure and the model as indicated by per residue free energy calculation.
The sidechain of hydrophobic residues, Val-59 and Ile-61 are oriented towards the protein hydrophobic core in (A) crystal 
structure, (B) exposed to the surfure in model. The beta sheet containing these residues is disordered in the generated model.
There is a H-bond between Lys-76 and Asp-11 in (D) model structure, which is absent in the (C) native structure.}
\label{fig:T0569_per_residue}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=4.9 in,height=4.0 in]{Target_538_compare.pdf}
\end{center}
\caption{Differences of the sidechain orientation in model 1 and crystallographic structure of CASP target T0538 as determined 
by per residue free energy calculation.  
(A) Two salt bridges between residues Glu-35-Arg-32 and Glu-28-Lys-24 exist in the native crystallographic structure. These are
absent in the model structure. Instead
(B) a new salt bridge is formed between residues Arg-32-Glu-28 in model. (C) Another salt bridge between
residues  Arg-26-Glu-50 in the native structure, (D) This salt bridge is absent in the model structure. In comparison to the crystallographic 
structure, orientation of Phe-51 differs in model1.}
\label{fig:T0538compare}
\end{figure}

\Ken{[This fig isn't very readable.  We need to focus into the sidechains, keeping only the nearby pieces of chain and leaving the rest out or making it fuzzy or something.  We need to see some point that you are making by looking at the figure.]}

\begin{thebibliography}{99}

\bibitem{Krivov2009}
Krivov, G.G.; Shapovalov, M.V.; Dunbrack, RL Jr. Improved prediction of protein side-chain conformations with SCWRL4.
Proteins. (2009) 77, 778-795.

\end{thebibliography}

%\end{suppinfo}
\end{document}

### 2013-04-24 Paper/paper.tex

@@
        (20 ns long each), where the harmonic restraint force constant was scaled from 0.00005
-       \Alberto{units??}
-       (mostly free) to 81.92 (frozen in one microstate). In this final restrained state, the
+       Kcal/Mol 
+       (mostly free) to 81.92 iKcal/Mol (frozen in one microstate). In this final restrained state, the
        rotational contribution to the free energy is frozen out
@@
 Wang, Q.; Vantasin, K.; Xu, D.; Shang, Y. MUFOLD-WQA: A new selective consensus method for quality assessment in protein structure prediction.
-Proteins, 2011, 79: 185–195.
+Proteins, 2011, 79: 185-195.
 

### 2013-04-24 Paper/paper.tex
changed CASP summery

@@
 Second, we looked at 6 target proteins from the CASP9 experiment.  For each target, we examined up to 5 submitted models.  We 
-computed the difference free energy between the true native and the best model.  Table~\ref{table:casp_control} shows that, in 5 
+computed the difference free energy between the true native and the best model. Figure~\ref{fig:summary_casp} and 
+Table~\ref{table:casp_control} shows that, in 5 
 out of 6 cases, the confinement method assigns a lower free energy to the experimentally determined structure than to any of the 
@@
  it is just an independent useful validation that the confinement calculations make sense.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.5 in,height=3.2 in]{casp.pdf}
+\end{center}
+\caption{In 5 out of 6 CASP Target, the confinement method assigns a lower free energy to the experimentally determined native structure.} 
+\label{fig:summary_casp}
+\end{figure}
 

### 2013-04-26 Paper/correction.tex
changed method & citation

--- 2013-04-24 Paper/paper.tex
@@
-% vim:set textwidth=100:
-% vim:set fo+=t:
+\bibitem{Handl2009}
+Handl, J.; Knowles, J.; Lovel, S.C. Artefacts and biases affecting the evaluation of scoring 
+functions on decoy sets for protein structure prediction. Bioinformatics, 2009, 25, 1271-1279.
 
-\documentclass[12pt]{article}
+\bibitem{Hornak2006}
+Hornak, V.; Abel, R.; Okur, A.; Strockbine, B.; Roitberg, A.; Simmerling, C. Comparison of multiple Amber force fields 
+and development of improved protein backbone parameters. Proteins., 2006, 65, 712-725.
 
-\usepackage{amsmath}
-\usepackage[nofiglist,notablist]{endfloat}
-\usepackage[usenames,dvipsnames]{color}
-\usepackage{color}
-\usepackage{authblk}
-\usepackage{graphicx}
-\usepackage{palatino}
-\usepackage[activate={true,nocompatibility},final]{microtype}
-\usepackage[super,sort&compress]{natbib}
-\pagenumbering{arabic}
-\parskip = 0.08in \parindent = 0.0in
-
-% Custom macros for author comments
-\newcommand{\Alberto}[1]{\color{ForestGreen}#1\normalcolor }
-\newcommand{\Justin}[1]{\color{blue}#1\normalcolor}
-\newcommand{\Arijit}[1]{\color{magenta}#1\normalcolor}
-\newcommand{\Ken}[1]{\color{red}#1\normalcolor}
-
-\author{Arijit~Roy}
-\author{Alberto~Perez}
-\author{Ken~A.~Dill}
-\author{Justin~L.~MacCallum}
-\affil{Laufer Center for Physical and Quantitative Biology\\
-    and Departments of Physics and Chemistry\\
-    Stony Brook University\\
-    Stony Brook, NY 11794-5252.}
-
-\title{Predicting the conformational preferences of proteins using a physics-based free energy
-method}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-
-Calculation of free energy differences is of central importance in the simulation of biochemical
-systems. It is particularly difficult to calculate between pairs of macromolecular conformations as
-well as a computationally expensive task with existing methods. In this work, confinement approach
-is used to calculate absolute free energies of biomolecular systems. This method provides two main
-advantages: it does not require a reaction coordinate or transition path and it is fast to compute.
-Free energy calculated can be decomposed into a per residue contribution in an approximate way. Per
-residue free energy allows us to identify the reason behind conformational preferences in
-biomolecules. Through out the article we show its use in different challenging modeling problems. In
-particular, we show its use in predicting the conformational preference of chamaleon sequences
-(sequences with high sequence identity and different folds). This sequence dependent conformational
-preferences and per residue free energy decomposition set the stage for the use of this method in
-protein design.
-
-\end{abstract}
-
-\Ken{[Guys - I recommend we get a better name than confinement.  It's totally uninformative.  In my view, better would be `confine-and-configure' or something.  
-Please give it some thought.]}
-
-\section{Introduction}
-
-In some problems of protein science, you want to know the relative stability of a protein's conformation $A$ compared to it's 
-conformation $B$.  We call this the \emph{difference free energy} (DFE). For example, in allosteric mechanisms, a protein often 
-adopts one conformation when a ligand is bound and another conformation when no ligand is bound \cite{Elber2007}. To understand the mechanism requires knowledge 
-of the relative free energies of the protein conformations.  Another example is in protein structure prediction, for example as practiced in 
-the CASP competition \cite{Moult2011}.  If you are using some computational model and have predicted two putative native structures, $A$ 
-and $B$, you want to compute which has the lower free energy, in order to know which is the more native-like.  In a third example, you may 
-want to know whether mutating a few amino acids in a protein could cause the protein to switch from one stable conformation to another, because 
-that could have important consequences for biological mechanisms and disease.  And fourth, sometimes binding a ligand induces a protein from 
-one conformation to another.  A quantitative understanding of such \emph{induced-fit} situations requires knowledge of the difference free 
-energy of the protein in the two states. In all these cases, it would be useful to have a computational method that can be used efficiently 
-with atomically detailed physical forcefield models to compute the free energy differences between different given protein conformations.
-
-Methods like thermodynamic integration \cite{Tironi1994},\cite{Meirovitch2007} which are successful in alchemical free energy calculation
-may not be as useful for calculation of conformational free energy.
-One widely explored strategy is to use molecular dynamics simulations along some putative reaction coordinate pathway from conformation $A$ 
-to $B$ \cite{Elber2007,West2007,Haas2007,Jonsson1998,E2007,Dellago2002,Cheng2006}. \Ken{we need 
-many refs here: cite Elber's milestoning, Jhih-Wei Chu, and the many other path-sampling approaches.]} The free energy along this reaction 
-coordinate can then be determined using methods such as umbrella sampling \cite{Torrie1977}. Such approaches have several limitations.  
-First, it is necessary to know a reaction pathway from $A$ to $B$. If conformations $A$ and $B$ are quite different, then it can be 
-challenging to find such paths. Second, these methods 
-are computationally slow. To get an accurate estimate of the total free energy difference $\Delta G = G_B - G_A$ requires accurate determinations 
-of the many small free energy differences $A \rightarrow 1 \rightarrow 2 \ldots \rightarrow B$, so each microstep requires substantial amounts 
-of sampling.  And third, these methods are prone to large errors, because the pathway error is a sum of many errors among the many microsteps.
-
-\Arijit{Will it be useful to add methods like PMF or replica exchange based free energy methods ?}
-
-\Ken{[Are all these methods below ones that don't involve paths?  Also, are there other key methods we should cite?  We need a good scholarly 
-list of all the relevant works.]} 
-The calculation of protein conformational free energy has been successfully attempted by a number of
-groups \cite{Shell2010,Ytreberg2006,Zheng2008,Spichty2010,Strajbl2000,Park2008,Tyka2006,Cecchini2009,Ovchinnikov2013}. 
-Some of these methods like reference system method \cite{Ytreberg2006}, deactivated morphing \cite{Park2008}, confinement 
-method \cite{Tyka2006,Cecchini2009,Ovchinnikov2013} take an alternative strategy for computing difference free energies 
-that does not require knowing a pathway from $A$ to $B$.  
-\Ken{[please add the several new Karplus papers from 2012-2013.]}.
-
-Here, we adopt the confinement method of Tyka et al \cite{Tyka2006} and Cecchini et al.\cite{Cecchini2009}. This method 
-also resembles the `confine-and-release' method \cite{Mobley2007}, \cite{Mobley2012}, \cite{Mobley2006}.  
-\Ken{[cite mobley's jctc 07 (ref 216 on our labsite), and mobley's recent j chem phys review, 
-and mobley (ref 199) j chem phys, and others that are relevant.]}. The method is based on the thermodynamic cycle showns in 
-Figure \ref{fig:method}.  To start, conformation $A$ is an ensemble.  We reduce ensemble $A$ to one particular microstate $A^\ast$ 
-in that ensemble by applying restraints in an MD simulation.  In the same way, we reduce conformational ensemble $B$ to one of its 
-microstates, $B^\ast$.  Then, we simply compute the free energy difference between the microstates 
-as $\Delta G_{A^\ast, B^\ast} = \Delta H_{A^\ast, B^\ast}$ (which is just an enthalpy difference since there is no conformational 
-entropy difference between the microstates).  $\Delta H_{A^\ast, B^\ast}$ is readily calculated using normal mode analysis. In 
-this way, we can calculate the free energy difference between the two end states without needing to define a physical path or reaction 
-coordinate connecting the two states.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.5 in,height=3.5 in]{method.pdf}
-\end{center}
-\caption{Graphical representation of the thermodynamic cycle involving confinement method.}
-\label{fig:method}
-\end{figure}
-
-\Ken{For this fig, let's make A and A* green and B and B* red or something, to distinguish them visually.
-Let's also blur the ensemble more relative to the microstate, so it's clearer the distinction.]}
-
-Previously, the confinement method has been validated on small model peptides \cite{Tyka2006,Cecchini2009}. Here, we show that the confinement 
-method can also be applied successfully to larger systems, such as the series of chameleon proteins designed by Orban and co-workers
-\cite{Alexander2007,He2008,Alexander2009,He2012}, which can switch between two completely different folds with only few-amino-acid changes 
-in sequence.  We also show that our computed difference free energies are useful in evaluating the quality of CASP target protein predictions.  
-This could be of significant value for ultimately improving the energetics in protein-structure models, not just the structures.  And, 
-finally, we show that we can approximately decompose difference free energies into individual amino acid components.  This offers the 
-opportunity for diagnosing the structural basis for difference free energies, which can be useful for interpreting biological mechanisms.
-
-\section{Results and Discussion}
-
-\subsection{The confinement method succeeds at some basic consistency checks}
-
-First, we validated that our implementation of the confinement method produces results compatible with previous calculations reported in the 
-literature.  The method has previously been applied to a 16 amino acid residue $\beta$-hairpin from protein G, known as 
-BHP \cite{Cecchini2009}. We calculated the free energy difference between the native conformation (called bhp1), which has 
-a two-stranded $\beta$-sheet and a three-stranded $\beta$-sheet called bhp3.  Our confinement calculation shows that bhp1 
-is more stable by 1.7 kcal/mol, consistent with 4 $\mu$s equilibrium simulations showing that bhp1 is more favorable configuration 
-by 1.8 kcal/mol, and in agreement with previous calculations \cite{Cecchini2009}.
-
-Second, we looked at 6 target proteins from the CASP9 experiment.  For each target, we examined up to 5 submitted models.  We 
-computed the difference free energy between the true native and the best model.  Table~\ref{table:casp_control} shows that, in 5 
-out of 6 cases, the confinement method assigns a lower free energy to the experimentally determined structure than to any of the 
-decoys.  Other well-known discriminators  can also successfully make this recognition \cite{Sheffler2009}; 
- it is just an independent useful validation that the confinement calculations make sense.
-
-\begin{table}
-\begin{center}
-\caption{The confinement method assigns a more favorable free energy to the experimentally
-    determined structure than to computer-generated predictions. For each target, we examined as
-    many as five predictions submitted by CASP participants. We report the free energy difference
-    between the most favorable decoy and the experimentally determined structure. Positive
-    $\Delta\Delta G$ values indicate that the experimental structure is predicted to be more
-favorable than any of the decoys.}
-\label{table:casp_control}
-\begin{tabular}{l l l}\hline
-    CASP Target  & PDB Identifier & $\Delta \Delta G = G_{best~decoy} - G_{native}$ (kcal/mol) \\ \hline
-     T0531       &    2KJX        &          $11.15 \pm 0.70$ \\ \hline
-     T0538       &    2L09        &          $-3.00 \pm 0.47$ \\ \hline
-     T0540       &    3MX7        &          $16.94 \pm 0.49$ \\ \hline
-     T0559       &    2L01        &          $2.10 \pm 0.24$ \\ \hline
-     T0560       &    2L02        &          $18.00 \pm 0.49$ \\ \hline
-     T0569       &    2KWY        &          $20.01 \pm 0.69$  \\ \hline
-\end{tabular}
-\end{center}
-\end{table}
-
-\Ken{[ I think a better notation is $\Delta \Delta G = G_{best~decoy} - G_{native}$, rather than having the arrow.]}
-
-\subsection{The confinement method correctly predicts the structures of chameleon sequences}
-
-We also tested the confinement method predictions for difference free energies on the chameleon sequences of Orban 
-et al. \cite{Alexander2007,He2008,Alexander2009,Bryan2010,He2012,Shortle2009}; these 
-are instances in which two highly similar sequences fold into remarkably different structures. Orban and co-workers have designed a 
-protein-G-like sequence of 56-residues that
-is marginally stable in one of two possible folds. By mutating key residues in this sequence they
-are able to stabilize one fold or the other (see Figure~\ref{fig:orban}). We call sequences ``$\beta$'' that fold into 
-the protein-G-like that fold into a $4 \beta + \alpha$ fold, and we call sequences ``$\alpha$'' that fold into $3 \alpha$ helical folds.  One pair 
-of sequences (GA88/GB88) is 88 percent identical in sequence, differing in seven positions. Another pair (GA95/GB95) is 95 percent identical, 
-differing in three positions. The third pair (GA98/GB98) differs in only a single tyrosine-to-alanine mutation. Accurately predicting the 
-structural preferences of these structures presents a serious challenge for computational methods 
-\cite{Alexander2007,He2008,Alexander2009,Bryan2010,He2012,Shortle2009}.
-
-\begin{figure}
-\includegraphics[width=5.0in]{orban.pdf}
-\label{fig:orban}
-\caption{Confinement method correctly predicts the structural preferences of four chameleon
-sequences. The top part of the figure represent four sequences used in this study along with the
-protein data bank id. Each sequence adopts either a three-helix bundle fold (denoted $\alpha$) or
-$4 \beta + \alpha$ fold (denoted $\beta$). The relative free energies of
-the two folds are reported for each sequence.}
-\end{figure}
-
-We initially approached this problem by making a model of each sequence with the same backbone
-structure as its partner chameleon sequence. For example, we took the sequence of GA88 and built a
-model with the same overall structure as GB88. We then used the confinement method to assess the
-free energy difference between the experimentally determined structure of GA88 and the model (with
-the GA88 sequence and the GB88 structure). The confinement method was able to predict the
-conformational preferences correctly for all four sequences (data not shown). This is, however, not
-surprising. It is well known [CITE]\Arijit{still searching this citation} that it is easy to distinguish 
-computational models from native structures. To avoid this potential problem, we instead computed relative free energy of two
-different computer generated models for each sequence. One model is based on the $\alpha$ structure and
-the other on the $\beta$ structure (see Supporting Information for details on the modeling procedure).
-This is a much more realistic test of the confinement method's ability to accurately calculate
-relative free energies.
-
-\Ken{[ It's not clear what this test is and why it's better, and why there are only 5 sequences.  Please clarify.]}
-\Arijit{ Now there are 4 sequences as they have the experimentally available structure. The fifth sequence is discussed 
-at the end of the paragraph.}
-
-Figure~\ref{fig:orban} shows that the confinement method identifies the correct structure for all four sequences. One hypothesis for fold switching of
-chameleon sequences is that the structural transitions require states with diminished stability
-\cite{Bryan2010}. It is believed that if the free energy of the native state and the alternative state are
-within a range of 5 kcal/mol, then it is possible for the native state to be destabilized relative
-to the alternative state with only small changes in sequence. The stability of the native state can
-decrease for a number reason ranging from chemical modification, breaking of disulphide bonds or
-mutations---as is the case here. The calculated free energy differences range from around 3.5 to 5.0
-kcal/mol, which is consistent with the above hypothesis \cite{He2008,Alexander2009,Bryan2010}. 
-In a more recent study\cite{He2012}, the amino acid residue at position 45 (Tyr for $\beta$ and Leu for $\alpha$) was
-found to be important for switching between  $\alpha$ and $\beta$ conformation. This inspired us to 
-introduce another mutation at 45 position. For our future reference this Tyr-45-Ala mutation of GB95 sequence will
-be known as GB98. Interestingly, with this mutation, the equilibrium shifts to $\alpha$ conformation as it gains 3.84 Kcal/Mol 
-stability compare to $\beta$ conformation and indicate the importance of Tyr-45 residue. It will be interesting to 
-know the actutal experimental structure which can validate our result.
-
-\subsection{Per-residue free energy calculation can identify the mechanistic detail behind
-conformational preferences} \label{sec:orban_per_residue}
-
-To understand the mechanism behind the conformational preferences of the different protein
-sequences, we decomposed the calculated free energy into per-residue contributions in
-an approximate way \cite{Tyka2006}. For this purpose, the confinement free energy, $\Delta G_{A,A*}$ and
-$\Delta G_{B,B*}$ of each residue is calculated in the usual numerical way as described
-the method section. We call this method approximate as we ignore the contribution from the normal 
-mode or quasiharmonic analysis. Instead, decomp module of amber is used to calculate the internal energy 
-of each residue from the final restrained trajectory.
-
-\Ken{[Arijit -- I didn't find this methods calculation.  We should at least put in a sentence or two here summarizing it.]}  
-This helps us to identify important residues that stabilize a particular
-conformation. The per residue free energy, $\Delta \Delta G ((\beta) - (\alpha))$ is
-shown in Figure~\ref{fig:perresidue_orban}. 
-In this plot, residues colored in blue favor the $\alpha$ structure, residues in red favor the $\beta$ structure, and white residues
-have no preference.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=4.5in]{delta_g.pdf}
-\end{center}
-\caption{The $\alpha$ and $\beta$ conformation associated with GA95 and GB95 sequences are colored according to the relative per residue
-free energy value. In this figure, residues colored in blue favor the $\alpha$ structure, residues in red favor the $\beta$ structure, and 
-white residues have no preference. Some of the mutated residues are also labelled.}
-\label{fig:perresidue_orban}
-\end{figure}
-
-Although the overall free energy difference between the two structures is small (within 3-5 Kcal/Mol), the individual residues 
-can show marked preferences for being in either $\alpha$ or $\beta$
-conformation. Such differences can be undestood by looking in detail at the local environemnt for
-those resiudes. For example, the region around residue 7 forms a random coil in the $\alpha$
-structure, whereas it is forming a beta sheet structure in the $\beta$ structure. These residues
-strongly favour the locally well packed and hydrogen bonded environment found in the $\beta$ sheet.
-Overall, favoring $\alpha$ or $\beta$ structure is a delicate balance, were the relative global free
-energy difference is small and the contributions of different per residue tendencies balance out. It
-is therefore very likely that by changing key resiude preferences the global fold preferences can be
-changed.
-
-Three residues are mutated between GA95 and GB95, at positions 20, 30, and 45. In GA95, these
-residues (L20, I30 and L45) stabilize the $\alpha$ structure. In GB95, two of these residues (F30
-and Y45) favor the $\beta$ structure, because they have large solvent exposed surface areas in the
-$\alpha$ structure, but are more buried in the $\beta$ structure. Additionally Y45 forms a hydrogen
-bond with D47 in the $\beta$ structure. On the other hand, residue A20 from GB95 still favors the
-$\alpha$ structure, although not as strongly as L20 in GA95.
-
-The $\alpha$ and $\beta$ sequences are nearly identical, so there are some common features observed for all
-sequences. Roles of some important residues in stabilizing either the $\alpha$ or $\beta$
-conformation and possible reasons for such conformational preferences are summerized in 
-supporting information Figure S1. The experimental observations
-classified the protein into two parts: Amino acids 9-51 are fully structured in both folds, whereas
-residues 1-8 and 52-56 are unstructured in the $\alpha$ fold, but form a $\beta$-strand in the
-$\beta$ fold. Most of the amino acid residues in the region 1-9 have negative per-residue free
-energies, which means that these residues favor the $\beta$ structure.
-
-In addition to the direct effects of the mutation, there are also indirect effects due to small
-perturbations due to the mutations. For example, the L20A mutation causes a slight repacking around
-residue 20. This causes large changes in the per-residue free energies of nearby residues T25 and
-A26. However, these differences largely cancel and the overall effect of the L20A mutation is weaker
-than either the I30F or L45Y mutations.
-
-Such per-residue free energy decompositions allow us to understand the driving forces behind protein
-folding and conformational change and may be useful for designing proteins with specific structures
-and functions.
-
-\subsection{Applcation of confinement method for structure prediction}
-
-We applied the confinement method as a ``meta-predictor'' for structure prediction. Here, the
-task is to correctly identify the most accurate models out of a set of ``decoys'' generated by
-different methods during the Critical Assessment of Structure Prediction (CASP) experiment. CASP is
-a blind test in which different groups apply methods to predict the 3-dimensional structure of
-proteins from their sequences. Each group is allowed to submit five
-possible structures, which they are supposed to rank from best to worst. We have performed two
-experiments centered on CASP.
-
-In the first experiment, we examined several targets and tried to rank-order the predictions
-made by the same group (same methodology used to generate the models), in others the
-structures were drawn from several different groups (different methodology used). The goal of this 
-experiment is to determine whether 
-the physics-based confinement method can correctly identify more native-like structures as having
-low free energies.
-
-Our second experiment was to see if the confinement method can identify structures that are missed
-by other meta-prediction servers. Most successful meta-prediction servers are based on the idea of
-consensus: if many different prediction methods produce similar results, then that is probably a
-correct prediction. This is often a powerful heuristic, but it can miss cases where there is a very
-good result that is only predicted by one method. We chose several cases where such structures were
-missed and assessed if the confinement method can correctly identify these accurate models.
-
-As is common in the CASP experiment, we assess our results in terms of Global Distance Test Total
-Score (GDT-TS) \cite{Zemla2003}, which is a C$\alpha$ based measure of structural accuracy. It can
-be understood roughly as the percentage of residues that are correctly positioned in the model
-(range 0 to 100, higher is better). The initial models for the test was chosen from different server 
-groups those have traditionally done well in past CASP events.
-
-\subsubsection{The confinement method can correctly rank different models generated by the same methodology
-and distinguish the native structure}
-
-Here, our task is to rank different submitted models of CASP9 targets that have been generated with
-the same methodology. In absence of any other reliable alternative, we compare all our free energy based 
-ranking with the GDT-TS based ranking which are available in CASP website.
-
-The first test case is a protein BVU3908 from Bacteroides vulgatus whose PDB id and CASP target 
-code are 2L01 and T0559 respectively.
-The best predictor group for this 69 amino acid long target was "BAKER-ROSETTASERVER". We initially checked similarity
-between the models submitted by them and discarded two from the analysis on
-the basis of being too similar to some of the other models. The GDT\_TS score and rmsd values as
-shown in Figure~\ref{fig:T0559} indicate that model 1 was predicted correctly, whereas the order of
-model 3 and 5 was wrong. The main difference between model 3 and rest of the models was the  
-orientation of the first alpha helix. On the other hand, as shown
-in Figure~\ref{fig:T0559} the confinement method not only can differentiate the native
-structure from the submitted models, the ranking also correlates perfectly with the GDT\_TS score.
-
-\Ken{Arijit- we could probably make a better visualization of this figure above by putting it onto a toy landscape: x-axis is rmsd, y-axis is DDG, and above each of the 4 points on this graph will be the ribbon diagram and the GDT-TS.
-I suggest we do that with the fig below too.
-Let's don't use 4 different colors for the fig below.
-Let's just show differences in structures in some different color, or show them all as green.]}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.8 in,height=3.2 in]{T0559.pdf}
-\end{center}
-\caption{The native and three submitted model structure, along with their GDT\_TS, RMSD and relative Free energy values
-of protein BVU3908 from Bacteroides vulgatus (PDB id: 2L01 and CASP code: T0559).}
-\label{fig:T0559}
-\end{figure}
-
-To further test the method we continued with the example of protein BT2368 from Bacteroides
-thetaiotaomicron. The PDB id of this 74 amino acid residue protein is 2L02 and CASP target code is
-T0560. We compared the free energy difference between the native structure and the two of the
-five submitted models from the group "Splicer". The remaining three models were discarded as 
-they were too similar to the rest of the models. As shown in Figure~\ref{fig:T0560}, we can identify 
-the native state correctly and our calculated free energy based ranking matches well with the
-GDT\_TS score.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=4.0 in,height=3.0 in]{T0560.pdf}
-\end{center}
-\caption{Native and two model structure of protein BT2368 from Bacteroides thetaiotaomicron (pdb id: 2L02 and CASP code: T0560). The two models
-were from the group "Splicer".}
-\label{fig:T0560}
-\end{figure}
-
-\subsubsection{The confinement method can rank models generated by different methods}
-
-Our next test was between models that are produced with different methodologies. 
-Here, our model system was a fas apoptosis inhibitory protein molecule whose
-pdb id and CASP target codes are 3MX7 and T0540 respectively. 
-This protein contain 8 beta strands and 90 amino acid residues. 
-The top models from groups "LTB" and "Mufold" was chosen for analysis.
-We will call this as Model 1 and Model 2 for future references. As summerized in
-Figure~\ref{fig:T0540}, the free energy based ranking once again correlates well with the 
-GDT\_TS based score.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=4.0 in,height=3.4 in]{T0540.pdf}
-\end{center}
-\caption{X-ray crystallographic structure and two submitted models of fas apoptosis inhibitory protein (pdb id: 3MX7 and CASP code: T0540).
-Model 1 and Model 2 in this analysis were submitted by the group LTB and MUFOLD respectively.}
-\label{fig:T0540}
-\end{figure}
-
-\subsubsection{Per residue free energy calculation identifies the residues that are responsible for
-differences between two conformations}
-
-\Ken{[Arijit - This and the following sections need to be shortened and made more concise.]}
-
-In section~\protect\ref{sec:orban_per_residue} we discussed how per residue free energy calculation
-can identify the mechanistic detail behind conformational preference of a chameleon sequence. In this section, our aim 
-is to apply the same method and try to understand whether per residue free energy can help us identify
-residues which stabilizes/destabilizes a particular region of a protein. The difference  
-here is that the protein has similar fold with only mismatch in some region. 
-To investigate this, we choose a domain of adhesion exoprotein from Pediococcus pentosaceus (pdb id=2KYW) from protein data bank 
-and its best predicted model of CASP9 (casp id=T0569, submitted by 'Mufold' group which had a GDT\_TS of 78).
-We first calculated the free energy of the whole protein and found that the native 
-structure is stabilized by 20 Kcal/Mol (see suppurting information Figure S2).
-
-Next we proceed with the per residue free energy calculation. Our effort clearly identifies two hydrophobic residues 
-Val-59 and Ile-61, which destabilizes the submitted model with respect to the crystallograohic structure.
-The sidechains of these hydrophobic residues are oriented towards the protein hydrophobic core in the
-native NMR structure but oriented outside protein and solvent exposed in the model. These residues are the primary reason
-why two of the beta sheets in the native structure are disordered in the model structure ( see Figure~\ref{fig:T0569_per_residue}
-and supporting information Figure S3). This method also identifies some other residues which stabilize/destabilize either 
-of the conformation. For
-example, Lys-76 stabilizes the model as it has a H-bond with Asp-11 which is missing in the native structure
-(Figure~\ref{fig:T0569_per_residue} and supporting information figue S3).
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=4.3 in,height=4 in]{T0569_perres1.pdf}
-\end{center}
-\caption{To better understand the reason behind difference in conformation, structure of CASP Target 569 (pdb id: 2KWY) and a 
-generated model was colored according to the per residue free energy value. The amino acid residues that are colored in deep red 
-and deep blue stabilizes the crystal and the model respectively, while the residues with light blue color does not have a preferance.
- Some of the important residues are also labelled.}   
-\label{fig:T0569_per_residue}
-\end{figure}
-
-\subsubsection{Failures in the confinement method}
-
-Despite the great success in most of the studied systems, there are few failures, specially when
-the GDT scores of the compared structures are very close. We observed this for a $54$ amino acid residue long
- engineered protein from Asr4154 protein (pdb id: 2L09 and CASP code T0538).
-Like previous cases, we kept the native structure and include models from the group "PconsR" (GDT\_TS=96.23, model1),
-"Shell" (GDT\_TS = 90.09, model2) and "FOLDIT" (GDT\_TS = 86.32, model3) for analysis. 
-Contrary to our expectation, model1 was found to be more stable than the crystal structure 
-(see Figure~\ref{fig:T0538}). In order to rationalize such obseravtion we calculates per residue free energy
-between crystal structure and model1. The results show us that despite the small variation at the backbone level (as
-shown by high GDT\_TS scores and low RMSDs), the sidechains are oriented in very different ways,
-giving rise to large differences in the stabilization of certain residues. In particular, some of
-the differences arise from different salt bridge patterns (Arg-32 with Glu-35, Glu-28 with Lys-24 and Arg-26 with Glu-50
-in crystal versus Arg-32 with Glu-28 in model 1) and certain flexible polar residues exposed to the
-surface (Lys-24 in model 1, which has an entropic gain from absence of salt bridge and is
-stabilized by interactions with the solvent) (See supporting information Figure S4). All in all, this
-unexpected result just shows us that this method is sensitive to local interactions such as those
-happening from side chain reorientation and also indicate the limitations of GDT\_TS based ranking.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.6 in,height=3.4 in]{T0538.pdf}
-\end{center}
-\caption{The native and three model structure of engineered protein from Asr4154 protein (PDB ID: 2L09 and CASP code:T0538). The model 1,2 and 3 are
-from the group PconsR, Shell and FOLDIT respectively.}
-\label{fig:T0538}
-\end{figure}
-
-\subsubsection{What can we say about low resolution models?}
-
-So far we have seen that the method is good at predicting preferences when the structures are not
-very far from the native. But the question remains how far from native can we go and still see 
-that the method produces correct result. In this section we explored this question with models from  
-extracellular domain of the jumping translocation breakpoint protein (pdb id: 2KJX). Most of the
-group could only generate low resolution models for this CASP9 target (id: T0531) . In our comparison, we choose five
-models by the group MUFOLD-MD, which was the best performing group for this target with their best model 
-had a GDT\_TS value of 44. The result presented in Figure~\ref{fig:T0531} shows: 1.) native is correctly identified as
-expected and 2.) Surprisingly there is a high level of correlation between the GDT ranking and the
-free energy ranking for model 1 and model 3, the rest three structures with GDT\_TS score less than
-35  are ordered incorrectly. It is worth to note that models 2 and 3 have the same GDT and very
-different free energies, meaning that the actual ordering could change a lot \cite{Perez2012}. It is
-encouraging that at least the method can pick out the best model even though it is got a low GDT
-score: 44.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.5 in,height=3.5 in]{T0531.pdf}
-\end{center}
-\caption{The native structure and 5 models of extracellular domain of the jumping translocation
-breakpoint protein (pdb id: 2KJX and the CASP code: T0531).}
-\label{fig:T0531}
-\end{figure}
-
-\subsubsection{Can the confinement method perform better quality assessment in protein structure prediction?}
-
-\Ken{[Arijit - let's make the following section more concise and focused.]}
-
-A part of the CASP experiment is dedicated to the quality assessment (QA) of predicted models
-\cite{Kryshtafovych2011}. Here, predictors were asked to score each model on a scale 
-(known as qmode) from 0 to 1, with higher values corresponding to better models \cite{Kryshtafovych2011}. 
-It will be interesting to know how confinement method perform in quality assessment compare to 
-the other groups in CASP9. We investigated this using couple of CASP Targets. Here, we present a 
-case, where confinement method perform well than the top performing group MUFOLD-WQA \cite{Wang2011} 
-of CASP9. We choose  
-two models from CASP taget T0538. They were model 3 submitted by PconsR (GDT\_TS = 96, qmode 1 = 0.5434) 
-and model 5 from the MULTICOM-NOVEL ( GDT\_TS = 83, qmode 1 = 0.5865). Both are server predicted models and
-the qmode value presented are from MUFOLD-WQA \cite{Wang2011}. We choose these two models as
-the model by PconsR was the most accurate predicted model for this target, whereas the other model 
-was predicted best by QA test of MUFOLD-WQA. The calculated free energy using confinement method indicate
-that the model by PconsR is more stable by $3.9 Kcal/Mol$ which support the GDT\_TS trend.
-There is no doubt that the consensus approach can predict the model quality
-in a faster manner. But we expect confinement method can predict the model quality in a
-relatively expensive but much more accurate way.
-
-\section{Conclusion}
-
-We have described a computational method called confinement for computing the difference free energy from a protein conformation $A$ to $B$.  We 
-show that it can give accurate values, even for relatively large conformational changes.  We have demonstrated that it can discriminate 
-which Orban chameleon peptides fold into a 3-helix bundle vs. a 4$\beta$+1 helix structure. We show that it can discriminate between good and bad CASP 
-models.  Perhaps most importantly, we have shown that it can be used to give residue-level insights into what are the dominant structural factors in a 
-protein that are responsible for the difference free energies.  A key advantage is that this method does not require any reaction coordinate, or sampling 
-a pathway from conformation $A$ to $B$. We have tested the method for structures of proteins having up to 100 amino acids.  The computational cost 
-for a 56-residue protein is about 4 hours [on 1 gpu] for 20 ns of confinement [Is 20ns all you need?].
-
-\Arijit{I need 21 (confinement run) x 20 ns for single confinement}
-%There is another issue, which is normal mode analysis calculations are Oh(N**3) and require large
-%amounts of memory, effectively limiting the size of the systems under study. In most cases
-%quasiharmonic analysis.
-
-\Ken{[If this is a relevant limitation, let's say something about it.]}
-\Arijit{I think Ken is pointing towards the normal mode issue and number of residues that we can handle}
+\bibitem{Mongan2006}
+Mongan, J.; Simmerling, C.; A. McCammon, J.; A. Case, D.; Onufriev, A. Generalized
+Born with a simple, robust molecular volume correction. J. Chem. Theory Comput.,
+2006, 3, 156-169.
 
@@
 Cecchini et  al. \cite{Cecchini2009}. The basic approach of the confinement approach is the same in
-both these papers. However, there are some technical differences. Here we briefly describe the
-procedure to compute free energy $\Delta G_{AB}$ between two conformations A and B of a protein. The
-basic idea to compute the free energy between conformations A and B is to perform a thermodynamic
-cycle:
+both these papers. A thermodynamic cycle is used to compute the free energy between conformations A 
+and B. However, there are some small technical differences between the two approaches. Here we briefly 
+describe the procedure that we used.
 
@@
 
-\item  Minimization of A and B. These minimized conformations (A* and B*)
-       are the reference conformation of that state. During this minimization the backbone is kept
-       restrained to the original position to avoid large conformational changes.
+\item  In the first step, a minimization of A and B conformations are performed. These minimized conformations (A* and B*)
+       are the reference conformation of that state.
 
@@
        harmonic restraints on all the atoms of the biomolecule. This is done by running 21 molecular dynamics simulation
-       (20 ns long each), where the harmonic restraint force constant was scaled from 0.00005
-       \Alberto{units??}
-       (mostly free) to 81.92 (frozen in one microstate). In this final restrained state, the
+       (each 20 ns long) for each leg of the thermodynamic cycle, where the harmonic restraint force constant was scaled from 0.00005
+       Kcal/Mol (mostly free) to 81.92 Kcal/Mol (frozen in one microstate). In this final restrained state, the
        rotational contribution to the free energy is frozen out
@@
 
-All calculations where performed with the amber 11 suit of programs in combination with ff99SB and
-GB/SA implicit solvent. Interestingly, we extend the method for calculation of per residue free
+All calculations where performed with the amber 11 suit of programs \cite{Case2012},\cite{Goetz2012} in 
+combination with ff99SB forcefield\cite{Hornak2006} and generalized born implicit solvent \cite{Mongan2006}.
+Interestingly, we extend the method for calculation of per residue free
 energy in an approximate way. For this purpose, the confinement energy, $\Delta G_{A,A*}$ and
-$\Delta G_{B,B*}$ of each residue is calculated in the usual numerical way as described above. The
-internal energy of each residue is calculated using the decomp module of amber from the final
-restrained trajectory. We call this method approximate as we ignore the contribution from the normal
+$\Delta G_{B,B*}$ of each residue is calculated in the usual numerical way as described in ref. by Tyka 
+et al. \cite{Tyka2006}. The internal energy of each residue is calculated using the decomp module of amber from the final
+restrained trajectory. We call this method approximate as we ignore the entropic contribution from the normal
 mode or quasiharmonic analysis. However, this contrbution to the total free energy is much smaller,
 which allow us to study the mechanistic details of conformational preference of each residue.
-
-%There are two big bottlenecks in this calculation. The first has to do with the amount of
-%simulations that have to be performed to confine the ensemble of conformations close to a
-%microstate into a single microstate, the second bottleneck has to do with the size of the system in
-%the normal mode calculation. We have tackled the first issue by using GPU (Graphical Processing
-%Unit) technology as opposed to the classical CPUs, which gives us a two order of magnitude boost in
-%computational efficiency. Normal mode analysis calculations are Oh(N**3) and require large amounts
-%of memory, effectively limiting the size of the systems under study. Where as this can be overcome
-%by larger memory machines, these are often not available, and so we propose an alternative by using
-%quasiharmonic analysis. This is still an Oh(N**3) problem, but we now the motions of atoms are
-%already determined in trajectories, and so we can choose to do quasiharmonic analysis excluding
-%hydrogens, which will reduce the number of atoms N to almost N/2. This means that we can use this
-%method for systems twice as large.
-
-\begin{thebibliography}{99}
-
-\bibitem{Elber2007}
-Elber, R. A Milestoning Study of the Kinetics of an Allosteric Transition: Atomically Detailed Simulations of Deoxy Scapharca
-Hemoglobin. Biophysical J., 2007, 92, 85-87.
-
-\bibitem{Moult2011}
-Moult, J.; Fidelis, K.; Kryshtafovych, A.; Tramontano, A. Critical assessment of methods of protein structure prediction (CASP)-round IX.
-Proteins, 2011, 79, 1-5.
-
-\bibitem{West2007}
-West, A.M.; Elber, R.; Shalloway, D. Extending molecular dynamics time scales with milestoning: example of complex kinetics
-in a solvated peptide. J Chem Phys. 2007, 126, 145104-1451014.
-
-\bibitem{Haas2007}
-Haas, K.; Chu, J.W. Decomposition of energy and free energy changes by following the flow of work along reaction path.
-J. Chem. Phys. 2009, 131, 144105-144111.
-
-\bibitem{Jonsson1998}
-Jónsson, H.; Mills, G.; Jacobsen, K.W. Nudged Elastic Band Method for Finding Minimum Energy Paths of Transitions,
-in Classical and Quantum Dynamics in Condensed Phase Simulations, Ed. B. J. Berne, G. Ciccotti and D. F.
-Coker, 385 (World Scientific, 1998).
-
-\bibitem{E2007}
-E, W.; Ren, W.; Vanden-Eijnden, E. Simplified and improved string method for computing the minimum energy paths in
-barrier-crossing eventsi. J. Chem. Phys. 2007, 126, 164103.
-
-\bibitem{Dellago2002}
-Dellago, C.; Bolhuis, P.G.; Geissler, P.L. Transition Path Sampling, Adv. Chem. Phys. 2002, 123, 1-84.
-
-\bibitem{Cheng2006}
-Cheng, X.; Wang, H.; Grant, B.; Sine, S.M.; McCammon, J.A. Targeted Molecular Dynamics Study of C-Loop Closure
-and Channel Gating in Nicotinic Receptors. 2006, 9, 134.
-
-\bibitem{Elber2005}
-Elber, R. Long-timescale simulation methods. Cur. Opi. in Str. Biol. 2005, 15, 151-156.
-
-\bibitem{Torrie1977}
-Torrie, G. M.; Valleau, J. P. Nonphysical sampling distributions in Monte Carlo free-energy estimation: Umbrella sampling
-(1977) J. Comput. Phys. 23, 187
-
-\bibitem{Tironi1994}
-Tironi, I.G.; van Gunsteren, W.F. A molecular-dynamics simulation study of chloroform. Mol. Phys. 1994, 83, 381-403.
-
-\bibitem{Meirovitch2007}
-Meirovitch, H. Recent developments in methodologies for calculating the entropy and free energy of biological systems by computer simulation.
-Current Opinion in Structural Biology, 2007, 17, 181-186.
-
-\bibitem{Mobley2007}
-Mobley, D.L.; Chodera, J.D.; Dill, K.A. The combining and release method: obtaining correct binding free energies in
-the presence of protein conformational change. Journal of Chemical Theory and Computation 2007, 3, 1231-1235.
-
-\bibitem{Mobley2012}
-Mobley, D.L.; Klimovich, P.V. Perspective: Alchemical free energy calculations for drug discovery.
-J. Chem. Phys. 2012, 137, 230901-12.
-
-\bibitem{Mobley2006}
-Mobley, D.L.; Chodera, J.D.; Dill, K.A. On the use of orientational restraints and symmetry corrections in alchemical
-free energy calculations. J. Chem. Phy. 2006, 125, 084902.
-
-\bibitem{Chipot2007}
-Chipot, C.; Shell, M.S.; Pohorille, A. Introduction, in Chipot, C., Pohorille, A., editors. Free Energy
-Calculations: Theory and Applications in Chemistry and Biology. Springer Series in Chemical
-Physics, vol. 86. Berlin and Heidelberg: Springer; 2007, p. 1–32.
-
-\bibitem{Jorgensen2004}
-Jorgensen, W.L. The many roles of computation in drug discovery, Science 2004, 303, 1813–8.
-
-\bibitem{Gilson2007}
-Gilson, M.K.; Zhou, H.X. Calculation of protein-ligand binding affinities. Annu Rev Biophys Biomol Struct. (2007) 36, 21-42.
-
-\bibitem{Dill1997}
-Dill, K.A.; H.S. Chan.  From Levinthal to Pathways to Funnels:  The "New View" of Protein Folding Kinetics.  Nature Structural Biology 4, 10-19 (1997)
-
-\bibitem{Dill2008}
-Dill, K.A.; Ozkan, S.B.; Shell, M.S.; Weikl, T.R. The protein folding problem. Annual Review of Biophysics (2008), 37, 289-316.
-
-\bibitem{Anfinsen1973}
-Anfinsen. C.B. Principles that Govern the Folding of Protein Chains. Science (1973) 181, 223-230.
-
-\bibitem{Christ2007}
-Christ, C.D.; van Gunsteren, W.F. Enveloping distribution sampling: A method to calculate free energy differences from a single simulation,
-J. Chem. Phys. (2007), 126, 184110.
-
-\bibitem{Ytreberg2006}
-Ytreberg, F.; Zuckerman, D. Simple estimation of absolute free energies for biomolecules. J. Chem. Phys. 2006, 124, 104105.
-
-\bibitem{Park2008}
-Park, S.; Lau, A.; Roux, B. Computing conformational free energy by deactivated morphing. J. Chem. Phys. 2008, 129, 134102
-
-\bibitem{Zheng2008}
-Zheng, L.; Chen, M.; Yang, W. Random walk in orthogonal space to achieve efficient free-energy simulation of complex systems, Proc. Natl. Acad. Sci. 200
-8, 105 (51), 20227.
-
-\bibitem{Shell2010}
-Shell, S.M. A replica-exchange approach to computing peptide conformational free energies. Mol. Sim. 2010, 7, 505-515.
-
-\bibitem{Tyka2006}Tyka, M.; Clarke, A.; Sessions, R. An Efficient, Path-Independent Method for Free-Energy Calculations. J.Phys.Chem. B 2006, 110, 17212-17220.
-
-\bibitem{Cecchini2009}
-Cecchini, M., Krivov, S.V., Spichty, M., Karplus, M. Calculation of free-energy differences by confinement simulations. Application to peptide conformers. J. Phys. Chem. B. 2009, 113, 9728-9740.
-
-\bibitem{Ovchinnikov2013}
-Ovchinnikov, V.; Cecchini, M.; Karplus, M. A Simplified Confinement Method for Calculating Absolute Free Energies
-and Free Energy and Entropy Differences. J. Phys. Chem. B. 2013, 117, 750-762.
-
-\bibitem{Spichty2010}Spichty, M.; Cecchini, M.; Karplus, M. Conformational Free-Energy Difference of a Miniprotein from Nonequilibrium
-Simulations. J. Phys. Chem. Lett., 2010, 1, 1922-1926.
-
-\bibitem{Strajbl2000}
-Strajbl, M.; Sham, Y.Y.; Villà, J.; Chu, Z.-T.; Warshel, A. Calculations of Activation Entropies of Chemical Reactions
-in Solution. (2000) 104, 4578-4584.
-
-\bibitem{Krivov2004}
-Krivov, S.; Karplus, M. Hidden complexity of free energy surfaces for peptide (protein) folding Proc. Natl. Acad. Sci. U.S.A. 2004, 101, (41), 14766.
-
-\bibitem{Alexander2007}Alexander, P.A.; He, Y.; Chen, Y.; Orban, J.
-Bryan, P.
-The design and characterization of two proteins with $88 \%$ sequence identity but differentstructure and function.
-Proc.
-Natl.
-Acad.
-Sci.
-2007, 104 (29), 11963-11968.
-
-\bibitem{He2008}
-He, Y.; Chen, Y.; Alexander, P.A.; Orban, J. NMR structures of two designed proteins with high sequence identity but different fold and function. Proc. Natl. Acad. Sci. 2008, 105 (38), 14412-14417.
-
-\bibitem{Alexander2009}
-Alexander, P.A.; He, Y.; Chen, Y.; Orban, J. Bryan, P. A minimal sequence code for switching protein structure and function. Proc. Natl. Acad. Sci. 2009
-, 106(50), 21149-21154.
-
-\bibitem{Bryan2010}
-Bryan, P.N.; Orban, J. Proteins that switch folds. Curr Opin Struct Biol. 2010, 20(4), 482-488.
-
-\bibitem{He2012}
-He, Y.; Chen, Y.; Alexander, P.A.; Bryan, P.N.; Orban, J. Mutational tipping points for switching protein folds and functions. Structure. 2012, 20(2), 2
-83-91.
-
-\bibitem{Shortle2009}
-Shortle, D. One sequence plus one mutation equals two folds. Proc. Natl. Acad. Sci. 2009, 106(50), 21011-21012.
-
-\bibitem{Sheffler2009}
-Sheffler, W.; Baker, D. RosettaHoles: Rapid assessment of protein core packing for structure prediction, refinement, design, and validation. Protein Sci
-ence. 2009, 18(1), 229-239.
-
-\bibitem{Case2012}
-D.A.
-Case, T.A.
-Darden, T.E.
-Cheatham, III, C.L.
-Simmerling, J.
-Wang, R.E.
-Duke, R.
-Luo, R.C.
-Walker, W.
-Zhang, K.M.
-Merz, B.
-Roberts, S.
-Hayik, A.
-Roitberg, G.
-Seabra, J.
-Swails, A.W.
-Goetz, I.
-Kolossváry, K.F.
-Wong, F.
-Paesani, J.
-Vanicek, R.M.
-Wolf, J.
-Liu, X.
-Wu, S.R.
-Brozell, T.
-Steinbrecher, H.
-Gohlke, Q.
-Cai, X.
-Ye, J.
-Wang, M.-J.
-Hsieh, G.
-Cui, D.R.
-Roe, D.H.
-Mathews, M.G.
-Seetin, R.
-Salomon-Ferrer, C.
-Sagui, V.
-Babin, T.
-Luchko, S.
-Gusarov, A.
-Kovalenko, and P.A.
-Kollman (2012), AMBER 12, University of California, San Francisco.
-
-\bibitem{Goetz2012}
-Goetz, A.W.; Williamson, M.J.; Xu, D.; Poole, D.; Le Grand, S.; Walker, R.C. Routine microsecond molecular dynamics simulations with AMBER - Part I: Generalized
-Born. J. Chem. Theory Comput. 2012, 8(5) 1542.
-
-\bibitem{MacCallum2011}
-MacCallum, J.; Perez, A.; Schnieders, MJ.; Hua, L.; Jacobson, M.P.; Dill, K.A. Assessment of protein structure refinement
-in CASP9. Proteins, 2011, 79, 74-90.
-
-\bibitem{Kryshtafovych2011}
-Kryshtafovych, A.; Fidelis, K; and Tramontano, A. Evaluation of model quality predictions in CASP9. Proteins, 2011, 79, 91–106
-
-\bibitem{Zemla2003}
-Zemla, A. LGA: a method for finding 3D similarities in protein structures. Nucleic Acids Res 2003, 31, 3370–3374.
-
-\bibitem{Perez2012}
-Perez, A.; Yang, Z.; Bahar, I.; Dill, K.A.; MacCallum, J.L.; FlexE: Using Elastic Network Models to Compare Models of Protein Structure. J. Chem. Theory Comput., 2012, 8, 3985-3991.
-
-\bibitem{Wang2011}
-Wang, Q.; Vantasin, K.; Xu, D.; Shang, Y. MUFOLD-WQA: A new selective consensus method for quality assessment in protein structure prediction.
-Proteins, 2011, 79: 185–195.
-
-\bibitem{Case2005}
-Case, D.A.; Cheatham, III, T.E.; Darden, T.; Gohlke, Luo, H.R.; Merz, Jr., K.M.;  Onufriev, A; Simmerling, C.;
-Wang, B.; R. Woods, R. The Amber biomolecular simulation programs. J. Computat. Chem. (20005) 26, 1668-1688.
-
-\end{thebibliography}
-
-\end{document}

### 2013-04-26 Paper/paper.tex
changed method & citation

@@
 is marginally stable in one of two possible folds. By mutating key residues in this sequence they
-are able to stabilize one fold or the other (see Figure~\ref{fig:orban}). We call sequences ``$\beta$'' that fold into 
-the protein-G-like that fold into a $4 \beta + \alpha$ fold, and we call sequences ``$\alpha$'' that fold into $3 \alpha$ helical folds.  One pair 
+are able to stabilize one fold or the other (see Figure~\ref{fig:orban}). We call conformation ``$\beta$'' that fold into 
+the protein-G-like that fold into a $4 \beta + \alpha$ fold, and we call conformation ``$\alpha$'' that fold into $3 \alpha$ helical folds.  One pair 
 of sequences (GA88/GB88) is 88 percent identical in sequence, differing in seven positions. Another pair (GA95/GB95) is 95 percent identical, 
-differing in three positions. The third pair (GA98/GB98) differs in only a single tyrosine-to-alanine mutation. Accurately predicting the 
+differing in three positions. 
+%The third pair (GA98/GB98) differs in only a single tyrosine-to-alanine mutation. 
+Accurately predicting the 
 structural preferences of these structures presents a serious challenge for computational methods 
@@
 conformational preferences correctly for all four sequences (data not shown). This is, however, not
-surprising. It is well known [CITE]\Arijit{still searching this citation} that it is easy to distinguish 
+surprising. It is well known \cite{Handl2009},\cite{Sheffler2009} that it is easy to distinguish 
 computational models from native structures. To avoid this potential problem, we instead computed relative free energy of two
@@
 \Ken{[If this is a relevant limitation, let's say something about it.]}
-\Arijit{I think Ken is pointing towards the normal mode issue and number of residues that we can handle}
+\Arijit{I think we need to write the normal mode issue and number of residues that we can handle. If 
+one of you contribute here that will be great.}
 
@@
 Cecchini et  al. \cite{Cecchini2009}. The basic approach of the confinement approach is the same in
-both these papers. However, there are some technical differences. Here we briefly describe the
-procedure to compute free energy $\Delta G_{AB}$ between two conformations A and B of a protein. The
-basic idea to compute the free energy between conformations A and B is to perform a thermodynamic
-cycle:
+both these papers. A thermodynamic cycle is used to compute the free energy between conformations A
+and B. However, there are some small technical differences between the two approaches. Here we briefly
+describe the procedure that we used.
 
@@
        harmonic restraints on all the atoms of the biomolecule. This is done by running 21 molecular dynamics simulation
-       (20 ns long each), where the harmonic restraint force constant was scaled from 0.00005
-       Kcal/Mol 
-       (mostly free) to 81.92 iKcal/Mol (frozen in one microstate). In this final restrained state, the
+       (each 20 ns long) for each leg of the thermodynamic cycle, where the harmonic restraint force constant was scaled from 0.00005
+       Kcal/Mol (mostly free) to 81.92 Kcal/Mol (frozen in one microstate). In this final restrained state, the
        rotational contribution to the free energy is frozen out
@@
 
-All calculations where performed with the amber 11 suit of programs in combination with ff99SB and
-GB/SA implicit solvent. Interestingly, we extend the method for calculation of per residue free
+All calculations where performed with the amber 11 suit of programs \cite{Case2012},\cite{Goetz2012} in
+combination with ff99SB forcefield\cite{Hornak2006} and generalized born implicit solvent \cite{Mongan2006}.
+Interestingly, we extend the method for calculation of per residue free
 energy in an approximate way. For this purpose, the confinement energy, $\Delta G_{A,A*}$ and
-$\Delta G_{B,B*}$ of each residue is calculated in the usual numerical way as described above. The
-internal energy of each residue is calculated using the decomp module of amber from the final
-restrained trajectory. We call this method approximate as we ignore the contribution from the normal
+$\Delta G_{B,B*}$ of each residue is calculated in the usual numerical way as described in ref. by Tyka 
+et al. \cite{Tyka2006}. The internal energy of each residue is calculated using the decomp module of amber from the final
+restrained trajectory. We call this method approximate as we ignore the entropic contribution from the normal
 mode or quasiharmonic analysis. However, this contrbution to the total free energy is much smaller,
@@
 Gilson, M.K.; Zhou, H.X. Calculation of protein-ligand binding affinities. Annu Rev Biophys Biomol Struct. (2007) 36, 21-42.
+
+\bibitem{Handl2009}
+Handl, J.; Knowles, J.; Lovel, S.C. Artefacts and biases affecting the evaluation of scoring
+functions on decoy sets for protein structure prediction. Bioinformatics, 2009, 25, 1271-1279.
 
@@
 
+\bibitem{Hornak2006}
+Hornak, V.; Abel, R.; Okur, A.; Strockbine, B.; Roitberg, A.; Simmerling, C. Comparison of multiple Amber force fields
+and development of improved protein backbone parameters. Proteins., 2006, 65, 712-725.
+
+\bibitem{Mongan2006}
+Mongan, J.; Simmerling, C.; A. McCammon, J.; A. Case, D.; Onufriev, A. Generalized
+Born with a simple, robust molecular volume correction. J. Chem. Theory Comput.,
+2006, 3, 156-169.
+
 \end{thebibliography}

### 2013-04-29 Paper/paper.tex
Made revisions

@@
 
-\Ken{[Guys - I recommend we get a better name than confinement.  It's totally uninformative.  In my view, better would be `confine-and-configure' or something.  
-Please give it some thought.]}
+\Ken{[Guys---I recommend we get a better name than confinement.  It's totally uninformative.  In my view, better would
+    be `confine-and-configure' or something.  Please give it some thought.]}
+
+\Justin{[Ken---I agree that the name isn't great, but I don't think we should change it because: it's already in the
+    literature and we didn't come up with the method. I think we should consider capitalizing it as Confinement Method
+    throughout.]}
 
@@
 
-In some problems of protein science, you want to know the relative stability of a protein's conformation $A$ compared to it's 
-conformation $B$.  We call this the \emph{difference free energy} (DFE). For example, in allosteric mechanisms, a protein often 
-adopts one conformation when a ligand is bound and another conformation when no ligand is bound \cite{Elber2007}. To understand the mechanism requires knowledge 
-of the relative free energies of the protein conformations.  Another example is in protein structure prediction, for example as practiced in 
-the CASP competition \cite{Moult2011}.  If you are using some computational model and have predicted two putative native structures, $A$ 
-and $B$, you want to compute which has the lower free energy, in order to know which is the more native-like.  In a third example, you may 
-want to know whether mutating a few amino acids in a protein could cause the protein to switch from one stable conformation to another, because 
-that could have important consequences for biological mechanisms and disease.  And fourth, sometimes binding a ligand induces a protein from 
-one conformation to another.  A quantitative understanding of such \emph{induced-fit} situations requires knowledge of the difference free 
-energy of the protein in the two states. In all these cases, it would be useful to have a computational method that can be used efficiently 
-with atomically detailed physical forcefield models to compute the free energy differences between different given protein conformations.
+In some problems of protein science, you want to know the relative stability of a protein's conformation $A$ compared to
+its conformation $B$.  We call this the \emph{difference free energy}. For example, in all esoteric mechanisms, a protein
+often adopts one conformation when a ligand is bound and another conformation when no ligand is bound~\cite{Elber2007}.
+To understand the mechanism requires knowledge of the relative free energies of the protein conformations.  Another
+example is in protein structure prediction, for example as practiced in the Critical Assessment of Structure Prediction
+(CASP) competition~\cite{Moult2011}. If you are using some computational model and have predicted two putative native
+structures, $A$ and $B$, you want to compute which has the lower free energy, in order to know which is the more
+native-like.  In a third example, you may want to know whether mutating a few amino acids in a protein could cause the
+protein to switch from one stable conformation to another, because that could have important consequences for biological
+mechanisms and disease. And fourth, sometimes binding a ligand induces a protein from one conformation to another. A
+quantitative understanding of such \emph{induced-fit} situations requires knowledge of the difference free energy of the
+protein in the two states. In all these cases, it would be useful to have a computational method that can be used
+efficiently with atomically detailed physical forcefield models to compute the free energy differences between different
+given protein conformations.
 
-Methods like thermodynamic integration \cite{Tironi1994},\cite{Meirovitch2007} which are successful in alchemical free energy calculation
-may not be as useful for calculation of conformational free energy.
-One widely explored strategy is to use molecular dynamics simulations along some putative reaction coordinate pathway from conformation $A$ 
-to $B$ \cite{Elber2007,West2007,Haas2007,Jonsson1998,E2007,Dellago2002,Cheng2006}. \Ken{we need 
-many refs here: cite Elber's milestoning, Jhih-Wei Chu, and the many other path-sampling approaches.]} The free energy along this reaction 
-coordinate can then be determined using methods such as umbrella sampling \cite{Torrie1977}. Such approaches have several limitations.  
-First, it is necessary to know a reaction pathway from $A$ to $B$. If conformations $A$ and $B$ are quite different, then it can be 
-challenging to find such paths. Second, these methods 
-are computationally slow. To get an accurate estimate of the total free energy difference $\Delta G = G_B - G_A$ requires accurate determinations 
-of the many small free energy differences $A \rightarrow 1 \rightarrow 2 \ldots \rightarrow B$, so each microstep requires substantial amounts 
-of sampling.  And third, these methods are prone to large errors, because the pathway error is a sum of many errors among the many microsteps.
+\Justin{[TI is actually the basis of the confinement step. We need to be more explicit about what we mean here.] Methods
+    like thermodynamic integration~\cite{Tironi1994},\cite{Meirovitch2007} which are successful in alchemical free
+    energy calculation may not be as useful for calculation of conformational free energy.} One widely explored strategy
+is to use molecular dynamics simulations along some putative reaction coordinate pathway from conformation $A$ to
+$B$~\cite{Elber2007,West2007,Haas2007,Jonsson1998,E2007,Dellago2002,Cheng2006}. \Ken{[we need many refs here: cite
+    Elber's milestoning, Jhih-Wei Chu, and the many other path-sampling approaches.]} The free energy along this
+reaction coordinate can then be determined using methods such as umbrella sampling \cite{Torrie1977}\Justin{[add ref
+    where this is used computing free energy differences between conformations]} or targeted MD \Justin{[refs]}. Such
+approaches have several limitations. First, it is necessary to know an efficient reaction pathway from $A$ to $B$. If
+conformations $A$ and $B$ are quite different, then it can be challenging to find such paths. Second, these methods are
+computationally slow. To get an accurate estimate of the total free energy difference $\Delta G = G_B - G_A$ requires
+accurate determinations of the many small free energy differences $A \rightarrow 1 \rightarrow 2 \ldots \rightarrow B$,
+so each step requires substantial amounts of sampling.  And third, these methods are prone to large errors, because the
+pathway error is a sum of many errors among the many steps.
 
-\Arijit{Will it be useful to add methods like PMF or replica exchange based free energy methods ?}
+\Ken{[Are all these methods below ones that don't involve paths?  Also, are there other key methods we should cite?  We
+    need a good scholarly list of all the relevant works.]}
 
-\Ken{[Are all these methods below ones that don't involve paths?  Also, are there other key methods we should cite?  We need a good scholarly 
-list of all the relevant works.]} 
-The calculation of protein conformational free energy has been successfully attempted by a number of
-groups \cite{Shell2010,Ytreberg2006,Zheng2008,Spichty2010,Strajbl2000,Park2008,Tyka2006,Cecchini2009,Ovchinnikov2013}. 
-Some of these methods like reference system method \cite{Ytreberg2006}, deactivated morphing \cite{Park2008}, confinement 
-method \cite{Tyka2006,Cecchini2009,Ovchinnikov2013} take an alternative strategy for computing difference free energies 
-that does not require knowing a pathway from $A$ to $B$.  
-\Ken{[please add the several new Karplus papers from 2012-2013.]}.
+The calculation of protein conformational free energy has been successfully attempted by a number of groups
+\cite{Shell2010,Ytreberg2006,Zheng2008,Spichty2010,Strajbl2000,Park2008,Tyka2006,Cecchini2009,Ovchinnikov2013}. Some of
+these methods like the reference system method~\cite{Ytreberg2006}, deactivated morphing~\cite{Park2008}, the
+confinement method~\cite{Tyka2006,Cecchini2009,Ovchinnikov2013} take an alternative strategy for computing difference
+free energies that does not require knowing a pathway from $A$ to $B$.  \Ken{[please add the several new Karplus papers
+    from 2012-2013.]}.
 
-Here, we adopt the confinement method of Tyka et al \cite{Tyka2006} and Cecchini et al.\cite{Cecchini2009}. This method 
-also resembles the `confine-and-release' method \cite{Mobley2007}, \cite{Mobley2012}, \cite{Mobley2006}.  
-\Ken{[cite mobley's jctc 07 (ref 216 on our labsite), and mobley's recent j chem phys review, 
-and mobley (ref 199) j chem phys, and others that are relevant.]}. The method is based on the thermodynamic cycle showns in 
-Figure \ref{fig:method}.  To start, conformation $A$ is an ensemble.  We reduce ensemble $A$ to one particular microstate $A^\ast$ 
-in that ensemble by applying restraints in an MD simulation.  In the same way, we reduce conformational ensemble $B$ to one of its 
-microstates, $B^\ast$.  Then, we simply compute the free energy difference between the microstates 
-as $\Delta G_{A^\ast, B^\ast} = \Delta H_{A^\ast, B^\ast}$ (which is just an enthalpy difference since there is no conformational 
-entropy difference between the microstates).  $\Delta H_{A^\ast, B^\ast}$ is readily calculated using normal mode analysis. In 
-this way, we can calculate the free energy difference between the two end states without needing to define a physical path or reaction 
-coordinate connecting the two states.
+Here, we adopt the confinement method of Tyka et al~\cite{Tyka2006} and Cecchini et al.~\cite{Cecchini2009}, which is
+based on the thermodynamic cycle shown in Figure~\ref{fig:method}. We start with an ensemble of related conformations,
+$A$. We then restrict $A$ to a much ``tighter'' ensemble $A^\ast$ by applying position restraints in a series of MD
+simulations. Similarly, a second ensemble $B$ is restricted to $B^\ast$. We then compute the free energy between
+$A^\ast$ and $B^\ast$ using either normal mode analysis or the quasiharmonic method [ref]. This step takes into account
+both the remaining conformational entropy in each ensemble and the remaining enthalpy difference. In this way, we can
+calculate the free energy difference between the two end states without needing to define a physical path or reaction
+coordinate connecting the two states. The confinement method shares some similarities with the `confine-and-release'
+method for computing ligand binding affinities~\cite{Mobley2007,Mobley2012,Mobley2006} \Ken{[cite mobley's jctc 07 (ref
+    216 on our labsite), and mobley's recent j chem phys review, and mobley (ref 199) j chem phys, and others that are
+    relevant.]}.
 
@@
     \begin{center}
-\includegraphics[width=3.5 in,height=3.5 in]{method.pdf}
+        \includegraphics[width=3.5 in]{method.pdf}\label{fig:method}
     \end{center}
     \caption{Graphical representation of the thermodynamic cycle involving confinement method.}
-\label{fig:method}
 \end{figure}
 
-\Ken{For this fig, let's make A and A* green and B and B* red or something, to distinguish them visually.
-Let's also blur the ensemble more relative to the microstate, so it's clearer the distinction.]}
-
-Previously, the confinement method has been validated on small model peptides \cite{Tyka2006,Cecchini2009}. Here, we show that the confinement 
-method can also be applied successfully to larger systems, such as the series of chameleon proteins designed by Orban and co-workers
-\cite{Alexander2007,He2008,Alexander2009,He2012}, which can switch between two completely different folds with only few-amino-acid changes 
-in sequence.  We also show that our computed difference free energies are useful in evaluating the quality of CASP target protein predictions.  
-This could be of significant value for ultimately improving the energetics in protein-structure models, not just the structures.  And, 
-finally, we show that we can approximately decompose difference free energies into individual amino acid components.  This offers the 
-opportunity for diagnosing the structural basis for difference free energies, which can be useful for interpreting biological mechanisms.
+Previously, the confinement method has been validated on small model peptides~\cite{Tyka2006,Cecchini2009}. Here, we
+show that the confinement method can also be applied successfully to larger systems, such as the series of chameleon
+proteins designed by Orban and co-workers~\cite{Alexander2007,He2008,Alexander2009,He2012}, which can switch between two
+completely different folds with only slight changes in sequence. We also show that our computed difference free energies
+are useful in evaluating the quality of CASP target protein predictions.  This could be of significant value for
+ultimately improving the energetics in protein-structure models, not just the structures. And, finally, we show that we
+can approximately decompose difference free energies into individual amino acid components. This offers the opportunity
+for diagnosing the structural basis for difference free energies, which can be useful for interpreting biological
+mechanisms.
 
@@
 
-First, we validated that our implementation of the confinement method produces results compatible with previous calculations reported in the 
-literature.  The method has previously been applied to a 16 amino acid residue $\beta$-hairpin from protein G, known as 
-BHP \cite{Cecchini2009}. We calculated the free energy difference between the native conformation (called bhp1), which has 
-a two-stranded $\beta$-sheet and a three-stranded $\beta$-sheet called bhp3.  Our confinement calculation shows that bhp1 
-is more stable by 1.7 kcal/mol, consistent with 4 $\mu$s equilibrium simulations showing that bhp1 is more favorable configuration 
-by 1.8 kcal/mol, and in agreement with previous calculations \cite{Cecchini2009}.
+First, we validated that our implementation of the confinement method produces results compatible with previous
+calculations reported in the literature.  The method has previously been applied to a 16 amino acid residue
+$\beta$-hairpin from protein G, known as BHP~\cite{Cecchini2009}. We calculated the free energy difference between the
+native conformation---called bhp1, which has a two-stranded $\beta$-sheet---and a non-native conformation---called bhp3,
+which has a three-stranded $\beta$-sheet. Our confinement calculation shows that bhp1 is more stable by 1.7 kcal/mol,
+consistent with 4 $\mu$s equilibrium simulations showing that bhp1 is more favorable configuration by 1.8 kcal/mol, and
+in agreement with previous calculations~\cite{Cecchini2009}.
 
-Second, we looked at 6 target proteins from the CASP9 experiment.  For each target, we examined up to 5 submitted models.  We 
-computed the difference free energy between the true native and the best model. Figure~\ref{fig:summary_casp} and 
-Table~\ref{table:casp_control} shows that, in 5 
-out of 6 cases, the confinement method assigns a lower free energy to the experimentally determined structure than to any of the 
-decoys.  Other well-known discriminators  can also successfully make this recognition \cite{Sheffler2009}; 
- it is just an independent useful validation that the confinement calculations make sense.
+Second, we looked at 6 target proteins from the CASP9 experiment [ref].  For each target, we examined up to 5 submitted
+models.  We computed the difference free energy between the true native and the best model.
+Figure~\ref{fig:summary_casp} and Table~\ref{table:casp_control} show that, in 5 out of 6 cases, the confinement method
+assigns a lower free energy to the experimentally determined structure than to any of the decoys.  Other well-known
+discriminators can also successfully tell native structures from computer generated models~\cite{Sheffler2009} [add
+rosetta holes, rosetta ff, dfire]; it is just an independent useful validation that the confinement calculations make
+sense.
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.5 in,height=3.2 in]{casp.pdf}
-\end{center}
-\caption{In 5 out of 6 CASP Target, the confinement method assigns a lower free energy to the experimentally determined native structure.} 
-\label{fig:summary_casp}
-\end{figure}
+\Justin{[Arijit, please add GDT of best decoy to this table.]}
 
@@
 \begin{center}
-\caption{The confinement method assigns a more favorable free energy to the experimentally
-    determined structure than to computer-generated predictions. For each target, we examined as
-    many as five predictions submitted by CASP participants. We report the free energy difference
-    between the most favorable decoy and the experimentally determined structure. Positive
-    $\Delta\Delta G$ values indicate that the experimental structure is predicted to be more
-favorable than any of the decoys.}
+\caption{The confinement method assigns a more favorable free energy to the experimentally determined structure than to
+    computer-generated predictions. For each target, we examined as many as five predictions submitted by CASP
+    participants. We report the free energy difference between the most favorable decoy and the experimentally
+    determined structure. Positive $\Delta\Delta G$ values indicate that the experimental structure is predicted to be
+    more favorable than any of the decoys.}
 \label{table:casp_control}
@@
 
-\Ken{[ I think a better notation is $\Delta \Delta G = G_{best~decoy} - G_{native}$, rather than having the arrow.]}
-
 \subsection{The confinement method correctly predicts the structures of chameleon sequences}
 
-We also tested the confinement method predictions for difference free energies on the chameleon sequences of Orban 
-et al. \cite{Alexander2007,He2008,Alexander2009,Bryan2010,He2012,Shortle2009}; these 
-are instances in which two highly similar sequences fold into remarkably different structures. Orban and co-workers have designed a 
-protein-G-like sequence of 56-residues that
-is marginally stable in one of two possible folds. By mutating key residues in this sequence they
-are able to stabilize one fold or the other (see Figure~\ref{fig:orban}). We call conformation ``$\beta$'' that fold into 
-the protein-G-like that fold into a $4 \beta + \alpha$ fold, and we call conformation ``$\alpha$'' that fold into $3 \alpha$ helical folds.  One pair 
-of sequences (GA88/GB88) is 88 percent identical in sequence, differing in seven positions. Another pair (GA95/GB95) is 95 percent identical, 
-differing in three positions. 
-%The third pair (GA98/GB98) differs in only a single tyrosine-to-alanine mutation. 
-Accurately predicting the 
-structural preferences of these structures presents a serious challenge for computational methods 
-\cite{Alexander2007,He2008,Alexander2009,Bryan2010,He2012,Shortle2009}.
+We also tested the confinement method predictions for difference free energies on the chameleon sequences of Orban et
+al.!\cite{Alexander2007,He2008,Alexander2009,Bryan2010,He2012,Shortle2009}. These are instances in which two highly
+similar sequences fold into remarkably different structures. Orban and co-workers have designed a protein-G-like
+sequence of 56-residues that is marginally stable in one of two possible folds. By mutating key residues in this
+sequence they are able to stabilize one fold or the other (see Figure~\ref{fig:orban}). We refer to the $4\beta +
+\alpha$ structure as the $\beta$ conformation, and we refer to the $3\alpha$ structure as the $\alpha$ conformation. One
+pair of sequences (GA88/GB88) is 88 percent identical in sequence, differing in seven positions. Another pair
+(GA95/GB95) is 95 percent identical, differing in three positions. Accurately predicting the structural preferences of
+these structures presents a serious challenge for computational
+methods~\cite{Alexander2007,He2008,Alexander2009,Bryan2010,He2012,Shortle2009}. \Justin{Do any of these citations try to
+    do any modeling or comment about how challenging this would be?}
 
 \begin{figure}
+    \begin{center}
     \includegraphics[width=5.0in]{orban.pdf}
         \label{fig:orban}
-\caption{Confinement method correctly predicts the structural preferences of four chameleon
-sequences. The top part of the figure represent four sequences used in this study along with the
-protein data bank id. Each sequence adopts either a three-helix bundle fold (denoted $\alpha$) or
-$4 \beta + \alpha$ fold (denoted $\beta$). The relative free energies of
-the two folds are reported for each sequence.}
+    \end{center}
+    \caption{The confinement method correctly predicts the structural preferences of four chameleon sequences. The top
+        part of the figure represent four sequences used in this study along with the protein data bank identifier. The
+        experimentally observed fold, computationally predicted fold, and difference free energy between the two folds
+        is reported for each sequence.}
 \end{figure}
 
-We initially approached this problem by making a model of each sequence with the same backbone
-structure as its partner chameleon sequence. For example, we took the sequence of GA88 and built a
-model with the same overall structure as GB88. We then used the confinement method to assess the
-free energy difference between the experimentally determined structure of GA88 and the model (with
-the GA88 sequence and the GB88 structure). The confinement method was able to predict the
-conformational preferences correctly for all four sequences (data not shown). This is, however, not
-surprising. It is well known \cite{Handl2009},\cite{Sheffler2009} that it is easy to distinguish 
-computational models from native structures. To avoid this potential problem, we instead computed relative free energy of two
-different computer generated models for each sequence. One model is based on the $\alpha$ structure and
-the other on the $\beta$ structure (see Supporting Information for details on the modeling procedure).
-This is a much more realistic test of the confinement method's ability to accurately calculate
-relative free energies.
+We initially approached this problem by making a model of each sequence with the same backbone structure as its partner
+chameleon sequence. For example, we took the sequence of GA88 and built a model with the same overall structure as GB88.
+We then used the confinement method to assess the free energy difference between the experimentally determined structure
+of GA88 and the model (with the GA88 sequence and the GB88 structure). The confinement method was able to predict the
+conformational preferences correctly for all four sequences (data not shown). This is, however, not surprising as it is
+often easy to distinguish computational models from native structures~\cite{Handl2009,Sheffler2009}. To avoid this
+potential problem, we instead computed relative free energy of two different computer generated models for each
+sequence. One model is based on the $\alpha$ structure and the other on the $\beta$ structure (see Supporting
+Information for details on the modeling procedure). This is a much more realistic test of the confinement method's
+ability to accurately calculate the difference free energy.
 
 \Ken{[ It's not clear what this test is and why it's better, and why there are only 5 sequences.  Please clarify.]}
+\Justin{[There are five pairs of sequences that can be arranged into 3 different pairings. Arijit has removed the 5th
+    sequence.]}
+
 \Arijit{ Now there are 4 sequences as they have the experimentally available structure. The fifth sequence is discussed
@@
 
-Figure~\ref{fig:orban} shows that the confinement method identifies the correct structure for all four sequences. One hypothesis for fold switching of
-chameleon sequences is that the structural transitions require states with diminished stability
-\cite{Bryan2010}. It is believed that if the free energy of the native state and the alternative state are
-within a range of 5 kcal/mol, then it is possible for the native state to be destabilized relative
-to the alternative state with only small changes in sequence. The stability of the native state can
-decrease for a number reason ranging from chemical modification, breaking of disulphide bonds or
-mutations---as is the case here. The calculated free energy differences range from around 3.5 to 5.0
-kcal/mol, which is consistent with the above hypothesis \cite{He2008,Alexander2009,Bryan2010}. 
-In a more recent study\cite{He2012}, the amino acid residue at position 45 (Tyr for $\beta$ and Leu for $\alpha$) was
-found to be important for switching between  $\alpha$ and $\beta$ conformation. This inspired us to 
-introduce another mutation at 45 position. For our future reference this Tyr-45-Ala mutation of GB95 sequence will
-be known as GB98. Interestingly, with this mutation, the equilibrium shifts to $\alpha$ conformation as it gains 3.84 Kcal/Mol 
-stability compare to $\beta$ conformation and indicate the importance of Tyr-45 residue. It will be interesting to 
-know the actutal experimental structure which can validate our result.
+\Justin{[Ken: it's not that hard to tell an experimental structure from a computer generated model. Almost all of the
+    knowledge-based scoring functions can do this a substantial fraction of the time. Comparing a near-native model with
+    another model is a much more challenging problem.]}
 
-\subsection{Per-residue free energy calculation can identify the mechanistic detail behind
-conformational preferences} \label{sec:orban_per_residue}
+Figure~\ref{fig:orban} shows that the confinement method identifies the correct structure for all four sequences. One
+hypothesis is that such structural transitions require states with diminished stability~\cite{Bryan2010}. It is believed
+that if the free energy of the native state and the alternative state are within a range of around 5 kcal/mol, then it
+is possible for the native state to be destabilized relative to the alternative state with only small changes in
+sequence~\cite{He2008,Alexander2009,Bryan2010}. The calculated free energy differences range from around 3.5 to 5.0
+kcal/mol, which is consistent with this hypothesis. In a more recent study\cite{He2012}, the amino acid residue at
+position 45 (Tyr for $\beta$ and Leu for $\alpha$) was found to be important for switching between $\alpha$ and $\beta$
+conformation. This inspired us to introduce another mutation at this position, Y45A, which we refer to as GA98. Our
+calculations predict that this mutation shifts the equilibrium to the $\alpha$ conformation, which is now more stable
+than the $\beta$ conformation by 3.8 kcal/mol. Although this result has not yet been confirmed experimentally, it is
+consistent with the previously observed effect of Y45L.
 
-To understand the mechanism behind the conformational preferences of the different protein
-sequences, we decomposed the calculated free energy into per-residue contributions in
-an approximate way \cite{Tyka2006}. For this purpose, the confinement free energy, $\Delta G_{A,A*}$ and
-$\Delta G_{B,B*}$ of each residue is calculated in the usual numerical way as described
-the method section. We call this method approximate as we ignore the contribution from the normal 
-mode or quasiharmonic analysis. Instead, decomp module of amber is used to calculate the internal energy 
-of each residue from the final restrained trajectory.
+\subsection{Per-residue free energy calculations can identify the mechanistic detail behind conformational preferences}
+\label{sec:orban_per_residue}
 
-\Ken{[Arijit -- I didn't find this methods calculation.  We should at least put in a sentence or two here summarizing it.]}  
-This helps us to identify important residues that stabilize a particular
-conformation. The per residue free energy, $\Delta \Delta G ((\beta) - (\alpha))$ is
-shown in Figure~\ref{fig:perresidue_orban}. 
-In this plot, residues colored in blue favor the $\alpha$ structure, residues in red favor the $\beta$ structure, and white residues
-have no preference.
+To better understand the mechanism behind these chameleon proteins, we decomposed the calculated free energy into
+per-residue contributions in an approximate way~\cite{Tyka2006}. We decompose each confinement step ($\Delta
+G_{A,A^\ast}$, $\Delta G_{B,B^\ast}$) into per-residue contributions. We can also decompose the remaining enthalpy in
+the confined ensemble into per-residue contributions. However, the method is approximate because we do not include the
+residual conformational entropy from the normal mode or quasiharmonic analysis steps. Computing per-residue
+contributions helps us to identify important residues that stabilize a particular conformation. The per residue free
+energy, $\Delta \Delta G ((\beta) - (\alpha))$ is shown in Figure~\ref{fig:perresidue_orban}.
 
@@
         \includegraphics[width=4.5in]{delta_g.pdf}
+        \label{fig:perresidue_orban}
     \end{center}
-\caption{The $\alpha$ and $\beta$ conformation associated with GA95 and GB95 sequences are colored according to the relative per residue
-free energy value. In this figure, residues colored in blue favor the $\alpha$ structure, residues in red favor the $\beta$ structure, and 
-white residues have no preference. Some of the mutated residues are also labelled.}
-\label{fig:perresidue_orban}
+    \caption{Per-residue difference free energies for the $\alpha$ and $\beta$ conformations of the GA95 and GB95
+        sequences. Residues colored in blue favor the $\alpha$ structure, residues in red favor the $\beta$ structure,
+        and white residues have no preference.}
 \end{figure}
 
-Although the overall free energy difference between the two structures is small (within 3-5 Kcal/Mol), the individual residues 
-can show marked preferences for being in either $\alpha$ or $\beta$
-conformation. Such differences can be undestood by looking in detail at the local environemnt for
-those resiudes. For example, the region around residue 7 forms a random coil in the $\alpha$
-structure, whereas it is forming a beta sheet structure in the $\beta$ structure. These residues
-strongly favour the locally well packed and hydrogen bonded environment found in the $\beta$ sheet.
-Overall, favoring $\alpha$ or $\beta$ structure is a delicate balance, were the relative global free
-energy difference is small and the contributions of different per residue tendencies balance out. It
-is therefore very likely that by changing key resiude preferences the global fold preferences can be
-changed.
+Although the overall free energy difference between the two structures is small (within 3--5 kcal/mol), individual
+residues can show marked preferences for being in either $\alpha$ or $\beta$ conformation. Such differences can be
+understood by looking in detail at the local environment for those residues. For example, the region around residue 7
+forms a random coil in the $\alpha$ structure, whereas it is forming a beta sheet structure in the $\beta$ structure.
+These residues strongly favour the locally well packed and hydrogen bonded environment found in the $\beta$ sheet.
+Overall, favoring $\alpha$ or $\beta$ structure is a delicate balance, were the relative global free energy difference
+is small and the contributions of different per residue tendencies balance out. It is therefore very likely that by
+changing key residue preferences the global fold preferences can be changed.
 
-Three residues are mutated between GA95 and GB95, at positions 20, 30, and 45. In GA95, these
-residues (L20, I30 and L45) stabilize the $\alpha$ structure. In GB95, two of these residues (F30
-and Y45) favor the $\beta$ structure, because they have large solvent exposed surface areas in the
-$\alpha$ structure, but are more buried in the $\beta$ structure. Additionally Y45 forms a hydrogen
-bond with D47 in the $\beta$ structure. On the other hand, residue A20 from GB95 still favors the
-$\alpha$ structure, although not as strongly as L20 in GA95.
+Three residues are mutated between GA95 and GB95, at positions 20, 30, and 45. In GA95, these residues (L20, I30 and
+L45) stabilize the $\alpha$ structure (compare the upper and lower panels on Figure~\ref{fig:perresidue_orban}). In
+GB95, two of these residues (F30 and Y45) favor the $\beta$ structure, because they have large solvent exposed surface
+areas in the $\alpha$ structure, but are more buried in the $\beta$ structure. Additionally Y45 forms a hydrogen bond
+with D47 in the $\beta$ structure. On the other hand, residue A20 from GB95 still favors the $\alpha$ structure,
+although less strongly than L20 in GA95.
 
-The $\alpha$ and $\beta$ sequences are nearly identical, so there are some common features observed for all
-sequences. Roles of some important residues in stabilizing either the $\alpha$ or $\beta$
-conformation and possible reasons for such conformational preferences are summerized in 
-supporting information Figure S1. The experimental observations
-classified the protein into two parts: Amino acids 9-51 are fully structured in both folds, whereas
-residues 1-8 and 52-56 are unstructured in the $\alpha$ fold, but form a $\beta$-strand in the
-$\beta$ fold. Most of the amino acid residues in the region 1-9 have negative per-residue free
-energies, which means that these residues favor the $\beta$ structure.
+The $\alpha$ and $\beta$ sequences are nearly identical, so there are some common features observed for all sequences.
+The experimental observations classified the protein into two parts: Amino acids 9--51 are fully structured in both
+folds, whereas residues 1--8 and 52--56 are unstructured in the $\alpha$ fold, but form $\beta$-strands in the $\beta$
+fold. Most of the amino acid residues in the region 1--9 have negative per-residue free energies, which means that these
+residues favor the $\beta$ structure. Roles of some other important residues in stabilizing either the $\alpha$ or
+$\beta$ conformation are summarized in Figure S1.
 
-In addition to the direct effects of the mutation, there are also indirect effects due to small
-perturbations due to the mutations. For example, the L20A mutation causes a slight repacking around
-residue 20. This causes large changes in the per-residue free energies of nearby residues T25 and
-A26. However, these differences largely cancel and the overall effect of the L20A mutation is weaker
-than either the I30F or L45Y mutations.
+In addition to the direct effects of the mutation, there are also indirect effects due to small perturbations in the
+environment around the mutations. For example, the L20A mutation causes a slight repacking around residue 20. This
+causes large changes in the per-residue free energies of nearby residues T25 and A26. However, the changes for these two
+residues have opposite signs and nearly cancel.
 
-Such per-residue free energy decompositions allow us to understand the driving forces behind protein
-folding and conformational change and may be useful for designing proteins with specific structures
-and functions.
+These per-residue free energy decompositions provide a great deal of insight into the driving forces behind protein
+folding and conformational change. We believe that such calculations may also be useful for protein design---designing
+proteins with specific structures and functions.
 
-\subsection{Applcation of confinement method for structure prediction}
+\subsection{The confinement method is a useful tool for structure prediction}
 
-We applied the confinement method as a ``meta-predictor'' for structure prediction. Here, the
-task is to correctly identify the most accurate models out of a set of ``decoys'' generated by
-different methods during the Critical Assessment of Structure Prediction (CASP) experiment. CASP is
-a blind test in which different groups apply methods to predict the 3-dimensional structure of
-proteins from their sequences. Each group is allowed to submit five
-possible structures, which they are supposed to rank from best to worst. We have performed two
-experiments centered on CASP.
+We have tested the ability of the confinement method to act as a ``meta-predictor'' for structure prediction. Here, the
+task is to correctly identify the most accurate models out of a set of ``decoys'' generated by different methods during
+the CASP experiment. CASP is a blind test in which different groups apply methods to predict the 3-dimensional structure
+of proteins from their sequences. Each group is allowed to submit five possible structures, which they are supposed to
+rank from best to worst.
 
-In the first experiment, we examined several targets and tried to rank-order the predictions
-made by the same group (same methodology used to generate the models), in others the
-structures were drawn from several different groups (different methodology used). The goal of this 
-experiment is to determine whether 
-the physics-based confinement method can correctly identify more native-like structures as having
-low free energies.
+We have performed two experiments centered on CASP. In the first experiment, we tried to rank-order predictions for
+several targets. For each target, the predictions were either produced by a single group---presumably using the same
+method for each prediction, or were produced by several different groups---using different methods. The goal of this
+experiment is to determine if the physics-based confinement method can correctly identify more native-like structures as
+having lower free energies. Our second experiment was to see if the confinement method can identify structures that are
+missed by other meta-prediction servers. Most successful meta-prediction servers are based on the idea of consensus: if
+many different prediction methods produce similar results, then that is probably a correct prediction [ref]. This is
+often a powerful heuristic, but it can miss cases where there is a very good result that is only predicted by one
+method. We chose several cases where such structures were missed by the best meta-predictors in CASP and assessed if the
+confinement method can correctly identify these accurate models.
 
-Our second experiment was to see if the confinement method can identify structures that are missed
-by other meta-prediction servers. Most successful meta-prediction servers are based on the idea of
-consensus: if many different prediction methods produce similar results, then that is probably a
-correct prediction. This is often a powerful heuristic, but it can miss cases where there is a very
-good result that is only predicted by one method. We chose several cases where such structures were
-missed and assessed if the confinement method can correctly identify these accurate models.
+As is common in the CASP experiment, we assess our results in terms of Global Distance Test Total Score (GDT-TS)
+\cite{Zemla2003}, which is a C$\alpha$ based measure of structural accuracy. It can be understood roughly as the
+percentage of residues that are correctly positioned in the model (range 0 to 100, higher is better). We did have enough
+computer model to analyze every model, so the initial models for the test were chosen from a selection of different
+server groups that have done well in past CASP events.
 
-As is common in the CASP experiment, we assess our results in terms of Global Distance Test Total
-Score (GDT-TS) \cite{Zemla2003}, which is a C$\alpha$ based measure of structural accuracy. It can
-be understood roughly as the percentage of residues that are correctly positioned in the model
-(range 0 to 100, higher is better). The initial models for the test was chosen from different server 
-groups those have traditionally done well in past CASP events.
-
-\subsubsection{The confinement method can correctly rank different models generated by the same methodology
-and distinguish the native structure}
-
-Here, our task is to rank different submitted models of CASP9 targets that have been generated with
-the same methodology. In absence of any other reliable alternative, we compare all our free energy based 
-ranking with the GDT-TS based ranking which are available in CASP website.
-
-The first test case is a protein BVU3908 from Bacteroides vulgatus whose PDB id and CASP target 
-code are 2L01 and T0559 respectively.
-The best predictor group for this 69 amino acid long target was "BAKER-ROSETTASERVER". We initially checked similarity
-between the models submitted by them and discarded two from the analysis on
-the basis of being too similar to some of the other models. The GDT\_TS score and rmsd values as
-shown in Figure~\ref{fig:T0559} indicate that model 1 was predicted correctly, whereas the order of
-model 3 and 5 was wrong. The main difference between model 3 and rest of the models was the  
-orientation of the first alpha helix. On the other hand, as shown
-in Figure~\ref{fig:T0559} the confinement method not only can differentiate the native
-structure from the submitted models, the ranking also correlates perfectly with the GDT\_TS score.
-
-\Ken{Arijit- we could probably make a better visualization of this figure above by putting it onto a toy landscape: x-axis is rmsd, y-axis is DDG, and above each of the 4 points on this graph will be the ribbon diagram and the GDT-TS.
-I suggest we do that with the fig below too.
-Let's don't use 4 different colors for the fig below.
-Let's just show differences in structures in some different color, or show them all as green.]}
+Overall, the confinement method performs well on our CASP tests. Figure~\ref{fig:summary_casp} shows that in almost
+every case, the native structure has the lowest free energy and the best model has the next lowest free energy. The
+confinement method appears to be a useful tool for ranking structure predictions.
 
@@
 \begin{center}
-\includegraphics[width=3.8 in,height=3.2 in]{T0559.pdf}
+\includegraphics[width=3.5 in]{casp.pdf}
 \end{center}
-\caption{The native and three submitted model structure, along with their GDT\_TS, RMSD and relative Free energy values
-of protein BVU3908 from Bacteroides vulgatus (PDB id: 2L01 and CASP code: T0559).}
-\label{fig:T0559}
+\caption{The confinement method is usually able to identify the native structure and the best model (the model with the
+    highest GDT-TS) from a set of decoys.}
+\label{fig:summary_casp}
 \end{figure}
 
-To further test the method we continued with the example of protein BT2368 from Bacteroides
-thetaiotaomicron. The PDB id of this 74 amino acid residue protein is 2L02 and CASP target code is
-T0560. We compared the free energy difference between the native structure and the two of the
-five submitted models from the group "Splicer". The remaining three models were discarded as 
-they were too similar to the rest of the models. As shown in Figure~\ref{fig:T0560}, we can identify 
-the native state correctly and our calculated free energy based ranking matches well with the
-GDT\_TS score.
+\subsubsection{The confinement method can correctly rank-order structure predictions}
+
+First, we examined the ability of the confinement method to rank different models for a target that have been generated
+by a single group using the same methodology for all predictions. We examined two targets: T0559 and T0560 (see Table
+?).
+
+\Justin{[Arijit: Let's move the protein names, organisms, pdb ids, etc to a separate table and just use the CASP ids.]}
+
+The first test case is target T0559. The best predictor group for this 69 amino acid target was ``BAKER-ROSETTASERVER''.
+To save computer time, we excluded two models that were very similar to other models that we did include. For this
+target, the difference free energy computed by the confinement method can be used to accurately rank-order all of the
+models and the native structure (Figure~\ref{fig:T0559}).
+
+\Ken{Arijit- we could probably make a better visualization of this figure above by putting it onto a toy landscape:
+    x-axis is rmsd, y-axis is DDG, and above each of the 4 points on this graph will be the ribbon diagram and the
+    GDT-TS.  I suggest we do that with the fig below too.  Let's don't use 4 different colors for the fig below.  Let's
+    just show differences in structures in some different color, or show them all as green.]}
+
+\Justin{Ken: I don't mind the 3d plots that are in here now, but I'm not sure I would like something that looks like a
+    free energy landscape.}
 
@@
     \begin{center}
-\includegraphics[width=4.0 in,height=3.0 in]{T0560.pdf}
+        \includegraphics[width=3.8 in]{T0559.pdf}
+        \label{fig:T0559}
     \end{center}
-\caption{Native and two model structure of protein BT2368 from Bacteroides thetaiotaomicron (pdb id: 2L02 and CASP code: T0560). The two models
-were from the group "Splicer".}
+    \caption{The confinement method can correctly rank the native structure and three predictions produced by a single
+        method.}
+\end{figure}
+
+\Justin{I'm not sure we need a second figure of this. Move it to the SI?}
+
+We performed a similar calculation for target T0560 with two models form the group ``Splicer''. The remaining three
+models were discarded as they were too similar to the rest of the models. Again, we can correctly identify the native
+sate and our calculated free energy based ranking matches well with GDT-TS (Figures S?).
+
+\begin{figure}
+    \begin{center}
+        \includegraphics[width=3.5 in]{T0560.pdf}
+    \end{center}
+    \caption{Native structure and two models (from group ``Splicer'') for target T0560.}
     \label{fig:T0560}
@@
 
-\subsubsection{The confinement method can rank models generated by different methods}
-
-Our next test was between models that are produced with different methodologies. 
-Here, our model system was a fas apoptosis inhibitory protein molecule whose
-pdb id and CASP target codes are 3MX7 and T0540 respectively. 
-This protein contain 8 beta strands and 90 amino acid residues. 
-The top models from groups "LTB" and "Mufold" was chosen for analysis.
-We will call this as Model 1 and Model 2 for future references. As summerized in
-Figure~\ref{fig:T0540}, the free energy based ranking once again correlates well with the 
-GDT\_TS based score.
+Next, we tested the ability of the confinement method to rank models for target T0540 that were produced by different
+prediction groups. The top models from groups ``LTB'' (Model 1) and ``Mufold'' (Model 2) were chosen for analysis. Once
+again, the free energy based ranking correlates well with the GDT-TS based score (Figure~\ref{fig:T0540}).
 
@@
     \begin{center}
-\includegraphics[width=4.0 in,height=3.4 in]{T0540.pdf}
+        \includegraphics[width=3.5 in]{T0540.pdf}
     \end{center}
-\caption{X-ray crystallographic structure and two submitted models of fas apoptosis inhibitory protein (pdb id: 3MX7 and CASP code: T0540).
-Model 1 and Model 2 in this analysis were submitted by the group LTB and MUFOLD respectively.}
+    \caption{The confinement method can correctly rank the native structure and two models submitted by different
+        prediction groups.}
     \label{fig:T0540}
@@
 
-\subsubsection{Per residue free energy calculation identifies the residues that are responsible for
-differences between two conformations}
+\subsubsection{The per-residue free energy is sensitive to small changes in protein conformation}
 
-\Ken{[Arijit - This and the following sections need to be shortened and made more concise.]}
+In section~\protect\ref{sec:orban_per_residue} we discussed how the per-residue free energy can reveal mechanistic
+detail behind the conformational preference of a chameleon sequence. In this section, our aim is to apply the same
+method and try to understand if the per-residue free energy can help us identify residues which stabilize or destabilize
+a particular region of a protein. In this case, the two conformations have similar folds with only small changes in
+localized regions. We chose CASP target T0569 and compared the experimental NMR structure with the best predicted model
+(GDT-TS=78; predicted by the ``Mufold'' group). The confinement method predicts that the native structure is more stable
+by 20 kcal/mol (Figure S2).
 
-In section~\protect\ref{sec:orban_per_residue} we discussed how per residue free energy calculation
-can identify the mechanistic detail behind conformational preference of a chameleon sequence. In this section, our aim 
-is to apply the same method and try to understand whether per residue free energy can help us identify
-residues which stabilizes/destabilizes a particular region of a protein. The difference  
-here is that the protein has similar fold with only mismatch in some region. 
-To investigate this, we choose a domain of adhesion exoprotein from Pediococcus pentosaceus (pdb id=2KYW) from protein data bank 
-and its best predicted model of CASP9 (casp id=T0569, submitted by 'Mufold' group which had a GDT\_TS of 78).
-We first calculated the free energy of the whole protein and found that the native 
-structure is stabilized by 20 Kcal/Mol (see suppurting information Figure S2).
+The confinement method clearly identifies two hydrophobic residues V59 and I61, which destabilize the predicted model
+with respect to the crystallographic structure (Figure~\ref{fig:T0569_per_residue} and Figure S3). The sidechains of
+these hydrophobic residues are oriented towards the protein hydrophobic core in the native NMR structure but oriented
+towards the exterior of the protein and are solvent exposed in the model. These residues also form part of a beta-sheet
+in the experimental structure, but do not form inter-strand hydrogen bonds in the predicted model. There is also a large
+difference around K76, which forms a salt-bridge with D11 in the predicted model, but not in the experimental structure.
+This suggests that salt-bridge interactions are too favorable for the combination of force field and implicit solvent
+model we use, which has been a problem noted in the past [ref carlos].
 
-Next we proceed with the per residue free energy calculation. Our effort clearly identifies two hydrophobic residues 
-Val-59 and Ile-61, which destabilizes the submitted model with respect to the crystallograohic structure.
-The sidechains of these hydrophobic residues are oriented towards the protein hydrophobic core in the
-native NMR structure but oriented outside protein and solvent exposed in the model. These residues are the primary reason
-why two of the beta sheets in the native structure are disordered in the model structure ( see Figure~\ref{fig:T0569_per_residue}
-and supporting information Figure S3). This method also identifies some other residues which stabilize/destabilize either 
-of the conformation. For
-example, Lys-76 stabilizes the model as it has a H-bond with Asp-11 which is missing in the native structure
-(Figure~\ref{fig:T0569_per_residue} and supporting information figue S3).
+\Justin{[Arijit: Figure~\ref{fig:T0569_per_residue} says crystal, but the text says NMR.]}
 
@@
     \begin{center}
-\includegraphics[width=4.3 in,height=4 in]{T0569_perres1.pdf}
+        \includegraphics[width=3.5in]{T0569_perres1.pdf}
     \end{center}
-\caption{To better understand the reason behind difference in conformation, structure of CASP Target 569 (pdb id: 2KWY) and a 
-generated model was colored according to the per residue free energy value. The amino acid residues that are colored in deep red 
-and deep blue stabilizes the crystal and the model respectively, while the residues with light blue color does not have a preferance.
- Some of the important residues are also labelled.}   
+    \caption{Per-residue difference free energy between the experimental NMR structure and the best prediction for CASP
+        target T0569. The amino acid residues that are colored in deep red and deep blue stabilizes the NMR structure
+        and the prediction, respectively; the residues with light blue color does not have a strong preference.}
     \label{fig:T0569_per_residue}
@@
 
-\subsubsection{Failures in the confinement method}
+\subsubsection{The confinement method occasionally produces incorrect results}
 
-Despite the great success in most of the studied systems, there are few failures, specially when
-the GDT scores of the compared structures are very close. We observed this for a $54$ amino acid residue long
- engineered protein from Asr4154 protein (pdb id: 2L09 and CASP code T0538).
-Like previous cases, we kept the native structure and include models from the group "PconsR" (GDT\_TS=96.23, model1),
-"Shell" (GDT\_TS = 90.09, model2) and "FOLDIT" (GDT\_TS = 86.32, model3) for analysis. 
-Contrary to our expectation, model1 was found to be more stable than the crystal structure 
-(see Figure~\ref{fig:T0538}). In order to rationalize such obseravtion we calculates per residue free energy
-between crystal structure and model1. The results show us that despite the small variation at the backbone level (as
-shown by high GDT\_TS scores and low RMSDs), the sidechains are oriented in very different ways,
-giving rise to large differences in the stabilization of certain residues. In particular, some of
-the differences arise from different salt bridge patterns (Arg-32 with Glu-35, Glu-28 with Lys-24 and Arg-26 with Glu-50
-in crystal versus Arg-32 with Glu-28 in model 1) and certain flexible polar residues exposed to the
-surface (Lys-24 in model 1, which has an entropic gain from absence of salt bridge and is
-stabilized by interactions with the solvent) (See supporting information Figure S4). All in all, this
-unexpected result just shows us that this method is sensitive to local interactions such as those
-happening from side chain reorientation and also indicate the limitations of GDT\_TS based ranking.
+\Justin{Arijit: let's move Figure~\ref{fig:T0538} to the supporting info and move Figure S4 into the main text.}
+
+Despite success for most of the studied systems, there are few failures, specially when the GDT-TS scores of the
+compared structures are very close. One example is Target T0538, where we compared the crystal structure with three
+models (Model 1: ``PconsR''---GDT-ts=96; Model 2: ``Shell''---GDT-TS=90; Model 3: ``FOLDIT''---GDT-TS=86). Contrary to
+our expectation, the confinement method predicts that Model 1 is more stable than the crystal structure (Figure ?).
+Per-residue free energy calculations (not shown) show that despite only small variations at the backbone level, the
+sidechains are oriented in very different ways (Figure ?), giving rise to large differences in the stabilization of
+certain residues. In particular, some of the differences arise from different salt bridge patterns and certain flexible
+polar residues exposed to the surface. This unexpected result shows that the confinement method is very sensitive to
+local interactions (including sidechain reorientation) and may indicate issues with the forcefield and implicit solvent
+models used in the calculation.
 
@@
 \begin{center}
-\includegraphics[width=3.6 in,height=3.4 in]{T0538.pdf}
+\includegraphics[width=3.5in]{T0538.pdf}
 \end{center}
-\caption{The native and three model structure of engineered protein from Asr4154 protein (PDB ID: 2L09 and CASP code:T0538). The model 1,2 and 3 are
-from the group PconsR, Shell and FOLDIT respectively.}
+\caption{The native and three model structure of engineered protein from Asr4154 protein (PDB ID: 2L09 and CASP
+    code:T0538). The model 1,2 and 3 are from the group PconsR, Shell and FOLDIT respectively.}
 \label{fig:T0538}
@@
 
-So far we have seen that the method is good at predicting preferences when the structures are not
-very far from the native. But the question remains how far from native can we go and still see 
-that the method produces correct result. In this section we explored this question with models from  
-extracellular domain of the jumping translocation breakpoint protein (pdb id: 2KJX). Most of the
-group could only generate low resolution models for this CASP9 target (id: T0531) . In our comparison, we choose five
-models by the group MUFOLD-MD, which was the best performing group for this target with their best model 
-had a GDT\_TS value of 44. The result presented in Figure~\ref{fig:T0531} shows: 1.) native is correctly identified as
-expected and 2.) Surprisingly there is a high level of correlation between the GDT ranking and the
-free energy ranking for model 1 and model 3, the rest three structures with GDT\_TS score less than
-35  are ordered incorrectly. It is worth to note that models 2 and 3 have the same GDT and very
-different free energies, meaning that the actual ordering could change a lot \cite{Perez2012}. It is
-encouraging that at least the method can pick out the best model even though it is got a low GDT
-score: 44.
+\Justin{I think we should consider dropping this whole section. We have a sample size of one and I don't expect that we
+    can generally do well ranking models with a GDT of 44. I think this is a fluke.}
+
+So far we have seen that the method is good at predicting preferences when the structures are not very far from the
+native. But the question remains how far from native can we go and still see that the method produces correct result. In
+this section we explored this question with models from  extracellular domain of the jumping translocation breakpoint
+protein (pdb id: 2KJX). Most of the group could only generate low resolution models for this CASP9 target (id: T0531) .
+In our comparison, we choose five models by the group MUFOLD-MD, which was the best performing group for this target
+with their best model had a GDT\_TS value of 44. The result presented in Figure~\ref{fig:T0531} shows: 1.) native is
+correctly identified as expected and 2.) Surprisingly there is a high level of correlation between the GDT ranking and
+the free energy ranking for model 1 and model 3, the rest three structures with GDT\_TS score less than 35  are ordered
+incorrectly. It is worth to note that models 2 and 3 have the same GDT and very different free energies, meaning that
+the actual ordering could change a lot \cite{Perez2012}. It is encouraging that at least the method can pick out the
+best model even though it is got a low GDT score: 44.
 
@@
 \begin{center}
-\includegraphics[width=3.5 in,height=3.5 in]{T0531.pdf}
+\includegraphics[width=3.5 in]{T0531.pdf}
 \end{center}
-\caption{The native structure and 5 models of extracellular domain of the jumping translocation
-breakpoint protein (pdb id: 2KJX and the CASP code: T0531).}
+\caption{The native structure and 5 models of extracellular domain of the jumping translocation breakpoint protein (pdb
+    id: 2KJX and the CASP code: T0531).}
 \label{fig:T0531}
@@
 
-\Ken{[Arijit - let's make the following section more concise and focused.]}
+\Justin{Arijit---is this the case where we chose a structure that was missed by the consensus predictors, if so, we
+    should try to indicate this. The point of this section should be that although consensus prediction strategies do
+    well, they can miss a good answer that is only predicted by a few methods. At least in this case, the confinement
+    method is able to capture this structure that was otherwise missed.}
 
-A part of the CASP experiment is dedicated to the quality assessment (QA) of predicted models
-\cite{Kryshtafovych2011}. Here, predictors were asked to score each model on a scale 
-(known as qmode) from 0 to 1, with higher values corresponding to better models \cite{Kryshtafovych2011}. 
-It will be interesting to know how confinement method perform in quality assessment compare to 
-the other groups in CASP9. We investigated this using couple of CASP Targets. Here, we present a 
-case, where confinement method perform well than the top performing group MUFOLD-WQA \cite{Wang2011} 
-of CASP9. We choose  
-two models from CASP taget T0538. They were model 3 submitted by PconsR (GDT\_TS = 96, qmode 1 = 0.5434) 
-and model 5 from the MULTICOM-NOVEL ( GDT\_TS = 83, qmode 1 = 0.5865). Both are server predicted models and
-the qmode value presented are from MUFOLD-WQA \cite{Wang2011}. We choose these two models as
-the model by PconsR was the most accurate predicted model for this target, whereas the other model 
-was predicted best by QA test of MUFOLD-WQA. The calculated free energy using confinement method indicate
-that the model by PconsR is more stable by $3.9 Kcal/Mol$ which support the GDT\_TS trend.
-There is no doubt that the consensus approach can predict the model quality
-in a faster manner. But we expect confinement method can predict the model quality in a
-relatively expensive but much more accurate way.
+A part of the CASP experiment is dedicated to the quality assessment (QA) of predicted models \cite{Kryshtafovych2011}.
+Here, predictors were asked to score each model on a scale (known as qmode) from 0 to 1, with higher values
+corresponding to better models \cite{Kryshtafovych2011}. It will be interesting to know how confinement method perform
+in quality assessment compare to the other groups in CASP9. We investigated this using couple of CASP Targets. Here, we
+present a case, where confinement method perform well than the top performing group MUFOLD-WQA \cite{Wang2011} of CASP9.
+We choose  two models from CASP taget T0538. They were model 3 submitted by PconsR (GDT\_TS = 96, qmode 1 = 0.5434) and
+model 5 from the MULTICOM-NOVEL ( GDT\_TS = 83, qmode 1 = 0.5865). Both are server predicted models and the qmode value
+presented are from MUFOLD-WQA \cite{Wang2011}. We choose these two models as the model by PconsR was the most accurate
+predicted model for this target, whereas the other model was predicted best by QA test of MUFOLD-WQA. The calculated
+free energy using confinement method indicate that the model by PconsR is more stable by $3.9 Kcal/Mol$ which support
+the GDT\_TS trend. There is no doubt that the consensus approach can predict the model quality in a faster manner. But
+we expect confinement method can predict the model quality in a relatively expensive but much more accurate way.
 
@@
 
-We have described a computational method called confinement for computing the difference free energy from a protein conformation $A$ to $B$.  We 
-show that it can give accurate values, even for relatively large conformational changes.  We have demonstrated that it can discriminate 
-which Orban chameleon peptides fold into a 3-helix bundle vs. a 4$\beta$+1 helix structure. We show that it can discriminate between good and bad CASP 
-models.  Perhaps most importantly, we have shown that it can be used to give residue-level insights into what are the dominant structural factors in a 
-protein that are responsible for the difference free energies.  A key advantage is that this method does not require any reaction coordinate, or sampling 
-a pathway from conformation $A$ to $B$. We have tested the method for structures of proteins having up to 100 amino acids.  The computational cost 
-for a 56-residue protein is about 4 hours [on 1 gpu] for 20 ns of confinement [Is 20ns all you need?].
+We have described a computational method called confinement for computing the difference free energy between two
+conformational ensembles. We show that the difference free energy can be calculated on ~100 residue sized proteins, even
+for large conformational changes. We have demonstrated that it can discriminate the folding preferences of a series of
+chameleon proteins. We show that the confinement method can discriminate between the native structure and structure
+predictions and can identify the best prediction reliably. Perhaps most importantly, we have shown that it can be used
+to give residue-level insights into what are the dominant structural factors in a protein that are responsible for the
+difference free energies. The confinement should be useful for protein design, structure prediction, and understanding
+the mechanism of conformational change.
+
+A key advantage is that this method does not require any reaction coordinate, or sampling a
+pathway from conformation $A$ to $B$. We have tested the method for structures of proteins having up to 100 amino acids.
+The computational cost for a 56-residue protein is about 4 hours [on 1 gpu] for 20 ns of confinement [Is 20ns all you
+need?].
 
 \Arijit{I need 21 (confinement run) x 20 ns for single confinement}
-%There is another issue, which is normal mode analysis calculations are Oh(N**3) and require large
-%amounts of memory, effectively limiting the size of the systems under study. In most cases
-%quasiharmonic analysis.
+%There is another issue, which is normal mode analysis calculations are Oh(N**3) and require large amounts of memory,
+%effectively limiting the size of the systems under study. In most cases quasiharmonic analysis.
 
 \Ken{[If this is a relevant limitation, let's say something about it.]}
-\Arijit{I think we need to write the normal mode issue and number of residues that we can handle. If 
-one of you contribute here that will be great.}
+
+\Arijit{I think we need to write the normal mode issue and number of residues that we can handle. If one of you
+    contribute here that will be great.}
+
+\Justin{The conclusion is not the place to talk about how long the method takes or what the technical challenges are.
+    Let's move this to methods section.}
 
@@
 
-\item  In the first step, a minimization of A and B conformations are performed. These minimized conformations (A* and B*)
-       are the reference conformation of that state.
+\item  In the first step, a minimization of A and B conformations are performed. These minimized conformations
+    ($A^{\ast\ast}$ and $B^{\ast\ast}$) are the reference conformation of that state.
 
-\item  The free energy of confining the ensemble (A or B) to a microstate (A* or B*) is
-       calculated. This is done by gradually applying larger and larger
-       harmonic restraints on all the atoms of the biomolecule. This is done by running 21 molecular dynamics simulation
-       (each 20 ns long) for each leg of the thermodynamic cycle, where the harmonic restraint force constant was scaled from 0.00005
-       Kcal/Mol (mostly free) to 81.92 Kcal/Mol (frozen in one microstate). In this final restrained state, the
-       rotational contribution to the free energy is frozen out
-       and the only remaining contribution is the vibrational part. The free energy for this step is
-       estimated from the fluctuations around
-       the reference structure using a numerical
-       approach developed by Tyka et. al. \cite{Tyka2006}. The confinement free energy calculated in
-       this way is recorded as
-       $\Delta G_{A,A*}$ and $\Delta G_{B,B*}$ as shown in Figure~\ref{fig:method}.
+\item The free energy of confining the ensemble (A or B) to a tighter ensemble ($A^\ast$ or $B^\ast$) is calculated.
+    This is done by gradually confining each atom to its position in the reference conformation ($A^{\ast\ast}$ or
+    $B^{\ast\ast}$) using a series of progressively stronger position restraints. This is done by running 21 molecular
+    dynamics simulation (each 20 ns long) for each leg of the thermodynamic cycle, where the harmonic restraint force
+    constant was scaled from 0.00005 Kcal/Mol (mostly free) to 81.92 kcal/mol (tightly restrained). The free energy for
+    this step is estimated using a thermodynamic integration approach developed by Tyka et. al. \cite{Tyka2006}. The
+    confinement free energies calculated this step are denoted as $\Delta G_{A,A^\ast}$ and $\Delta G_{B,B^\ast}$ in
+    Figure~\ref{fig:method}.
 
-\item  Finally the thermodynamic cycle is closed by calculating the free energy between the final
-       restrained state A* and B* using normal mode analysis or quasiharmonic analysis. The free energy calculated in
-       this way is shown as $\Delta G_{A*,B*}$ in Figure~\ref{fig:method}.
+\item The thermodynamic cycle is closed by calculating the free energy between the final restrained states $A^\ast$ and
+    $B^\ast$ using normal mode analysis or quasiharmonic analysis. The free energy calculated in this way is shown as
+    $\Delta G_{A^\ast,B^\ast}$ in Figure~\ref{fig:method}.
 
-\item  The full free energy, $\Delta G_{A,B}$ between the two state A and B is calculated using the equation
-       $\Delta G_{A,B}$ = $\Delta G_{A,A*}$ - $\Delta G_{B,B*}$ + $\Delta G_{A*,B*}$
+\item  The full free energy, $\Delta G_{A,B}$ between the two state A and B is calculated as
+    $\Delta G{A,B} = \Delta G{A,A^\ast} - \Delta G{B,B^\ast} + \Delta G{A^\ast,B^\ast}$.
 
@@
 
-All calculations where performed with the amber 11 suit of programs \cite{Case2012},\cite{Goetz2012} in
-combination with ff99SB forcefield\cite{Hornak2006} and generalized born implicit solvent \cite{Mongan2006}.
-Interestingly, we extend the method for calculation of per residue free
-energy in an approximate way. For this purpose, the confinement energy, $\Delta G_{A,A*}$ and
-$\Delta G_{B,B*}$ of each residue is calculated in the usual numerical way as described in ref. by Tyka 
-et al. \cite{Tyka2006}. The internal energy of each residue is calculated using the decomp module of amber from the final
-restrained trajectory. We call this method approximate as we ignore the entropic contribution from the normal
-mode or quasiharmonic analysis. However, this contrbution to the total free energy is much smaller,
-which allow us to study the mechanistic details of conformational preference of each residue.
+All calculations where performed with the Amber 11 suit of programs~\cite{Case2012,Goetz2012} in combination with ff99SB
+forcefield~\cite{Hornak2006} and generalized born implicit solvent \cite{Mongan2006}. Interestingly, we extend the
+method for calculation of per residue free energy in an approximate way. For this purpose, the confinement energy,
+$\Delta G_{A,A*}$ and $\Delta G_{B,B*}$ of each residue is calculated in the usual numerical way as described in ref. by
+Tyka et al. \cite{Tyka2006}. The internal energy of each residue is calculated using the decomp module of amber from the
+final restrained trajectory. We call this method approximate as we ignore the entropic contribution from the normal mode
+or quasiharmonic analysis. However, this contrbution to the total free energy is much smaller, which allow us to study
+the mechanistic details of conformational preference of each residue.
 
-%There are two big bottlenecks in this calculation. The first has to do with the amount of
-%simulations that have to be performed to confine the ensemble of conformations close to a
-%microstate into a single microstate, the second bottleneck has to do with the size of the system in
-%the normal mode calculation. We have tackled the first issue by using GPU (Graphical Processing
-%Unit) technology as opposed to the classical CPUs, which gives us a two order of magnitude boost in
-%computational efficiency. Normal mode analysis calculations are Oh(N**3) and require large amounts
-%of memory, effectively limiting the size of the systems under study. Where as this can be overcome
-%by larger memory machines, these are often not available, and so we propose an alternative by using
-%quasiharmonic analysis. This is still an Oh(N**3) problem, but we now the motions of atoms are
-%already determined in trajectories, and so we can choose to do quasiharmonic analysis excluding
-%hydrogens, which will reduce the number of atoms N to almost N/2. This means that we can use this
-%method for systems twice as large.
+%There are two big bottlenecks in this calculation. The first has to do with the amount of simulations that have to be
+%performed to confine the ensemble of conformations close to a microstate into a single microstate, the second
+%bottleneck has to do with the size of the system in the normal mode calculation. We have tackled the first issue by
+%using GPU (Graphical Processing Unit) technology as opposed to the classical CPUs, which gives us a two order of
+%magnitude boost in computational efficiency. Normal mode analysis calculations are Oh(N**3) and require large amounts
+%of memory, effectively limiting the size of the systems under study. Where as this can be overcome by larger memory
+%machines, these are often not available, and so we propose an alternative by using quasiharmonic analysis. This is
+%still an Oh(N**3) problem, but we now the motions of atoms are already determined in trajectories, and so we can choose
+%to do quasiharmonic analysis excluding hydrogens, which will reduce the number of atoms N to almost N/2. This means
+%that we can use this method for systems twice as large.
 

### 2013-05-03 Paper/supporting_information.tex
changed 3May

@@
 
-We prepared computer generated models in both $\alpha$ and $\beta$ conformation for all five chamelon sequences mentioned 
-in the main text. Crystallographic structure of GA95 and GB95 sequence have $\alpha$ and $\beta$ fold respectively. When,
-we attempt to prepare the $\alpha$ and $\beta$ fold with GA95 sequence we start with the crystallographic 
+We prepared computer generated models in both $\alpha$ and $\beta$ conformations for all five chamelon sequences mentioned 
+in the main text. Crystallographic structure of GA95 and GB95 sequence have $\alpha$ and $\beta$ fold respectively. When
+we attempt to prepare the $\alpha$ and $\beta$ fold with GA95 sequence, we start with the crystallographic 
 structure of GA95 and GB95 with GA95 sequence. However, we remove the sidechains at three mutated positions in both the fold 
@@
 
+\setcounter{figure}{0}
+\makeatletter 
+\renewcommand{\thefigure}{S\@arabic\c@figure}
 \begin{figure}
@@
 residues control the free energy equilibrium to either $\beta$ or
-$\alpha$ conformation. All figures in the left are in $\beta$ conformation, whereas the right ones are in $\alpha$ conformation except in Figure (F).
-Sidechains of hydrophobic residues Leu-7 in (A) and Ala-26 in (B) are oriented towards the protein interior in $\beta$ 
+$\alpha$ conformation. All figures in the left are in $\beta$ conformation, whereas the right ones are in $\alpha$ conformation except in Figure (F) 
+where both are in $\alpha$ conformation.
+Sidechains of hydrophobic residues L7 in (A) and A26 in (B) are oriented towards the protein interior in $\beta$ 
 conformation. In such conformation they have hydrophobic interactions with other hydrophobic residues colored in white. In $\alpha$ conformation these residues are 
 exposed to the protein surface.  
-(C) Tyr-45 forms H-bond with Asp-47 with GB95 sequence and $\beta$ conformation. (D) Gln-11 and Glu-15 form a H-bond in $\beta$ conformation which is absent
-in the $\alpha$ conformation. (E) Hydrophobic Ile-49 is exposed to the protein surface in $\beta$ conformation, whereas it is inside protein hydrophobic core in
-$\alpha$ conformation. (F) Sidechain of Ile-30 is inside the protein hydrophobic core in as it has a smaller sidechain with GA95 sequence in $\alpha$ conformation. 
-In contrast, Phe-30 with larger sidechain is exposed to the protein surface with GB95 sequence.}
+(C) Y45 forms H-bond with D47 with GB95 sequence and $\beta$ conformation. (D) Q11 and E15 forms a salt bridge in $\beta$ conformation which is absent
+in the $\alpha$ conformation. (E) Hydrophobic I49 is exposed to the protein surface in $\beta$ conformation, whereas in $\alpha$ conformation it 
+is inside protein hydrophobic core. (F) Sidechain of I30 is inside the protein hydrophobic core as it has a smaller sidechain with GA95 sequence 
+and $\alpha$ conformation. In contrast, Phe-30 with larger sidechain is exposed to the protein surface with GB95 sequence and $\beta$ conformation.}
 \end{figure}
@@
     \begin{center}
-\includegraphics[width=3.6 in,height=3.4 in]{T0569.pdf}
+        \includegraphics[width=3.5 in]{T0560.pdf}
     \end{center}
-\caption{Native and best model structure of a domain of adhesion exoprotein from Pediococcus pentosaceus (pdb id: 2KWY and CASP code: T0569).
-The best model was from the group "Mufold".}
+    \caption{Native structure and two models (from group ``Splicer'') for target T0560.}
+    \label{fig:T0560}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.8 in,height=3.4 in]{T0569.pdf}
+\end{center}
+\caption{Native and best computer generated model structure of CASP Target T0569.} 
 \label{fig:T0569}
@@
 \end{center}
-\caption{Key differences between native crystal structure and the model as indicated by per residue free energy calculation.
-The sidechain of hydrophobic residues, Val-59 and Ile-61 are oriented towards the protein hydrophobic core in (A) crystal 
+\caption{Key differences between experimental structure and the computer generated model model as indicated by per residue free energy calculation.
+The sidechain of hydrophobic residues, V59 and I61 are oriented towards the protein hydrophobic core in (A)  
 structure, (B) exposed to the surfure in model. The beta sheet containing these residues is disordered in the generated model.
-There is a H-bond between Lys-76 and Asp-11 in (D) model structure, which is absent in the (C) native structure.}
+A salt bridge between K76 and D11 in (D) computer generated structure, which is absent in the (C) native structure.}
 \label{fig:T0569_per_residue}
@@
 \begin{center}
-\includegraphics[width=4.9 in,height=4.0 in]{Target_538_compare.pdf}
+\includegraphics[width=3.5in]{T0538.pdf}
 \end{center}
-\caption{Differences of the sidechain orientation in model 1 and crystallographic structure of CASP target T0538 as determined 
-by per residue free energy calculation.  
-(A) Two salt bridges between residues Glu-35-Arg-32 and Glu-28-Lys-24 exist in the native crystallographic structure. These are
-absent in the model structure. Instead
-(B) a new salt bridge is formed between residues Arg-32-Glu-28 in model. (C) Another salt bridge between
-residues  Arg-26-Glu-50 in the native structure, (D) This salt bridge is absent in the model structure. In comparison to the crystallographic 
-structure, orientation of Phe-51 differs in model1.}
-\label{fig:T0538compare}
+\caption{The native and three model structure of engineered protein from Asr4154 protein (PDB ID: 2L09 and CASP
+    code:T0538). The model 1,2 and 3 are from the group PconsR, Shell and FOLDIT respectively.}
+\label{fig:T0538}
 \end{figure}
 
-\Ken{[This fig isn't very readable.
-We need to focus into the sidechains, keeping only the nearby pieces of chain and leaving the rest out or making it fuzzy or something.
-We need to see some point that you are making by looking at the figure.]}
+\section{What can we say about low resolution models from CASP experiments?}
+
+In the main text, we discussed how well confinement method predicted preferences when the structures are not very far from the
+native. But the question remains how far from native can we go and still see that the method produces correct result. In
+this section we explored this question with models from  extracellular domain of the jumping translocation breakpoint
+protein (pdb id: 2KJX). Most of the group could only generate low resolution models for this CASP9 target (id: T0531) .
+In our comparison, we choose five models by the group MUFOLD-MD, which was the best performing group for this target
+with their best model had a GDT\_TS value of 44. The result presented in Figure~\ref{fig:T0531} shows: 1.) native is
+correctly identified as expected and 2.) Surprisingly there is a high level of correlation between the GDT ranking and
+the free energy ranking for model 1 and model 3, the rest three structures with GDT\_TS score less than 35  are ordered
+incorrectly. It is worth to note that models 2 and 3 have the same GDT and very different free energies, meaning that
+the actual ordering could change a lot \cite{Perez2012}. It is encouraging that at least the method can pick out the
+best model even though it is got a low GDT score: 44.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.5 in]{T0531.pdf}
+\end{center}
+\caption{The native structure and 5 models of extracellular domain of the jumping translocation breakpoint protein (pdb
+    id: 2KJX and the CASP code: T0531).}
+\label{fig:T0531}
+\end{figure}
 
@@
 
+\bibitem{Perez2012}
+Perez, A.; Yang, Z.; Bahar, I.; Dill, K.A.; MacCallum, J.L.; FlexE: Using Elastic Network Models to Compare Models of Protein Structure. J. Chem. Theory Comput., 2012, 8, 3985-3991.
+
 \end{thebibliography}

### 2013-05-03 Paper/tt.tex
changed 3May

--- 2013-04-29 Paper/paper.tex
@@
 
-\Ken{[Guys---I recommend we get a better name than confinement.  It's totally uninformative.  In my view, better would
-    be `confine-and-configure' or something.  Please give it some thought.]}
-
-\Justin{[Ken---I agree that the name isn't great, but I don't think we should change it because: it's already in the
-    literature and we didn't come up with the method. I think we should consider capitalizing it as Confinement Method
-    throughout.]}
-
-\section{Introduction}
+\section*{Introduction}
 
@@
 is to use molecular dynamics simulations along some putative reaction coordinate pathway from conformation $A$ to
-$B$~\cite{Elber2007,West2007,Haas2007,Jonsson1998,E2007,Dellago2002,Cheng2006}. \Ken{[we need many refs here: cite
-    Elber's milestoning, Jhih-Wei Chu, and the many other path-sampling approaches.]} The free energy along this
-reaction coordinate can then be determined using methods such as umbrella sampling \cite{Torrie1977}\Justin{[add ref
-    where this is used computing free energy differences between conformations]} or targeted MD \Justin{[refs]}. Such
+$B$~\cite{Elber2007,West2007,Haas2007,Jonsson1998,E2007,Dellago2002,Cheng2006}. The free energy along this
+reaction coordinate can then be determined using methods such as umbrella sampling \cite{Torrie1977,Mascarenhas2013}  
+and weighted histogram analysis method (WHAM) \cite{Kumar1992}. Such
 approaches have several limitations. First, it is necessary to know an efficient reaction pathway from $A$ to $B$. If
@@
 
-\Ken{[Are all these methods below ones that don't involve paths?  Also, are there other key methods we should cite?  We
-    need a good scholarly list of all the relevant works.]}
-
 The calculation of protein conformational free energy has been successfully attempted by a number of groups
@@
 confinement method~\cite{Tyka2006,Cecchini2009,Ovchinnikov2013} take an alternative strategy for computing difference
-free energies that does not require knowing a pathway from $A$ to $B$.  \Ken{[please add the several new Karplus papers
-    from 2012-2013.]}.
+free energies that does not require knowing a pathway from $A$ to $B$.
 
@@
 simulations. Similarly, a second ensemble $B$ is restricted to $B^\ast$. We then compute the free energy between
-$A^\ast$ and $B^\ast$ using either normal mode analysis or the quasiharmonic method [ref]. This step takes into account
+$A^\ast$ and $B^\ast$ using either normal mode analysis \cite{Brooks1983,Case1994} or the quasiharmonic 
+method \cite{Karplus1981,Levy1984}. This step takes into account
 both the remaining conformational entropy in each ensemble and the remaining enthalpy difference. In this way, we can
@@
 coordinate connecting the two states. The confinement method shares some similarities with the `confine-and-release'
-method for computing ligand binding affinities~\cite{Mobley2007,Mobley2012,Mobley2006} \Ken{[cite mobley's jctc 07 (ref
-    216 on our labsite), and mobley's recent j chem phys review, and mobley (ref 199) j chem phys, and others that are
-    relevant.]}.
+method for computing ligand binding affinities~\cite{Mobley2007,Mobley2012,Mobley2006}.
 
@@
 
-\section{Results and Discussion}
+\section*{Results and Discussion}
 
-\subsection{The confinement method succeeds at some basic consistency checks}
+\subsection*{The confinement method succeeds at some basic consistency checks}
 
@@
 
-Second, we looked at 6 target proteins from the CASP9 experiment [ref].  For each target, we examined up to 5 submitted
+Second, we looked at 6 target proteins from the CASP9 experiment~\cite{Moult2011}.  For each target, we examined up to 5 submitted
 models.  We computed the difference free energy between the true native and the best model.
@@
 
-\Justin{[Arijit, please add GDT of best decoy to this table.]}
-
 \begin{table}
@@
     participants. We report the free energy difference between the most favorable decoy and the experimentally
-    determined structure. Positive $\Delta\Delta G$ values indicate that the experimental structure is predicted to be
+    determined structure. Positive $\Delta\Delta G (=G_{best~decoy} - G_{native})$ values indicate that the experimental structure is predicted to be
     more favorable than any of the decoys.}
 \label{table:casp_control}
-\begin{tabular}{l l l}\hline
-    CASP Target  & PDB Identifier & $\Delta \Delta G = G_{best~decoy} - G_{native}$ (kcal/mol) \\ \hline
-     T0531       &    2KJX        &          $11.15 \pm 0.70$ \\ \hline
-     T0538       &    2L09        &          $-3.00 \pm 0.47$ \\ \hline
-     T0540       &    3MX7        &          $16.94 \pm 0.49$ \\ \hline
-     T0559       &    2L01        &          $2.10 \pm 0.24$ \\ \hline
-     T0560       &    2L02        &          $18.00 \pm 0.49$ \\ \hline
-     T0569       &    2KWY        &          $20.01 \pm 0.69$  \\ \hline
+\begin{tabular}{l l l l}\hline
+    CASP Target  & PDB Identifier &  organism type  \\ \hline
+     T0531       &    2KJX        &  extracellular domain of the jumping translocation breakpoint protein &  $44$ \\ \hline
+     T0538       &    2L09        &  protein asr4154 from Nostoc sp. PCC7120        &  $96$ \\ \hline
+     T0540       &    3MX7        &  human Fas apoptotic inhibitory molecule    &  $70$ \\ \hline
+     T0559       &    2L01        &          $2.10 \pm 0.24$     &  $94$ \\ \hline
+     T0560       &    2L02        &          $18.00 \pm 0.49$    &  $94$ \\ \hline
+     T0569       &    2KWY        &          $20.01 \pm 0.69$    &  $78$ \\ \hline
 \end{tabular}
@@
 
-\subsection{The confinement method correctly predicts the structures of chameleon sequences}
+\subsection*{The confinement method correctly predicts the structures of chameleon sequences}
 
 We also tested the confinement method predictions for difference free energies on the chameleon sequences of Orban et
-al.!\cite{Alexander2007,He2008,Alexander2009,Bryan2010,He2012,Shortle2009}. These are instances in which two highly
+al.\cite{Alexander2007,He2008,Alexander2009,Bryan2010,He2012,Shortle2009}. These are instances in which two highly
 similar sequences fold into remarkably different structures. Orban and co-workers have designed a protein-G-like
@@
 these structures presents a serious challenge for computational
-methods~\cite{Alexander2007,He2008,Alexander2009,Bryan2010,He2012,Shortle2009}. \Justin{Do any of these citations try to
-    do any modeling or comment about how challenging this would be?}
+methods \cite{Allison2011}.
 
@@
 
-\Ken{[ It's not clear what this test is and why it's better, and why there are only 5 sequences.  Please clarify.]}
-\Justin{[There are five pairs of sequences that can be arranged into 3 different pairings. Arijit has removed the 5th
-    sequence.]}
-
-\Arijit{ Now there are 4 sequences as they have the experimentally available structure. The fifth sequence is discussed
-    at the end of the paragraph.}
-
-\Justin{[Ken: it's not that hard to tell an experimental structure from a computer generated model. Almost all of the
-    knowledge-based scoring functions can do this a substantial fraction of the time. Comparing a near-native model with
-    another model is a much more challenging problem.]}
-
 Figure~\ref{fig:orban} shows that the confinement method identifies the correct structure for all four sequences. One
@@
 
-\subsection{Per-residue free energy calculations can identify the mechanistic detail behind conformational preferences}
+\subsection*{Per-residue free energy calculations can identify the mechanistic detail behind conformational preferences}
 \label{sec:orban_per_residue}
@@
 contributions helps us to identify important residues that stabilize a particular conformation. The per residue free
-energy, $\Delta \Delta G ((\beta) - (\alpha))$ is shown in Figure~\ref{fig:perresidue_orban}.
+energy, $\Delta \Delta G (\beta - \alpha)$ is shown in Figure~\ref{fig:perresidue_orban}.
 
@@
 
-\subsection{The confinement method is a useful tool for structure prediction}
+\subsection*{The confinement method is a useful tool for structure prediction}
 
@@
 
-\subsubsection{The confinement method can correctly rank-order structure predictions}
+\subsubsection*{The confinement method can correctly rank-order structure predictions}
 
@@
 
+\begin{table}
+\begin{center}
+\caption{The confinement method assigns a more favorable free energy to the experimentally determined structure than to
+    computer-generated predictions. For each target, we examined as many as five predictions submitted by CASP
+    participants. We report the free energy difference between the most favorable decoy and the experimentally
+    determined structure. Positive $\Delta\Delta G (=G_{best~decoy} - G_{native})$ values indicate that the experimental structure is predicted to be
+    more favorable than any of the decoys.}
+\label{table:casp_control}
+\begin{tabular}{l l l l}\hline
+    CASP Target  & PDB Identifier & $\Delta \Delta G$ (kcal/mol) & GDT-TS \\ \hline
+     T0531       &    2KJX        &          $11.15 \pm 0.70$    &  $44$ \\ \hline
+     T0538       &    2L09        &          $-3.00 \pm 0.47$    &  $96$ \\ \hline
+     T0540       &    3MX7        &          $16.94 \pm 0.49$    &  $70$ \\ \hline
+     T0559       &    2L01        &          $2.10 \pm 0.24$     &  $94$ \\ \hline
+     T0560       &    2L02        &          $18.00 \pm 0.49$    &  $94$ \\ \hline
+     T0569       &    2KWY        &          $20.01 \pm 0.69$    &  $78$ \\ \hline
+\end{tabular}
+\end{center}
+\end{table}
 \Justin{[Arijit: Let's move the protein names, organisms, pdb ids, etc to a separate table and just use the CASP ids.]}
@@
 
-\subsubsection{The per-residue free energy is sensitive to small changes in protein conformation}
+\subsubsection*{The per-residue free energy is sensitive to small changes in protein conformation:}
 
-In section~\protect\ref{sec:orban_per_residue} we discussed how the per-residue free energy can reveal mechanistic
+Previously, we discussed how the per-residue free energy can reveal mechanistic
 detail behind the conformational preference of a chameleon sequence. In this section, our aim is to apply the same
@@
 
-\subsubsection{The confinement method occasionally produces incorrect results}
+\subsubsection*{The confinement method occasionally produces incorrect results:}
 
@@
 
-\subsubsection{What can we say about low resolution models?}
+\subsubsection*{What can we say about low resolution models?}
 
@@
 \item The thermodynamic cycle is closed by calculating the free energy between the final restrained states $A^\ast$ and
-    $B^\ast$ using normal mode analysis or quasiharmonic analysis. The free energy calculated in this way is shown as
+    $B^\ast$ using normal mode analysis \cite{Brooks1983,Case1994} or quasiharmonic analysis \cite{Karplus1981,Levy1984}. The free energy calculated in this way is shown as
     $\Delta G_{A^\ast,B^\ast}$ in Figure~\ref{fig:method}.
@@
 
+\bibitem{Mascarenhas2013}
+Mascarenhas, N. M.; Kastner, J. How maltose influences structural changes to bind to maltose-binding protein: Results from umbrella sampling simulation. Proteins. 2013, 81, 185-198.
+
+\bibitem{Kumar1992}
+Kumar, S.; Rosenberg, J.M.; Bouzida, D.; Swendsen, R.H.; Kollman, P.A. The weighted histogram analysis method for free-energy calculations on biomolecules. I. The 
+method. J Comput Chem. 1992, 13, 1011-1021.
+
 \bibitem{Mobley2007}
@@
 \bibitem{Krivov2004}
-Krivov, S.; Karplus, M. Hidden complexity of free energy surfaces for peptide (protein) folding Proc. Natl. Acad. Sci. U.S.A. 2004, 101, (41), 14766.
+Krivov, S.; Karplus, M. Hidden complexity of free energy surfaces for peptide (protein) folding Proc. Natl. Acad. Sci. U.S.A. 2004, 101, 14766.
 
-\bibitem{Alexander2007}Alexander, P.A.; He, Y.; Chen, Y.; Orban, J.
-Bryan, P.
-The design and characterization of two proteins with $88 \%$ sequence identity but differentstructure and function.
-Proc.
-Natl.
-Acad.
-Sci.
-2007, 104 (29), 11963-11968.
+\bibitem{Brooks1983}
+Brooks, B. R.; Karplus, M. Harmonic dynamics of proteins: Normal modes and fluctuations in bovine pancreatic trypsin inhibitor. Proc. Natl. Acad. Sci. U.S.A. 1983, 80, 6571-6575.
+
+\bibitem{Case1994}
+Case, D. Normal-mode analysis of protein dynamics Curr. Opin. Struct. Biol. 1994, 4, 285-290.
+
+\bibitem{Karplus1981}
+Karplus, M.; Kushick, J. Method for estimating the configurational entropy of macromolecules Macromolecules. 1981, 14, 325-332.
+
+\bibitem{Levy1984}
+Levy, R.; Karplus, M.; Kushick, J.; Perahia, D. Evaluation of the configurational entropy for proteins: application to molecular dynamics simulations of 
+an $\alpha$-helix Macromolecules 1984, 17, 1370-1374.
+
+\bibitem{Alexander2007}
+Alexander, P.A.; He, Y.; Chen, Y.; Orban, J. Bryan, P. The design and characterization of two proteins with $88 \%$ sequence identity but differentstructure and 
+function. Proc. Natl. Acad. Sci. 2007, 104, 11963-11968.
 
 \bibitem{He2008}
-He, Y.; Chen, Y.; Alexander, P.A.; Orban, J. NMR structures of two designed proteins with high sequence identity but different fold and function. Proc. Natl. Acad. Sci. 2008, 105 (38), 14412-14417.
+He, Y.; Chen, Y.; Alexander, P.A.; Orban, J. NMR structures of two designed proteins with high sequence identity but different fold and function. Proc. Natl. Acad. Sci. 2008, 105, 14412-14417.
 
@@
 Alexander, P.A.; He, Y.; Chen, Y.; Orban, J. Bryan, P. A minimal sequence code for switching protein structure and function. Proc. Natl. Acad. Sci. 2009
-, 106(50), 21149-21154.
+, 106, 21149-21154.
 
 \bibitem{Bryan2010}
-Bryan, P.N.; Orban, J. Proteins that switch folds. Curr Opin Struct Biol. 2010, 20(4), 482-488.
+Bryan, P.N.; Orban, J. Proteins that switch folds. Curr Opin Struct Biol. 2010, 20, 482-488.
 
 \bibitem{He2012}
-He, Y.; Chen, Y.; Alexander, P.A.; Bryan, P.N.; Orban, J. Mutational tipping points for switching protein folds and functions. Structure. 2012, 20(2), 2
+He, Y.; Chen, Y.; Alexander, P.A.; Bryan, P.N.; Orban, J. Mutational tipping points for switching protein folds and functions. Structure. 2012, 20,
 83-91.
@@
 \bibitem{Shortle2009}
-Shortle, D. One sequence plus one mutation equals two folds. Proc. Natl. Acad. Sci. 2009, 106(50), 21011-21012.
+Shortle, D. One sequence plus one mutation equals two folds. Proc. Natl. Acad. Sci. 2009, 106, 21011-21012.
+
+\bibitem{Allison2011}
+Allison, J.R.; Bergeler, M.; van Gunsteren, W.F. Can computer modeling explain why two highly similar sequences fold into different structures ?
+Biochemistry. 2011, 50, 10965-10973.
 

### 2013-05-03 Paper/corrected.tex

--- 2013-05-03 Paper/tt.tex
@@
 In some problems of protein science, you want to know the relative stability of a protein's conformation $A$ compared to
-its conformation $B$.  We call this the \emph{difference free energy}. For example, in all esoteric mechanisms, a protein
+its conformation $B$.  We call this the \emph{difference free energy}. For example, in all allosteric mechanisms, a protein
 often adopts one conformation when a ligand is bound and another conformation when no ligand is bound~\cite{Elber2007}.
@@
 
-\Justin{[TI is actually the basis of the confinement step. We need to be more explicit about what we mean here.] Methods
-    like thermodynamic integration~\cite{Tironi1994},\cite{Meirovitch2007} which are successful in alchemical free
-    energy calculation may not be as useful for calculation of conformational free energy.} One widely explored strategy
+One widely explored strategy
 is to use molecular dynamics simulations along some putative reaction coordinate pathway from conformation $A$ to
-$B$~\cite{Elber2007,West2007,Haas2007,Jonsson1998,E2007,Dellago2002,Cheng2006}. The free energy along this
+$B$~\cite{Elber2007,West2007,Haas2007,Jonsson1998,E2007,Dellago2002,Cheng2006,Elber2005,Chipot2007}. The free energy along this
 reaction coordinate can then be determined using methods such as umbrella sampling \cite{Torrie1977,Mascarenhas2013}  
@@
     \begin{center}
-        \includegraphics[width=3.5 in]{method.pdf}\label{fig:method}
+        \includegraphics[width=3.5 in]{method.pdf}
     \end{center}
     \caption{Graphical representation of the thermodynamic cycle involving confinement method.}
+\label{fig:method}
 \end{figure}
@@
 show that the confinement method can also be applied successfully to larger systems, such as the series of chameleon
-proteins designed by Orban and co-workers~\cite{Alexander2007,He2008,Alexander2009,He2012}, which can switch between two
+proteins designed by Orban and co-workers~\cite{Alexander2007,He2008,Alexander2009,Bryan2010,He2012}, which can switch between two
 completely different folds with only slight changes in sequence. We also show that our computed difference free energies
@@
 assigns a lower free energy to the experimentally determined structure than to any of the decoys.  Other well-known
-discriminators can also successfully tell native structures from computer generated models~\cite{Sheffler2009} [add
-rosetta holes, rosetta ff, dfire]; it is just an independent useful validation that the confinement calculations make
+discriminators can also successfully tell native structures from computer generated models~\cite{Sheffler2009,Zhou2002} 
+; it is just an independent useful validation that the confinement calculations make
 sense.
@@
 \begin{tabular}{l l l l}\hline
-    CASP Target  & PDB Identifier &  organism type  \\ \hline
-     T0531       &    2KJX        &  extracellular domain of the jumping translocation breakpoint protein &  $44$ \\ \hline
-     T0538       &    2L09        &  protein asr4154 from Nostoc sp. PCC7120        &  $96$ \\ \hline
-     T0540       &    3MX7        &  human Fas apoptotic inhibitory molecule    &  $70$ \\ \hline
+    CASP Target  & PDB Identifier & $\Delta \Delta G$ (kcal/mol) & GDT-TS \\ \hline
+     T0531       &    2KJX        &          $11.15 \pm 0.70$    &  $44$ \\ \hline
+     T0538       &    2L09        &          $-3.00 \pm 0.47$    &  $96$ \\ \hline
+     T0540       &    3MX7        &          $16.94 \pm 0.49$    &  $70$ \\ \hline
      T0559       &    2L01        &          $2.10 \pm 0.24$     &  $94$ \\ \hline
      T0560       &    2L02        &          $18.00 \pm 0.49$    &  $94$ \\ \hline
-     T0569       &    2KWY        &          $20.01 \pm 0.69$    &  $78$ \\ \hline
+     T0569       &    2KYW        &          $20.01 \pm 0.69$    &  $78$ \\ \hline
 \end{tabular}
@@
 We also tested the confinement method predictions for difference free energies on the chameleon sequences of Orban et
-al.\cite{Alexander2007,He2008,Alexander2009,Bryan2010,He2012,Shortle2009}. These are instances in which two highly
+al.\cite{Alexander2007,He2008,Alexander2009,Bryan2010,He2012}. These are instances in which two highly
 similar sequences fold into remarkably different structures. Orban and co-workers have designed a protein-G-like
@@
     \includegraphics[width=5.0in]{orban.pdf}
-        \label{fig:orban}
     \end{center}
@@
         is reported for each sequence.}
+\label{fig:orban}
 \end{figure}
@@
 \subsection*{Per-residue free energy calculations can identify the mechanistic detail behind conformational preferences}
-\label{sec:orban_per_residue}
 
@@
         \includegraphics[width=4.5in]{delta_g.pdf}
-        \label{fig:perresidue_orban}
     \end{center}
@@
         and white residues have no preference.}
+\label{fig:perresidue_orban}
 \end{figure}
@@
 These residues strongly favour the locally well packed and hydrogen bonded environment found in the $\beta$ sheet.
-Overall, favoring $\alpha$ or $\beta$ structure is a delicate balance, were the relative global free energy difference
+Overall, favoring $\alpha$ or $\beta$ structure is a delicate balance, where the relative global free energy difference
 is small and the contributions of different per residue tendencies balance out. It is therefore very likely that by
-changing key residue preferences the global fold preferences can be changed.
+changing key residue the global fold preferences can be changed.
 
@@
 missed by other meta-prediction servers. Most successful meta-prediction servers are based on the idea of consensus: if
-many different prediction methods produce similar results, then that is probably a correct prediction [ref]. This is
+many different prediction methods produce similar results, then that is probably a correct prediction \cite{Kryshtafovych2011,Wang2011}. This is
 often a powerful heuristic, but it can miss cases where there is a very good result that is only predicted by one
@@
 
-\subsubsection*{The confinement method can correctly rank-order structure predictions}
+\subsection*{The confinement method can correctly rank-order structure predictions}
 
 First, we examined the ability of the confinement method to rank different models for a target that have been generated
-by a single group using the same methodology for all predictions. We examined two targets: T0559 and T0560 (see Table
-?).
+by a single group using the same methodology for all predictions. We examined two targets: T0559 and T0560 (see Table~\ref{table:casp_testcase}).
 
@@
 \begin{center}
-\caption{The confinement method assigns a more favorable free energy to the experimentally determined structure than to
-    computer-generated predictions. For each target, we examined as many as five predictions submitted by CASP
-    participants. We report the free energy difference between the most favorable decoy and the experimentally
-    determined structure. Positive $\Delta\Delta G (=G_{best~decoy} - G_{native})$ values indicate that the experimental structure is predicted to be
-    more favorable than any of the decoys.}
-\label{table:casp_control}
-\begin{tabular}{l l l l}\hline
-    CASP Target  & PDB Identifier & $\Delta \Delta G$ (kcal/mol) & GDT-TS \\ \hline
-     T0531       &    2KJX        &          $11.15 \pm 0.70$    &  $44$ \\ \hline
-     T0538       &    2L09        &          $-3.00 \pm 0.47$    &  $96$ \\ \hline
-     T0540       &    3MX7        &          $16.94 \pm 0.49$    &  $70$ \\ \hline
-     T0559       &    2L01        &          $2.10 \pm 0.24$     &  $94$ \\ \hline
-     T0560       &    2L02        &          $18.00 \pm 0.49$    &  $94$ \\ \hline
-     T0569       &    2KWY        &          $20.01 \pm 0.69$    &  $78$ \\ \hline
+\caption{The six different CASP targets that are used in this study. The corresponding PDB Identifier and the 
+         organism type are also listed.}
+\label{table:casp_testcase}
+\begin{tabular}{l l l}\hline
+    CASP Target  & PDB Identifier &  organism type  \\ \hline
+     T0531       &    2KJX        &  extracellular domain of the jumping  \\
+                 &                &  translocation breakpoint protein  \\ \hline
+     T0538       &    2L09        &  protein asr4154 from Nostoc sp. PCC7120     \\ \hline
+     T0540       &    3MX7        &  human Fas apoptotic inhibitory molecule     \\ \hline
+     T0559       &    2L01        &  protein BVU3908 from Bacteroides vulgatus   \\ \hline
+     T0560       &    2L02        &  protein BT2368 from Bacteroides  \\ 
+                 &                &  thetaiotaomicron         \\ \hline
+     T0569       &    2KYW        &  domain of adhesion exoprotein from \\ 
+                 &                &   Pediococcus pentosaceus \\ \hline
 \end{tabular}
@@
 \end{table}
-\Justin{[Arijit: Let's move the protein names, organisms, pdb ids, etc to a separate table and just use the CASP ids.]}
 
@@
 target, the difference free energy computed by the confinement method can be used to accurately rank-order all of the
-models and the native structure (Figure~\ref{fig:T0559}).
-
-\Ken{Arijit- we could probably make a better visualization of this figure above by putting it onto a toy landscape:
-    x-axis is rmsd, y-axis is DDG, and above each of the 4 points on this graph will be the ribbon diagram and the
-    GDT-TS.  I suggest we do that with the fig below too.  Let's don't use 4 different colors for the fig below.  Let's
-    just show differences in structures in some different color, or show them all as green.]}
-
-\Justin{Ken: I don't mind the 3d plots that are in here now, but I'm not sure I would like something that looks like a
-    free energy landscape.}
+models and the native experimental structure (Figure~\ref{fig:T0559}).
 
@@
         \includegraphics[width=3.8 in]{T0559.pdf}
-        \label{fig:T0559}
     \end{center}
     \caption{The confinement method can correctly rank the native structure and three predictions produced by a single
-        method.}
+        method for target T0559.}
+\label{fig:T0559}
 \end{figure}
 
-\Justin{I'm not sure we need a second figure of this. Move it to the SI?}
-
-We performed a similar calculation for target T0560 with two models form the group ``Splicer''. The remaining three
+We performed a similar calculation for target T0560 with two models from the group ``Splicer''. The remaining three
 models were discarded as they were too similar to the rest of the models. Again, we can correctly identify the native
-sate and our calculated free energy based ranking matches well with GDT-TS (Figures S?).
-
-\begin{figure}
-    \begin{center}
-        \includegraphics[width=3.5 in]{T0560.pdf}
-    \end{center}
-    \caption{Native structure and two models (from group ``Splicer'') for target T0560.}
-    \label{fig:T0560}
-\end{figure}
+sate and our calculated free energy based ranking matches well with GDT-TS (Figures S2 in the supporting information).
 
@@
 
-\subsubsection*{The per-residue free energy is sensitive to small changes in protein conformation:}
+\subsection*{The per-residue free energy is sensitive to small changes in protein conformation:}
 
-Previously, we discussed how the per-residue free energy can reveal mechanistic
+In a previous section, we discussed how the per-residue free energy can reveal mechanistic
 detail behind the conformational preference of a chameleon sequence. In this section, our aim is to apply the same
@@
 localized regions. We chose CASP target T0569 and compared the experimental NMR structure with the best predicted model
-(GDT-TS=78; predicted by the ``Mufold'' group). The confinement method predicts that the native structure is more stable
+(GDT-TS=78; predicted by the ``Mufold'' group). The confinement method predicts that the experimental structure is more stable
 by 20 kcal/mol (Figure S2).
@@
 these hydrophobic residues are oriented towards the protein hydrophobic core in the native NMR structure but oriented
-towards the exterior of the protein and are solvent exposed in the model. These residues also form part of a beta-sheet
-in the experimental structure, but do not form inter-strand hydrogen bonds in the predicted model. There is also a large
+towards the exterior of the protein and are solvent exposed in the model. These residues are part of a beta-sheet
+in the experimental structure, but because of their sidechain orientation the corresponding beta-sheet become disordered in the predicted model 
+(Figure~\ref{fig:T0569_per_residue} and Figure S4). There is also a large
 difference around K76, which forms a salt-bridge with D11 in the predicted model, but not in the experimental structure.
 This suggests that salt-bridge interactions are too favorable for the combination of force field and implicit solvent
-model we use, which has been a problem noted in the past [ref carlos].
-
-\Justin{[Arijit: Figure~\ref{fig:T0569_per_residue} says crystal, but the text says NMR.]}
+model we use, which has been a problem noted in the past \cite{Roe2007}.
 
@@
 
-\subsubsection*{The confinement method occasionally produces incorrect results:}
-
-\Justin{Arijit: let's move Figure~\ref{fig:T0538} to the supporting info and move Figure S4 into the main text.}
+\subsection*{The confinement method occasionally produces incorrect results:}
 
 Despite success for most of the studied systems, there are few failures, specially when the GDT-TS scores of the
-compared structures are very close. One example is Target T0538, where we compared the crystal structure with three
-models (Model 1: ``PconsR''---GDT-ts=96; Model 2: ``Shell''---GDT-TS=90; Model 3: ``FOLDIT''---GDT-TS=86). Contrary to
-our expectation, the confinement method predicts that Model 1 is more stable than the crystal structure (Figure ?).
+compared structures are very close. One example is Target T0538, where we compared the experimental structure with three
+models (Model 1: ``PconsR''---GDT-TS=96; Model 2: ``Shell''---GDT-TS=90; Model 3: ``FOLDIT''---GDT-TS=86). Contrary to
+our expectation, the confinement method predicts that computer generated Model 1 is more stable than the crystal structure (Figure S5).
 Per-residue free energy calculations (not shown) show that despite only small variations at the backbone level, the
-sidechains are oriented in very different ways (Figure ?), giving rise to large differences in the stabilization of
+sidechains are oriented in very different ways (Figure~\ref{fig:T0538compare}), giving rise to large differences in the stabilization of
 certain residues. In particular, some of the differences arise from different salt bridge patterns and certain flexible
@@
 \begin{center}
-\includegraphics[width=3.5in]{T0538.pdf}
+\includegraphics[width=4.9 in,height=4.0 in]{Target_538_compare.pdf}
 \end{center}
-\caption{The native and three model structure of engineered protein from Asr4154 protein (PDB ID: 2L09 and CASP
-    code:T0538). The model 1,2 and 3 are from the group PconsR, Shell and FOLDIT respectively.}
-\label{fig:T0538}
+\caption{The per residue free energy calculation found some differences of the sidechain orientation in computer generated model 1 and 
+experimental structure of CASP target T0538. 
+(A) Two salt bridges between residues E35-R32 and E28-K24 exist in the native structure. These are
+absent in the computer generated model structure. Instead
+(B) a new salt bridge is formed between residues R32-E28 in model. (C) Another salt bridge between
+residues  R26-E50 in the experimental structure. (D) In comparison to the native 
+structure, orientation of F51 differs in the model.}
+\label{fig:T0538compare}
 \end{figure}
 
-\subsubsection*{What can we say about low resolution models?}
-
-\Justin{I think we should consider dropping this whole section. We have a sample size of one and I don't expect that we
-    can generally do well ranking models with a GDT of 44. I think this is a fluke.}
-
-So far we have seen that the method is good at predicting preferences when the structures are not very far from the
-native. But the question remains how far from native can we go and still see that the method produces correct result. In
-this section we explored this question with models from  extracellular domain of the jumping translocation breakpoint
-protein (pdb id: 2KJX). Most of the group could only generate low resolution models for this CASP9 target (id: T0531) .
-In our comparison, we choose five models by the group MUFOLD-MD, which was the best performing group for this target
-with their best model had a GDT\_TS value of 44. The result presented in Figure~\ref{fig:T0531} shows: 1.) native is
-correctly identified as expected and 2.) Surprisingly there is a high level of correlation between the GDT ranking and
-the free energy ranking for model 1 and model 3, the rest three structures with GDT\_TS score less than 35  are ordered
-incorrectly. It is worth to note that models 2 and 3 have the same GDT and very different free energies, meaning that
-the actual ordering could change a lot \cite{Perez2012}. It is encouraging that at least the method can pick out the
-best model even though it is got a low GDT score: 44.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.5 in]{T0531.pdf}
-\end{center}
-\caption{The native structure and 5 models of extracellular domain of the jumping translocation breakpoint protein (pdb
-    id: 2KJX and the CASP code: T0531).}
-\label{fig:T0531}
-\end{figure}
-
-\subsubsection{Can the confinement method perform better quality assessment in protein structure prediction?}
-
-\Justin{Arijit---is this the case where we chose a structure that was missed by the consensus predictors, if so, we
-    should try to indicate this. The point of this section should be that although consensus prediction strategies do
-    well, they can miss a good answer that is only predicted by a few methods. At least in this case, the confinement
-    method is able to capture this structure that was otherwise missed.}
+\subsection*{Can the confinement method perform better quality assessment in protein structure prediction?}
 
@@
 Here, predictors were asked to score each model on a scale (known as qmode) from 0 to 1, with higher values
-corresponding to better models \cite{Kryshtafovych2011}. It will be interesting to know how confinement method perform
-in quality assessment compare to the other groups in CASP9. We investigated this using couple of CASP Targets. Here, we
-present a case, where confinement method perform well than the top performing group MUFOLD-WQA \cite{Wang2011} of CASP9.
-We choose  two models from CASP taget T0538. They were model 3 submitted by PconsR (GDT\_TS = 96, qmode 1 = 0.5434) and
-model 5 from the MULTICOM-NOVEL ( GDT\_TS = 83, qmode 1 = 0.5865). Both are server predicted models and the qmode value
-presented are from MUFOLD-WQA \cite{Wang2011}. We choose these two models as the model by PconsR was the most accurate
+corresponding to better models \cite{Kryshtafovych2011}. Most of the groups use consensus strategies in such experiments.
+Here, we chose two computer generated models from CASP taget T0538, where the top performing group "MUFOLD-WQA" \cite{Wang2011} failed.
+One of the model was from "PconsR" (GDT-TS = 83, qmode 1 = 0.5865) and another from "MULTICOM-NOVEL" ( GDT-TS = 83, qmode 1 = 0.5865). We 
+choose these two models as the model by PconsR was the most accurate
 predicted model for this target, whereas the other model was predicted best by QA test of MUFOLD-WQA. The calculated
 free energy using confinement method indicate that the model by PconsR is more stable by $3.9 Kcal/Mol$ which support
-the GDT\_TS trend. There is no doubt that the consensus approach can predict the model quality in a faster manner. But
-we expect confinement method can predict the model quality in a relatively expensive but much more accurate way.
+the GDT\_TS trend.  Although consensus prediction strategies do well specially when dealing with large numbers of structures, they 
+can miss a good answer that is only predicted by a few methods. At least in this case, the confinement
+method is able to capture this structure that was otherwise missed.
 
-\section{Conclusion}
+\section*{Conclusion}
 
@@
 A key advantage is that this method does not require any reaction coordinate, or sampling a
-pathway from conformation $A$ to $B$. We have tested the method for structures of proteins having up to 100 amino acids.
-The computational cost for a 56-residue protein is about 4 hours [on 1 gpu] for 20 ns of confinement [Is 20ns all you
-need?].
-
-\Arijit{I need 21 (confinement run) x 20 ns for single confinement}
-%There is another issue, which is normal mode analysis calculations are Oh(N**3) and require large amounts of memory,
-%effectively limiting the size of the systems under study. In most cases quasiharmonic analysis.
-
-\Ken{[If this is a relevant limitation, let's say something about it.]}
-
-\Arijit{I think we need to write the normal mode issue and number of residues that we can handle. If one of you
-    contribute here that will be great.}
-
-\Justin{The conclusion is not the place to talk about how long the method takes or what the technical challenges are.
-    Let's move this to methods section.}
+pathway from conformation $A$ to $B$.
 
@@
 \item  In the first step, a minimization of A and B conformations are performed. These minimized conformations
-    ($A^{\ast\ast}$ and $B^{\ast\ast}$) are the reference conformation of that state.
+    ($A^{\ast}$ and $B^{\ast}$) are the reference conformation of that state.
 
 \item The free energy of confining the ensemble (A or B) to a tighter ensemble ($A^\ast$ or $B^\ast$) is calculated.
-    This is done by gradually confining each atom to its position in the reference conformation ($A^{\ast\ast}$ or
-    $B^{\ast\ast}$) using a series of progressively stronger position restraints. This is done by running 21 molecular
+    This is done by gradually confining each atom to its position in the reference conformation ($A^{\ast}$ or
+    $B^{\ast}$) using a series of progressively stronger position restraints. This is done by running 21 molecular
     dynamics simulation (each 20 ns long) for each leg of the thermodynamic cycle, where the harmonic restraint force
@@
 
-All calculations where performed with the Amber 11 suit of programs~\cite{Case2012,Goetz2012} in combination with ff99SB
-forcefield~\cite{Hornak2006} and generalized born implicit solvent \cite{Mongan2006}. Interestingly, we extend the
+One advantage of confinement method is that
+none of the simulation during restraining step depends on each other. Therefore, with available computer resources it can be fast 
+to compute. The computational cost for a $56$-residue proteins with chameleon sequence that is reported in this article is only $4$ hours 
+in a single gpu computer. All calculations were performed with the Amber 11 suit of programs~\cite{Case2012,Goetz2012} in combination with ff99SB
+forcefield~\cite{Hornak2006} and generalized born implicit solvent \cite{Mongan2006}.
+
+Interestingly, we extend the
 method for calculation of per residue free energy in an approximate way. For this purpose, the confinement energy,
@@
 the mechanistic details of conformational preference of each residue.
-
-%There are two big bottlenecks in this calculation. The first has to do with the amount of simulations that have to be
-%performed to confine the ensemble of conformations close to a microstate into a single microstate, the second
-%bottleneck has to do with the size of the system in the normal mode calculation. We have tackled the first issue by
-%using GPU (Graphical Processing Unit) technology as opposed to the classical CPUs, which gives us a two order of
-%magnitude boost in computational efficiency. Normal mode analysis calculations are Oh(N**3) and require large amounts
-%of memory, effectively limiting the size of the systems under study. Where as this can be overcome by larger memory
-%machines, these are often not available, and so we propose an alternative by using quasiharmonic analysis. This is
-%still an Oh(N**3) problem, but we now the motions of atoms are already determined in trajectories, and so we can choose
-%to do quasiharmonic analysis excluding hydrogens, which will reduce the number of atoms N to almost N/2. This means
-%that we can use this method for systems twice as large.
 
@@
 \bibitem{Jonsson1998}
-Jónsson, H.; Mills, G.; Jacobsen, K.W. Nudged Elastic Band Method for Finding Minimum Energy Paths of Transitions,
+Jonsson, H.; Mills, G.; Jacobsen, K.W. Nudged Elastic Band Method for Finding Minimum Energy Paths of Transitions,
 in Classical and Quantum Dynamics in Condensed Phase Simulations, Ed. B. J. Berne, G. Ciccotti and D. F.
@@
 
+\bibitem{Chipot2007}
+Chipot, C.; Shell, M.S.; Pohorille, A. Introduction, in Chipot, C., Pohorille, A., editors. Free Energy
+Calculations: Theory and Applications in Chemistry and Biology. Springer Series in Chemical
+Physics, vol. 86. Berlin and Heidelberg: Springer; 2007, p. 1–32.
+
 \bibitem{Torrie1977}
@@
 (1977) J. Comput. Phys. 23, 187
-
-\bibitem{Tironi1994}
-Tironi, I.G.; van Gunsteren, W.F. A molecular-dynamics simulation study of chloroform. Mol. Phys. 1994, 83, 381-403.
-
-\bibitem{Meirovitch2007}
-Meirovitch, H. Recent developments in methodologies for calculating the entropy and free energy of biological systems by computer simulation.
-Current Opinion in Structural Biology, 2007, 17, 181-186.
 
@@
 method. J Comput Chem. 1992, 13, 1011-1021.
-
-\bibitem{Mobley2007}
-Mobley, D.L.; Chodera, J.D.; Dill, K.A. The combining and release method: obtaining correct binding free energies in
-the presence of protein conformational change. Journal of Chemical Theory and Computation 2007, 3, 1231-1235.
-
-\bibitem{Mobley2012}
-Mobley, D.L.; Klimovich, P.V. Perspective: Alchemical free energy calculations for drug discovery.
-J. Chem. Phys. 2012, 137, 230901-12.
-
-\bibitem{Mobley2006}
-Mobley, D.L.; Chodera, J.D.; Dill, K.A. On the use of orientational restraints and symmetry corrections in alchemical
-free energy calculations. J. Chem. Phy. 2006, 125, 084902.
-
-\bibitem{Chipot2007}
-Chipot, C.; Shell, M.S.; Pohorille, A. Introduction, in Chipot, C., Pohorille, A., editors. Free Energy
-Calculations: Theory and Applications in Chemistry and Biology. Springer Series in Chemical
-Physics, vol. 86. Berlin and Heidelberg: Springer; 2007, p. 1–32.
-
-\bibitem{Jorgensen2004}
-Jorgensen, W.L. The many roles of computation in drug discovery, Science 2004, 303, 1813–8.
-
-\bibitem{Gilson2007}
-Gilson, M.K.; Zhou, H.X. Calculation of protein-ligand binding affinities. Annu Rev Biophys Biomol Struct. (2007) 36, 21-42.
-
-\bibitem{Handl2009}
-Handl, J.; Knowles, J.; Lovel, S.C. Artefacts and biases affecting the evaluation of scoring
-functions on decoy sets for protein structure prediction. Bioinformatics, 2009, 25, 1271-1279.
-
-\bibitem{Dill1997}
-Dill, K.A.; H.S. Chan.  From Levinthal to Pathways to Funnels:  The "New View" of Protein Folding Kinetics.  Nature Structural Biology 4, 10-19 (1997)
-
-\bibitem{Dill2008}
-Dill, K.A.; Ozkan, S.B.; Shell, M.S.; Weikl, T.R. The protein folding problem. Annual Review of Biophysics (2008), 37, 289-316.
-
-\bibitem{Anfinsen1973}
-Anfinsen. C.B. Principles that Govern the Folding of Protein Chains. Science (1973) 181, 223-230.
-
-\bibitem{Christ2007}
-Christ, C.D.; van Gunsteren, W.F. Enveloping distribution sampling: A method to calculate free energy differences from a single simulation,
-J. Chem. Phys. (2007), 126, 184110.
 
@@
 
-\bibitem{Krivov2004}
-Krivov, S.; Karplus, M. Hidden complexity of free energy surfaces for peptide (protein) folding Proc. Natl. Acad. Sci. U.S.A. 2004, 101, 14766.
-
 \bibitem{Brooks1983}
@@
 an $\alpha$-helix Macromolecules 1984, 17, 1370-1374.
+
+\bibitem{Mobley2007}
+Mobley, D.L.; Chodera, J.D.; Dill, K.A. The combining and release method: obtaining correct binding free energies in
+the presence of protein conformational change. Journal of Chemical Theory and Computation 2007, 3, 1231-1235.
+
+\bibitem{Mobley2012}
+Mobley, D.L.; Klimovich, P.V. Perspective: Alchemical free energy calculations for drug discovery.
+J. Chem. Phys. 2012, 137, 230901-12.
+
+\bibitem{Mobley2006}
+Mobley, D.L.; Chodera, J.D.; Dill, K.A. On the use of orientational restraints and symmetry corrections in alchemical
+free energy calculations. J. Chem. Phy. 2006, 125, 084902.
 
@@
 
-\bibitem{Shortle2009}
-Shortle, D. One sequence plus one mutation equals two folds. Proc. Natl. Acad. Sci. 2009, 106, 21011-21012.
+\bibitem{Sheffler2009}
+Sheffler, W.; Baker, D. RosettaHoles: Rapid assessment of protein core packing for structure prediction, refinement, design, and validation. Protein Sci
+ence. 2009, 18(1), 229-239.
+
+\bibitem{Zhou2002}
+Zhou H, Zhou Y. Distance-scaled, finite ideal-gas reference state improves structure-derived potentials of mean force for structure selection and
+stability prediction. Protein Sci. 2002, 11, 2714-2726.
 
@@
 
-\bibitem{Sheffler2009}
-Sheffler, W.; Baker, D. RosettaHoles: Rapid assessment of protein core packing for structure prediction, refinement, design, and validation. Protein Sci
-ence. 2009, 18(1), 229-239.
+\bibitem{Handl2009}
+Handl, J.; Knowles, J.; Lovel, S.C. Artefacts and biases affecting the evaluation of scoring
+functions on decoy sets for protein structure prediction. Bioinformatics, 2009, 25, 1271-1279.
+
+\bibitem{Kryshtafovych2011}
+Kryshtafovych, A.; Fidelis, K; and Tramontano, A. Evaluation of model quality predictions in CASP9. Proteins, 2011, 79, 91–106
+
+\bibitem{Wang2011}
+Wang, Q.; Vantasin, K.; Xu, D.; Shang, Y. MUFOLD-WQA: A new selective consensus method for quality assessment in protein structure prediction.
+Proteins, 2011, 79: 185-195.
+
+\bibitem{Zemla2003}
+Zemla, A. LGA: a method for finding 3D similarities in protein structures. Nucleic Acids Res 2003, 31, 3370–3374.
 
@@
 
-\bibitem{MacCallum2011}
-MacCallum, J.; Perez, A.; Schnieders, MJ.; Hua, L.; Jacobson, M.P.; Dill, K.A. Assessment of protein structure refinement
-in CASP9. Proteins, 2011, 79, 74-90.
-
-\bibitem{Kryshtafovych2011}
-Kryshtafovych, A.; Fidelis, K; and Tramontano, A. Evaluation of model quality predictions in CASP9. Proteins, 2011, 79, 91–106
-
-\bibitem{Zemla2003}
-Zemla, A. LGA: a method for finding 3D similarities in protein structures. Nucleic Acids Res 2003, 31, 3370–3374.
-
-\bibitem{Perez2012}
-Perez, A.; Yang, Z.; Bahar, I.; Dill, K.A.; MacCallum, J.L.; FlexE: Using Elastic Network Models to Compare Models of Protein Structure. J. Chem. Theory Comput., 2012, 8, 3985-3991.
-
-\bibitem{Wang2011}
-Wang, Q.; Vantasin, K.; Xu, D.; Shang, Y. MUFOLD-WQA: A new selective consensus method for quality assessment in protein structure prediction.
-Proteins, 2011, 79: 185-195.
+\bibitem{Roe2007}
+Roe, D.R.; Okur, A.; Wickstrom, L.; Hornak, V.; Simmerling, C. Secondary structure bias in generalized Born solvent models: comparison of conformational ensembles 
+and free energy of solvent polarization from explicit and implicit solvation. J Phys Chem B. 2007, 111, 1846-1857.
 

### 2013-05-03 Paper/paper.tex

@@
 
-\Ken{[Guys---I recommend we get a better name than confinement.  It's totally uninformative.  In my view, better would
-    be `confine-and-configure' or something.  Please give it some thought.]}
-
-\Justin{[Ken---I agree that the name isn't great, but I don't think we should change it because: it's already in the
-    literature and we didn't come up with the method. I think we should consider capitalizing it as Confinement Method
-    throughout.]}
-
-\section{Introduction}
+\section*{Introduction}
 
 In some problems of protein science, you want to know the relative stability of a protein's conformation $A$ compared to
-its conformation $B$.  We call this the \emph{difference free energy}. For example, in all esoteric mechanisms, a protein
+its conformation $B$.  We call this the \emph{difference free energy}. For example, in all allosteric mechanisms, a protein
 often adopts one conformation when a ligand is bound and another conformation when no ligand is bound~\cite{Elber2007}.
@@
 
-\Justin{[TI is actually the basis of the confinement step. We need to be more explicit about what we mean here.] Methods
-    like thermodynamic integration~\cite{Tironi1994},\cite{Meirovitch2007} which are successful in alchemical free
-    energy calculation may not be as useful for calculation of conformational free energy.} One widely explored strategy
+One widely explored strategy
 is to use molecular dynamics simulations along some putative reaction coordinate pathway from conformation $A$ to
-$B$~\cite{Elber2007,West2007,Haas2007,Jonsson1998,E2007,Dellago2002,Cheng2006}. \Ken{[we need many refs here: cite
-    Elber's milestoning, Jhih-Wei Chu, and the many other path-sampling approaches.]} The free energy along this
-reaction coordinate can then be determined using methods such as umbrella sampling \cite{Torrie1977}\Justin{[add ref
-    where this is used computing free energy differences between conformations]} or targeted MD \Justin{[refs]}. Such
+$B$~\cite{Elber2007,West2007,Haas2007,Jonsson1998,E2007,Dellago2002,Cheng2006,Elber2005,Chipot2007}. The free energy along this
+reaction coordinate can then be determined using methods such as umbrella sampling \cite{Torrie1977,Mascarenhas2013}  
+and weighted histogram analysis method (WHAM) \cite{Kumar1992}. Such
 approaches have several limitations. First, it is necessary to know an efficient reaction pathway from $A$ to $B$. If
@@
 assigns a lower free energy to the experimentally determined structure than to any of the decoys.  Other well-known
-discriminators can also successfully tell native structures from computer generated models~\cite{Sheffler2009} [add
-rosetta holes, rosetta ff, dfire]; it is just an independent useful validation that the confinement calculations make
+discriminators can also successfully tell native structures from computer generated models~\cite{Sheffler2009,Zhou2002} 
+; it is just an independent useful validation that the confinement calculations make
 sense.
-
-\Justin{[Arijit, please add GDT of best decoy to this table.]}
 
@@
     participants. We report the free energy difference between the most favorable decoy and the experimentally
-    determined structure. Positive $\Delta\Delta G$ values indicate that the experimental structure is predicted to be
+    determined structure. Positive $\Delta\Delta G (=G_{best~decoy} - G_{native})$ values indicate that the experimental structure is predicted to be
     more favorable than any of the decoys.}
 \label{table:casp_control}
-\begin{tabular}{l l l}\hline
-    CASP Target  & PDB Identifier & $\Delta \Delta G = G_{best~decoy} - G_{native}$ (kcal/mol) \\ \hline
-     T0531       &    2KJX        &          $11.15 \pm 0.70$ \\ \hline
-     T0538       &    2L09        &          $-3.00 \pm 0.47$ \\ \hline
-     T0540       &    3MX7        &          $16.94 \pm 0.49$ \\ \hline
-     T0559       &    2L01        &          $2.10 \pm 0.24$ \\ \hline
-     T0560       &    2L02        &          $18.00 \pm 0.49$ \\ \hline
-     T0569       &    2KWY        &          $20.01 \pm 0.69$  \\ \hline
+\begin{tabular}{l l l l}\hline
+    CASP Target  & PDB Identifier & $\Delta \Delta G$ (kcal/mol) & GDT-TS \\ \hline
+     T0531       &    2KJX        &          $11.15 \pm 0.70$    &  $44$ \\ \hline
+     T0538       &    2L09        &          $-3.00 \pm 0.47$    &  $96$ \\ \hline
+     T0540       &    3MX7        &          $16.94 \pm 0.49$    &  $70$ \\ \hline
+     T0559       &    2L01        &          $2.10 \pm 0.24$     &  $94$ \\ \hline
+     T0560       &    2L02        &          $18.00 \pm 0.49$    &  $94$ \\ \hline
+     T0569       &    2KYW        &          $20.01 \pm 0.69$    &  $78$ \\ \hline
 \end{tabular}
@@
 
-\subsection{The confinement method correctly predicts the structures of chameleon sequences}
+\subsection*{The confinement method correctly predicts the structures of chameleon sequences}
 
 We also tested the confinement method predictions for difference free energies on the chameleon sequences of Orban et
-al.!\cite{Alexander2007,He2008,Alexander2009,Bryan2010,He2012,Shortle2009}. These are instances in which two highly
+al.\cite{Alexander2007,He2008,Alexander2009,Bryan2010,He2012}. These are instances in which two highly
 similar sequences fold into remarkably different structures. Orban and co-workers have designed a protein-G-like
@@
 
-\subsection{Per-residue free energy calculations can identify the mechanistic detail behind conformational preferences}
-\label{sec:orban_per_residue}
+\subsection*{Per-residue free energy calculations can identify the mechanistic detail behind conformational preferences}
 
@@
 
-\subsubsection{The confinement method can correctly rank-order structure predictions}
+\subsection*{The confinement method can correctly rank-order structure predictions}
 
 First, we examined the ability of the confinement method to rank different models for a target that have been generated
-by a single group using the same methodology for all predictions. We examined two targets: T0559 and T0560 (see Table
-?).
+by a single group using the same methodology for all predictions. We examined two targets: T0559 and T0560 (see Table~\ref{table:casp_testcase}).
 
-\Justin{[Arijit: Let's move the protein names, organisms, pdb ids, etc to a separate table and just use the CASP ids.]}
+\begin{table}
+\begin{center}
+\caption{The six different CASP targets that are used in this study. The corresponding PDB Identifier and the 
+         organism type are also listed.}
+\label{table:casp_testcase}
+\begin{tabular}{l l l}\hline
+    CASP Target  & PDB Identifier &  organism type  \\ \hline
+     T0531       &    2KJX        &  extracellular domain of the jumping  \\
+                 &                &  translocation breakpoint protein  \\ \hline
+     T0538       &    2L09        &  protein asr4154 from Nostoc sp. PCC7120     \\ \hline
+     T0540       &    3MX7        &  human Fas apoptotic inhibitory molecule     \\ \hline
+     T0559       &    2L01        &  protein BVU3908 from Bacteroides vulgatus   \\ \hline
+     T0560       &    2L02        &  protein BT2368 from Bacteroides  \\ 
+                 &                &  thetaiotaomicron         \\ \hline
+     T0569       &    2KYW        &  domain of adhesion exoprotein from \\ 
+                 &                &   Pediococcus pentosaceus \\ \hline
+\end{tabular}
+\end{center}
+\end{table}
 
@@
 
-\subsubsection{The per-residue free energy is sensitive to small changes in protein conformation}
+\subsection*{The per-residue free energy is sensitive to small changes in protein conformation:}
 
-In section~\protect\ref{sec:orban_per_residue} we discussed how the per-residue free energy can reveal mechanistic
+In a previous section, we discussed how the per-residue free energy can reveal mechanistic
 detail behind the conformational preference of a chameleon sequence. In this section, our aim is to apply the same
@@
 The confinement method clearly identifies two hydrophobic residues V59 and I61, which destabilize the predicted model
-with respect to the crystallographic structure (Figure~\ref{fig:T0569_per_residue} and Figure S3). The sidechains of
+with respect to the experimental structure (Figure~\ref{fig:T0569_per_residue} and Figure S3). The sidechains of
 these hydrophobic residues are oriented towards the protein hydrophobic core in the native NMR structure but oriented
-towards the exterior of the protein and are solvent exposed in the model. These residues also form part of a beta-sheet
-in the experimental structure, but do not form inter-strand hydrogen bonds in the predicted model. There is also a large
+towards the exterior of the protein and are solvent exposed in the model. These residues are part of a beta-sheet
+in the experimental structure, but because of their sidechain orientation the corresponding beta-sheet become disordered in the predicted model 
+(Figure~\ref{fig:T0569_per_residue} and Figure S4). There is also a large
 difference around K76, which forms a salt-bridge with D11 in the predicted model, but not in the experimental structure.
 This suggests that salt-bridge interactions are too favorable for the combination of force field and implicit solvent
-model we use, which has been a problem noted in the past [ref carlos].
-
-\Justin{[Arijit: Figure~\ref{fig:T0569_per_residue} says crystal, but the text says NMR.]}
+model we use, which has been a problem noted in the past \cite{Roe2007}.
 
@@
 
-\subsubsection{The confinement method occasionally produces incorrect results}
-
-\Justin{Arijit: let's move Figure~\ref{fig:T0538} to the supporting info and move Figure S4 into the main text.}
+\subsection*{The confinement method occasionally produces incorrect results:}
 
 Despite success for most of the studied systems, there are few failures, specially when the GDT-TS scores of the
-compared structures are very close. One example is Target T0538, where we compared the crystal structure with three
-models (Model 1: ``PconsR''---GDT-ts=96; Model 2: ``Shell''---GDT-TS=90; Model 3: ``FOLDIT''---GDT-TS=86). Contrary to
-our expectation, the confinement method predicts that Model 1 is more stable than the crystal structure (Figure ?).
+compared structures are very close. One example is Target T0538, where we compared the experimental structure with three
+models (Model 1: ``PconsR''---GDT-TS=96; Model 2: ``Shell''---GDT-TS=90; Model 3: ``FOLDIT''---GDT-TS=86). Contrary to
+our expectation, the confinement method predicts that computer generated Model 1 is more stable than the crystal structure (Figure S5).
 Per-residue free energy calculations (not shown) show that despite only small variations at the backbone level, the
-sidechains are oriented in very different ways (Figure ?), giving rise to large differences in the stabilization of
+sidechains are oriented in very different ways (Figure~\ref{fig:T0538compare}), giving rise to large differences in the stabilization of
 certain residues. In particular, some of the differences arise from different salt bridge patterns and certain flexible
@@
 \begin{center}
-\includegraphics[width=3.5in]{T0538.pdf}
+\includegraphics[width=4.9 in,height=4.0 in]{Target_538_compare.pdf}
 \end{center}
-\caption{The native and three model structure of engineered protein from Asr4154 protein (PDB ID: 2L09 and CASP
-    code:T0538). The model 1,2 and 3 are from the group PconsR, Shell and FOLDIT respectively.}
-\label{fig:T0538}
+\caption{The per residue free energy calculation found some differences of the sidechain orientation in computer generated model 1 and 
+experimental structure of CASP target T0538. 
+(A) Two salt bridges between residues E35-R32 and E28-K24 exist in the native structure. These are
+absent in the computer generated model structure. Instead
+(B) a new salt bridge is formed between residues R32-E28 in model. (C) Another salt bridge between
+residues  R26-E50 in the experimental structure. (D) In comparison to the native 
+structure, orientation of F51 differs in the model.}
+\label{fig:T0538compare}
 \end{figure}
 
-\subsubsection{What can we say about low resolution models?}
-
-\Justin{I think we should consider dropping this whole section. We have a sample size of one and I don't expect that we
-    can generally do well ranking models with a GDT of 44. I think this is a fluke.}
-
-So far we have seen that the method is good at predicting preferences when the structures are not very far from the
-native. But the question remains how far from native can we go and still see that the method produces correct result. In
-this section we explored this question with models from  extracellular domain of the jumping translocation breakpoint
-protein (pdb id: 2KJX). Most of the group could only generate low resolution models for this CASP9 target (id: T0531) .
-In our comparison, we choose five models by the group MUFOLD-MD, which was the best performing group for this target
-with their best model had a GDT\_TS value of 44. The result presented in Figure~\ref{fig:T0531} shows: 1.) native is
-correctly identified as expected and 2.) Surprisingly there is a high level of correlation between the GDT ranking and
-the free energy ranking for model 1 and model 3, the rest three structures with GDT\_TS score less than 35  are ordered
-incorrectly. It is worth to note that models 2 and 3 have the same GDT and very different free energies, meaning that
-the actual ordering could change a lot \cite{Perez2012}. It is encouraging that at least the method can pick out the
-best model even though it is got a low GDT score: 44.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.5 in]{T0531.pdf}
-\end{center}
-\caption{The native structure and 5 models of extracellular domain of the jumping translocation breakpoint protein (pdb
-    id: 2KJX and the CASP code: T0531).}
-\label{fig:T0531}
-\end{figure}
-
-\subsubsection{Can the confinement method perform better quality assessment in protein structure prediction?}
-
-\Justin{Arijit---is this the case where we chose a structure that was missed by the consensus predictors, if so, we
-    should try to indicate this. The point of this section should be that although consensus prediction strategies do
-    well, they can miss a good answer that is only predicted by a few methods. At least in this case, the confinement
-    method is able to capture this structure that was otherwise missed.}
+\subsection*{Can the confinement method perform better quality assessment in protein structure prediction?}
 
@@
 
-\bibitem{Torrie1977}
-Torrie, G. M.; Valleau, J. P. Nonphysical sampling distributions in Monte Carlo free-energy estimation: Umbrella sampling
-(1977) J. Comput. Phys. 23, 187
-
-\bibitem{Tironi1994}
-Tironi, I.G.; van Gunsteren, W.F. A molecular-dynamics simulation study of chloroform. Mol. Phys. 1994, 83, 381-403.
-
-\bibitem{Meirovitch2007}
-Meirovitch, H. Recent developments in methodologies for calculating the entropy and free energy of biological systems by computer simulation.
-Current Opinion in Structural Biology, 2007, 17, 181-186.
-
-\bibitem{Mobley2007}
-Mobley, D.L.; Chodera, J.D.; Dill, K.A. The combining and release method: obtaining correct binding free energies in
-the presence of protein conformational change. Journal of Chemical Theory and Computation 2007, 3, 1231-1235.
-
-\bibitem{Mobley2012}
-Mobley, D.L.; Klimovich, P.V. Perspective: Alchemical free energy calculations for drug discovery.
-J. Chem. Phys. 2012, 137, 230901-12.
-
-\bibitem{Mobley2006}
-Mobley, D.L.; Chodera, J.D.; Dill, K.A. On the use of orientational restraints and symmetry corrections in alchemical
-free energy calculations. J. Chem. Phy. 2006, 125, 084902.
-
 \bibitem{Chipot2007}
@@
 
-\bibitem{Jorgensen2004}
-Jorgensen, W.L. The many roles of computation in drug discovery, Science 2004, 303, 1813–8.
+\bibitem{Torrie1977}
+Torrie, G. M.; Valleau, J. P. Nonphysical sampling distributions in Monte Carlo free-energy estimation: Umbrella sampling
+(1977) J. Comput. Phys. 23, 187
 
-\bibitem{Gilson2007}
-Gilson, M.K.; Zhou, H.X. Calculation of protein-ligand binding affinities. Annu Rev Biophys Biomol Struct. (2007) 36, 21-42.
+\bibitem{Mascarenhas2013}
+Mascarenhas, N. M.; Kastner, J. How maltose influences structural changes to bind to maltose-binding protein: Results from umbrella sampling simulation. Proteins. 2013, 81, 185-198.
 
-\bibitem{Handl2009}
-Handl, J.; Knowles, J.; Lovel, S.C. Artefacts and biases affecting the evaluation of scoring
-functions on decoy sets for protein structure prediction. Bioinformatics, 2009, 25, 1271-1279.
-
-\bibitem{Dill1997}
-Dill, K.A.; H.S. Chan.  From Levinthal to Pathways to Funnels:  The "New View" of Protein Folding Kinetics.  Nature Structural Biology 4, 10-19 (1997)
-
-\bibitem{Dill2008}
-Dill, K.A.; Ozkan, S.B.; Shell, M.S.; Weikl, T.R. The protein folding problem. Annual Review of Biophysics (2008), 37, 289-316.
-
-\bibitem{Anfinsen1973}
-Anfinsen. C.B. Principles that Govern the Folding of Protein Chains. Science (1973) 181, 223-230.
-
-\bibitem{Christ2007}
-Christ, C.D.; van Gunsteren, W.F. Enveloping distribution sampling: A method to calculate free energy differences from a single simulation,
-J. Chem. Phys. (2007), 126, 184110.
+\bibitem{Kumar1992}
+Kumar, S.; Rosenberg, J.M.; Bouzida, D.; Swendsen, R.H.; Kollman, P.A. The weighted histogram analysis method for free-energy calculations on biomolecules. I. The 
+method. J Comput Chem. 1992, 13, 1011-1021.
 
@@
 
-\bibitem{Krivov2004}
-Krivov, S.; Karplus, M. Hidden complexity of free energy surfaces for peptide (protein) folding Proc. Natl. Acad. Sci. U.S.A. 2004, 101, (41), 14766.
+\bibitem{Brooks1983}
+Brooks, B. R.; Karplus, M. Harmonic dynamics of proteins: Normal modes and fluctuations in bovine pancreatic trypsin inhibitor. Proc. Natl. Acad. Sci. U.S.A. 1983, 80, 6571-6575.
 
-\bibitem{Alexander2007}Alexander, P.A.; He, Y.; Chen, Y.; Orban, J.
-Bryan, P.
-The design and characterization of two proteins with $88 \%$ sequence identity but differentstructure and function.
-Proc.
-Natl.
-Acad.
-Sci.
-2007, 104 (29), 11963-11968.
+\bibitem{Case1994}
+Case, D. Normal-mode analysis of protein dynamics Curr. Opin. Struct. Biol. 1994, 4, 285-290.
+
+\bibitem{Karplus1981}
+Karplus, M.; Kushick, J. Method for estimating the configurational entropy of macromolecules Macromolecules. 1981, 14, 325-332.
+
+\bibitem{Levy1984}
+Levy, R.; Karplus, M.; Kushick, J.; Perahia, D. Evaluation of the configurational entropy for proteins: application to molecular dynamics simulations of 
+an $\alpha$-helix Macromolecules 1984, 17, 1370-1374.
+
+\bibitem{Mobley2007}
+Mobley, D.L.; Chodera, J.D.; Dill, K.A. The combining and release method: obtaining correct binding free energies in
+the presence of protein conformational change. Journal of Chemical Theory and Computation 2007, 3, 1231-1235.
+
+\bibitem{Mobley2012}
+Mobley, D.L.; Klimovich, P.V. Perspective: Alchemical free energy calculations for drug discovery.
+J. Chem. Phys. 2012, 137, 230901-12.
+
+\bibitem{Mobley2006}
+Mobley, D.L.; Chodera, J.D.; Dill, K.A. On the use of orientational restraints and symmetry corrections in alchemical
+free energy calculations. J. Chem. Phy. 2006, 125, 084902.
+
+\bibitem{Alexander2007}
+Alexander, P.A.; He, Y.; Chen, Y.; Orban, J. Bryan, P. The design and characterization of two proteins with $88 \%$ sequence identity but differentstructure and 
+function. Proc. Natl. Acad. Sci. 2007, 104, 11963-11968.
 
 \bibitem{He2008}
-He, Y.; Chen, Y.; Alexander, P.A.; Orban, J. NMR structures of two designed proteins with high sequence identity but different fold and function. Proc. Natl. Acad. Sci. 2008, 105 (38), 14412-14417.
+He, Y.; Chen, Y.; Alexander, P.A.; Orban, J. NMR structures of two designed proteins with high sequence identity but different fold and function. Proc. Natl. Acad. Sci. 2008, 105, 14412-14417.
 
@@
 Alexander, P.A.; He, Y.; Chen, Y.; Orban, J. Bryan, P. A minimal sequence code for switching protein structure and function. Proc. Natl. Acad. Sci. 2009
-, 106(50), 21149-21154.
+, 106, 21149-21154.
 
 \bibitem{Bryan2010}
-Bryan, P.N.; Orban, J. Proteins that switch folds. Curr Opin Struct Biol. 2010, 20(4), 482-488.
+Bryan, P.N.; Orban, J. Proteins that switch folds. Curr Opin Struct Biol. 2010, 20, 482-488.
 
 \bibitem{He2012}
-He, Y.; Chen, Y.; Alexander, P.A.; Bryan, P.N.; Orban, J. Mutational tipping points for switching protein folds and functions. Structure. 2012, 20(2), 2
+He, Y.; Chen, Y.; Alexander, P.A.; Bryan, P.N.; Orban, J. Mutational tipping points for switching protein folds and functions. Structure. 2012, 20,
 83-91.
-
-\bibitem{Shortle2009}
-Shortle, D. One sequence plus one mutation equals two folds. Proc. Natl. Acad. Sci. 2009, 106(50), 21011-21012.
 
@@
 ence. 2009, 18(1), 229-239.
+
+\bibitem{Zhou2002}
+Zhou H, Zhou Y. Distance-scaled, finite ideal-gas reference state improves structure-derived potentials of mean force for structure selection and
+stability prediction. Protein Sci. 2002, 11, 2714-2726.
+
+\bibitem{Allison2011}
+Allison, J.R.; Bergeler, M.; van Gunsteren, W.F. Can computer modeling explain why two highly similar sequences fold into different structures ?
+Biochemistry. 2011, 50, 10965-10973.
+
+\bibitem{Handl2009}
+Handl, J.; Knowles, J.; Lovel, S.C. Artefacts and biases affecting the evaluation of scoring
+functions on decoy sets for protein structure prediction. Bioinformatics, 2009, 25, 1271-1279.
+
+\bibitem{Kryshtafovych2011}
+Kryshtafovych, A.; Fidelis, K; and Tramontano, A. Evaluation of model quality predictions in CASP9. Proteins, 2011, 79, 91–106
+
+\bibitem{Wang2011}
+Wang, Q.; Vantasin, K.; Xu, D.; Shang, Y. MUFOLD-WQA: A new selective consensus method for quality assessment in protein structure prediction.
+Proteins, 2011, 79: 185-195.
+
+\bibitem{Zemla2003}
+Zemla, A. LGA: a method for finding 3D similarities in protein structures. Nucleic Acids Res 2003, 31, 3370–3374.
 

### 2013-05-05 Paper/paper.tex
Revisions on manuscript and SI may 5, Alberto

@@
 
-Calculation of free energy differences is of central importance in the simulation of biochemical
+The calculation of free energy differences is of central importance in the simulation of biochemical
 systems. It is particularly difficult to calculate between pairs of macromolecular conformations as
-well as a computationally expensive task with existing methods. In this work, confinement approach
+well as a computationally expensive task with existing methods. In this work, the confinement approach
 is used to calculate absolute free energies of biomolecular systems. This method provides two main
 advantages: it does not require a reaction coordinate or transition path and it is fast to compute.
-Free energy calculated can be decomposed into a per residue contribution in an approximate way. Per
+The calculated free energy can be decomposed into a per residue contribution in an approximate way. Per
 residue free energy allows us to identify the reason behind conformational preferences in
@@
 \end{table}
+\Alberto{Should we comment on T0560? Why is it giving such a high value with such high gdt? It's not
+the point to make here, but it is what jumps at me.}
 
@@
 sequence they are able to stabilize one fold or the other (see Figure~\ref{fig:orban}). We refer to the $4\beta +
-\alpha$ structure as the $\beta$ conformation, and we refer to the $3\alpha$ structure as the $\alpha$ conformation. One
+\alpha$ structure as the $\beta$ conformation, and we refer to the $3\alpha$ structure as the
+$\alpha$ conformation. We denote sequences that prefer the $\alpha$ fold as GA and sequences that
+prefer $\beta$ as GB. One
 pair of sequences (GA88/GB88) is 88 percent identical in sequence, differing in seven positions. Another pair
@@
 \cite{Zemla2003}, which is a C$\alpha$ based measure of structural accuracy. It can be understood roughly as the
-percentage of residues that are correctly positioned in the model (range 0 to 100, higher is better). We did have enough
-computer model to analyze every model, so the initial models for the test were chosen from a selection of different
+percentage of residues that are correctly positioned in the model (range 0 to 100, higher is
+better). We did not have enough
+computer power to analyze every model, so the initial models for the test were chosen from a selection of different
 server groups that have done well in past CASP events.
@@
 
-First, we examined the ability of the confinement method to rank different models for a target that have been generated
+First, we examined the ability of the confinement method to rank different models for a target that
+has been generated
 by a single group using the same methodology for all predictions. We examined two targets: T0559 and T0560 (see Table~\ref{table:casp_testcase}).
@@
 \begin{tabular}{l l l}\hline
-    CASP Target  & PDB Identifier &  organism type  \\ \hline
+    CASP Target  & PDB Identifier &  Organism type  \\ \hline
      T0531       &    2KJX        &  extracellular domain of the jumping  \\
@@
 \end{figure}
-
+\Alberto{if the drop down bars/point are a different color it will be easier to see when the crystal
+is on top of the axis}
 \subsection*{The per-residue free energy is sensitive to small changes in protein conformation:}
@@
 The confinement method clearly identifies two hydrophobic residues V59 and I61, which destabilize the predicted model
-with respect to the experimental structure (Figure~\ref{fig:T0569_per_residue} and Figure S3). The sidechains of
+with respect to the experimental structure (Figure~\ref{fig:T0569_per_residue} and Figure S3). The side chains of
 these hydrophobic residues are oriented towards the protein hydrophobic core in the native NMR structure but oriented
 towards the exterior of the protein and are solvent exposed in the model. These residues are part of a beta-sheet
-in the experimental structure, but because of their sidechain orientation the corresponding beta-sheet become disordered in the predicted model 
+in the experimental structure, but because of their side chain orientation the corresponding
+beta-sheet becomes disordered in the predicted model 
 (Figure~\ref{fig:T0569_per_residue} and Figure S4). There is also a large
@@
 Per-residue free energy calculations (not shown) show that despite only small variations at the backbone level, the
-sidechains are oriented in very different ways (Figure~\ref{fig:T0538compare}), giving rise to large differences in the stabilization of
+side chains are oriented in very different ways (Figure~\ref{fig:T0538compare}), giving rise to large differences in the stabilization of
 certain residues. In particular, some of the differences arise from different salt bridge patterns and certain flexible
 polar residues exposed to the surface. This unexpected result shows that the confinement method is very sensitive to
-local interactions (including sidechain reorientation) and may indicate issues with the forcefield and implicit solvent
+local interactions (including side chain reorientation) and may indicate issues with the forcefield and implicit solvent
 models used in the calculation.
@@
 \end{center}
-\caption{The per residue free energy calculation found some differences of the sidechain orientation in computer generated model 1 and 
+\caption{The per residue free energy calculation found some differences of the side chain orientation in computer generated model 1 and 
 experimental structure of CASP target T0538. 
@@
 predicted model for this target, whereas the other model was predicted best by QA test of MUFOLD-WQA. The calculated
-free energy using confinement method indicate that the model by PconsR is more stable by $3.9 Kcal/Mol$ which support
+free energy using confinement method indicate that the model by PconsR is more stable by $3.9
+Kcal/Mol$ which supports
 the GDT\_TS trend.  Although consensus prediction strategies do well specially when dealing with large numbers of structures, they 
@@
 
-The confinement method has been described in details in ref. by Tyka et al. \cite{Tyka2006} and
+The confinement method has been described in details by Tyka et al. in reference \cite{Tyka2006} and
 Cecchini et  al. \cite{Cecchini2009}. The basic approach of the confinement approach is the same in
@@
 
-\item  In the first step, a minimization of A and B conformations are performed. These minimized conformations
+\item  In the first step, minimizations of A and B conformations are performed. These minimized conformations
     ($A^{\ast}$ and $B^{\ast}$) are the reference conformation of that state.
@@
     this step is estimated using a thermodynamic integration approach developed by Tyka et. al. \cite{Tyka2006}. The
-    confinement free energies calculated this step are denoted as $\Delta G_{A,A^\ast}$ and $\Delta G_{B,B^\ast}$ in
+    confinement free energies calculated in this step are denoted as $\Delta G_{A,A^\ast}$ and $\Delta G_{B,B^\ast}$ in
     Figure~\ref{fig:method}.
@@
 One advantage of confinement method is that
-none of the simulation during restraining step depends on each other. Therefore, with available computer resources it can be fast 
-to compute. The computational cost for a $56$-residue proteins with chameleon sequence that is reported in this article is only $4$ hours 
-in a single gpu computer. All calculations were performed with the Amber 11 suit of programs~\cite{Case2012,Goetz2012} in combination with ff99SB
-forcefield~\cite{Hornak2006} and generalized born implicit solvent \cite{Mongan2006}.
+none of the simulations during the restraining step depends on each other. Therefore, with available computer resources it can be fast 
+to compute. The computational cost for the $56$-residue protein with chameleon sequence that is reported in this article is only $4$ hours 
+in a single gpu computer\Alberto{all the different confinements in one GPU?}. All calculations were
+performed with the Amber 11 suit of programs~\cite{Case2012,Goetz2012} in combination with the ff99SB
+forcefield~\cite{Hornak2006} and the generalized born implicit solvent \cite{Mongan2006}.
 
@@
 method for calculation of per residue free energy in an approximate way. For this purpose, the confinement energy,
-$\Delta G_{A,A*}$ and $\Delta G_{B,B*}$ of each residue is calculated in the usual numerical way as described in ref. by
-Tyka et al. \cite{Tyka2006}. The internal energy of each residue is calculated using the decomp module of amber from the
+$\Delta G_{A,A*}$ and $\Delta G_{B,B*}$ of each residue is calculated in the usual numerical way as described by
+Tyka et al. in reference \cite{Tyka2006}. The internal energy of each residue is calculated using
+amber's "decomp" module using the
 final restrained trajectory. We call this method approximate as we ignore the entropic contribution from the normal mode
-or quasiharmonic analysis. However, this contrbution to the total free energy is much smaller, which allow us to study
+or quasiharmonic analysis. However, this contribution to the total free energy is much smaller, which allow us to study
 the mechanistic details of conformational preference of each residue.

### 2013-05-05 Paper/supporting_information.tex
Revisions on manuscript and SI may 5, Alberto

@@
 
-\section{Preparation of input model for chamelon sequences}
+\section*{Preparation of input models for chamelon sequences}
 
 We prepared computer generated models in both $\alpha$ and $\beta$ conformations for all five chamelon sequences mentioned 
-in the main text. Crystallographic structure of GA95 and GB95 sequence have $\alpha$ and $\beta$ fold respectively. When
-we attempt to prepare the $\alpha$ and $\beta$ fold with GA95 sequence, we start with the crystallographic 
-structure of GA95 and GB95 with GA95 sequence. However, we remove the sidechains at three mutated positions in both the fold 
-and call the program SCWRL4 \cite{Krivov2009} to generate the sidechain conformation of the mutated residues and its neighbors. Followed by 
-this step, we do long molecular dynamics
-minimization to remove any possible bad contact. This procedure was followed to generate all $\alpha$ and 
-$\beta$ conformations of five sequences that are used for free energy calculation.
+in the main text. The crystallographic structure of GA95 and GB95 sequences have $\alpha$ and $\beta$ fold respectively.
+To prepare the $\alpha$ and $\beta$ fold for the GA95 sequence, we start with the crystallographic
+conformation of the native GA95 and GB95 sequences. 
+We then remove the sidechains at three mutated positions in both folds
+and call the program SCWRL4 \cite{Krivov2009} to generate the side chain conformations of the
+mutated residues and their neighbors. Followed by 
+this step, we perform extensive molecular dynamics
+minimizations to remove any possible bad contact. This procedure was used to generate all $\alpha$ and 
+$\beta$ conformations for the five sequences that were used in the free energy calculation.
 
@@
 \label{fig:orban_full}
-\caption{Per residue free energy calculation indicate significant differences of sidechain orientation of key amino acid residues. These amino acid
-residues control the free energy equilibrium to either $\beta$ or
+\caption{Per residue free energy calculations indicate significant differences of side chain orientation of key amino acid residues. These amino acid
+residues displace the free energy equilibrium to either $\beta$ or
 $\alpha$ conformation. All figures in the left are in $\beta$ conformation, whereas the right ones are in $\alpha$ conformation except in Figure (F) 
@@
 \end{center}
-\caption{Key differences between experimental structure and the computer generated model model as indicated by per residue free energy calculation.
-The sidechain of hydrophobic residues, V59 and I61 are oriented towards the protein hydrophobic core in (A)  
-structure, (B) exposed to the surfure in model. The beta sheet containing these residues is disordered in the generated model.
-A salt bridge between K76 and D11 in (D) computer generated structure, which is absent in the (C) native structure.}
+\caption{Key differences between the experimental structure and the computer generated models as
+    indicated by the per residue free energy calculation.
+The side chains of hydrophobic residues, V59 and I61 are oriented towards the protein hydrophobic core in (A)  
+, in (B) they are exposed to the surface. The beta sheet containing these residues is disordered in
+the generated model due to
+a salt bridge between K76 and D11 in (D) the computer generated structure, which is absent in the (C) native structure.}
 \label{fig:T0569_per_residue}
@@
 \end{center}
-\caption{The native and three model structure of engineered protein from Asr4154 protein (PDB ID: 2L09 and CASP
-    code:T0538). The model 1,2 and 3 are from the group PconsR, Shell and FOLDIT respectively.}
+\caption{The native and three model structures for an engineered protein (PDB ID: 2L09 and CASP
+    code:T0538). Models 1,2 and 3 are from the group PconsR, Shell and FOLDIT respectively.}
 \label{fig:T0538}
@@
 
-\section{What can we say about low resolution models from CASP experiments?}
+\section*{What can we say about low resolution models from CASP experiments?}
 
@@
 native. But the question remains how far from native can we go and still see that the method produces correct result. In
-this section we explored this question with models from  extracellular domain of the jumping translocation breakpoint
-protein (pdb id: 2KJX). Most of the group could only generate low resolution models for this CASP9 target (id: T0531) .
-In our comparison, we choose five models by the group MUFOLD-MD, which was the best performing group for this target
-with their best model had a GDT\_TS value of 44. The result presented in Figure~\ref{fig:T0531} shows: 1.) native is
+this section we explored this question with models from the extracellular domain of the jumping translocation breakpoint
+protein (pdb id: 2KJX). Most of the groups could only generate low resolution models for this CASP9 target (id: T0531).
+In our comparison, we chose five models by the MUFOLD-MD group, which was the best performing group for this target
+even though their best model had a GDT\_TS value of only 44. The results presented in Figure~\ref{fig:T0531} show: 1.) native is
 correctly identified as expected and 2.) Surprisingly there is a high level of correlation between the GDT ranking and
-the free energy ranking for model 1 and model 3, the rest three structures with GDT\_TS score less than 35  are ordered
+the free energy ranking for model 1 and model 3, the other three structures with GDT\_TS score less than 35  are ordered
 incorrectly. It is worth to note that models 2 and 3 have the same GDT and very different free energies, meaning that

### 2013-05-08 Paper/paper.tex
upload May5th
May8th

@@
 The calculation of protein conformational free energy has been successfully attempted by a number of groups
-\cite{Shell2010,Ytreberg2006,Zheng2008,Spichty2010,Strajbl2000,Park2008,Tyka2006,Cecchini2009,Ovchinnikov2013}. Some of
+\cite{Ytreber2006a,Shell2010,Ytreberg2006,Zheng2008,Spichty2010,Strajbl2000,Park2008,Tyka2006,Cecchini2009,Ovchinnikov2013}. Some of
 these methods like the reference system method~\cite{Ytreberg2006}, deactivated morphing~\cite{Park2008}, the
@@
 the point to make here, but it is what jumps at me.}
+\Arijit{I think we have made it clear that crystal structures are often more stable due to salt-bridge effects. Also, we have shown
+in other cases that sidechains orientation may play important roles inspite of close GDT-TS. You will see free energy difference between 
+two computer generated model is not so high. So, I believe sidechain may play some role here.}
 
@@
     \caption{The confinement method can correctly rank the native structure and three predictions produced by a single
-        method for target T0559.}
+        method for target T0559. The backbone region of the predicted structures which differ substantially compare to the
+        experimental structure upon superposition are colored brown.}
 \label{fig:T0559}
@@
 is on top of the axis}
+\Arijit{I will wait for Sarina's figure. I have changed the caption of figure 5 to say why some region are colored brown.}
+
 \subsection*{The per-residue free energy is sensitive to small changes in protein conformation:}
@@
 \item  The full free energy, $\Delta G_{A,B}$ between the two state A and B is calculated as
-    $\Delta G{A,B} = \Delta G{A,A^\ast} - \Delta G{B,B^\ast} + \Delta G{A^\ast,B^\ast}$.
+    $\Delta G_{A,B} = \Delta G_{A,A^\ast} - \Delta G_{B,B^\ast} + \Delta G_{A^\ast,B^\ast}$.
 
@@
 none of the simulations during the restraining step depends on each other. Therefore, with available computer resources it can be fast 
-to compute. The computational cost for the $56$-residue protein with chameleon sequence that is reported in this article is only $4$ hours 
-in a single gpu computer\Alberto{all the different confinements in one GPU?}. All calculations were
+to compute. The computational cost of a single 20 ns simulation (along the confinement step) of $56$-residue protein 
+is only $4$ hours in a single gpu computer  \Alberto{all the different confinements in one GPU?}. All calculations were
 performed with the Amber 11 suit of programs~\cite{Case2012,Goetz2012} in combination with the ff99SB
@@
 \bibitem{Jonsson1998}
-Jonsson, H.; Mills, G.; Jacobsen, K.W. Nudged Elastic Band Method for Finding Minimum Energy Paths of Transitions,
+J\'{on}sson, H.; Mills, G.; Jacobsen, K.W. Nudged Elastic Band Method for Finding Minimum Energy Paths of Transitions,
 in Classical and Quantum Dynamics in Condensed Phase Simulations, Ed. B. J. Berne, G. Ciccotti and D. F.
@@
 E, W.; Ren, W.; Vanden-Eijnden, E. Simplified and improved string method for computing the minimum energy paths in
-barrier-crossing eventsi. J. Chem. Phys. 2007, 126, 164103.
+barrier-crossing events. J. Chem. Phys. 2007, 126, 164103.
 
@@
 
+\bibitem{Ytreber2006a}
+Ytreberg, F.M.; Swendsen, R.H.; Zuckerman, D.M. Comparison of free energy methods for molecular systems. J Chem Phys. 2006, 125, 184114.
+
 \bibitem{Ytreberg2006}
@@
 
-\bibitem{Tyka2006}Tyka, M.; Clarke, A.; Sessions, R. An Efficient, Path-Independent Method for Free-Energy Calculations. J.Phys.Chem. B 2006, 110, 17212-17220.
+\bibitem{Tyka2006}
+Tyka, M.; Clarke, A.; Sessions, R. An Efficient, Path-Independent Method for Free-Energy Calculations. J.Phys.Chem. B 2006, 110, 17212-17220.
 

### 2013-05-08 Paper/Average/analysis/make_b_factor.py

@@ -1,91 +1,0 @@
-#!/usr/bin/env python
-# encoding: utf-8
-
-import prody
-import numpy
-import math
-
-
-def smooth_b_factors(protein, cutoff):
-    return protein
-
-    energies = []
-
-    for i, atom_i in enumerate(protein):
-        w_total = 0.
-        energy_total = 0.
-        for j, atom_j in enumerate(protein):
-            r = numpy.linalg.norm(atom_i.getCoords() - atom_j.getCoords())
-            w = math.exp(-r ** 2 / (0.5 * cutoff ** 2))
-            energy_total += w * atom_j.getBeta()
-            w_total += w
-        energies.append(energy_total / w_total)
-
-    protein.setBetas(energies)
-    return protein
-
-
-def set_betas(energies, protein):
-    for i, atom in enumerate(protein):
-        res_index = atom.getResnum() - 1
-        atom.setBeta(energies[res_index])
-    return protein
-
-
-def get_per_residue(filename):
-    energies = numpy.loadtxt(filename)[:, 1:]
-    energies[:, -1] = 0.0072 * energies[:, -1]
-    return numpy.sum(energies, axis=1)
-
-energies_ga_a = -1 * numpy.loadtxt('confinement_Alpha_GA95')[:, 1] + 1 * numpy.loadtxt('enthalpy_spring_Alpha_GA95')[:, 1]
-energies_ga_a += get_per_residue('residue_Alpha_GA95')
-alpha_ga = prody.parsePDB('../Alpha/GA95/A95_min2.pdb', subset='CA')
-alpha_ga = set_betas(energies_ga_a, alpha_ga)
-alpha_ga = smooth_b_factors(alpha_ga, 5.0)
-
-energies_ga_b = -1 * numpy.loadtxt('confinement_Beta_GA95')[:, 1] + 1 * numpy.loadtxt('enthalpy_spring_Beta_GA95')[:, 1]
-energies_ga_b += 1 * get_per_residue('residue_Beta_GA95')
-beta_ga = prody.parsePDB('../Beta/GA95/A95_min.pdb', subset='CA')
-beta_ga = set_betas(energies_ga_b, beta_ga)
[…]

### 2013-05-08 Paper/paper.tex
May8th

@@
 none of the simulations during the restraining step depends on each other. Therefore, with available computer resources it can be fast 
-to compute. The computational cost of a single 20 ns simulation (along the confinement step) of $56$-residue protein 
+to compute. The computational cost for a single 20 ns simulation (along the confinement step) of $56$-residue protein 
 is only $4$ hours in a single gpu computer  \Alberto{all the different confinements in one GPU?}. All calculations were

### 2013-05-09 Paper/paper.tex
May8th_2
9may

@@
 which has a three-stranded $\beta$-sheet. Our confinement calculation shows that bhp1 is more stable by 1.7 kcal/mol,
-consistent with 4 $\mu$s equilibrium simulations showing that bhp1 is more favorable configuration by 1.8 kcal/mol, and
+consistent with 200 $\mu$s equilibrium simulations showing that bhp1 is more favorable configuration by 1.8 kcal/mol, and
 in agreement with previous calculations~\cite{Cecchini2009}.
@@
 Here, we chose two computer generated models from CASP taget T0538, where the top performing group "MUFOLD-WQA" \cite{Wang2011} failed.
-One of the model was from "PconsR" (GDT-TS = 83, qmode 1 = 0.5865) and another from "MULTICOM-NOVEL" ( GDT-TS = 83, qmode 1 = 0.5865). We 
+One of the model was from "PconsR" (GDT-TS = 96, qmode 1 = 0.5434) and another from "MULTICOM-NOVEL" ( GDT-TS = 83, qmode 1 = 0.5865). We 
 choose these two models as the model by PconsR was the most accurate

### 2013-05-09 Paper/paper.tex
9may_1

@@
 
-Second, we looked at 6 target proteins from the CASP9 experiment~\cite{Moult2011}.  For each target, we examined up to 5 submitted
+Second, we looked at a series of protein structure which has a similar sequence but different fold and function. We 
+investigated whether confinement method can predict their conformational preferences.
+Third, we looked at 6 target proteins from the CASP9 experiment~\cite{Moult2011}.  For each target, we examined up to 5 submitted
 models.  We computed the difference free energy between the true native and the best model.

### 2013-05-12 Paper/supporting_information.tex

@@
 \end{center}
-\caption{The native and three model structures for an engineered protein (PDB ID: 2L09 and CASP
-    code:T0538). Models 1,2 and 3 are from the group PconsR, Shell and FOLDIT respectively.}
+\caption{The free energy based ranking of the experimental and three predicted structures for an engineered protein (PDB ID: 2L09 and CASP
+    code:T0538). Models 1,2 and 3 are from the group PconsR, Shell and FOLDIT respectively. In this case, one of the predicted structure 
+    was found to be more stable than the experimental strucre.}
 \label{fig:T0538}

### 2013-05-12 Paper/paper.tex
Further tweaks.

--- 2013-05-08 Paper/paper.tex
@@
 
-The calculation of free energy differences is of central importance in the simulation of biochemical
-systems. It is particularly difficult to calculate between pairs of macromolecular conformations as
-well as a computationally expensive task with existing methods. In this work, the confinement approach
-is used to calculate absolute free energies of biomolecular systems. This method provides two main
-advantages: it does not require a reaction coordinate or transition path and it is fast to compute.
-The calculated free energy can be decomposed into a per residue contribution in an approximate way. Per
-residue free energy allows us to identify the reason behind conformational preferences in
-biomolecules. Through out the article we show its use in different challenging modeling problems. In
-particular, we show its use in predicting the conformational preference of chamaleon sequences
-(sequences with high sequence identity and different folds). This sequence dependent conformational
-preferences and per residue free energy decomposition set the stage for the use of this method in
-protein design.
+Understanding the free energy differences between different different conformations of a protein is important for
+understanding mechanism in biochemistry. The calculation of such conformational free energy differences is difficult and
+computational demanding. In this work, we explore the use of ``the confinement method'' to calculate free energy
+differences between protein conformations. This method provides two main advantages: (1) it does not require a reaction
+coordinate or transition path and (2) it is fast to compute. The free energy difference can be approximately composed
+into per-residue contributions, which provides insight into the physical basis for the preference of one conformation
+over another. We show that the confinement method can correctly predict the conformational preferences of chameleon
+sequences (nearly identical sequences with different structures), can identify the most native-like structures produced
+by structure prediction algorithms, and can yield detailed insight into the mechanisms behind these preferences.
 
@@
 
-In some problems of protein science, you want to know the relative stability of a protein's conformation $A$ compared to
-its conformation $B$.  We call this the \emph{difference free energy}. For example, in all allosteric mechanisms, a protein
-often adopts one conformation when a ligand is bound and another conformation when no ligand is bound~\cite{Elber2007}.
-To understand the mechanism requires knowledge of the relative free energies of the protein conformations.  Another
-example is in protein structure prediction, for example as practiced in the Critical Assessment of Structure Prediction
-(CASP) competition~\cite{Moult2011}. If you are using some computational model and have predicted two putative native
-structures, $A$ and $B$, you want to compute which has the lower free energy, in order to know which is the more
-native-like.  In a third example, you may want to know whether mutating a few amino acids in a protein could cause the
-protein to switch from one stable conformation to another, because that could have important consequences for biological
-mechanisms and disease. And fourth, sometimes binding a ligand induces a protein from one conformation to another. A
-quantitative understanding of such \emph{induced-fit} situations requires knowledge of the difference free energy of the
-protein in the two states. In all these cases, it would be useful to have a computational method that can be used
-efficiently with atomically detailed physical forcefield models to compute the free energy differences between different
-given protein conformations.
+In some problems of protein science, we want to know the relative stability of a protein's conformation $A$ compared to
+another of its conformations $B$. We call this the \emph{difference free energy}. For example, in allosteric mechanisms,
+a protein often adopts one conformation when a ligand is bound and another conformation when no ligand is
+bound~\cite{Elber2007}. To understand the mechanism requires knowledge of the relative free energies of the different
+protein conformations.  Another example is in protein structure prediction, such as is in the Critical Assessment of
+Structure Prediction (CASP) competition~\cite{Moult2011}. If you are using some computational model and have predicted
+two putative native structures, $A$ and $B$, you want to compute which has the lower free energy, in order to know which
+is the more native-like.  In a third example, you may want to know whether mutating a few amino acids in a protein could
+cause the protein to switch from one stable conformation to another, because that could have important consequences for
+biological mechanisms and disease. Fourth, sometimes binding a ligand induces a protein from one conformation to
+another. A quantitative understanding of such \emph{induced-fit} situations requires knowledge of the difference free
+energy of the protein in the two states. In all these cases, it would be useful to have a computational method that can
+be used efficiently with atomically detailed physical forcefield models to compute the free energy differences between
+different given protein conformations.
 
-One widely explored strategy
-is to use molecular dynamics simulations along some putative reaction coordinate pathway from conformation $A$ to
-$B$~\cite{Elber2007,West2007,Haas2007,Jonsson1998,E2007,Dellago2002,Cheng2006,Elber2005,Chipot2007}. The free energy along this
-reaction coordinate can then be determined using methods such as umbrella sampling \cite{Torrie1977,Mascarenhas2013}  
-and weighted histogram analysis method (WHAM) \cite{Kumar1992}. Such
-approaches have several limitations. First, it is necessary to know an efficient reaction pathway from $A$ to $B$. If
-conformations $A$ and $B$ are quite different, then it can be challenging to find such paths. Second, these methods are
+One widely explored strategy is to use molecular dynamics simulations along some putative reaction coordinate pathway
+from conformation $A$ to
+$B$~\cite{Elber2007,West2007,Haas2007,Jonsson1998,E2007,Dellago2002,Cheng2006,Elber2005,Chipot2007}. The free energy
+along this reaction coordinate can then be determined using methods such as umbrella sampling
+\cite{Torrie1977,Mascarenhas2013} and the weighted histogram analysis method (WHAM) \cite{Kumar1992}. Such approaches
+have several limitations. First, it is necessary to know an efficient reaction pathway from $A$ to $B$. If conformations
+$A$ and $B$ are quite different, then it can be challenging to find such paths. Second, these methods are
 computationally slow. To get an accurate estimate of the total free energy difference $\Delta G = G_B - G_A$ requires
 accurate determinations of the many small free energy differences $A \rightarrow 1 \rightarrow 2 \ldots \rightarrow B$,
-so each step requires substantial amounts of sampling.  And third, these methods are prone to large errors, because the
-pathway error is a sum of many errors among the many steps.
+and each step requires substantial amounts of sampling.  Third, large errors can accumlate, because the pathway error is
+a sum of many errors along the many steps.
 
 The calculation of protein conformational free energy has been successfully attempted by a number of groups
-\cite{Ytreber2006a,Shell2010,Ytreberg2006,Zheng2008,Spichty2010,Strajbl2000,Park2008,Tyka2006,Cecchini2009,Ovchinnikov2013}. Some of
-these methods like the reference system method~\cite{Ytreberg2006}, deactivated morphing~\cite{Park2008}, the
-confinement method~\cite{Tyka2006,Cecchini2009,Ovchinnikov2013} take an alternative strategy for computing difference
-free energies that does not require knowing a pathway from $A$ to $B$.
+\cite{Ytreber2006a,Shell2010,Ytreberg2006,Zheng2008,Spichty2010,Strajbl2000,Park2008,Tyka2006,Cecchini2009,Ovchinnikov2013}.
+Some of these methods, like the reference system method~\cite{Ytreberg2006}, deactivated morphing~\cite{Park2008}, and
+the confinement method~\cite{Tyka2006,Cecchini2009,Ovchinnikov2013}, take an alternative strategy for computing
+difference free energies that does not require knowing a pathway from $A$ to $B$.
 
@@
 simulations. Similarly, a second ensemble $B$ is restricted to $B^\ast$. We then compute the free energy between
-$A^\ast$ and $B^\ast$ using either normal mode analysis \cite{Brooks1983,Case1994} or the quasiharmonic 
-method \cite{Karplus1981,Levy1984}. This step takes into account
-both the remaining conformational entropy in each ensemble and the remaining enthalpy difference. In this way, we can
-calculate the free energy difference between the two end states without needing to define a physical path or reaction
-coordinate connecting the two states. The confinement method shares some similarities with the `confine-and-release'
-method for computing ligand binding affinities~\cite{Mobley2007,Mobley2012,Mobley2006}.
+$A^\ast$ and $B^\ast$ using either normal mode analysis \cite{Brooks1983,Case1994} or the quasiharmonic method
+\cite{Karplus1981,Levy1984}. This step takes into account both the remaining conformational entropy in each ensemble and
+the remaining enthalpy difference. In this way, we can calculate the free energy difference between the two end states
+without needing to define a physical path or reaction coordinate connecting the two states. The confinement method
+shares some similarities with the `confine-and-release' method for computing ligand binding
+affinities~\cite{Mobley2007,Mobley2012,Mobley2006}.
 
@@
     \end{center}
-    \caption{Graphical representation of the thermodynamic cycle involving confinement method.}
+    \caption{Graphical representation of the thermodynamic cycle employed in the confinement method.}
 \label{fig:method}
@@
 show that the confinement method can also be applied successfully to larger systems, such as the series of chameleon
-proteins designed by Orban and co-workers~\cite{Alexander2007,He2008,Alexander2009,Bryan2010,He2012}, which can switch between two
-completely different folds with only slight changes in sequence. We also show that our computed difference free energies
-are useful in evaluating the quality of CASP target protein predictions.  This could be of significant value for
-ultimately improving the energetics in protein-structure models, not just the structures. And, finally, we show that we
-can approximately decompose difference free energies into individual amino acid components. This offers the opportunity
-for diagnosing the structural basis for difference free energies, which can be useful for interpreting biological
-mechanisms.
+proteins designed by Orban and co-workers~\cite{Alexander2007,He2008,Alexander2009,Bryan2010,He2012}, which can switch
+between two completely different folds with only slight changes in sequence. We also show that our computed difference
+free energies are useful in evaluating the quality of CASP target protein predictions.  This could be of significant
+value for ultimately improving the energetics in protein-structure models, not just the structures. And, finally, we
+show that we can approximately decompose difference free energies into individual amino acid components. This offers the
+opportunity for diagnosing the structural basis for difference free energies, which can be useful for interpreting
+biological mechanisms.
 
@@
 
-First, we validated that our implementation of the confinement method produces results compatible with previous
+First, we verified that our implementation of the confinement method produces results compatible with previous
 calculations reported in the literature.  The method has previously been applied to a 16 amino acid residue
 $\beta$-hairpin from protein G, known as BHP~\cite{Cecchini2009}. We calculated the free energy difference between the
-native conformation---called bhp1, which has a two-stranded $\beta$-sheet---and a non-native conformation---called bhp3,
-which has a three-stranded $\beta$-sheet. Our confinement calculation shows that bhp1 is more stable by 1.7 kcal/mol,
-consistent with 4 $\mu$s equilibrium simulations showing that bhp1 is more favorable configuration by 1.8 kcal/mol, and
-in agreement with previous calculations~\cite{Cecchini2009}.
+native conformation (called bhp1, which has a two-stranded $\beta$-sheet) and a non-native conformation (called bhp3,
+which has a three-stranded $\beta$-sheet). Our confinement calculation shows that bhp1 is more stable by 1.7 kcal/mol,
+which is consistent with 200 $\mu$s equilibrium simulations showing that bhp1 is more favorable configuration by 1.8
+kcal/mol, and in agreement with previous calculations~\cite{Cecchini2009}.
 
-Second, we looked at 6 target proteins from the CASP9 experiment~\cite{Moult2011}.  For each target, we examined up to 5 submitted
-models.  We computed the difference free energy between the true native and the best model.
+Second, we looked at 6 target proteins from the CASP9 experiment~\cite{Moult2011}. For each target, we examined up to 5
+submitted models.  We computed the difference free energy between the experimental native structure and the best model.
 Figure~\ref{fig:summary_casp} and Table~\ref{table:casp_control} show that, in 5 out of 6 cases, the confinement method
 assigns a lower free energy to the experimentally determined structure than to any of the decoys. Other well-known
-discriminators can also successfully tell native structures from computer generated models~\cite{Sheffler2009,Zhou2002} 
-; it is just an independent useful validation that the confinement calculations make
-sense.
+discriminators can also successfully tell native structures from computer generated models~\cite{Sheffler2009,Zhou2002};
+it is just an independent useful validation that the confinement calculations make sense.
 
@@
     participants. We report the free energy difference between the most favorable decoy and the experimentally
-    determined structure. Positive $\Delta\Delta G (=G_{best~decoy} - G_{native})$ values indicate that the experimental structure is predicted to be
-    more favorable than any of the decoys.}
+    determined structure. Positive $\Delta\Delta G (=\Delta G_{best~decoy} - \Delta G_{native})$ values indicate that
+    the experimental structure is predicted to be more favorable than any of the decoys.}
 \label{table:casp_control}
@@
 \end{table}
-\Alberto{Should we comment on T0560? Why is it giving such a high value with such high gdt? It's not
-the point to make here, but it is what jumps at me.}
-\Arijit{I think we have made it clear that crystal structures are often more stable due to salt-bridge effects. Also, we have shown
-in other cases that sidechains orientation may play important roles inspite of close GDT-TS. You will see free energy difference between 
-two computer generated model is not so high. So, I believe sidechain may play some role here.}
 
@@
 We also tested the confinement method predictions for difference free energies on the chameleon sequences of Orban et
-al.\cite{Alexander2007,He2008,Alexander2009,Bryan2010,He2012}. These are instances in which two highly
-similar sequences fold into remarkably different structures. Orban and co-workers have designed a protein-G-like
-sequence of 56-residues that is marginally stable in one of two possible folds. By mutating key residues in this
-sequence they are able to stabilize one fold or the other (see Figure~\ref{fig:orban}). We refer to the $4\beta +
-\alpha$ structure as the $\beta$ conformation, and we refer to the $3\alpha$ structure as the
-$\alpha$ conformation. We denote sequences that prefer the $\alpha$ fold as GA and sequences that
-prefer $\beta$ as GB. One
-pair of sequences (GA88/GB88) is 88 percent identical in sequence, differing in seven positions. Another pair
-(GA95/GB95) is 95 percent identical, differing in three positions. Accurately predicting the structural preferences of
-these structures presents a serious challenge for computational
-methods \cite{Allison2011}.
+al.\cite{Alexander2007,He2008,Alexander2009,Bryan2010,He2012}. These are instances in which two highly similar sequences
+fold into remarkably different structures. Orban and co-workers have designed a protein-G-like sequence of 56-residues
+that is marginally stable in one of two possible folds. By mutating key residues in this sequence they are able to
+stabilize one fold or the other (see Figure~\ref{fig:orban}). We refer to the $4\beta + \alpha$ structure as the $\beta$
+conformation, and we refer to the $3\alpha$ structure as the $\alpha$ conformation. We denote sequences that prefer the
+$\alpha$ fold as GA and sequences that prefer $\beta$ as GB. One pair of sequences (GA88/GB88) is 88 percent identical
+in sequence, differing in seven positions. Another pair (GA95/GB95) is 95 percent identical, differing in three
+positions. Accurately predicting the structural preferences such similar sequences poses a difficult challenge for
+computational methods \cite{Allison2011}.
 
@@
 chameleon sequence. For example, we took the sequence of GA88 and built a model with the same overall structure as GB88.
-We then used the confinement method to assess the free energy difference between the experimentally determined structure
-of GA88 and the model (with the GA88 sequence and the GB88 structure). The confinement method was able to predict the
-conformational preferences correctly for all four sequences (data not shown). This is, however, not surprising as it is
-often easy to distinguish computational models from native structures~\cite{Handl2009,Sheffler2009}. To avoid this
-potential problem, we instead computed relative free energy of two different computer generated models for each
-sequence. One model is based on the $\alpha$ structure and the other on the $\beta$ structure (see Supporting
-Information for details on the modeling procedure). This is a much more realistic test of the confinement method's
-ability to accurately calculate the difference free energy.
+We then used the confinement method to assess the free energy difference between these two structures. In other words,
+we are comparing the experimentally determined GA88/$\alpha$ structure with the computer-generated GA88/$\beta$
+structure. The confinement method was able to predict the conformational preferences correctly for all four sequences
+(data not shown). This is, however, not surprising as it is often easy to distinguish computational models from native
+structures~\cite{Handl2009,Sheffler2009}. To avoid this potential problem, we instead computed relative free energy of
+two different computer generated models for each sequence. One model is based on the $\alpha$ structure and the other on
+the $\beta$ structure (see Supporting Information for details on the modeling procedure). This is a much more realistic
+test of the confinement method's ability to accurately calculate the difference free energy.
 
@@
 
-To better understand the mechanism behind these chameleon proteins, we decomposed the calculated free energy into
-per-residue contributions in an approximate way~\cite{Tyka2006}. We decompose each confinement step ($\Delta
-G_{A,A^\ast}$, $\Delta G_{B,B^\ast}$) into per-residue contributions. We can also decompose the remaining enthalpy in
-the confined ensemble into per-residue contributions. However, the method is approximate because we do not include the
-residual conformational entropy from the normal mode or quasiharmonic analysis steps. Computing per-residue
-contributions helps us to identify important residues that stabilize a particular conformation. The per residue free
-energy, $\Delta \Delta G (\beta - \alpha)$ is shown in Figure~\ref{fig:perresidue_orban}.
+To better understand the mechanism behind these chameleon proteins, we approximately decomposed~\cite{Tyka2006} the free
+energy into per-residue contributions. We decompose each confinement step ($\Delta G_{A,A^\ast}$, $\Delta G_{B,B^\ast}$)
+into per-residue contributions. We can also decompose the remaining enthalpy in the confined ensemble into per-residue
+contributions. However, the method is approximate because we do not include the residual conformational entropy from the
+normal mode or quasiharmonic analysis steps. Computing per-residue contributions helps us to identify important residues
+that stabilize a particular conformation. The per-residue difference free energy, $\Delta \Delta G (\beta - \alpha)$, is
+shown in Figure~\ref{fig:perresidue_orban}.
 
@@
 understood by looking in detail at the local environment for those residues. For example, the region around residue 7
-forms a random coil in the $\alpha$ structure, whereas it is forming a beta sheet structure in the $\beta$ structure.
-These residues strongly favour the locally well packed and hydrogen bonded environment found in the $\beta$ sheet.
-Overall, favoring $\alpha$ or $\beta$ structure is a delicate balance, where the relative global free energy difference
-is small and the contributions of different per residue tendencies balance out. It is therefore very likely that by
-changing key residue the global fold preferences can be changed.
+forms a random coil in the $\alpha$ structure, whereas it forms a beta-sheet structure in the $\beta$ structure. These
+residues strongly favour the locally well-packed and hydrogen bonded environment found in the $\beta$ sheet, but in the
+alpha conformation there are no other residues to form a beta sheet with. The global conformation of the protein has
+``overridden'' the local preference of these residues. Overall, favoring the $\alpha$ or $\beta$ structure is a delicate
+balance, where the global free energy difference is small and the contributions of different per-residue tendencies
+largely cancel out. For these chameleon proteins, it is therefore very likely that by changing key residue the global
+fold preferences can be changed.
 
 Three residues are mutated between GA95 and GB95, at positions 20, 30, and 45. In GA95, these residues (L20, I30 and
-L45) stabilize the $\alpha$ structure (compare the upper and lower panels on Figure~\ref{fig:perresidue_orban}). In
-GB95, two of these residues (F30 and Y45) favor the $\beta$ structure, because they have large solvent exposed surface
-areas in the $\alpha$ structure, but are more buried in the $\beta$ structure. Additionally Y45 forms a hydrogen bond
-with D47 in the $\beta$ structure. On the other hand, residue A20 from GB95 still favors the $\alpha$ structure,
-although less strongly than L20 in GA95.
+L45) stabilize the $\alpha$ structure (Figure~\ref{fig:perresidue_orban}, upper panel). In GB95 (lower panel), two of
+these residues (F30 and Y45) favor the $\beta$ structure, because they have large solvent exposed surface areas in the
+$\alpha$ structure, but are more buried in the $\beta$ structure. Additionally, Y45 forms a hydrogen bond with D47 in
+the $\beta$ structure. On the other hand, residue A20 from GB95 still favors the $\alpha$ structure, although less
+strongly than L20 in GA95.
 
-The $\alpha$ and $\beta$ sequences are nearly identical, so there are some common features observed for all sequences.
+The GA and GB sequences are nearly identical, so there are some common features observed for all sequences.
 The experimental observations classified the protein into two parts: Amino acids 9--51 are fully structured in both
@@
 These per-residue free energy decompositions provide a great deal of insight into the driving forces behind protein
-folding and conformational change. We believe that such calculations may also be useful for protein design---designing
-proteins with specific structures and functions.
+folding and conformational change. We believe that such calculations may also be useful for designing proteins with
+specific structures and functions.
 
@@
 missed by other meta-prediction servers. Most successful meta-prediction servers are based on the idea of consensus: if
-many different prediction methods produce similar results, then that is probably a correct prediction \cite{Kryshtafovych2011,Wang2011}. This is
-often a powerful heuristic, but it can miss cases where there is a very good result that is only predicted by one
-method. We chose several cases where such structures were missed by the best meta-predictors in CASP and assessed if the
-confinement method can correctly identify these accurate models.
+many different prediction methods produce similar results, then that is probably a correct prediction
+\cite{Kryshtafovych2011,Wang2011}. This is often a powerful heuristic, but it can miss cases where there is a very good
+result that is only predicted by one method. We chose several cases where such structures were missed by the best
+meta-predictors in CASP and assessed if the confinement method can correctly identify these accurate models. All of our
+calculations were performed after the CASP experiment; they are not blind predictions.
 
@@
 \cite{Zemla2003}, which is a C$\alpha$ based measure of structural accuracy. It can be understood roughly as the
-percentage of residues that are correctly positioned in the model (range 0 to 100, higher is
-better). We did not have enough
-computer power to analyze every model, so the initial models for the test were chosen from a selection of different
+percentage of residues that are correctly positioned in the model (range 0 to 100, higher is better). We did not have
+enough computer power to analyze every model, so we chose a small subset of examples from a selection of different
 server groups that have done well in past CASP events.
@@
 
-First, we examined the ability of the confinement method to rank different models for a target that
-has been generated
-by a single group using the same methodology for all predictions. We examined two targets: T0559 and T0560 (see Table~\ref{table:casp_testcase}).
+First, we examined the ability of the confinement method to rank different models for a target that has been generated
+by a single group using the same methodology for all predictions. We examined two targets: T0559 and T0560 (see
+Table~\ref{table:casp_testcase}).
 
@@
 \begin{center}
-\caption{The six different CASP targets that are used in this study. The corresponding PDB Identifier and the 
-         organism type are also listed.}
+\caption{The six different CASP targets that are used in this study. The corresponding PDB Identifier and description
+    are also listed.}
 \label{table:casp_testcase}
 \begin{tabular}{l l l}\hline
-    CASP Target  & PDB Identifier &  Organism type  \\ \hline
+    CASP Target  & PDB Identifier &  Description \\ \hline
      T0531       &    2KJX        &  extracellular domain of the jumping  \\
@@
 models were discarded as they were too similar to the rest of the models. Again, we can correctly identify the native
-sate and our calculated free energy based ranking matches well with GDT-TS (Figures S2 in the supporting information).
+sate and our calculated free energy based ranking matches well with GDT-TS (Figure S2 in the supporting information).
 
@@
 \end{figure}
-\Alberto{if the drop down bars/point are a different color it will be easier to see when the crystal
-is on top of the axis}
-\Arijit{I will wait for Sarina's figure. I have changed the caption of figure 5 to say why some region are colored brown.}
 
-\subsection*{The per-residue free energy is sensitive to small changes in protein conformation:}
+\subsection*{The per-residue free energy is sensitive to small changes in protein conformation}
 
-In a previous section, we discussed how the per-residue free energy can reveal mechanistic
-detail behind the conformational preference of a chameleon sequence. In this section, our aim is to apply the same
-method and try to understand if the per-residue free energy can help us identify residues which stabilize or destabilize
-a particular region of a protein. In this case, the two conformations have similar folds with only small changes in
-localized regions. We chose CASP target T0569 and compared the experimental NMR structure with the best predicted model
-(GDT-TS=78; predicted by the ``Mufold'' group). The confinement method predicts that the experimental structure is more stable
-by 20 kcal/mol (Figure S2).
+In a previous section, we discussed how the per-residue free energy can reveal mechanistic detail behind the
+conformational preference of chameleon sequences. In this section, our aim is to apply the same method and try to
+understand if the per-residue free energy can help us identify problematic residues in structure predictions. We chose
+CASP target T0569 and compared the experimental NMR structure with the best predicted model (GDT-TS=78; predicted by the
+``Mufold'' group). The confinement method predicts that the experimental structure is more stable by 20 kcal/mol (Figure
+S2).
 
 The confinement method clearly identifies two hydrophobic residues V59 and I61, which destabilize the predicted model
-with respect to the experimental structure (Figure~\ref{fig:T0569_per_residue} and Figure S3). The side chains of
-these hydrophobic residues are oriented towards the protein hydrophobic core in the native NMR structure but oriented
-towards the exterior of the protein and are solvent exposed in the model. These residues are part of a beta-sheet
-in the experimental structure, but because of their side chain orientation the corresponding
-beta-sheet becomes disordered in the predicted model 
-(Figure~\ref{fig:T0569_per_residue} and Figure S4). There is also a large
-difference around K76, which forms a salt-bridge with D11 in the predicted model, but not in the experimental structure.
-This suggests that salt-bridge interactions are too favorable for the combination of force field and implicit solvent
-model we use, which has been a problem noted in the past \cite{Roe2007}.
+with respect to the experimental structure (Figure~\ref{fig:T0569_per_residue} and Figure S3). The side chains of these
+hydrophobic residues are oriented towards the protein hydrophobic core in the native NMR structure but oriented towards
+the exterior of the protein and are solvent exposed in the model. These residues are part of a beta-sheet in the
+experimental structure, but because of their side chain orientation the corresponding beta-sheet becomes disordered in
+the predicted model (Figure~\ref{fig:T0569_per_residue} and Figure S4). There is also a large difference around K76,
+which forms a salt-bridge with D11 in the predicted model, but not in the experimental structure. This suggests that
+salt-bridge interactions are too favorable for the combination of force field and implicit solvent model we use, which
+has been a problem noted in the past \cite{Roe2007}.
 
@@
         target T0569. The amino acid residues that are colored in deep red and deep blue stabilizes the NMR structure
-        and the prediction, respectively; the residues with light blue color does not have a strong preference.}
+        and the prediction, respectively; the residues with light blue color do not have a strong preference.}
 \label{fig:T0569_per_residue}
@@
 
-\subsection*{The confinement method occasionally produces incorrect results:}
+\subsection*{The confinement method occasionally produces incorrect results}
 
@@
 models (Model 1: ``PconsR''---GDT-TS=96; Model 2: ``Shell''---GDT-TS=90; Model 3: ``FOLDIT''---GDT-TS=86). Contrary to
-our expectation, the confinement method predicts that computer generated Model 1 is more stable than the crystal structure (Figure S5).
-Per-residue free energy calculations (not shown) show that despite only small variations at the backbone level, the
-side chains are oriented in very different ways (Figure~\ref{fig:T0538compare}), giving rise to large differences in the stabilization of
-certain residues. In particular, some of the differences arise from different salt bridge patterns and certain flexible
-polar residues exposed to the surface. This unexpected result shows that the confinement method is very sensitive to
-local interactions (including side chain reorientation) and may indicate issues with the forcefield and implicit solvent
-models used in the calculation.
+our expectation, the confinement method predicts that computer generated Model 1 is more stable than the crystal
+structure (Figure S5). Per-residue free energy calculations (not shown) show that despite only small variations at the
+backbone level, the side chains are oriented in very different ways (Figure~\ref{fig:T0538compare}), giving rise to
+large differences in the stabilization of certain residues. In particular, some of the differences arise from different
+salt bridge patterns and certain flexible polar residues exposed to the surface. This unexpected result shows that the
+confinement method is very sensitive to local interactions (including side chain reorientation) and may indicate issues
+with the forcefield and implicit solvent models used in the calculation.
 
@@
 \end{center}
-\caption{The per residue free energy calculation found some differences of the side chain orientation in computer generated model 1 and 
-experimental structure of CASP target T0538. 
-(A) Two salt bridges between residues E35-R32 and E28-K24 exist in the native structure. These are
-absent in the computer generated model structure. Instead
-(B) a new salt bridge is formed between residues R32-E28 in model. (C) Another salt bridge between
-residues  R26-E50 in the experimental structure. (D) In comparison to the native 
-structure, orientation of F51 differs in the model.}
+\caption{Small differences in sidechain orientation can cause large differences in free energy. We compare the
+    experimental structure and a computer generated model for CASP target T0538. (A) Two salt bridges between residues
+    E35-R32 and E28-K24 exist in the native structure. These are absent in the computer generated model structure.
+    Instead, (B) a new salt bridge is formed between residues R32-E28 in model. (C) Another salt bridge between residues
+    R26-E50 in the experimental structure. (D) In comparison to the native structure, orientation of F51 differs in the
+    model.}
 \label{fig:T0538compare}
@@
 A part of the CASP experiment is dedicated to the quality assessment (QA) of predicted models \cite{Kryshtafovych2011}.
-Here, predictors were asked to score each model on a scale (known as qmode) from 0 to 1, with higher values
-corresponding to better models \cite{Kryshtafovych2011}. Most of the groups use consensus strategies in such experiments.
-Here, we chose two computer generated models from CASP taget T0538, where the top performing group "MUFOLD-WQA" \cite{Wang2011} failed.
-One of the model was from "PconsR" (GDT-TS = 83, qmode 1 = 0.5865) and another from "MULTICOM-NOVEL" ( GDT-TS = 83, qmode 1 = 0.5865). We 
-choose these two models as the model by PconsR was the most accurate
-predicted model for this target, whereas the other model was predicted best by QA test of MUFOLD-WQA. The calculated
-free energy using confinement method indicate that the model by PconsR is more stable by $3.9
-Kcal/Mol$ which supports
-the GDT\_TS trend.  Although consensus prediction strategies do well specially when dealing with large numbers of structures, they 
-can miss a good answer that is only predicted by a few methods. At least in this case, the confinement
-method is able to capture this structure that was otherwise missed.
+Here, predictors were asked to produce an overall score (called QMODE1) for each model on a scale from 0 to 1, with
+higher values corresponding to better models \cite{Kryshtafovych2011}. Many of the groups use consensus strategies in
+such experiments. Here, we chose two computer generated models from CASP target T0538, where the top performing group
+``MUFOLD-WQA'' \cite{Wang2011} failed to identify the best model. We examined two models, one from ``PconsR'' (GDT-TS =
+96), which MUFOLD-WQA gave a QMODE1 = 0.54. The second model was from ``MULTICOM-NOVEL'' (GDT-TS=83), which is a poorer
+model, but MUFOLD-WQA assigned a higher QMODE1 = 0.59. The confinement method gives a difference free energy that favors
+the PconsR model by 3.9 kcal/mol, which correctly identifies the more accurate model. Although consensus prediction
+strategies are very effective, they can miss a good prediction that is only predicted by a few methods. At least in this
+case, the confinement method is able to capture this structure that was otherwise missed.
 
@@
 
-We have described a computational method called confinement for computing the difference free energy between two
-conformational ensembles. We show that the difference free energy can be calculated on ~100 residue sized proteins, even
-for large conformational changes. We have demonstrated that it can discriminate the folding preferences of a series of
-chameleon proteins. We show that the confinement method can discriminate between the native structure and structure
+We have described a computational method called the confinement method for computing the difference free energy between
+two conformational ensembles. We show that the difference free energy can be calculated on ~100 residue sized proteins,
+even for large conformational changes. We have demonstrated that it can discriminate the folding preferences of a series
+of chameleon proteins. We show that the confinement method can discriminate between the native structure and structure
 predictions and can identify the best prediction reliably. Perhaps most importantly, we have shown that it can be used
@@
 
-A key advantage is that this method does not require any reaction coordinate, or sampling a
-pathway from conformation $A$ to $B$.
-
 \section{Method}
 
-The confinement method has been described in details by Tyka et al. in reference \cite{Tyka2006} and
-Cecchini et  al. \cite{Cecchini2009}. The basic approach of the confinement approach is the same in
-both these papers. A thermodynamic cycle is used to compute the free energy between conformations A
-and B. However, there are some small technical differences between the two approaches. Here we briefly
-describe the procedure that we used.
+The confinement method has been described in detail by Tyka et al. in reference \cite{Tyka2006} and Cecchini et al.
+\cite{Cecchini2009}. The basic approach of the confinement approach is the same in both these papers. A thermodynamic
+cycle is used to compute the free energy between conformations A and B. However, there are some small technical
+differences between the two approaches. Here, we briefly describe the procedure that we used.
 
@@
     dynamics simulation (each 20 ns long) for each leg of the thermodynamic cycle, where the harmonic restraint force
-    constant was scaled from 0.00005 Kcal/Mol (mostly free) to 81.92 kcal/mol (tightly restrained). The free energy for
+    constant was scaled from 0.00005 kcal/mol (mostly free) to 81.92 kcal/mol (tightly restrained). The free energy for
     this step is estimated using a thermodynamic integration approach developed by Tyka et. al. \cite{Tyka2006}. The
@@
 \item The thermodynamic cycle is closed by calculating the free energy between the final restrained states $A^\ast$ and
-    $B^\ast$ using normal mode analysis \cite{Brooks1983,Case1994} or quasiharmonic analysis \cite{Karplus1981,Levy1984}. The free energy calculated in this way is shown as
-    $\Delta G_{A^\ast,B^\ast}$ in Figure~\ref{fig:method}.
+    $B^\ast$ using normal mode analysis \cite{Brooks1983,Case1994} or quasiharmonic analysis
+    \cite{Karplus1981,Levy1984}. The free energy calculated in this way is shown as $\Delta G_{A^\ast,B^\ast}$ in
+    Figure~\ref{fig:method}.
 
@@
 
-One advantage of confinement method is that
-none of the simulations during the restraining step depends on each other. Therefore, with available computer resources it can be fast 
-to compute. The computational cost for a single 20 ns simulation (along the confinement step) of $56$-residue protein 
-is only $4$ hours in a single gpu computer  \Alberto{all the different confinements in one GPU?}. All calculations were
-performed with the Amber 11 suit of programs~\cite{Case2012,Goetz2012} in combination with the ff99SB
-forcefield~\cite{Hornak2006} and the generalized born implicit solvent \cite{Mongan2006}.
+One advantage of confinement method is that none of the simulations during the restraining step depends on each other.
+Therefore, with available computer resources it can be fast to compute. We ran each confinement calculation on a single
+graphics processing unit (GPU). For a 56-residue protein, this leads to a calculation time of only 4 hours on 40 GPUs (1
+GPU per confinement calculation x 20 calculations per structure x 2 structures). All calculations were performed with
+the Amber 11 suit of programs~\cite{Case2012,Goetz2012} in combination with the ff99SB forcefield~\cite{Hornak2006} and
+the GBneck generalized born implicit solvent model \cite{Mongan2006}.
 
-Interestingly, we extend the
-method for calculation of per residue free energy in an approximate way. For this purpose, the confinement energy,
-$\Delta G_{A,A*}$ and $\Delta G_{B,B*}$ of each residue is calculated in the usual numerical way as described by
-Tyka et al. in reference \cite{Tyka2006}. The internal energy of each residue is calculated using
-amber's "decomp" module using the
-final restrained trajectory. We call this method approximate as we ignore the entropic contribution from the normal mode
-or quasiharmonic analysis. However, this contribution to the total free energy is much smaller, which allow us to study
-the mechanistic details of conformational preference of each residue.
+We calculated the approximate per-residue free energy as follows. The confinement energy, $\Delta G_{A,A*}$ and $\Delta
+G_{B,B*}$ of each residue is calculated in the usual numerical way as described by Tyka et al. \cite{Tyka2006}. The
+internal energy of each residue is calculated using amber's "decomp" module using the final restrained trajectory. We
+call this method approximate as we ignore the entropic contribution from the normal mode or quasiharmonic analysis.
+However, this contribution to the total free energy is much smaller, which allow us to study the mechanistic details of
+conformational preference of each residue.
 

### 2013-05-16 Paper/old/bib.tex

--- 2013-04-29 Paper/paper.tex
@@
-% vim:set textwidth=100:
-% vim:set fo+=t:
-
-\documentclass[12pt]{article}
-
-\usepackage{amsmath}
-\usepackage[nofiglist,notablist]{endfloat}
-\usepackage[usenames,dvipsnames]{color}
-\usepackage{color}
-\usepackage{authblk}
-\usepackage{graphicx}
-\usepackage{palatino}
-\usepackage[activate={true,nocompatibility},final]{microtype}
-\usepackage[super,sort&compress]{natbib}
-\pagenumbering{arabic}
-\parskip = 0.08in \parindent = 0.0in
-
-% Custom macros for author comments
-\newcommand{\Alberto}[1]{\color{ForestGreen}#1\normalcolor }
-\newcommand{\Justin}[1]{\color{blue}#1\normalcolor}
-\newcommand{\Arijit}[1]{\color{magenta}#1\normalcolor}
-\newcommand{\Ken}[1]{\color{red}#1\normalcolor}
-
-\author{Arijit~Roy}
-\author{Alberto~Perez}
-\author{Ken~A.~Dill}
-\author{Justin~L.~MacCallum}
-\affil{Laufer Center for Physical and Quantitative Biology\\
-    and Departments of Physics and Chemistry\\
-    Stony Brook University\\
-    Stony Brook, NY 11794-5252.}
-
-\title{Predicting the conformational preferences of proteins using a physics-based free energy
-method}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-
-Calculation of free energy differences is of central importance in the simulation of biochemical
-systems. It is particularly difficult to calculate between pairs of macromolecular conformations as
-well as a computationally expensive task with existing methods. In this work, confinement approach
-is used to calculate absolute free energies of biomolecular systems. This method provides two main
-advantages: it does not require a reaction coordinate or transition path and it is fast to compute.
-Free energy calculated can be decomposed into a per residue contribution in an approximate way. Per
-residue free energy allows us to identify the reason behind conformational preferences in
-biomolecules. Through out the article we show its use in different challenging modeling problems. In
-particular, we show its use in predicting the conformational preference of chamaleon sequences
-(sequences with high sequence identity and different folds). This sequence dependent conformational
-preferences and per residue free energy decomposition set the stage for the use of this method in
-protein design.
-
-\end{abstract}
-
-\Ken{[Guys---I recommend we get a better name than confinement.  It's totally uninformative.  In my view, better would
-    be `confine-and-configure' or something.  Please give it some thought.]}
-
-\Justin{[Ken---I agree that the name isn't great, but I don't think we should change it because: it's already in the
-    literature and we didn't come up with the method. I think we should consider capitalizing it as Confinement Method
-    throughout.]}
-
-\section{Introduction}
-
-In some problems of protein science, you want to know the relative stability of a protein's conformation $A$ compared to
-its conformation $B$.  We call this the \emph{difference free energy}. For example, in all esoteric mechanisms, a protein
-often adopts one conformation when a ligand is bound and another conformation when no ligand is bound~\cite{Elber2007}.
-To understand the mechanism requires knowledge of the relative free energies of the protein conformations.  Another
-example is in protein structure prediction, for example as practiced in the Critical Assessment of Structure Prediction
-(CASP) competition~\cite{Moult2011}. If you are using some computational model and have predicted two putative native
-structures, $A$ and $B$, you want to compute which has the lower free energy, in order to know which is the more
-native-like.  In a third example, you may want to know whether mutating a few amino acids in a protein could cause the
-protein to switch from one stable conformation to another, because that could have important consequences for biological
-mechanisms and disease. And fourth, sometimes binding a ligand induces a protein from one conformation to another. A
-quantitative understanding of such \emph{induced-fit} situations requires knowledge of the difference free energy of the
-protein in the two states. In all these cases, it would be useful to have a computational method that can be used
-efficiently with atomically detailed physical forcefield models to compute the free energy differences between different
-given protein conformations.
-
-\Justin{[TI is actually the basis of the confinement step. We need to be more explicit about what we mean here.] Methods
-    like thermodynamic integration~\cite{Tironi1994},\cite{Meirovitch2007} which are successful in alchemical free
-    energy calculation may not be as useful for calculation of conformational free energy.} One widely explored strategy
-is to use molecular dynamics simulations along some putative reaction coordinate pathway from conformation $A$ to
-$B$~\cite{Elber2007,West2007,Haas2007,Jonsson1998,E2007,Dellago2002,Cheng2006}. \Ken{[we need many refs here: cite
-    Elber's milestoning, Jhih-Wei Chu, and the many other path-sampling approaches.]} The free energy along this
-reaction coordinate can then be determined using methods such as umbrella sampling \cite{Torrie1977}\Justin{[add ref
-    where this is used computing free energy differences between conformations]} or targeted MD \Justin{[refs]}. Such
-approaches have several limitations. First, it is necessary to know an efficient reaction pathway from $A$ to $B$. If
-conformations $A$ and $B$ are quite different, then it can be challenging to find such paths. Second, these methods are
-computationally slow. To get an accurate estimate of the total free energy difference $\Delta G = G_B - G_A$ requires
-accurate determinations of the many small free energy differences $A \rightarrow 1 \rightarrow 2 \ldots \rightarrow B$,
-so each step requires substantial amounts of sampling.  And third, these methods are prone to large errors, because the
-pathway error is a sum of many errors among the many steps.
-
-\Ken{[Are all these methods below ones that don't involve paths?  Also, are there other key methods we should cite?  We
-    need a good scholarly list of all the relevant works.]}
-
-The calculation of protein conformational free energy has been successfully attempted by a number of groups
-\cite{Shell2010,Ytreberg2006,Zheng2008,Spichty2010,Strajbl2000,Park2008,Tyka2006,Cecchini2009,Ovchinnikov2013}. Some of
-these methods like the reference system method~\cite{Ytreberg2006}, deactivated morphing~\cite{Park2008}, the
-confinement method~\cite{Tyka2006,Cecchini2009,Ovchinnikov2013} take an alternative strategy for computing difference
-free energies that does not require knowing a pathway from $A$ to $B$.  \Ken{[please add the several new Karplus papers
-    from 2012-2013.]}.
-
-Here, we adopt the confinement method of Tyka et al~\cite{Tyka2006} and Cecchini et al.~\cite{Cecchini2009}, which is
-based on the thermodynamic cycle shown in Figure~\ref{fig:method}. We start with an ensemble of related conformations,
-$A$. We then restrict $A$ to a much ``tighter'' ensemble $A^\ast$ by applying position restraints in a series of MD
-simulations. Similarly, a second ensemble $B$ is restricted to $B^\ast$. We then compute the free energy between
-$A^\ast$ and $B^\ast$ using either normal mode analysis or the quasiharmonic method [ref]. This step takes into account
-both the remaining conformational entropy in each ensemble and the remaining enthalpy difference. In this way, we can
-calculate the free energy difference between the two end states without needing to define a physical path or reaction
-coordinate connecting the two states. The confinement method shares some similarities with the `confine-and-release'
-method for computing ligand binding affinities~\cite{Mobley2007,Mobley2012,Mobley2006} \Ken{[cite mobley's jctc 07 (ref
-    216 on our labsite), and mobley's recent j chem phys review, and mobley (ref 199) j chem phys, and others that are
-    relevant.]}.
-
-\begin{figure}
-    \begin{center}
-        \includegraphics[width=3.5 in]{method.pdf}\label{fig:method}
-    \end{center}
-    \caption{Graphical representation of the thermodynamic cycle involving confinement method.}
-\end{figure}
-
-Previously, the confinement method has been validated on small model peptides~\cite{Tyka2006,Cecchini2009}. Here, we
-show that the confinement method can also be applied successfully to larger systems, such as the series of chameleon
-proteins designed by Orban and co-workers~\cite{Alexander2007,He2008,Alexander2009,He2012}, which can switch between two
-completely different folds with only slight changes in sequence. We also show that our computed difference free energies
-are useful in evaluating the quality of CASP target protein predictions.  This could be of significant value for
-ultimately improving the energetics in protein-structure models, not just the structures. And, finally, we show that we
-can approximately decompose difference free energies into individual amino acid components. This offers the opportunity
-for diagnosing the structural basis for difference free energies, which can be useful for interpreting biological
-mechanisms.
-
-\section{Results and Discussion}
-
-\subsection{The confinement method succeeds at some basic consistency checks}
-
-First, we validated that our implementation of the confinement method produces results compatible with previous
-calculations reported in the literature.  The method has previously been applied to a 16 amino acid residue
-$\beta$-hairpin from protein G, known as BHP~\cite{Cecchini2009}. We calculated the free energy difference between the
-native conformation---called bhp1, which has a two-stranded $\beta$-sheet---and a non-native conformation---called bhp3,
-which has a three-stranded $\beta$-sheet. Our confinement calculation shows that bhp1 is more stable by 1.7 kcal/mol,
-consistent with 4 $\mu$s equilibrium simulations showing that bhp1 is more favorable configuration by 1.8 kcal/mol, and
-in agreement with previous calculations~\cite{Cecchini2009}.
-
-Second, we looked at 6 target proteins from the CASP9 experiment [ref].  For each target, we examined up to 5 submitted
-models.  We computed the difference free energy between the true native and the best model.
-Figure~\ref{fig:summary_casp} and Table~\ref{table:casp_control} show that, in 5 out of 6 cases, the confinement method
-assigns a lower free energy to the experimentally determined structure than to any of the decoys.  Other well-known
-discriminators can also successfully tell native structures from computer generated models~\cite{Sheffler2009} [add
-rosetta holes, rosetta ff, dfire]; it is just an independent useful validation that the confinement calculations make
-sense.
-
-\Justin{[Arijit, please add GDT of best decoy to this table.]}
-
-\begin{table}
-\begin{center}
-\caption{The confinement method assigns a more favorable free energy to the experimentally determined structure than to
-    computer-generated predictions. For each target, we examined as many as five predictions submitted by CASP
-    participants. We report the free energy difference between the most favorable decoy and the experimentally
-    determined structure. Positive $\Delta\Delta G$ values indicate that the experimental structure is predicted to be
-    more favorable than any of the decoys.}
-\label{table:casp_control}
-\begin{tabular}{l l l}\hline
-    CASP Target  & PDB Identifier & $\Delta \Delta G = G_{best~decoy} - G_{native}$ (kcal/mol) \\ \hline
-     T0531       &    2KJX        &          $11.15 \pm 0.70$ \\ \hline
-     T0538       &    2L09        &          $-3.00 \pm 0.47$ \\ \hline
-     T0540       &    3MX7        &          $16.94 \pm 0.49$ \\ \hline
-     T0559       &    2L01        &          $2.10 \pm 0.24$ \\ \hline
-     T0560       &    2L02        &          $18.00 \pm 0.49$ \\ \hline
-     T0569       &    2KWY        &          $20.01 \pm 0.69$  \\ \hline
-\end{tabular}
-\end{center}
-\end{table}
-
-\subsection{The confinement method correctly predicts the structures of chameleon sequences}
-
-We also tested the confinement method predictions for difference free energies on the chameleon sequences of Orban et
-al.!\cite{Alexander2007,He2008,Alexander2009,Bryan2010,He2012,Shortle2009}. These are instances in which two highly
-similar sequences fold into remarkably different structures. Orban and co-workers have designed a protein-G-like
-sequence of 56-residues that is marginally stable in one of two possible folds. By mutating key residues in this
-sequence they are able to stabilize one fold or the other (see Figure~\ref{fig:orban}). We refer to the $4\beta +
-\alpha$ structure as the $\beta$ conformation, and we refer to the $3\alpha$ structure as the $\alpha$ conformation. One
-pair of sequences (GA88/GB88) is 88 percent identical in sequence, differing in seven positions. Another pair
-(GA95/GB95) is 95 percent identical, differing in three positions. Accurately predicting the structural preferences of
-these structures presents a serious challenge for computational
-methods~\cite{Alexander2007,He2008,Alexander2009,Bryan2010,He2012,Shortle2009}. \Justin{Do any of these citations try to
-    do any modeling or comment about how challenging this would be?}
-
-\begin{figure}
-    \begin{center}
-    \includegraphics[width=5.0in]{orban.pdf}
-        \label{fig:orban}
-    \end{center}
-    \caption{The confinement method correctly predicts the structural preferences of four chameleon sequences. The top
-        part of the figure represent four sequences used in this study along with the protein data bank identifier. The
-        experimentally observed fold, computationally predicted fold, and difference free energy between the two folds
-        is reported for each sequence.}
-\end{figure}
-
-We initially approached this problem by making a model of each sequence with the same backbone structure as its partner
-chameleon sequence. For example, we took the sequence of GA88 and built a model with the same overall structure as GB88.
-We then used the confinement method to assess the free energy difference between the experimentally determined structure
-of GA88 and the model (with the GA88 sequence and the GB88 structure). The confinement method was able to predict the
-conformational preferences correctly for all four sequences (data not shown). This is, however, not surprising as it is
-often easy to distinguish computational models from native structures~\cite{Handl2009,Sheffler2009}. To avoid this
-potential problem, we instead computed relative free energy of two different computer generated models for each
-sequence. One model is based on the $\alpha$ structure and the other on the $\beta$ structure (see Supporting
-Information for details on the modeling procedure). This is a much more realistic test of the confinement method's
-ability to accurately calculate the difference free energy.
-
-\Ken{[ It's not clear what this test is and why it's better, and why there are only 5 sequences.  Please clarify.]}
-\Justin{[There are five pairs of sequences that can be arranged into 3 different pairings. Arijit has removed the 5th
-    sequence.]}
-
-\Arijit{ Now there are 4 sequences as they have the experimentally available structure. The fifth sequence is discussed
-    at the end of the paragraph.}
-
-\Justin{[Ken: it's not that hard to tell an experimental structure from a computer generated model. Almost all of the
-    knowledge-based scoring functions can do this a substantial fraction of the time. Comparing a near-native model with
-    another model is a much more challenging problem.]}
-
-Figure~\ref{fig:orban} shows that the confinement method identifies the correct structure for all four sequences. One
-hypothesis is that such structural transitions require states with diminished stability~\cite{Bryan2010}. It is believed
-that if the free energy of the native state and the alternative state are within a range of around 5 kcal/mol, then it
-is possible for the native state to be destabilized relative to the alternative state with only small changes in
-sequence~\cite{He2008,Alexander2009,Bryan2010}. The calculated free energy differences range from around 3.5 to 5.0
-kcal/mol, which is consistent with this hypothesis. In a more recent study\cite{He2012}, the amino acid residue at
-position 45 (Tyr for $\beta$ and Leu for $\alpha$) was found to be important for switching between $\alpha$ and $\beta$
-conformation. This inspired us to introduce another mutation at this position, Y45A, which we refer to as GA98. Our
-calculations predict that this mutation shifts the equilibrium to the $\alpha$ conformation, which is now more stable
-than the $\beta$ conformation by 3.8 kcal/mol. Although this result has not yet been confirmed experimentally, it is
-consistent with the previously observed effect of Y45L.
-
-\subsection{Per-residue free energy calculations can identify the mechanistic detail behind conformational preferences}
-\label{sec:orban_per_residue}
-
-To better understand the mechanism behind these chameleon proteins, we decomposed the calculated free energy into
-per-residue contributions in an approximate way~\cite{Tyka2006}. We decompose each confinement step ($\Delta
-G_{A,A^\ast}$, $\Delta G_{B,B^\ast}$) into per-residue contributions. We can also decompose the remaining enthalpy in
-the confined ensemble into per-residue contributions. However, the method is approximate because we do not include the
-residual conformational entropy from the normal mode or quasiharmonic analysis steps. Computing per-residue
-contributions helps us to identify important residues that stabilize a particular conformation. The per residue free
-energy, $\Delta \Delta G ((\beta) - (\alpha))$ is shown in Figure~\ref{fig:perresidue_orban}.
-
-\begin{figure}
-    \begin{center}
-        \includegraphics[width=4.5in]{delta_g.pdf}
-        \label{fig:perresidue_orban}
-    \end{center}
-    \caption{Per-residue difference free energies for the $\alpha$ and $\beta$ conformations of the GA95 and GB95
-        sequences. Residues colored in blue favor the $\alpha$ structure, residues in red favor the $\beta$ structure,
-        and white residues have no preference.}
-\end{figure}
-
-Although the overall free energy difference between the two structures is small (within 3--5 kcal/mol), individual
-residues can show marked preferences for being in either $\alpha$ or $\beta$ conformation. Such differences can be
-understood by looking in detail at the local environment for those residues. For example, the region around residue 7
-forms a random coil in the $\alpha$ structure, whereas it is forming a beta sheet structure in the $\beta$ structure.
-These residues strongly favour the locally well packed and hydrogen bonded environment found in the $\beta$ sheet.
-Overall, favoring $\alpha$ or $\beta$ structure is a delicate balance, were the relative global free energy difference
-is small and the contributions of different per residue tendencies balance out. It is therefore very likely that by
-changing key residue preferences the global fold preferences can be changed.
-
-Three residues are mutated between GA95 and GB95, at positions 20, 30, and 45. In GA95, these residues (L20, I30 and
-L45) stabilize the $\alpha$ structure (compare the upper and lower panels on Figure~\ref{fig:perresidue_orban}). In
-GB95, two of these residues (F30 and Y45) favor the $\beta$ structure, because they have large solvent exposed surface
-areas in the $\alpha$ structure, but are more buried in the $\beta$ structure. Additionally Y45 forms a hydrogen bond
-with D47 in the $\beta$ structure. On the other hand, residue A20 from GB95 still favors the $\alpha$ structure,
-although less strongly than L20 in GA95.
-
-The $\alpha$ and $\beta$ sequences are nearly identical, so there are some common features observed for all sequences.
-The experimental observations classified the protein into two parts: Amino acids 9--51 are fully structured in both
-folds, whereas residues 1--8 and 52--56 are unstructured in the $\alpha$ fold, but form $\beta$-strands in the $\beta$
-fold. Most of the amino acid residues in the region 1--9 have negative per-residue free energies, which means that these
-residues favor the $\beta$ structure. Roles of some other important residues in stabilizing either the $\alpha$ or
-$\beta$ conformation are summarized in Figure S1.
-
-In addition to the direct effects of the mutation, there are also indirect effects due to small perturbations in the
-environment around the mutations. For example, the L20A mutation causes a slight repacking around residue 20. This
-causes large changes in the per-residue free energies of nearby residues T25 and A26. However, the changes for these two
-residues have opposite signs and nearly cancel.
-
-These per-residue free energy decompositions provide a great deal of insight into the driving forces behind protein
-folding and conformational change. We believe that such calculations may also be useful for protein design---designing
-proteins with specific structures and functions.
-
-\subsection{The confinement method is a useful tool for structure prediction}
-
-We have tested the ability of the confinement method to act as a ``meta-predictor'' for structure prediction. Here, the
-task is to correctly identify the most accurate models out of a set of ``decoys'' generated by different methods during
-the CASP experiment. CASP is a blind test in which different groups apply methods to predict the 3-dimensional structure
-of proteins from their sequences. Each group is allowed to submit five possible structures, which they are supposed to
-rank from best to worst.
-
-We have performed two experiments centered on CASP. In the first experiment, we tried to rank-order predictions for
-several targets. For each target, the predictions were either produced by a single group---presumably using the same
-method for each prediction, or were produced by several different groups---using different methods. The goal of this
-experiment is to determine if the physics-based confinement method can correctly identify more native-like structures as
-having lower free energies. Our second experiment was to see if the confinement method can identify structures that are
-missed by other meta-prediction servers. Most successful meta-prediction servers are based on the idea of consensus: if
-many different prediction methods produce similar results, then that is probably a correct prediction [ref]. This is
-often a powerful heuristic, but it can miss cases where there is a very good result that is only predicted by one
-method. We chose several cases where such structures were missed by the best meta-predictors in CASP and assessed if the
-confinement method can correctly identify these accurate models.
-
-As is common in the CASP experiment, we assess our results in terms of Global Distance Test Total Score (GDT-TS)
-\cite{Zemla2003}, which is a C$\alpha$ based measure of structural accuracy. It can be understood roughly as the
-percentage of residues that are correctly positioned in the model (range 0 to 100, higher is better). We did have enough
-computer model to analyze every model, so the initial models for the test were chosen from a selection of different
-server groups that have done well in past CASP events.
-
-Overall, the confinement method performs well on our CASP tests. Figure~\ref{fig:summary_casp} shows that in almost
-every case, the native structure has the lowest free energy and the best model has the next lowest free energy. The
-confinement method appears to be a useful tool for ranking structure predictions.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.5 in]{casp.pdf}
-\end{center}
-\caption{The confinement method is usually able to identify the native structure and the best model (the model with the
-    highest GDT-TS) from a set of decoys.}
-\label{fig:summary_casp}
-\end{figure}
-
-\subsubsection{The confinement method can correctly rank-order structure predictions}
-
-First, we examined the ability of the confinement method to rank different models for a target that have been generated
-by a single group using the same methodology for all predictions. We examined two targets: T0559 and T0560 (see Table
-?).
-
-\Justin{[Arijit: Let's move the protein names, organisms, pdb ids, etc to a separate table and just use the CASP ids.]}
-
-The first test case is target T0559. The best predictor group for this 69 amino acid target was ``BAKER-ROSETTASERVER''.
-To save computer time, we excluded two models that were very similar to other models that we did include. For this
-target, the difference free energy computed by the confinement method can be used to accurately rank-order all of the
-models and the native structure (Figure~\ref{fig:T0559}).
-
-\Ken{Arijit- we could probably make a better visualization of this figure above by putting it onto a toy landscape:
-    x-axis is rmsd, y-axis is DDG, and above each of the 4 points on this graph will be the ribbon diagram and the
-    GDT-TS.  I suggest we do that with the fig below too.  Let's don't use 4 different colors for the fig below.  Let's
-    just show differences in structures in some different color, or show them all as green.]}
-
-\Justin{Ken: I don't mind the 3d plots that are in here now, but I'm not sure I would like something that looks like a
-    free energy landscape.}
-
-\begin{figure}
-    \begin{center}
-        \includegraphics[width=3.8 in]{T0559.pdf}
-        \label{fig:T0559}
-    \end{center}
-    \caption{The confinement method can correctly rank the native structure and three predictions produced by a single
-        method.}
-\end{figure}
-
-\Justin{I'm not sure we need a second figure of this. Move it to the SI?}
-
-We performed a similar calculation for target T0560 with two models form the group ``Splicer''. The remaining three
-models were discarded as they were too similar to the rest of the models. Again, we can correctly identify the native
-sate and our calculated free energy based ranking matches well with GDT-TS (Figures S?).
-
-\begin{figure}
-    \begin{center}
-        \includegraphics[width=3.5 in]{T0560.pdf}
-    \end{center}
-    \caption{Native structure and two models (from group ``Splicer'') for target T0560.}
-    \label{fig:T0560}
-\end{figure}
-
-Next, we tested the ability of the confinement method to rank models for target T0540 that were produced by different
-prediction groups. The top models from groups ``LTB'' (Model 1) and ``Mufold'' (Model 2) were chosen for analysis. Once
-again, the free energy based ranking correlates well with the GDT-TS based score (Figure~\ref{fig:T0540}).
-
-\begin{figure}
-    \begin{center}
-        \includegraphics[width=3.5 in]{T0540.pdf}
-    \end{center}
-    \caption{The confinement method can correctly rank the native structure and two models submitted by different
-        prediction groups.}
-    \label{fig:T0540}
-\end{figure}
-
-\subsubsection{The per-residue free energy is sensitive to small changes in protein conformation}
-
-In section~\protect\ref{sec:orban_per_residue} we discussed how the per-residue free energy can reveal mechanistic
-detail behind the conformational preference of a chameleon sequence. In this section, our aim is to apply the same
-method and try to understand if the per-residue free energy can help us identify residues which stabilize or destabilize
-a particular region of a protein. In this case, the two conformations have similar folds with only small changes in
-localized regions. We chose CASP target T0569 and compared the experimental NMR structure with the best predicted model
-(GDT-TS=78; predicted by the ``Mufold'' group). The confinement method predicts that the native structure is more stable
-by 20 kcal/mol (Figure S2).
-
-The confinement method clearly identifies two hydrophobic residues V59 and I61, which destabilize the predicted model
-with respect to the crystallographic structure (Figure~\ref{fig:T0569_per_residue} and Figure S3). The sidechains of
-these hydrophobic residues are oriented towards the protein hydrophobic core in the native NMR structure but oriented
-towards the exterior of the protein and are solvent exposed in the model. These residues also form part of a beta-sheet
-in the experimental structure, but do not form inter-strand hydrogen bonds in the predicted model. There is also a large
-difference around K76, which forms a salt-bridge with D11 in the predicted model, but not in the experimental structure.
-This suggests that salt-bridge interactions are too favorable for the combination of force field and implicit solvent
-model we use, which has been a problem noted in the past [ref carlos].
-
-\Justin{[Arijit: Figure~\ref{fig:T0569_per_residue} says crystal, but the text says NMR.]}
-
-\begin{figure}
-    \begin{center}
-        \includegraphics[width=3.5in]{T0569_perres1.pdf}
-    \end{center}
-    \caption{Per-residue difference free energy between the experimental NMR structure and the best prediction for CASP
-        target T0569. The amino acid residues that are colored in deep red and deep blue stabilizes the NMR structure
-        and the prediction, respectively; the residues with light blue color does not have a strong preference.}
-    \label{fig:T0569_per_residue}
-\end{figure}
-
-\subsubsection{The confinement method occasionally produces incorrect results}
-
-\Justin{Arijit: let's move Figure~\ref{fig:T0538} to the supporting info and move Figure S4 into the main text.}
-
-Despite success for most of the studied systems, there are few failures, specially when the GDT-TS scores of the
-compared structures are very close. One example is Target T0538, where we compared the crystal structure with three
-models (Model 1: ``PconsR''---GDT-ts=96; Model 2: ``Shell''---GDT-TS=90; Model 3: ``FOLDIT''---GDT-TS=86). Contrary to
-our expectation, the confinement method predicts that Model 1 is more stable than the crystal structure (Figure ?).
-Per-residue free energy calculations (not shown) show that despite only small variations at the backbone level, the
-sidechains are oriented in very different ways (Figure ?), giving rise to large differences in the stabilization of
-certain residues. In particular, some of the differences arise from different salt bridge patterns and certain flexible
-polar residues exposed to the surface. This unexpected result shows that the confinement method is very sensitive to
-local interactions (including sidechain reorientation) and may indicate issues with the forcefield and implicit solvent
-models used in the calculation.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.5in]{T0538.pdf}
-\end{center}
-\caption{The native and three model structure of engineered protein from Asr4154 protein (PDB ID: 2L09 and CASP
-    code:T0538). The model 1,2 and 3 are from the group PconsR, Shell and FOLDIT respectively.}
-\label{fig:T0538}
-\end{figure}
-
-\subsubsection{What can we say about low resolution models?}
-
-\Justin{I think we should consider dropping this whole section. We have a sample size of one and I don't expect that we
-    can generally do well ranking models with a GDT of 44. I think this is a fluke.}
-
-So far we have seen that the method is good at predicting preferences when the structures are not very far from the
-native. But the question remains how far from native can we go and still see that the method produces correct result. In
-this section we explored this question with models from  extracellular domain of the jumping translocation breakpoint
-protein (pdb id: 2KJX). Most of the group could only generate low resolution models for this CASP9 target (id: T0531) .
-In our comparison, we choose five models by the group MUFOLD-MD, which was the best performing group for this target
-with their best model had a GDT\_TS value of 44. The result presented in Figure~\ref{fig:T0531} shows: 1.) native is
-correctly identified as expected and 2.) Surprisingly there is a high level of correlation between the GDT ranking and
-the free energy ranking for model 1 and model 3, the rest three structures with GDT\_TS score less than 35  are ordered
-incorrectly. It is worth to note that models 2 and 3 have the same GDT and very different free energies, meaning that
-the actual ordering could change a lot \cite{Perez2012}. It is encouraging that at least the method can pick out the
-best model even though it is got a low GDT score: 44.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.5 in]{T0531.pdf}
-\end{center}
-\caption{The native structure and 5 models of extracellular domain of the jumping translocation breakpoint protein (pdb
-    id: 2KJX and the CASP code: T0531).}
-\label{fig:T0531}
-\end{figure}
-
-\subsubsection{Can the confinement method perform better quality assessment in protein structure prediction?}
-
-\Justin{Arijit---is this the case where we chose a structure that was missed by the consensus predictors, if so, we
-    should try to indicate this. The point of this section should be that although consensus prediction strategies do
-    well, they can miss a good answer that is only predicted by a few methods. At least in this case, the confinement
-    method is able to capture this structure that was otherwise missed.}
-
-A part of the CASP experiment is dedicated to the quality assessment (QA) of predicted models \cite{Kryshtafovych2011}.
-Here, predictors were asked to score each model on a scale (known as qmode) from 0 to 1, with higher values
-corresponding to better models \cite{Kryshtafovych2011}. It will be interesting to know how confinement method perform
-in quality assessment compare to the other groups in CASP9. We investigated this using couple of CASP Targets. Here, we
-present a case, where confinement method perform well than the top performing group MUFOLD-WQA \cite{Wang2011} of CASP9.
-We choose  two models from CASP taget T0538. They were model 3 submitted by PconsR (GDT\_TS = 96, qmode 1 = 0.5434) and
-model 5 from the MULTICOM-NOVEL ( GDT\_TS = 83, qmode 1 = 0.5865). Both are server predicted models and the qmode value
-presented are from MUFOLD-WQA \cite{Wang2011}. We choose these two models as the model by PconsR was the most accurate
-predicted model for this target, whereas the other model was predicted best by QA test of MUFOLD-WQA. The calculated
-free energy using confinement method indicate that the model by PconsR is more stable by $3.9 Kcal/Mol$ which support
-the GDT\_TS trend. There is no doubt that the consensus approach can predict the model quality in a faster manner. But
-we expect confinement method can predict the model quality in a relatively expensive but much more accurate way.
-
-\section{Conclusion}
-
-We have described a computational method called confinement for computing the difference free energy between two
-conformational ensembles. We show that the difference free energy can be calculated on ~100 residue sized proteins, even
-for large conformational changes. We have demonstrated that it can discriminate the folding preferences of a series of
-chameleon proteins. We show that the confinement method can discriminate between the native structure and structure
-predictions and can identify the best prediction reliably. Perhaps most importantly, we have shown that it can be used
-to give residue-level insights into what are the dominant structural factors in a protein that are responsible for the
-difference free energies. The confinement should be useful for protein design, structure prediction, and understanding
-the mechanism of conformational change.
-
-A key advantage is that this method does not require any reaction coordinate, or sampling a
-pathway from conformation $A$ to $B$. We have tested the method for structures of proteins having up to 100 amino acids.
-The computational cost for a 56-residue protein is about 4 hours [on 1 gpu] for 20 ns of confinement [Is 20ns all you
-need?].
-
-\Arijit{I need 21 (confinement run) x 20 ns for single confinement}
-%There is another issue, which is normal mode analysis calculations are Oh(N**3) and require large amounts of memory,
-%effectively limiting the size of the systems under study. In most cases quasiharmonic analysis.
-
-\Ken{[If this is a relevant limitation, let's say something about it.]}
-
-\Arijit{I think we need to write the normal mode issue and number of residues that we can handle. If one of you
-    contribute here that will be great.}
-
-\Justin{The conclusion is not the place to talk about how long the method takes or what the technical challenges are.
-    Let's move this to methods section.}
-
-\section{Method}
-
-The confinement method has been described in details in ref. by Tyka et al. \cite{Tyka2006} and
-Cecchini et  al. \cite{Cecchini2009}. The basic approach of the confinement approach is the same in
-both these papers. A thermodynamic cycle is used to compute the free energy between conformations A
-and B. However, there are some small technical differences between the two approaches. Here we briefly
-describe the procedure that we used.
-
-\begin{enumerate}
-
-\item  In the first step, a minimization of A and B conformations are performed. These minimized conformations
-    ($A^{\ast\ast}$ and $B^{\ast\ast}$) are the reference conformation of that state.
-
-\item The free energy of confining the ensemble (A or B) to a tighter ensemble ($A^\ast$ or $B^\ast$) is calculated.
-    This is done by gradually confining each atom to its position in the reference conformation ($A^{\ast\ast}$ or
-    $B^{\ast\ast}$) using a series of progressively stronger position restraints. This is done by running 21 molecular
-    dynamics simulation (each 20 ns long) for each leg of the thermodynamic cycle, where the harmonic restraint force
-    constant was scaled from 0.00005 Kcal/Mol (mostly free) to 81.92 kcal/mol (tightly restrained). The free energy for
-    this step is estimated using a thermodynamic integration approach developed by Tyka et. al. \cite{Tyka2006}. The
-    confinement free energies calculated this step are denoted as $\Delta G_{A,A^\ast}$ and $\Delta G_{B,B^\ast}$ in
-    Figure~\ref{fig:method}.
-
-\item The thermodynamic cycle is closed by calculating the free energy between the final restrained states $A^\ast$ and
-    $B^\ast$ using normal mode analysis or quasiharmonic analysis. The free energy calculated in this way is shown as
-    $\Delta G_{A^\ast,B^\ast}$ in Figure~\ref{fig:method}.
-
-\item  The full free energy, $\Delta G_{A,B}$ between the two state A and B is calculated as
-    $\Delta G{A,B} = \Delta G{A,A^\ast} - \Delta G{B,B^\ast} + \Delta G{A^\ast,B^\ast}$.
-
-\end{enumerate}
-
-All calculations where performed with the Amber 11 suit of programs~\cite{Case2012,Goetz2012} in combination with ff99SB
-forcefield~\cite{Hornak2006} and generalized born implicit solvent \cite{Mongan2006}. Interestingly, we extend the
-method for calculation of per residue free energy in an approximate way. For this purpose, the confinement energy,
-$\Delta G_{A,A*}$ and $\Delta G_{B,B*}$ of each residue is calculated in the usual numerical way as described in ref. by
-Tyka et al. \cite{Tyka2006}. The internal energy of each residue is calculated using the decomp module of amber from the
-final restrained trajectory. We call this method approximate as we ignore the entropic contribution from the normal mode
-or quasiharmonic analysis. However, this contrbution to the total free energy is much smaller, which allow us to study
-the mechanistic details of conformational preference of each residue.
-
-%There are two big bottlenecks in this calculation. The first has to do with the amount of simulations that have to be
-%performed to confine the ensemble of conformations close to a microstate into a single microstate, the second
-%bottleneck has to do with the size of the system in the normal mode calculation. We have tackled the first issue by
-%using GPU (Graphical Processing Unit) technology as opposed to the classical CPUs, which gives us a two order of
-%magnitude boost in computational efficiency. Normal mode analysis calculations are Oh(N**3) and require large amounts
-%of memory, effectively limiting the size of the systems under study. Where as this can be overcome by larger memory
-%machines, these are often not available, and so we propose an alternative by using quasiharmonic analysis. This is
-%still an Oh(N**3) problem, but we now the motions of atoms are already determined in trajectories, and so we can choose
-%to do quasiharmonic analysis excluding hydrogens, which will reduce the number of atoms N to almost N/2. This means
-%that we can use this method for systems twice as large.
-
 \begin{thebibliography}{99}
@@
 
-\bibitem{Handl2009}
-Handl, J.; Knowles, J.; Lovel, S.C. Artefacts and biases affecting the evaluation of scoring
-functions on decoy sets for protein structure prediction. Bioinformatics, 2009, 25, 1271-1279.
-
 \bibitem{Dill1997}
@@
 \bibitem{Zheng2008}
-Zheng, L.; Chen, M.; Yang, W. Random walk in orthogonal space to achieve efficient free-energy simulation of complex systems, Proc. Natl. Acad. Sci. 200
-8, 105 (51), 20227.
+Zheng, L.; Chen, M.; Yang, W. Random walk in orthogonal space to achieve efficient free-energy simulation of complex systems, Proc. Natl. Acad. Sci. 2008, 105 (51), 20227.
+
+\bibitem{Christ2007}
+Christ, C.D.; van Gunsteren, W.F. Enveloping distribution sampling: A method to calculate free energy differences from a 
+single simulation. 2007, J. Chem. Phys. 126, 184110.
 
@@
 
-\bibitem{Spichty2010}Spichty, M.; Cecchini, M.; Karplus, M. Conformational Free-Energy Difference of a Miniprotein from Nonequilibrium
+\bibitem{Spichty2010}
+Spichty, M.; Cecchini, M.; Karplus, M. Conformational Free-Energy Difference of a Miniprotein from Nonequilibrium 
 Simulations. J. Phys. Chem. Lett., 2010, 1, 1922-1926.
@@
 
-\bibitem{Alexander2007}Alexander, P.A.; He, Y.; Chen, Y.; Orban, J.
-Bryan, P.
-The design and characterization of two proteins with $88 \%$ sequence identity but differentstructure and function.
-Proc.
-Natl.
-Acad.
-Sci.
-2007, 104 (29), 11963-11968.
+\bibitem{Alexander2007}
+Alexander, P.A.; He, Y.; Chen, Y.; Orban, J. Bryan, P. The design and characterization of two proteins with $88 \%$ sequence identity but different 
+structure and function. Proc. Natl. Acad. Sci. 2007, 104 (29), 11963-11968.
 
@@
 \bibitem{Alexander2009}
-Alexander, P.A.; He, Y.; Chen, Y.; Orban, J. Bryan, P. A minimal sequence code for switching protein structure and function. Proc. Natl. Acad. Sci. 2009
-, 106(50), 21149-21154.
+Alexander, P.A.; He, Y.; Chen, Y.; Orban, J. Bryan, P. A minimal sequence code for switching protein structure and function. Proc. Natl. Acad. Sci. 2009, 106(50), 21149-21154.
 
@@
 \bibitem{He2012}
-He, Y.; Chen, Y.; Alexander, P.A.; Bryan, P.N.; Orban, J. Mutational tipping points for switching protein folds and functions. Structure. 2012, 20(2), 2
-83-91.
+He, Y.; Chen, Y.; Alexander, P.A.; Bryan, P.N.; Orban, J. Mutational tipping points for switching protein folds and functions. Structure. 2012, 20(2), 283-91.
 
@@
 \bibitem{Sheffler2009}
-Sheffler, W.; Baker, D. RosettaHoles: Rapid assessment of protein core packing for structure prediction, refinement, design, and validation. Protein Sci
-ence. 2009, 18(1), 229-239.
+Sheffler, W.; Baker, D. RosettaHoles: Rapid assessment of protein core packing for structure prediction, refinement, design, and validation. Protein Science. 2009, 18(1), 229-239.
 
@@
 Wang, Q.; Vantasin, K.; Xu, D.; Shang, Y. MUFOLD-WQA: A new selective consensus method for quality assessment in protein structure prediction. 
-Proteins, 2011, 79: 185-195.
+Proteins, 2011, 79: 185–195.
 
@@
 
-\bibitem{Hornak2006}
-Hornak, V.; Abel, R.; Okur, A.; Strockbine, B.; Roitberg, A.; Simmerling, C. Comparison of multiple Amber force fields
-and development of improved protein backbone parameters. Proteins., 2006, 65, 712-725.
-
-\bibitem{Mongan2006}
-Mongan, J.; Simmerling, C.; A. McCammon, J.; A. Case, D.; Onufriev, A. Generalized
-Born with a simple, robust molecular volume correction. J. Chem. Theory Comput.,
-2006, 3, 156-169.
-
 \end{thebibliography}

### 2013-05-16 Paper/old/paper.tex

--- 2013-05-12 Paper/paper.tex
@@
 
-Understanding the free energy differences between different different conformations of a protein is important for
-understanding mechanism in biochemistry. The calculation of such conformational free energy differences is difficult and
-computational demanding. In this work, we explore the use of ``the confinement method'' to calculate free energy
-differences between protein conformations. This method provides two main advantages: (1) it does not require a reaction
-coordinate or transition path and (2) it is fast to compute. The free energy difference can be approximately composed
-into per-residue contributions, which provides insight into the physical basis for the preference of one conformation
+Understanding the free energy differences between different conformations of a protein is important for
+understanding biochemical mechanisms. The calculation of such free energy differences is difficult and
+computationally demanding. In this work, we explore the use of the "confinement method'' to calculate free energy
+differences between conformations. This method provides two main advantages: (1) it does not require a reaction
+coordinate or transition path, and (2) it is fast to compute. The free energy difference can be approximately composed
+into per-residue contributions, providing insight into the physical basis for the preference of one conformation
 over another. We show that the confinement method can correctly predict the conformational preferences of chameleon
-sequences (nearly identical sequences with different structures), can identify the most native-like structures produced
-by structure prediction algorithms, and can yield detailed insight into the mechanisms behind these preferences.
+sequences (nearly identical sequences with different structures); can identify the most native-like structures produced
+by structure prediction algorithms; and can yield detailed insight into the mechanisms behind these preferences.
 
@@
 
+%Can you briefly identify how the fourth item differs from the first in the next para? Maybe put them together as alternative mechanisms?
+%\Arijit{I have removed the line: Fourth, sometimes binding a ligand induces a protein to switch from one conformation to another.}
+
 In some problems of protein science, we want to know the relative stability of a protein's conformation $A$ compared to
@@
 bound~\cite{Elber2007}. To understand the mechanism requires knowledge of the relative free energies of the different
-protein conformations.  Another example is in protein structure prediction, such as is in the Critical Assessment of
-Structure Prediction (CASP) competition~\cite{Moult2011}. If you are using some computational model and have predicted
+protein conformations.  Another example is in protein structure prediction, as in the Critical Assessment of
+Structure Prediction (CASP) competition~\cite{Moult2011}. If you have used a computational model to predict
 two putative native structures, $A$ and $B$, you want to compute which has the lower free energy, in order to know which
-is the more native-like.  In a third example, you may want to know whether mutating a few amino acids in a protein could
-cause the protein to switch from one stable conformation to another, because that could have important consequences for
-biological mechanisms and disease. Fourth, sometimes binding a ligand induces a protein from one conformation to
-another. A quantitative understanding of such \emph{induced-fit} situations requires knowledge of the difference free
-energy of the protein in the two states. In all these cases, it would be useful to have a computational method that can
-be used efficiently with atomically detailed physical forcefield models to compute the free energy differences between
-different given protein conformations.
+is more native-like. Third, you may want to know if mutating a few amino acids in a protein could
+cause a protein to switch from one stable conformation to another, because that could have important consequences for
+biological mechanisms and disease. A quantitative understanding of such change of conformation requires knowledge of the different free
+energies of the protein in the two states. In all these cases, it would be useful to have a computational method that can
+be used efficiently with atomically detailed physical forcefield models to compute the difference free energy between
+protein conformations.
 
@@
 accurate determinations of the many small free energy differences $A \rightarrow 1 \rightarrow 2 \ldots \rightarrow B$,
-and each step requires substantial amounts of sampling.  Third, large errors can accumlate, because the pathway error is
+and each step requires substantial amounts of sampling.  Third, large errors can accumulate, because the pathway error is
 a sum of many errors along the many steps.
@@
 Some of these methods, like the reference system method~\cite{Ytreberg2006}, deactivated morphing~\cite{Park2008}, and
-the confinement method~\cite{Tyka2006,Cecchini2009,Ovchinnikov2013}, take an alternative strategy for computing
-difference free energies that does not require knowing a pathway from $A$ to $B$.
+the confinement method~\cite{Tyka2006,Cecchini2009,Ovchinnikov2013}, take an alternative strategy that does not require knowing a pathway from $A$ to $B$ to compute a difference free energy.
 
-Here, we adopt the confinement method of Tyka et al~\cite{Tyka2006} and Cecchini et al.~\cite{Cecchini2009}, which is
+Here, we adopt the confinement method of Tyka et al~\cite{Tyka2006} and Cecchini et al~\cite{Cecchini2009}, which is
 based on the thermodynamic cycle shown in Figure~\ref{fig:method}. We start with an ensemble of related conformations,
-$A$. We then restrict $A$ to a much ``tighter'' ensemble $A^\ast$ by applying position restraints in a series of MD
+$A$. We then restrict $A$ to a much "tighter'' ensemble $A^\ast$ by applying position restraints in a series of MD
 simulations. Similarly, a second ensemble $B$ is restricted to $B^\ast$. We then compute the free energy between
-$A^\ast$ and $B^\ast$ using either normal mode analysis \cite{Brooks1983,Case1994} or the quasiharmonic method
+$A^\ast$ and $B^\ast$ using either normal mode analysis \cite{Brooks1983,Case1994} or the quasi-harmonic method
 \cite{Karplus1981,Levy1984}. This step takes into account both the remaining conformational entropy in each ensemble and
 the remaining enthalpy difference. In this way, we can calculate the free energy difference between the two end states
-without needing to define a physical path or reaction coordinate connecting the two states. The confinement method
+without defining a physical path or reaction coordinate. The confinement method
 shares some similarities with the `confine-and-release' method for computing ligand binding
@@
 between two completely different folds with only slight changes in sequence. We also show that our computed difference
-free energies are useful in evaluating the quality of CASP target protein predictions.  This could be of significant
-value for ultimately improving the energetics in protein-structure models, not just the structures. And, finally, we
+free energies are useful in evaluating the quality of CASP target protein predictions.  This could ultimately be of significant
+value for improving the energetics in protein-structure models, not just the structures. And, finally, we
 show that we can approximately decompose difference free energies into individual amino acid components. This offers the
-opportunity for diagnosing the structural basis for difference free energies, which can be useful for interpreting
+opportunity for diagnosing the structural basis for difference free energies, which may be useful for interpreting
 biological mechanisms.
@@
 
-First, we verified that our implementation of the confinement method produces results compatible with previous
-calculations reported in the literature.  The method has previously been applied to a 16 amino acid residue
+First, we verified that our implementation of the confinement method produces results compatible with
+calculations previously reported in the literature. The method has been applied to a 16 amino acid residue
 $\beta$-hairpin from protein G, known as BHP~\cite{Cecchini2009}. We calculated the free energy difference between the
@@
 which has a three-stranded $\beta$-sheet). Our confinement calculation shows that bhp1 is more stable by 1.7 kcal/mol,
-which is consistent with 200 $\mu$s equilibrium simulations showing that bhp1 is more favorable configuration by 1.8
-kcal/mol, and in agreement with previous calculations~\cite{Cecchini2009}.
+which is consistent with 200 $\mu$s equilibrium simulations showing that bhp1 is favored by 1.8
+kcal/mol, and also in agreement with previous calculations~\cite{Cecchini2009}.
 
@@
 discriminators can also successfully tell native structures from computer generated models~\cite{Sheffler2009,Zhou2002};
-it is just an independent useful validation that the confinement calculations make sense.
+providing independent validation that the confinement calculations make sense.
 
@@
 fold into remarkably different structures. Orban and co-workers have designed a protein-G-like sequence of 56-residues
-that is marginally stable in one of two possible folds. By mutating key residues in this sequence they are able to
+that is marginally stable in one of two possible folds. By mutating key residues in this sequence, they are able to
 stabilize one fold or the other (see Figure~\ref{fig:orban}). We refer to the $4\beta + \alpha$ structure as the $\beta$
@@
 in sequence, differing in seven positions. Another pair (GA95/GB95) is 95 percent identical, differing in three
-positions. Accurately predicting the structural preferences such similar sequences poses a difficult challenge for
+positions. Accurately predicting the structural preferences of such similar sequences poses a difficult challenge for
 computational methods \cite{Allison2011}.
+
+%The phrase "protein data bank identifier" is a bit confusing.
+I can't tell if it means the abbreviation that identifies the structure, or if you mean the structure.
+If you mean the experimentally determined structure, then I suggest saying "the structure determined by nmr/x-ray crystallography." I suppose that you will want to give the resolution of the experimental structures somewhere...
+%\Arijit{ I have not changed anything here ... may be we can just write (pdb id)}
 
@@
     \caption{The confinement method correctly predicts the structural preferences of four chameleon sequences. The top
-        part of the figure represent four sequences used in this study along with the protein data bank identifier. The
+        part of the figure represents four sequences used in this study along with the protein data bank identifier. The
         experimentally observed fold, computationally predicted fold, and difference free energy between the two folds
@@
 We then used the confinement method to assess the free energy difference between these two structures. In other words,
-we are comparing the experimentally determined GA88/$\alpha$ structure with the computer-generated GA88/$\beta$
-structure. The confinement method was able to predict the conformational preferences correctly for all four sequences
+we compared the experimentally determined GA88/$\alpha$ structure with the computer-generated GA88/$\beta$
+structure. The confinement method predicted the conformational preferences correctly for all four sequences
 (data not shown). This is, however, not surprising as it is often easy to distinguish computational models from native
-structures~\cite{Handl2009,Sheffler2009}. To avoid this potential problem, we instead computed relative free energy of
+structures~\cite{Handl2009,Sheffler2009}. To better test our method, we instead computed the relative free energies of
 two different computer generated models for each sequence. One model is based on the $\alpha$ structure and the other on
-the $\beta$ structure (see Supporting Information for details on the modeling procedure). This is a much more realistic
-test of the confinement method's ability to accurately calculate the difference free energy.
+the $\beta$ (see Supporting Information for details on the modeling procedure). This is a more stringent
+test of the confinement method's accuracy in predicting the difference free energy.
 
-Figure~\ref{fig:orban} shows that the confinement method identifies the correct structure for all four sequences. One
+Figure~\ref{fig:orban} shows that the confinement method identified the correct structure for all four sequences. One
 hypothesis is that such structural transitions require states with diminished stability~\cite{Bryan2010}. It is believed
@@
 position 45 (Tyr for $\beta$ and Leu for $\alpha$) was found to be important for switching between $\alpha$ and $\beta$
-conformation. This inspired us to introduce another mutation at this position, Y45A, which we refer to as GA98. Our
-calculations predict that this mutation shifts the equilibrium to the $\alpha$ conformation, which is now more stable
-than the $\beta$ conformation by 3.8 kcal/mol. Although this result has not yet been confirmed experimentally, it is
+conformations. This inspired us to introduce another mutation at this position, Y45A, which we refer to as GA98. Our
+calculations predicted that this mutation shifts the equilibrium to the $\alpha$ conformation, which is now more stable
+than the $\beta$ by 3.8 kcal/mol. Although this result has not yet been confirmed experimentally, it is
 consistent with the previously observed effect of Y45L.
 
-\subsection*{Per-residue free energy calculations can identify the mechanistic detail behind conformational preferences}
+\subsection*{Per-residue free energy calculations can identify mechanistic details behind conformational preferences}
 
-To better understand the mechanism behind these chameleon proteins, we approximately decomposed~\cite{Tyka2006} the free
-energy into per-residue contributions. We decompose each confinement step ($\Delta G_{A,A^\ast}$, $\Delta G_{B,B^\ast}$)
-into per-residue contributions. We can also decompose the remaining enthalpy in the confined ensemble into per-residue
+To better understand the switching mechanism of these chameleon proteins, we approximately decomposed~\cite{Tyka2006} the free
+energy into per-residue contributions to each confinement step ($\Delta G_{A,A^\ast}$, $\Delta G_{B,B^\ast}$). 
+We can also decompose the remaining enthalpy in the confined ensemble into per-residue
 contributions. However, the method is approximate because we do not include the residual conformational entropy from the
-normal mode or quasiharmonic analysis steps. Computing per-residue contributions helps us to identify important residues
+normal mode or quasi-harmonic analysis steps. Computing per-residue contributions helps us to identify important residues
 that stabilize a particular conformation. The per-residue difference free energy, $\Delta \Delta G (\beta - \alpha)$, is
@@
     \caption{Per-residue difference free energies for the $\alpha$ and $\beta$ conformations of the GA95 and GB95
-        sequences. Residues colored in blue favor the $\alpha$ structure, residues in red favor the $\beta$ structure,
+        sequences. Residues colored in blue favor the $\alpha$ structure; residues in red favor the $\beta$ structure;
         and white residues have no preference.}
@@
 residues can show marked preferences for being in either $\alpha$ or $\beta$ conformation. Such differences can be
-understood by looking in detail at the local environment for those residues. For example, the region around residue 7
+understood by looking at the details of the local environments for those residues. For example, the region around residue 7
 forms a random coil in the $\alpha$ structure, whereas it forms a beta-sheet structure in the $\beta$ structure. These
-residues strongly favour the locally well-packed and hydrogen bonded environment found in the $\beta$ sheet, but in the
+residues strongly favor the locally well-packed and hydrogen bonded environment found in the $\beta$ sheet, but in the
 alpha conformation there are no other residues to form a beta sheet with. The global conformation of the protein has
-``overridden'' the local preference of these residues. Overall, favoring the $\alpha$ or $\beta$ structure is a delicate
-balance, where the global free energy difference is small and the contributions of different per-residue tendencies
-largely cancel out. For these chameleon proteins, it is therefore very likely that by changing key residue the global
-fold preferences can be changed.
+"overridden'' the local preference of these residues. Overall, favoring the $\alpha$ or $\beta$ structure is a delicate
+balance, where the global free energy difference is small, and the contributions of different per-residue tendencies
+largely cancel out. For these chameleon proteins, it is therefore very likely that the global
+fold preferences can be changed by changing key residues.
 
@@
 these residues (F30 and Y45) favor the $\beta$ structure, because they have large solvent exposed surface areas in the
-$\alpha$ structure, but are more buried in the $\beta$ structure. Additionally, Y45 forms a hydrogen bond with D47 in
+$\alpha$ structure, while they are more buried in the $\beta$ structure. Additionally, Y45 forms a hydrogen bond with D47 in
 the $\beta$ structure. On the other hand, residue A20 from GB95 still favors the $\alpha$ structure, although less
@@
 folds, whereas residues 1--8 and 52--56 are unstructured in the $\alpha$ fold, but form $\beta$-strands in the $\beta$
-fold. Most of the amino acid residues in the region 1--9 have negative per-residue free energies, which means that these
+fold. Most of the amino acid residues in the 
+%structured?
+%\Arijit {yes}
+region 1--8 have negative per-residue free energies, which means that these
 residues favor the $\beta$ structure. Roles of some other important residues in stabilizing either the $\alpha$ or
@@
 
-We have tested the ability of the confinement method to act as a ``meta-predictor'' for structure prediction. Here, the
-task is to correctly identify the most accurate models out of a set of ``decoys'' generated by different methods during
-the CASP experiment. CASP is a blind test in which different groups apply methods to predict the 3-dimensional structure
-of proteins from their sequences. Each group is allowed to submit five possible structures, which they are supposed to
-rank from best to worst.
+Having tested the ability of the confinement method to act as a "meta-predictor'' for structure prediction, the next
+task was to correctly identify the most accurate models out of a set of "decoys'' generated by different methods during
+the CASP experiment. CASP is a blind test in which different groups apply computational methods to predict the 3-dimensional structures
+of proteins from their sequences. Each group is allowed to submit five possible structures, ranked by that group from best to worst.
 
-We have performed two experiments centered on CASP. In the first experiment, we tried to rank-order predictions for
-several targets. For each target, the predictions were either produced by a single group---presumably using the same
-method for each prediction, or were produced by several different groups---using different methods. The goal of this
-experiment is to determine if the physics-based confinement method can correctly identify more native-like structures as
-having lower free energies. Our second experiment was to see if the confinement method can identify structures that are
+We performed two experiments based on CASP. In the first experiment, we tried to rank-order predictions for
+several targets. For each target, the predictions were either produced by a single group—presumably using the same
+method for each prediction, or were produced by several different groups—using different methods. The goal of this
+experiment was to determine if the physics-based confinement method could correctly identify more native-like structures as
+having lower free energies. Our second experiment was to see if the confinement method could identify structures that are
 missed by other meta-prediction servers. Most successful meta-prediction servers are based on the idea of consensus: if
@@
 result that is only predicted by one method. We chose several cases where such structures were missed by the best
-meta-predictors in CASP and assessed if the confinement method can correctly identify these accurate models. All of our
-calculations were performed after the CASP experiment; they are not blind predictions.
+meta-predictors in CASP and assessed if the confinement method could correctly identify these accurate models. All of our
+calculations were performed after the CASP experiment; they were not blind predictions.
 
@@
 
-First, we examined the ability of the confinement method to rank different models for a target that has been generated
+First, we examined the ability of the confinement method to rank different models  generated
 by a single group using the same methodology for all predictions. We examined two targets: T0559 and T0560 (see
@@
 
-The first test case is target T0559. The best predictor group for this 69 amino acid target was ``BAKER-ROSETTASERVER''.
+The first test case is target T0559. The best predictor group for this 69 amino acid target was "BAKER-ROSETTASERVER''.
 To save computer time, we excluded two models that were very similar to other models that we did include. For this
-target, the difference free energy computed by the confinement method can be used to accurately rank-order all of the
+target, the difference free energy computed by the confinement method could be used to accurately rank-order all of the
 models and the native experimental structure (Figure~\ref{fig:T0559}).
@@
     \caption{The confinement method can correctly rank the native structure and three predictions produced by a single
-        method for target T0559. The backbone region of the predicted structures which differ substantially compare to the
-        experimental structure upon superposition are colored brown.}
+        method for target T0559. The backbone regions of the predicted structures that differ substantially from the
+        experimental structures upon superposition are colored brown.}
 \label{fig:T0559}
@@
 
-We performed a similar calculation for target T0560 with two models from the group ``Splicer''. The remaining three
+We performed a similar calculation for target T0560 with two models from the group "Splicer''. The remaining three
 models were discarded as they were too similar to the rest of the models. Again, we can correctly identify the native
-sate and our calculated free energy based ranking matches well with GDT-TS (Figure S2 in the supporting information).
+state, and our calculated free energy based ranking matches well with GDT-TS (Figure S2 in the supporting information).
 
-Next, we tested the ability of the confinement method to rank models for target T0540 that were produced by different
-prediction groups. The top models from groups ``LTB'' (Model 1) and ``Mufold'' (Model 2) were chosen for analysis. Once
+Next, we tested the ability of the confinement method to rank the models for target T0540 that were produced by different
+prediction groups. The top models from groups "LTB'' (Model 1) and "Mufold'' (Model 2) were chosen for analysis. Once
 again, the free energy based ranking correlates well with the GDT-TS based score (Figure~\ref{fig:T0540}).
@@
 
-In a previous section, we discussed how the per-residue free energy can reveal mechanistic detail behind the
-conformational preference of chameleon sequences. In this section, our aim is to apply the same method and try to
-understand if the per-residue free energy can help us identify problematic residues in structure predictions. We chose
+We discussed how the per-residue free energy can reveal mechanistic detail behind the
+conformational preference of chameleon sequences. Here, we applied the same method to
+see if the per-residue free energy can help us identify problematic residues in structure predictions. We chose
 CASP target T0569 and compared the experimental NMR structure with the best predicted model (GDT-TS=78; predicted by the
-``Mufold'' group). The confinement method predicts that the experimental structure is more stable by 20 kcal/mol (Figure
+"Mufold'' group). The confinement method predicts that the experimental structure is more stable by 20 kcal/mol (Figure
 S2).
-
-The confinement method clearly identifies two hydrophobic residues V59 and I61, which destabilize the predicted model
-with respect to the experimental structure (Figure~\ref{fig:T0569_per_residue} and Figure S3). The side chains of these
-hydrophobic residues are oriented towards the protein hydrophobic core in the native NMR structure but oriented towards
-the exterior of the protein and are solvent exposed in the model. These residues are part of a beta-sheet in the
-experimental structure, but because of their side chain orientation the corresponding beta-sheet becomes disordered in
+%The annoying rule is that "side chain" is correct for the noun usage, and side-chain for the adjective.
+The confinement method clearly identified the two hydrophobic residues V59 and I61 as destabilizing the predicted model
+relative to the experimental structure (Figure~\ref{fig:T0569_per_residue} and Figure S3). The side chains of these
+hydrophobic residues are oriented towards the protein hydrophobic core in the native NMR structure but are oriented towards
+the exterior of the protein, exposing them to solvent, in the model. These residues are part of a beta-sheet in the
+experimental structure, but because of their side-chain orientation, the corresponding beta-sheet becomes disordered in
 the predicted model (Figure~\ref{fig:T0569_per_residue} and Figure S4). There is also a large difference around K76,
 which forms a salt-bridge with D11 in the predicted model, but not in the experimental structure. This suggests that
-salt-bridge interactions are too favorable for the combination of force field and implicit solvent model we use, which
+salt-bridge interactions are too favorable for the combination of forcefield and implicit solvent model we use, which
 has been a problem noted in the past \cite{Roe2007}.
@@
 
-Despite success for most of the studied systems, there are few failures, specially when the GDT-TS scores of the
-compared structures are very close. One example is Target T0538, where we compared the experimental structure with three
-models (Model 1: ``PconsR''---GDT-TS=96; Model 2: ``Shell''---GDT-TS=90; Model 3: ``FOLDIT''---GDT-TS=86). Contrary to
+Despite success for most of the systems we studied, there were few failures, especially when the GDT-TS scores of the
+compared structures were very close. One example is Target T0538, where we compared the experimental structure with three
+models (Model 1: "PconsR''---GDT-TS=96; Model 2: "Shell''---GDT-TS=90; Model 3: "FOLDIT''---GDT-TS=86). Contrary to
 our expectation, the confinement method predicts that computer generated Model 1 is more stable than the crystal
@@
 \end{center}
-\caption{Small differences in sidechain orientation can cause large differences in free energy. We compare the
+\caption{Small differences in side-chain orientation can cause large differences in free energy. We compare the
     experimental structure and a computer generated model for CASP target T0538. (A) Two salt bridges between residues
@@
 
-\subsection*{Can the confinement method perform better quality assessment in protein structure prediction?}
+\subsection*{Can the confinement method improve the assessment of quality in protein structure prediction?}
 
 A part of the CASP experiment is dedicated to the quality assessment (QA) of predicted models \cite{Kryshtafovych2011}.
-Here, predictors were asked to produce an overall score (called QMODE1) for each model on a scale from 0 to 1, with
+Predictors are asked to produce an overall score (called QMODE1) for each model on a scale from 0 to 1, with
 higher values corresponding to better models \cite{Kryshtafovych2011}. Many of the groups use consensus strategies in
 such experiments. Here, we chose two computer generated models from CASP target T0538, where the top performing group
-``MUFOLD-WQA'' \cite{Wang2011} failed to identify the best model. We examined two models, one from ``PconsR'' (GDT-TS =
-96), which MUFOLD-WQA gave a QMODE1 = 0.54. The second model was from ``MULTICOM-NOVEL'' (GDT-TS=83), which is a poorer
-model, but MUFOLD-WQA assigned a higher QMODE1 = 0.59. The confinement method gives a difference free energy that favors
-the PconsR model by 3.9 kcal/mol, which correctly identifies the more accurate model. Although consensus prediction
-strategies are very effective, they can miss a good prediction that is only predicted by a few methods. At least in this
-case, the confinement method is able to capture this structure that was otherwise missed.
+"MUFOLD-WQA'' \cite{Wang2011} failed to identify the best model. We examined two models, one from "PconsR'' (GDT-TS =
+96), which MUFOLD-WQA gave a QMODE1 = 0.54. The second model was from "MULTICOM-NOVEL'' (GDT-TS=83), which is a poorer
+model, but MUFOLD-WQA assigned a higher QMODE1 = 0.59. The confinement method gave a difference free energy that favored
+the PconsR model by 3.9 kcal/mol, which correctly identified the more accurate model. Although consensus prediction
+strategies are very effective, they can miss a good prediction if it is predicted by only a few methods. At least in this
+case, the confinement method was able to capture a structure that was otherwise missed.
 
@@
 We have described a computational method called the confinement method for computing the difference free energy between
-two conformational ensembles. We show that the difference free energy can be calculated on ~100 residue sized proteins,
-even for large conformational changes. We have demonstrated that it can discriminate the folding preferences of a series
-of chameleon proteins. We show that the confinement method can discriminate between the native structure and structure
-predictions and can identify the best prediction reliably. Perhaps most importantly, we have shown that it can be used
-to give residue-level insights into what are the dominant structural factors in a protein that are responsible for the
-difference free energies. The confinement should be useful for protein design, structure prediction, and understanding
+two conformational ensembles. We showed that the difference free energy can be calculated on proteins of ~100 residues,
+even for large conformational changes; that it can discriminate the folding preferences of a series
+of chameleon proteins; it can discriminate between the native structure and structure
+predictions, and can identify the best prediction reliably. Perhaps most importantly, we have shown that it can be used
+to give residue-level insights into the dominant structural factors that are responsible for the
+difference free energies in conformations of a protein. The confinement method should be useful for protein design, structure prediction, and understanding
 the mechanism of conformational change.
@@
 \item  In the first step, minimizations of A and B conformations are performed. These minimized conformations
-    ($A^{\ast}$ and $B^{\ast}$) are the reference conformation of that state.
-
+    ($A^{\ast}$ and $B^{\ast}$) are the reference conformations.
+% I changed et. al. to et al for consistency.
 \item The free energy of confining the ensemble (A or B) to a tighter ensemble ($A^\ast$ or $B^\ast$) is calculated.
@@
     dynamics simulation (each 20 ns long) for each leg of the thermodynamic cycle, where the harmonic restraint force
-    constant was scaled from 0.00005 kcal/mol (mostly free) to 81.92 kcal/mol (tightly restrained). The free energy for
-    this step is estimated using a thermodynamic integration approach developed by Tyka et. al. \cite{Tyka2006}. The
+    constant is scaled from 0.00005 kcal/mol (mostly free) to 81.92 kcal/mol (tightly restrained). The free energy for
+    this step is estimated using a thermodynamic integration approach developed by Tyka et al \cite{Tyka2006}. The
     confinement free energies calculated in this step are denoted as $\Delta G_{A,A^\ast}$ and $\Delta G_{B,B^\ast}$ in
@@
 \item The thermodynamic cycle is closed by calculating the free energy between the final restrained states $A^\ast$ and
-    $B^\ast$ using normal mode analysis \cite{Brooks1983,Case1994} or quasiharmonic analysis
+    $B^\ast$ using normal mode analysis \cite{Brooks1983,Case1994} or quasi-harmonic analysis
     \cite{Karplus1981,Levy1984}. The free energy calculated in this way is shown as $\Delta G_{A^\ast,B^\ast}$ in
@@
 
-One advantage of confinement method is that none of the simulations during the restraining step depends on each other.
-Therefore, with available computer resources it can be fast to compute. We ran each confinement calculation on a single
+One advantage of the confinement method is that none of the simulations during the restraining step depends on any other.
+Therefore, it can be fast to compute with available computer resources. We ran each confinement calculation on a single
 graphics processing unit (GPU). For a 56-residue protein, this leads to a calculation time of only 4 hours on 40 GPUs (1
 GPU per confinement calculation x 20 calculations per structure x 2 structures). All calculations were performed with
-the Amber 11 suit of programs~\cite{Case2012,Goetz2012} in combination with the ff99SB forcefield~\cite{Hornak2006} and
+the Amber 11 suite of programs~\cite{Case2012,Goetz2012} in combination with the ff99SB forcefield~\cite{Hornak2006} and
 the GBneck generalized born implicit solvent model \cite{Mongan2006}.
@@
 We calculated the approximate per-residue free energy as follows. The confinement energy, $\Delta G_{A,A*}$ and $\Delta
-G_{B,B*}$ of each residue is calculated in the usual numerical way as described by Tyka et al. \cite{Tyka2006}. The
-internal energy of each residue is calculated using amber's "decomp" module using the final restrained trajectory. We
-call this method approximate as we ignore the entropic contribution from the normal mode or quasiharmonic analysis.
-However, this contribution to the total free energy is much smaller, which allow us to study the mechanistic details of
-conformational preference of each residue.
+G_{B,B*}$ of each residue was calculated numerically as described by Tyka et al. \cite{Tyka2006}. The
+internal energy of each residue was calculated with Amber's "decomp" module using the final restrained trajectory. We
+call this method approximate as we ignore the entropic contribution from the normal mode or quasi-harmonic analysis.
+However, the entropic 
+% the entropic
+%\Arijit{ I have inserted "the entropic"}
+contribution to the total free energy is much smaller, which allow us to study the details of
+the conformational preferences of each residue.
 
 \begin{thebibliography}{99}
+%You may want to check your styles for creating the references against the style of the Journal to which you are submitting the paper.
+Also, you may want to check your abbreviations for some of the journal titles against a list such as the one at: http://library.caltech.edu/reference/abbreviations/.
+E.g.
+J.
+Chem.
+Phys.
+or J Chem Phys...Proc.
+Natl.
+Acad.
+Sci.
+U.S.A.
+or Proc.
+Natl.
+Acad.
+Sci.
 
 \bibitem{Elber2007}
-Elber, R. A Milestoning Study of the Kinetics of an Allosteric Transition: Atomically Detailed Simulations of Deoxy Scapharca
+Elber, R. A. Milestoning Study of the Kinetics of an Allosteric Transition: Atomically Detailed Simulations of Deoxy Scapharca
 Hemoglobin. Biophysical J., 2007, 92, 85-87.
@@
 \bibitem{Jonsson1998}
-J\'{on}sson, H.; Mills, G.; Jacobsen, K.W. Nudged Elastic Band Method for Finding Minimum Energy Paths of Transitions,
+J\'{on}sson, H.; Mills, G.; Jacobsen, K.W. Nudged Elastic Band Method for Finding Minimum Energy Paths of Transitions
 in Classical and Quantum Dynamics in Condensed Phase Simulations, Ed. B. J. Berne, G. Ciccotti and D. F.
@@
 \bibitem{E2007}
+%Is E, W.; an error?
+%\Arijit{E is the last name}
 E, W.; Ren, W.; Vanden-Eijnden, E. Simplified and improved string method for computing the minimum energy paths in
@@
 \bibitem{Elber2005}
-Elber, R. Long-timescale simulation methods. Cur. Opi. in Str. Biol. 2005, 15, 151-156.
+Elber, R. Long-timescale simulation methods. Cur. Opin. in Str. Biol. 2005, 15, 151-156.
 
@@
 Calculations: Theory and Applications in Chemistry and Biology. Springer Series in Chemical
-Physics, vol. 86. Berlin and Heidelberg: Springer; 2007, p. 1–32.
+Physics, vol. 86. Berlin and Heidelberg: Springer; 2007, p. 1-32.
 
@@
 \bibitem{Zheng2008}
-Zheng, L.; Chen, M.; Yang, W. Random walk in orthogonal space to achieve efficient free-energy simulation of complex systems, Proc. Natl. Acad. Sci. 200
-8, 105 (51), 20227.
+Zheng, L.; Chen, M.; Yang, W. Random walk in orthogonal space to achieve efficient free-energy simulation of complex systems, Proc. Natl. Acad. Sci. 2008, 105 (51), 20227.
 
@@
 
-\bibitem{Spichty2010}Spichty, M.; Cecchini, M.; Karplus, M. Conformational Free-Energy Difference of a Miniprotein from Nonequilibrium
+\bibitem{Spichty2010}
+Spichty, M.; Cecchini, M.; Karplus, M. Conformational Free-Energy Difference of a Miniprotein from Nonequilibrium
 Simulations. J. Phys. Chem. Lett., 2010, 1, 1922-1926.
@@
 \bibitem{Strajbl2000}
-Strajbl, M.; Sham, Y.Y.; Villà, J.; Chu, Z.-T.; Warshel, A. Calculations of Activation Entropies of Chemical Reactions
+Strajbl, M.; Sham, Y.Y.; Vill√†, J.; Chu, Z.-T.; Warshel, A. Calculations of Activation Entropies of Chemical Reactions
 in Solution. (2000) 104, 4578-4584.
@@
 \bibitem{Case1994}
-Case, D. Normal-mode analysis of protein dynamics Curr. Opin. Struct. Biol. 1994, 4, 285-290.
+Case, D. Normal-mode analysis of protein dynamics. Curr. Opin. Struct. Biol. 1994, 4, 285-290.
 
 \bibitem{Karplus1981}
-Karplus, M.; Kushick, J. Method for estimating the configurational entropy of macromolecules Macromolecules. 1981, 14, 325-332.
+Karplus, M.; Kushick, J. Method for estimating the configurational entropy of macromolecules. Macromolecules. 1981, 14, 325-332.
 
@@
 Levy, R.; Karplus, M.; Kushick, J.; Perahia, D. Evaluation of the configurational entropy for proteins: application to molecular dynamics simulations of 
-an $\alpha$-helix Macromolecules 1984, 17, 1370-1374.
+an $\alpha$-helix Macromolecules. 1984, 17, 1370-1374.
 
@@
 Mobley, D.L.; Chodera, J.D.; Dill, K.A. The combining and release method: obtaining correct binding free energies in
-the presence of protein conformational change. Journal of Chemical Theory and Computation 2007, 3, 1231-1235.
+the presence of protein conformational change. Journal of Chemical Theory and Computation. 2007, 3, 1231-1235.
 
@@
 Mobley, D.L.; Chodera, J.D.; Dill, K.A. On the use of orientational restraints and symmetry corrections in alchemical
-free energy calculations. J. Chem. Phy. 2006, 125, 084902.
+free energy calculations. J. Chem. Phys. 2006, 125, 084902.
 
 \bibitem{Alexander2007}
-Alexander, P.A.; He, Y.; Chen, Y.; Orban, J. Bryan, P. The design and characterization of two proteins with $88 \%$ sequence identity but differentstructure and 
+Alexander, P.A.; He, Y.; Chen, Y.; Orban, J. Bryan, P. The design and characterization of two proteins with $88 \%$ sequence identity but different structure and 
 function. Proc. Natl. Acad. Sci. 2007, 104, 11963-11968.
@@
 \bibitem{Kryshtafovych2011}
-Kryshtafovych, A.; Fidelis, K; and Tramontano, A. Evaluation of model quality predictions in CASP9. Proteins, 2011, 79, 91–106
+Kryshtafovych, A.; Fidelis, K; and Tramontano, A. Evaluation of model quality predictions in CASP9. Proteins, 2011, 79, 91-106
 
@@
 \bibitem{Zemla2003}
-Zemla, A. LGA: a method for finding 3D similarities in protein structures. Nucleic Acids Res 2003, 31, 3370–3374.
+Zemla, A. LGA: a method for finding 3D similarities in protein structures. Nucleic Acids Res 2003, 31, 3370-3374.
 
+%I think that Case2012 is the only bibitem where you have the initials before the last name; and the date in parentheses.
+%\Arijit{Done it .... but not sure whether it was required to change}
 \bibitem{Case2012}
-D.A.
-Case, T.A.
-Darden, T.E.
-Cheatham, III, C.L.
-Simmerling, J.
-Wang, R.E.
-Duke, R.
-Luo, R.C.
-Walker, W.
-Zhang, K.M.
-Merz, B.
-Roberts, S.
-Hayik, A.
-Roitberg, G.
-Seabra, J.
-Swails, A.W.
-Goetz, I.
-Kolossváry, K.F.
-Wong, F.
-Paesani, J.
-Vanicek, R.M.
-Wolf, J.
-Liu, X.
-Wu, S.R.
-Brozell, T.
-Steinbrecher, H.
-Gohlke, Q.
-Cai, X.
-Ye, J.
-Wang, M.-J.
-Hsieh, G.
-Cui, D.R.
-Roe, D.H.
-Mathews, M.G.
-Seetin, R.
-Salomon-Ferrer, C.
-Sagui, V.
-Babin, T.
-Luchko, S.
-Gusarov, A.
-Kovalenko, and P.A.
-Kollman (2012), AMBER 12, University of California, San Francisco.
+Case, D.A.; Darden, T.A.; Cheatham, III, T.E.; Simmerling, C.L.; Wang, J.; Duke, R.E.; Luo, R.; Walker, R.C.; Zhang, W.; Merz, K.M.; Roberts, B.; Hayik, S.; 
+Roitberg, A.; Seabra, J.; Swails, A.W.; Goetz, A.W.; KolossvA°ry, I.; Wong, K.F.; Paesani, F.; Vanicek, J.; Wolf, R.M.; Liu, J.; Wu, X.; Brozell, S.R.; Steinbrecher, T.;
+Gohlke, H.; Cai, Q.; Ye, X.; Wang, J.; Hsieh, M.-J.; Cui, G.;  Roe, D.H.; Mathews, D.H.; Seetin, M.G.; Salomon-Ferrer, R.; Sagui, C.; 
+Babin, V.; Luchko, T.; Gusarov, S.; Kovalenko, A.; Kollman, P.A. 2012, Amber12, University of California, San Francisco.
 
@@
 Hornak, V.; Abel, R.; Okur, A.; Strockbine, B.; Roitberg, A.; Simmerling, C. Comparison of multiple Amber force fields
-and development of improved protein backbone parameters. Proteins., 2006, 65, 712-725.
+and development of improved protein backbone parameters. Proteins. 2006, 65, 712-725.
 
@@
 Mongan, J.; Simmerling, C.; A. McCammon, J.; A. Case, D.; Onufriev, A. Generalized
-Born with a simple, robust molecular volume correction. J. Chem. Theory Comput.,
+Born with a simple, robust molecular volume correction. J. Chem. Theory Comput.
 2006, 3, 156-169.

### 2013-05-16 Paper/paper.tex
May 16_3

@@
 We calculated the approximate per-residue free energy as follows. The confinement energy, $\Delta G_{A,A*}$ and $\Delta
-G_{B,B*}$ of each residue is calculated in the usual numerical way as described by Tyka et al. \cite{Tyka2006}. The
-internal energy of each residue is calculated using amber's "decomp" module using the final restrained trajectory. We
-call this method approximate as we ignore the entropic contribution from the normal mode or quasiharmonic analysis.
-However, this contribution to the total free energy is much smaller, which allow us to study the mechanistic details of
-conformational preference of each residue.
+G_{B,B*}$ of each residue was calculated numerically as described by Tyka et al. \cite{Tyka2006}. The
+internal energy of each residue was calculated with Amber's "decomp" module using the final restrained trajectory. We
+call this method approximate as we ignore the entropic contribution from the normal mode or quasi-harmonic analysis 
+at the highly restrained state. However, the entropic 
+% the entropic
+%\Arijit{ I have inserted "the entropic"}
+contribution to the total free energy is much smaller, which allow us to study the details of
+the conformational preferences of each residue.
 
 \begin{thebibliography}{99}
+%You may want to check your styles for creating the references against the style of the Journal to which you are submitting the paper.
+Also, you may want to check your abbreviations for some of the journal titles against a list such as the one at: http://library.caltech.edu/reference/abbreviations/.
+E.g.
+J.
+Chem.
+Phys.
+or J Chem Phys...Proc.
+Natl.
+Acad.
+Sci.
+U.S.A.
+or Proc.
+Natl.
+Acad.
+Sci.
 
 \bibitem{Elber2007}
-Elber, R. A Milestoning Study of the Kinetics of an Allosteric Transition: Atomically Detailed Simulations of Deoxy Scapharca
+Elber, R. A. Milestoning Study of the Kinetics of an Allosteric Transition: Atomically Detailed Simulations of Deoxy Scapharca
 Hemoglobin. Biophysical J., 2007, 92, 85-87.

### 2013-06-17 Paper/paper.tex
correction_June17

@@
 residues favor the $\beta$ structure. Roles of some other important residues in stabilizing either the $\alpha$ or
-$\beta$ conformation are summarized in Figure S1.
+$\beta$ conformation are summarized in Figure S1 and S2-S7.
 
@@
 models were discarded as they were too similar to the rest of the models. Again, we can correctly identify the native
-state, and our calculated free energy based ranking matches well with GDT-TS (Figure S2 in the supporting information).
+state, and our calculated free energy based ranking matches well with GDT-TS (Figure S8 in the supporting information).
 
@@
 The confinement method clearly identified the two hydrophobic residues V59 and I61 as destabilizing the predicted model
-relative to the experimental structure (Figure~\ref{fig:T0569_per_residue} and Figure S3). The side chains of these
+relative to the experimental structure (Figure~\ref{fig:T0569_per_residue} and Figure S9). The side chains of these
 hydrophobic residues are oriented towards the protein hydrophobic core in the native NMR structure but are oriented towards
@@
 experimental structure, but because of their side-chain orientation, the corresponding beta-sheet becomes disordered in
-the predicted model (Figure~\ref{fig:T0569_per_residue} and Figure S4). There is also a large difference around K76,
+the predicted model (Figure~\ref{fig:T0569_per_residue} and Figure S10). There is also a large difference around K76,
 which forms a salt-bridge with D11 in the predicted model, but not in the experimental structure. This suggests that
@@
 our expectation, the confinement method predicts that computer generated Model 1 is more stable than the crystal
-structure (Figure S5). Per-residue free energy calculations (not shown) show that despite only small variations at the
+structure (Figure S11). Per-residue free energy calculations (not shown) show that despite only small variations at the
 backbone level, the side chains are oriented in very different ways (Figure~\ref{fig:T0538compare}), giving rise to
@@
 \begin{center}
-\includegraphics[width=4.9 in,height=4.0 in]{Target_538_compare.pdf}
+\includegraphics[width=4.8in]{Target_538_compare.pdf}
 \end{center}
 \caption{Small differences in side-chain orientation can cause large differences in free energy. We compare the
-    experimental structure and a computer generated model for CASP target T0538. (A) Two salt bridges between residues
+    experimental structure and a computer generated model for CASP target T0538. Two salt bridges between residues
     E35-R32 and E28-K24 exist in the native structure. These are absent in the computer generated model structure.
-    Instead, (B) a new salt bridge is formed between residues R32-E28 in model. (C) Another salt bridge between residues
-    R26-E50 in the experimental structure. (D) In comparison to the native structure, orientation of F51 differs in the
-    model.}
+    Instead, a new salt bridge is formed between residues R32-E28 in model. } 
 \label{fig:T0538compare}
